### 2015-09-29 manuscript/cpcrw_manuscript.Rmd
Manuscript started

---
title: "cpcrw_ms"
author: "Ben Bond-Lamberty"
date: "September 29, 2015"
output: html_document
---

#### Abstract

Abstract.

#### Introduction

Introduction.

#### Methods

CPCRW text.

Thirty-nine soil cores, 30 cm high by TODO cm wide, were taken using a soil augur on 3-5 August 2015. We sampled from the bottom (within 0-2 cm) of the active layer, which averaged 80 cm depth. Sample points were randomly located in the transition zone described above, and separated by 2-5 m. The soil varied between TODO and TODO, and on averaged had ~20 cm of organic material over the mineral soil. Cores were kept cool in the field before being packed in dry ice and shipped to the lab in Richland, WA.

In the lab, the soil cores were stored in a TODO &deg;C cooler for several weeks. At that point (late August 2015), three fragmented or otherwise damaged cores were discarded. Six random cores were removed and destructively analyzed for TODO. The others were randomly assigned to one of five groups (each N=6). These groups included two incubation temperatures (4 &deg;C and 22 &deg;C) each with two moisture treatments: one in which soil moisture was maintained at field conditions, and a drought treatment in which no water was added. The fifth group was a 22 &deg;C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 &deg;C "drought" group.

The incubation experiment began on 31 August 2015.

Analysis.

#### Results

Results.

#### Discussion

Discussion.

### Acknowledgments

Acknowledgments.

### 2015-09-29 manuscript/cpcrw_manuscript.Rmd
Manuscript: methods

@@
 
+Boreal forests, permafrost, importance of active layer.
+
+Moisture likely to be a strong control of C oxidation. Which form? Importance of CO2 / CH4 balance.
+
 #### Methods
 
-CPCRW text.
+The field component of this research took place in Caribou-Poker Creeks Research Watershed (CPCRW), part of the Bonanza Creek LTER (TODO: URL).
+Annual average air temperature was 􏰀3.3°C, and the average temperature in July was 16.4°C, which was the maximum in this watershed.
+Annual average precipitation was 269 mm, and the average precipitation of 47 mm in August was the highest in this area.
 
-Thirty-nine soil cores, 30 cm high by TODO cm wide, were taken using a soil augur on 3-5 August 2015.
+SE slope, transition zone
+
+of this study site was more than TODO years according to the tree rings at stem base.
+Stand density of dominant Picea mariana and glauca trees (>1.3 m in height) was TODO trees ha^-1^, with a basal area of TODO m^2^ ha^-1^.
+
+Thirty-nine soil cores, 30 cm high by 7.5 cm wide, were taken using a soil augur on 3-5 August 2015.
 We sampled from the bottom (within 0-2 cm) of the active layer, which averaged 80 cm depth.
 Sample points were randomly located in the transition zone described above, and separated by 2-5 m.
-The soil varied between TODO and TODO, and on averaged had ~20 cm of organic material over the mineral soil.
+The soil was a TODO, and on average had ~20 cm of organic material over the mineral soil.
 Cores were kept cool in the field before being packed in dry ice and shipped to the lab in Richland, WA.
@@
 The others were randomly assigned to one of five groups (each N=6).
-These groups included two incubation temperatures (4 &deg;C and 22 &deg;C) each with two moisture treatments: one in which soil moisture was maintained at field conditions, and a drought treatment in which no water was added.
+These groups included two incubation temperatures (4 &deg;C and 22 &deg;C) each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by mass), and a drought treatment in which no water was added.
 The fifth group was a 22 &deg;C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 &deg;C "drought" group.
@@
 The incubation experiment began on 31 August 2015.
+Cores were put into one of two growth chambers (Conviron BDW80, Winnipeg, Canada) maintaining the 4 and 22 &deg;C temperatures.
+TODO: day/night cycle in chambers?
+We measured the cores' mass and GHG (CH~4~ and CO~2~) evolution three times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation.
+
+The cores were mounted on 0.05-MPa ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action.
+We placed the pore plates in contact with water when necessary to maintain the field moisture treatment groups at ~28% moisture, and to moderate the drying of the  22 &deg;C "controlled drought" group.
+
+For each measurement, a six-core treatment group was connected to a TODO multiplexer that was in turn connected to a Picarro TODO.
+Dry GHG concentration were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group.
+Ambient air was measured between treatment groups, and before starting measurements in a chamber.
 

### 2015-09-30 manuscript/cpcrw_manuscript.Rmd
Update cpcrw_manuscript.Rmd
Based on 9/29/15 email from Peyton.

--- 2015-09-29 manuscript/cpcrw_manuscript.Rmd
@@
 
+Boreal forests, permafrost, importance of active layer.
+
+Moisture likely to be a strong control of C oxidation. Which form? Importance of CO2 / CH4 balance.
+
 #### Methods
 
-CPCRW text.
+The field component of this research took place in Caribou-Poker Creeks Research Watershed (CPCRW), part of the Bonanza Creek LTER (http://www.lter.uaf.edu/default.cfm).
+Annual average air temperature was 􏰀3.3°C, and the average temperature in July was 16.4°C, which was the maximum in this watershed.
+Annual average precipitation was 269 mm, and the average precipitation of 47 mm in August was the highest in this area.
 
-Thirty-nine soil cores, 30 cm high by TODO cm wide, were taken using a soil augur on 3-5 August 2015.
+SE slope, transition zone
+
+of this study site was more than TODO years according to the tree rings at stem base.
+Stand density of dominant Picea mariana and glauca trees (>1.3 m in height) was TODO trees ha^-1^, with a basal area of TODO m^2^ ha^-1^.
+
+Thirty-nine soil cores, each 30 cm high by 7.5 cm wide, were taken using a soil recovery augur (AMS Inc., American Falls, ID) on 3-5 August 2015.
 We sampled from the bottom (within 0-2 cm) of the active layer, which averaged 80 cm depth.
 Sample points were randomly located in the transition zone described above, and separated by 2-5 m.
-The soil varied between TODO and TODO, and on averaged had ~20 cm of organic material over the mineral soil.
-Cores were kept cool in the field before being packed in dry ice and shipped to the lab in Richland, WA.
+The soil is characterized as a poorly-drained silt loam, and on average had ~20 cm of organic material over the mineral soil.
+Cores were kept cool in the field before being packed in dry ice and shipped to the lab in Richland, WA within 48-72 hours of collection.
 
-In the lab, the soil cores were stored in a TODO &deg;C cooler for several weeks.
-At that point (late August 2015), three fragmented or otherwise damaged cores were discarded.
-Six random cores were removed and destructively analyzed for TODO.
-The others were randomly assigned to one of five groups (each N=6).
-These groups included two incubation temperatures (4 &deg;C and 22 &deg;C) each with two moisture treatments: one in which soil moisture was maintained at field conditions, and a drought treatment in which no water was added.
-The fifth group was a 22 &deg;C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 &deg;C "drought" group.
+In the lab, the soil cores were stored at 4 &deg;C for several days until they were weighed and prepared for incubation.
+At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (each N=6).
+These groups included two incubation temperatures (4 &deg;C and 20 &deg;C) each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
+The fifth group was a 20 &deg;C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 &deg;C "drought" group.
+The final group includedsix cores for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density.
+Subsamples were collected and stored at -20 &deg;C for potential enzyme activity and ATP.
 
-The incubation experiment began on 31 August 2015.
+On 18 August 2015 cores were placed into one of two growth chambers (Conviron CMP6050 Control Systems, Winnipeg, Canada) maintained at 4 and 20 &deg;C temperatures and 70% relative humidity and allowed to equilibrate for two weeks.
+Starting on 31 August 2015 we measured the cores' mass and GHG (CH~4~ and CO~2~) evolution four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation.
+Throughout the incubation, cores were mounted on 0.05-MPa ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action and to moderate the drying of the 20 &deg;C "controlled drought" group.
+To maintain the field moisture treatment groups at ~28% moisture, the cores sat on saturated pore plates and received additional wetting from the top of the cores to supplement for when evaporation exceeded imbibition.
+
+For each measurement, a six-core treatment group was connected to a Picarro A0311 multiplexer that was in turn connected to a Picarro G2301 GHG Analyzer (Picarro Inc., Santa Clara, CA, USA).
+Dry GHG concentration were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group.
+Ambient air was measured between treatment groups, and before starting measurements in a chamber.
 

### 2015-10-06 4-qc.R
Improved raw data diagnostics

@@ -10,7 +10,7 @@ PROBLEM       <- FALSE
 SUMMARYDATA  <- file.path(OUTPUT_DIR, "summarydata_clean.csv")  # output from script 3
 RAWDATA      <- file.path(OUTPUT_DIR, "rawdata_samples.csv.gz")  # output from script 1
 
-save_diagnostic <- function(p, pname, ...) {
+save_diagnostic <- function(p, pname, printit = TRUE, ...) {
   print(p)
   printlog("Saving diagnostic for", pname)
   ggsave(paste0("qc_plots/", pname, ".png"))
@@ -132,7 +132,7 @@ for(i in seq_len(nrow(core_cv))) {
   if(i <= MAX_COMBINED_PLOTS) {
     rdata_final <- rbind(rdata_final, rdata)
     sdata_final <- rbind(sdata_final, sdata)
-    save_diagnostic(p, paste0("coreCV_", i))
+    save_diagnostic(p, paste0("coreCV_", i), printit = FALSE)
   } else {
     save_plot(paste0("coreCV_", i), p = p)
   }
@@ -176,14 +176,18 @@ p <- qplot(Date, CH4_ppb_s, data=smry) +
 save_diagnostic(p, "CH4_time")
 
 # Individual cores over time
-p <- qplot(incday, CO2_ppm_s, data=summarydata, color=paste(Treatment, Temperature))
+# Reorder the 'Core' factor so that facets are ordered by Treatment and Temperature
+summarydata$TT <- paste(summarydata$Treatment, summarydata$Temperature)
+sd_no_A <- summarydata %>% filter(Core != "Ambient4" & Core != "Ambient22") %>% arrange(TT)
+sd_no_A$Core <- factor(sd_no_A$Core, levels = unique(as.character(sd_no_A$Core)))
+p <- qplot(incday, CO2_ppm_s, data=sd_no_A, color=TT)
 p <- p + ggtitle("CO2 fluxes (uncorrected) by rep, core, incubation day") + scale_color_discrete("")
-p <- p + facet_wrap(~Core)
+p <- p + facet_wrap(~Core, ncol = 6)
 save_diagnostic(p, "CO2_incday")
 
-p <- qplot(incday, CH4_ppb_s, data=summarydata, color=paste(Treatment, Temperature))
+p <- qplot(incday, CH4_ppb_s, data=sd_no_A, color=paste(Treatment, Temperature))
[…]

### 2015-10-07 manuscript/cpcrw_manuscript.Rmd
Tweaks
Introduction outline
Re issue #14.

@@
 
-Introduction.
+* Boreal biome and climate change.
+ * Climate change and warming boreal biome.
+ * Variety of changes occurring - Alaska in particular.
+ * Vegetation transitions (setting up for transition zone sampling).
 
-Boreal forests, permafrost, importance of active layer.
+* Potential soils and permafrost C feedback. 
+ * Soils hold a lot of C, particularly at high latitudes.
+ * Importance of active layer.
+ * Uncertainty of climate feedback.
 
-Moisture likely to be a strong control of C oxidation. Which form? Importance of CO2 / CH4 balance.
+* Moisture likely to be a strong control of C oxidation.
+ * Cite theoretical, experimental, modeling work here. 
+ * Soil drainage/anoxia is *highly* uncertain.
+ * Will determine both magnitude and form (CO2/CH4) of oxidation.
+ * This will have different feedback effects.
+
+* Goals of this study.
+ * Sample AL soils at maximum AL depth, vegetation transition zone (high enough to warm, still lots of C).
+ * Manipulative experiment (drought and temperature) examining CO2/CH4 emissions.
+ * Link to structural/biological measurements.
 
@@
 
+**Laboratory iincubation*
+
 In the lab, the soil cores were stored at 4 &deg;C for several days until they were weighed and prepared for incubation.
@@
 
+Note open experiment...
+
+**Statistical analysis**
+
 Analysis.

### 2015-10-09 manuscript/cpcrw_manuscript.Rmd
Update cpcrw_manuscript.Rmd
Added Peyton's text to intro

@@
 
+The goal of this study was to characterize the chemical and biological properties of sub-Arctic, active layer soils subjected to changes in temperature and moisture.
+We also sought to determine the relative control temperature and moisture have on the fate of C and greenhouse gas emissions in active layer soils lying directly above permafrost.
+We hypothesized that the field moist 20˚C cores have a greater potential to lose C as both CO2 and CH4 than cores held at 4˚C due to increased aerobic and anaerobic microbial activity.
+We also hypothesized that drought-induced cores would have a higher potential to lose more ‘complex’ carbon compounds (aromatic C) as these compounds are desorbed due to increased surface tension.
+
 #### Methods
+
+**Field sampling**
 
@@
 
-SE slope, transition zone
+SE slope, transition zone...
 
@@
 
-**Laboratory iincubation*
+**Laboratory incubation**
 
@@
 
-Note open experiment...
-
 **Statistical analysis**
 
-Analysis.
+Analysis. Outlier exclusion.
+
+**Data and code availability**
+
+Note open experiment...
 

### 2015-10-09 manuscript/cpcrw_manuscript.Rmd
Update cpcrw_manuscript.Rmd

@@
 #### Abstract
+
+Peyton text to use in abstract:
+>The goal of this study was to characterize the chemical and biological properties of sub-Arctic, active layer soils subjected to changes in temperature and moisture.
+We also sought to determine the relative control temperature and moisture have on the fate of C and greenhouse gas emissions in active layer soils lying directly above permafrost.
+We measured carbon dioxide and methane emissions on soil at two different temperatures (4 and 20C) and moisture contents (field moist and drought-induced) throughout a 100-day incubation.
+We also determined soil structure, extracellular enzyme activity, total nitrogen, carbon, and the molecular composition of C pre and post-incubation cores.
+We hypothesized that the field moist 20˚C cores have a greater potential to lose C as both CO2 and CH4 than cores held at 4˚C due to increased aerobic and anaerobic microbial activity.
+We also hypothesize that drought-induced cores will have a higher potential to lose more ‘complex’ carbon compounds (aromatic-C) as these compounds are desorbed due to increased surface tension.
 

### 2015-12-15 R/1-data.R
Documentation updates

@@ -1,5 +1,6 @@
 # Process Picarro data for our CPCRW lab experiment
-# This script reads all available data, writing to a `rawdata` file
+# This script reads all available Picarro outputs in `data/picarro/`,
+# concatenating and writing to an `outputs/rawdata.csv.gz` file.
 # Ben Bond-Lamberty July 2015
 
 source("R/0-functions.R")

### 2015-12-15 R/2-summarize.R

@@ -1,7 +1,10 @@
 # Process Picarro data for Peyton's DWP lab experiment
-# This script summarizes individual Picarro observations to summaries
-# of "samples" (groups of observations made from a given core at some
-# point in time)
+# This workhorse script summarizes individual (raw) Picarro observations to 
+# summaries of "samples" (groups of consecutive observations made from a given 
+# core at a point in time). It computes gas concentration changes, performs 
+# some QC, merges the Picarro data with valve map and other ancillary data,
+# and writes `outputs/summarydata.csv`.
+# 
 # Ben Bond-Lamberty July 2015
 
 source("R/0-functions.R")

### 2015-12-15 R/2.5-remove.R

@@ -1,10 +1,11 @@
 # Process Picarro data for CPCRW incubation
-# Remove bad/freaky observations
+# This script remove bad/freaky/outlier observations from the data, based
+# on the `data/removals.csv` list. It then writes `outputs/summarydata_clean.csv`.
 # Ben Bond-Lamberty September 2015
 
 source("R/0-functions.R")
 
-SCRIPTNAME  	<- "3-remove.R"
+SCRIPTNAME  	<- "2.5-remove.R"
 PROBLEM       <- FALSE
 
 SUMMARYDATA  <- file.path(OUTPUT_DIR, "summarydata.csv")  # output from script 2

### 2015-12-15 R/3-qc.R

@@ -1,10 +1,11 @@
 # Process Picarro data for CPCRW incubation
-# Quality control
+# This script produces a wide range of diagnostic and QC plots, 
+# of both the raw and (clean) summary datasets.
 # Ben Bond-Lamberty September 2015
 
 source("R/0-functions.R")
 
-SCRIPTNAME  	<- "4-qc.R"
+SCRIPTNAME  	<- "3-qc.R"
 PROBLEM       <- FALSE
 
 SUMMARYDATA  <- file.path(OUTPUT_DIR, "summarydata_clean.csv")  # output from script 3

### 2015-12-15 R/3.5-investigate.R

@@ -1,12 +1,13 @@
 # Process Picarro data for CPCRW incubation
-# Investigate treatment summaries with particularly high variability
-# Note this interactive script is only used to generate figures and statistics;
+# This script is used to investigate treatment summaries with particularly high 
+# variability. Note this interactive script is only used to generate figures and 
+# statistics, identifying candidates to add to the `removals.csv` file; 
 # it doesn't make any changes to the datasets.
 # Ben Bond-Lamberty December 2015
 
 source("R/0-functions.R")
 
-SCRIPTNAME  	<- "5-highvar.R"
+SCRIPTNAME  	<- "3.5-investigate.R"
 PROBLEM       <- FALSE
 
 SUMMARYDATA  <- file.path(OUTPUT_DIR, "summarydata_clean.csv")  # output from script 3

### 2015-12-15 R/4-fluxes.R

@@ -1,11 +1,11 @@
 # NOTE THIS DOESN'T WORK WITH CURRENT DATASET YET
 
-# Process Picarro data for Peyton's DWP lab experiment
+# Computes fluxes. In progress.
 # Ben Bond-Lamberty April 2015
 
 source("R/0-functions.R")
 
-SCRIPTNAME  	<- "3-fluxes.R"
+SCRIPTNAME  	<- "4-fluxes.R"
 summarydata      <- file.path(OUTPUT_DIR, "summarydata.csv")  # output from script 2
 
 

### 2015-12-15 R/6-plots.R

@@ -1,6 +1,4 @@
 # NOTE THIS DOESN'T WORK WITH CURRENT DATASET YET
-
-# Process Peyton's Picarro data for DWP lab experiment
 # Ben Bond-Lamberty July 2015
 
 source("0-functions.R")

### 2015-12-15 R/2.5-remove.R
Removals script now recognizes removals vs. exclusions
Exclusions are based on data being identified as outliers; removals are
because of known instrument error or something like that. Addresses
issue #25.

@@ -7,6 +7,8 @@ source("R/0-functions.R")
 
 SCRIPTNAME  	<- "2.5-remove.R"
 PROBLEM       <- FALSE
+CATEGORY_ERROR <- "Error"
+CATEGORY_OUTLIER <- "Outlier"
 
 SUMMARYDATA  <- file.path(OUTPUT_DIR, "summarydata.csv")  # output from script 2
 REMOVALS     <- file.path("data/", "removals.csv")
@@ -28,36 +30,56 @@ printlog("Reading in removal data...")
 removals <- read_csv(REMOVALS, skip = 1)
 
 summarydata_clean <- summarydata
-removed_all <- data.frame()
+removed_data <- data.frame()
+excluded_data <- data.frame()
 for(i in seq_len(nrow(removals))) {
+  removed <- data.frame()
+  excluded <- data.frame()
   rem <- removals[i,]
   printlog(SEPARATOR)
   printlog("Removal", i)
   printlog("Reason:", rem$Reason)
-  
-  if(is.na(rem$Reason)) {
-    printlog("SKIPPED - 'reason' is required, but not provided!")
+  printlog("Category:", rem$Category)
+  if(is.na(rem$Reason) | is.na(rem$Category)) {
+    printlog("SKIPPED - 'reason' and `category` are both required, but not provided!")
     next
   }
   
   if(all(is.na(c(rem$min_CO2_dry, rem$max_CO2_dry, rem$min_CH4_dry, rem$max_CH4_dry)))) {
-    printlog("Removed entire sample", rem$samplenum)
-    removed <- subset(summarydata_clean, samplenum == rem$samplenum)
+    
+    if(rem$Category == CATEGORY_ERROR) {
+      printlog("Removed entire sample", rem$samplenum)
+      removed <- subset(summarydata_clean, samplenum == rem$samplenum)
+    } else if(rem$Category == CATEGORY_OUTLIER) {
+      printlog("Excluded entire sample", rem$samplenum)
+      excluded <- subset(summarydata_clean, samplenum == rem$samplenum)
+    } else {
+      stop("Unknown category!")
+    }
     summarydata_clean <- subset(summarydata_clean, samplenum != rem$samplenum)
   } else {
-    stop("Not implemented yet!")
+    stop("Method not implemented yet!")
   }
   
   printlog("Summary data observations removed:", nrow(removed))
-  removed_all <- rbind(removed_all, removed)
+  removed_data <- rbind(removed_data, removed)
[…]

### 2015-12-18 R/4-fluxes.R
Tweaks to data summary and graphs

@@ -106,20 +106,28 @@ p2 <- ggplot(fluxdata, aes(incday, cumCH4_flux_mgC, group = Core)) +
 print(p2)
 save_plot("cumulative_CH4")
 
-printlog("Flux summary plot...")
+printlog("Calculating flux summary datasets...")
 fluxdata %>%
   melt(id.vars = c("Treatment", "Temperature", "Core", "incday"), 
        measure.vars = c("cumCO2_flux_mgC", "cumCH4_flux_mgC"),
        variable.name = "Gas") %>%
+  mutate(Gas = substr(Gas, start = 4, stop = 6)) %>%   # rename
   group_by(Treatment, Temperature, Gas, Core) %>%
   arrange(incday) %>%
-  summarise(cum_flux_mgC = last(value)) %>%
+  summarise(cum_flux_mgC = last(value)) -> 
+  fd_summary_core
+
+save_data(fd_summary_core, scriptfolder = FALSE)
+
+fd_summary_core %>%   # already grouped by Treatment, Temperature, Gas
   summarise(cum_flux_mgC_sd = sd(cum_flux_mgC),
             cum_flux_mgC = mean(cum_flux_mgC)) ->
   fd_summary
 
+
+printlog("Flux summary plot...")
+
 # Tweak the data set - factors, names, etc.
-fd_summary$Gas <- substr(fd_summary$Gas, start = 4, stop = 6)
 fd_summary$Treatment <- factor(fd_summary$Treatment, levels = c("Drought", "Controlled drought", "Field moisture"))
 fd_summary$Temperature <- as.factor(fd_summary$Temperature)
 
@@ -134,7 +142,9 @@ fd_summary <- rbind(fd_summary, dmy)
 
 p3 <- ggplot(fd_summary, aes(Temperature, cum_flux_mgC, fill = Treatment)) + 
   geom_bar(stat='identity', position = position_dodge()) +
-  geom_errorbar(aes(color = Treatment, ymin = cum_flux_mgC * .9, ymax = cum_flux_mgC + cum_flux_mgC_sd), 
+  geom_errorbar(aes(color = Treatment, 
+                    ymin = cum_flux_mgC * 0.9, 
+                    ymax = cum_flux_mgC + cum_flux_mgC_sd), 
                 position = position_dodge(0.9), width = 0.4) +  
   facet_grid(Gas~., scales = "free") +
   ylab(paste("Cumulative C (mg) over", floor(max(fluxdata$incday)), "days")) +

### 2015-12-19 manuscript/cpcrw_manuscript.Rmd
Outlier exclusion citation
Thinking about possible tables and figures

@@
 
-Analysis. Outlier exclusion.
+AOutlier exclusion based on 3.2 mean absolute deviations from treatment and temperature mean. We excluded X of Y (Z%) of the summarized samples for this reason.
+
+Citation: Davies, P.L. and Gather, U. (1993). "The identification of multiple outliers" (with discussion). J. Amer. Statist. Assoc., 88, 782-801.
+
+nalysis. Outlier exclusion.
 
@@
 Acknowledgments.
+
+Table X. Summary of treatment effects on gas fluxes.
+
+Figure X. A faceted (Temperature ~ Treatment) plot showing fnormalized lux means and errors by incubation time.
+ (Basically the figure already in the README diagnostics.)
+TFigur X. 
+Core water content across the course of the experiment. (Basically the figure already in the README diagnostics.)

### 2015-12-19 R/4-fluxes.R
Fluxes plot tweaks

@@ -30,6 +30,9 @@ summarydata %>%
 # TODO: ask Peyton, what's the "Volume" column in the CoreData.xlsx file?
 fluxdata$Headspace_cm <- 5
 
+# Compute water content
+fluxdata$WC_fraction <- with(fluxdata, (Mass_g - DryMass_g) / DryMass_g)
+
 # At this point, `fluxdata` has slopes (CO2 ppm/s and CH4 ppb/s).
 # We want to convert this to mg C/g soil/s, using
 # A = dC/dt * V/M * Pa/RT (cf. Steduto et al. 2002), where
@@ -91,7 +94,7 @@ fluxdata %>%
 save_data(fluxdata, scriptfolder = FALSE)
 
 # A few diagnostic plots
-printlog("Flux diagnostic plots...")
+printlog("Cumulative flux diagnostic plots...")
 p1 <- ggplot(fluxdata, aes(incday, cumCO2_flux_mgC, group = Core)) + 
   geom_line() + 
   facet_grid(Temperature ~ Treatment) +
@@ -115,32 +118,32 @@ fluxdata %>%
   group_by(Treatment, Temperature, Gas, Core) %>%
   arrange(incday) %>%
   summarise(cum_flux_mgC = last(value)) -> 
-  fd_summary_core
+  fd_cumulative_core
 
-save_data(fd_summary_core, scriptfolder = FALSE)
+save_data(fd_cumulative_core, scriptfolder = FALSE)
 
-fd_summary_core %>%   # already grouped by Treatment, Temperature, Gas
+fd_cumulative_core %>%   # already grouped by Treatment, Temperature, Gas
   summarise(cum_flux_mgC_sd = sd(cum_flux_mgC),
             cum_flux_mgC = mean(cum_flux_mgC)) ->
-  fd_summary
+  fd_cumulative
 
 
 printlog("Flux summary plot...")
 
 # Tweak the data set - factors, names, etc.
-fd_summary$Treatment <- factor(fd_summary$Treatment, levels = c("Drought", "Controlled drought", "Field moisture"))
-fd_summary$Temperature <- as.factor(fd_summary$Temperature)
+fd_cumulative$Treatment <- factor(fd_cumulative$Treatment, levels = c("Drought", "Controlled drought", "Field moisture"))
+fd_cumulative$Temperature <- as.factor(fd_cumulative$Temperature)
 
 # Add a few dummy rows so graph bars are spaced correctly
[…]

### 2015-12-19 R/5-analyze.R
Beginning analysis script

@@ -1,0 +1,66 @@
+# Computes fluxes. In progress.
+# Ben Bond-Lamberty December 2015
+
+source("R/0-functions.R")
+
+SCRIPTNAME  	<- "5-analyze.R"
+FLUXDATA   <- file.path(outputdir(scriptfolder = FALSE), "fluxdata.csv")
+
+# ==============================================================================
+# Main 
+
+openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # open log
+
+printlog("Welcome to", SCRIPTNAME)
+
+fluxdata_orig <- read_csv(FLUXDATA)
+
+printlog("Transforming...")
+fluxdata_orig %>%
+  select(Core, Treatment, Temperature, incday, WC_fraction, CO2_flux_µmol_g_s, CH4_flux_µmol_g_s) %>%
+  melt(measure.vars = c("CO2_flux_µmol_g_s", "CH4_flux_µmol_g_s"),
+       value.name = "flux_µmol_g_s") %>%
+  mutate(Gas = substr(variable, 1, 3)) %>%
+  select(-variable) ->
+  fluxdata
+
+save_data(fluxdata, fname = "fluxdata_long")
+
+# -----------------------------------------------------------------------------
+# Does WC affect gas fluxes?
+fluxdata %>%
+  group_by(Treatment, Temperature, Gas) %>%
+  do(mod = lm(flux_µmol_g_s ~ WC_fraction, data = .)) ->
+  models
+
+models %>% 
+  do(data.frame(    # from the 'do' help page
+    var = names(coef(.$mod)),
+    coef(summary(.$mod)))) %>%
+  filter(var == "WC_fraction") %>%
+  mutate(p.value = round(`Pr...t..`, 3),
+         signif = p.value < 0.05) %>%
+  select(-var, -`Pr...t..`, -Estimate, -`Std..Error`) ->
+  modelsummary
+
+fluxdata %>%
+  group_by(Treatment, Temperature, Gas) %>% 
+  summarise(n=n()) %>% 
+  cbind(modelsummary) ->
+  WC_effect
+
+print(WC_effect)
+save_data(WC_effect)
+
+
+# m_co2 <- lm(cum_flux_mgC ~ Treatment * Temperature, 
+#             data=subset(fd_summary_core, Gas=="CO2"))
+# summary(m_co2)
+# 
[…]

### 2015-12-19 R/5-analyze.R
Basic linear models to assess statistical significance

@@ -24,6 +24,8 @@ fluxdata_orig %>%
   select(-variable) ->
   fluxdata
 
+fluxdata$Treatment <- factor(fluxdata$Treatment, levels = c("Field moisture", "Controlled drought", "Drought"))
+
 save_data(fluxdata, fname = "fluxdata_long")
 
 # -----------------------------------------------------------------------------
@@ -53,13 +55,22 @@ print(WC_effect)
 save_data(WC_effect)
 
 
-# m_co2 <- lm(cum_flux_mgC ~ Treatment * Temperature, 
-#             data=subset(fd_summary_core, Gas=="CO2"))
-# summary(m_co2)
-# 
-# m_ch4 <- lm(cum_flux_mgC ~ Treatment * Temperature, 
-#             data=subset(fd_summary_core, Gas=="CH4"))
-# summary(m_ch4)
+# -----------------------------------------------------------------------------
+# Main treatment effects
+
+# Fitting a non-mixed-effects model for now
+# TODO: later, might want to explore having Core as a random effect
+
+# Fit gases separately, just for simplicity
+printlog("Fitting CO2 model...")
+m_co2 <- lm(flux_µmol_g_s ~ Treatment * Temperature + WC_fraction,
+            data = fluxdata, subset = Gas == "CO2")
+print(summary(m_co2))
+
+printlog("Fitting CH4 model...")
+m_co2 <- lm(flux_µmol_g_s ~ Treatment * Temperature + WC_fraction,
+            data = fluxdata, subset = Gas == "CH4")
+print(summary(m_co2))
 
 
 printlog("All done with", SCRIPTNAME)

### 2016-01-31 manuscript/cpcrw_manuscript.Rmd
Filled out and tweaks methods text

@@
 The field component of this research took place in Caribou-Poker Creeks Research Watershed (CPCRW), part of the Bonanza Creek LTER (http://www.lter.uaf.edu/default.cfm).
-Annual average air temperature was 􏰀3.3°C, and the average temperature in July was 16.4°C, which was the maximum in this watershed.
-Annual average precipitation was 269 mm, and the average precipitation of 47 mm in August was the highest in this area.
+Annual average air temperature was -2.5 °C, and annual average precipitation was 400 mm {Petrone, 2006 #4313}.
+The watershed's lowlands and north-facing slopes are dominated by black spruce (*Picea mariana*), feathermoss (*Pleurozium schreberi* and others), and *Sphagnum* spp.; the drier south slopes tend to be deciduous with a mixture of quaking aspen (*Populus tremuloides*), paper birch (*Betula neoalaskana*), and patches of alder (*Alnus crispa*).
 
-SE slope, transition zone...
-
-of this study site was more than TODO years according to the tree rings at stem base.
-Stand density of dominant Picea mariana and glauca trees (>1.3 m in height) was TODO trees ha^-1^, with a basal area of TODO m^2^ ha^-1^.
+We sampled soils from a southeast slope (65.1620 °N, 147.4874 °W), in a 60 m transition zone between lowland *Picea mariana* Mill BSP.
+and upland *Betula neoalaskana*, with significant white spruce (*Picea glauca*) presence.
+Stand density in this transition zone was 4060 ± 2310 trees ha^-1^, with basal area of 27.9 ± 7.0 m^2^ ha^-1^.
+The forest was at least 85 years old according to tree rings at the stem base.
 
@@
 
-In the lab, the soil cores were stored at 4 &deg;C for several days until they were weighed and prepared for incubation.
-At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (each N=6).
-These groups included two incubation temperatures (4 &deg;C and 20 &deg;C) each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
-The fifth group was a 20 &deg;C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 &deg;C "drought" group.
-The final group includedsix cores for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density.
-Subsamples were collected and stored at -20 &deg;C for potential enzyme activity and ATP.
+In the lab, the soil cores were stored at 4 °C for several days until they were weighed and prepared for incubation.
+At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (N=6 in each group).
+These groups included two incubation temperatures (4 °C and 20 °C) each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
+The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 °C "drought" group.
+The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density.
+Subsamples were collected and stored at -20 °C for potential enzyme activity and ATP.
 
-On 18 August 2015 cores were placed into one of two growth chambers (Conviron CMP6050 Control Systems, Winnipeg, Canada) maintained at 4 and 20 &deg;C temperatures and 70% relative humidity and allowed to equilibrate for two weeks.
+On 18 August 2015 cores were placed into one of two growth chambers (Conviron CMP6050 Control Systems, Winnipeg, Canada) maintained at 4 and 20 °C temperatures and 70% relative humidity and allowed to equilibrate for two weeks.
 Starting on 31 August 2015 we measured the cores' mass and GHG (CH~4~ and CO~2~) evolution four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation.
-Throughout the incubation, cores were mounted on 0.05-MPa ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action and to moderate the drying of the 20 &deg;C "controlled drought" group.
-To maintain the field moisture treatment groups at ~28% moisture, the cores sat on saturated pore plates and received additional wetting from the top of the cores to supplement for when evaporation exceeded imbibition.
+Throughout the incubation, cores were mounted on 0.05-MPa ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action.
+This was done to moderate the drying of the 20 °C "controlled drought" group.
+The 4 and 20 °C "field moisture" groups received additional wetting from the top of the cores to maintain their water status.
 
 For each measurement, a six-core treatment group was connected to a Picarro A0311 multiplexer that was in turn connected to a Picarro G2301 GHG Analyzer (Picarro Inc., Santa Clara, CA, USA).
-Dry GHG concentration were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group.
+Dry CH~4~ and CO~2~ concentrations were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group.
+Cores were weighed immediately after (PEYTON) gas measurements.
 Ambient air was measured between treatment groups, and before starting measurements in a chamber.
+The incubation experiment concluded on 10 December 2015.
 
@@
 
-AOutlier exclusion based on 3.2 mean absolute deviations from treatment and temperature mean. We excluded X of Y (Z%) of the summarized samples for this reason.
+Outlier exclusion based on 3.2 mean absolute deviations from treatment and temperature mean. We excluded X of Y (Z%) of the summarized samples for this reason.
 

### 2016-02-01 manuscript/cpcrw_manuscript.Rmd
Introduction draft

@@
 ---
-title: "cpcrw_ms"
-author: "Ben Bond-Lamberty"
-date: "September 29, 2015"
+title: "Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils"
+author: "Ben Bond-Lamberty, Peyton Smith, Vanessa Bailey"
+#date: "February 1, 2016"
 output: html_document
@@
 
-Peyton text to use in abstract:
+Peyton potential text to use in abstract:
+
 >The goal of this study was to characterize the chemical and biological properties of sub-Arctic, active layer soils subjected to changes in temperature and moisture.
@@
 
-* Boreal biome and climate change.
- * Climate change and warming boreal biome.
- * Variety of changes occurring - Alaska in particular.
- * Vegetation transitions (setting up for transition zone sampling).
+High latitude ecosystems are being subjected to rapid changes in climate {IPCC, 2013 #3916} and fire occurrence and intensity {Kasischke, 2010 #2631}.
+This has a wide variety of effects on ecosystems and society, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}.
+As a result, rising temperatures and increasing fire will result in permafrost degradation {Pastick, 2015 #4311} and changes in soil temperature and hydrology that are expected to increase soil greenhouse gas (GHG) fluxes to the atmosphere, promoting a positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
 
-* Potential soils and permafrost C feedback. 
- * Soils hold a lot of C, particularly at high latitudes.
- * Importance of active layer.
- * Uncertainty of climate feedback.
+The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such a feedback remain highly uncertain {Schuur, 2015 #4319}.
+While northern soils hold enormous quantities {Tarnocai, 2009 #2231} of potentially mineralizable soil organic carbon (SOC), vegetation succession dynamics promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}.
+Such dynamics may however be disrupted by increased fire disturbance {Johnstone, 2010 #2465}.
+In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon and nitrogen pools {Weiss, 2015 #4317}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
 
-* Moisture likely to be a strong control of C oxidation.
- * Cite theoretical, experimental, modeling work here. 
- * Soil drainage/anoxia is *highly* uncertain.
- * Will determine both magnitude and form (CO2/CH4) of oxidation.
- * This will have different feedback effects.
+Temperature and moisture typically have strong and often interactive influences on soil GHG emissions.
+A long literature of laboratory incubations, field observations, and meta-analyses has documented increased CO2 fluxes with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
+GHG responses to wetting and thawing dynamics are much less certain, with substantial variability between studies {Kim, 2012 #2863}.
+The anaerobic conditions common following permafrost thaw lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and interact with soil temperature effects in long-term incubations {Elberling, 2013 #3720}.
+Such laboratory studies have weaknesses, but their controllled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
 
-* Goals of this study.
- * Sample AL soils at maximum AL depth, vegetation transition zone (high enough to warm, still lots of C).
- * Manipulative experiment (drought and temperature) examining CO2/CH4 emissions.
- * Link to structural/biological measurements.
-
-The goal of this study was to characterize the chemical and biological properties of sub-Arctic, active layer soils subjected to changes in temperature and moisture.
-We also sought to determine the relative control temperature and moisture have on the fate of C and greenhouse gas emissions in active layer soils lying directly above permafrost.
-We hypothesized that the field moist 20˚C cores have a greater potential to lose C as both CO2 and CH4 than cores held at 4˚C due to increased aerobic and anaerobic microbial activity.
-We also hypothesized that drought-induced cores would have a higher potential to lose more ‘complex’ carbon compounds (aromatic C) as these compounds are desorbed due to increased surface tension.
+The goal of this study was to examine how temperature and moisture control C loss and GHG (CO~2~ and CH~4~) emissions from soils sampled from the bottom of the annual active layer--i.e., directly above permafrost--in an Alaskan boreal forest.
+Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
+We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at difference temperature and moisture levels.We hypothesized that soils maintained at field moisture and high temperature would lose more C, both as CO~2~ and CH~4~, than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity.
+In addition, cores subject to drought (reduced soil moisture) would have a higher potential to lose more ‘complex’ carbon compounds (aromatic C) as these compounds are desorbed due to increased surface tension.
+TODO: maybe change/reword this last sentence.
 
@@
 
-We sampled soils from a southeast slope (65.1620 °N, 147.4874 °W), in a 60 m transition zone between lowland *Picea mariana* Mill BSP.
+We sampled soils from a southeast slope (65.1620 °N, 147.4874 °W) at CPCRW, in a 60 m transition zone between lowland *Picea mariana* Mill BSP.
 and upland *Betula neoalaskana*, with significant white spruce (*Picea glauca*) presence.
 Stand density in this transition zone was 4060 ± 2310 trees ha^-1^, with basal area of 27.9 ± 7.0 m^2^ ha^-1^.
-The forest was at least 85 years old according to tree rings at the stem base.
+The forest was at least 90 years old {cf.
+\\Morishita, 2014 #4316} according to tree rings taken at the stem base of several of the largest white spruce.
+The soil is characterized as a poorly-drained silt loam, and on average had ~20 cm of organic material over the mineral soil.
 
@@
 Sample points were randomly located in the transition zone described above, and separated by 2-5 m.
-The soil is characterized as a poorly-drained silt loam, and on average had ~20 cm of organic material over the mineral soil.
 Cores were kept cool in the field before being packed in dry ice and shipped to the lab in Richland, WA within 48-72 hours of collection.
@@
 At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (N=6 in each group).
-These groups included two incubation temperatures (4 °C and 20 °C) each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
+These groups included two incubation temperatures (4 and 20 °C), each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
 The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 °C "drought" group.
 The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density.
-Subsamples were collected and stored at -20 °C for potential enzyme activity and ATP.
+Subsamples were collected and stored at -20 °C for potential enzyme activity and ATP analysis.
 
@@
 Ambient air was measured between treatment groups, and before starting measurements in a chamber.
-The incubation experiment concluded on 10 December 2015.
+
+The incubation experiment concluded on 10 December 2015. PEYTON: text re C, N, etc. analyses.
 
@@
 
-Outlier exclusion based on 3.2 mean absolute deviations from treatment and temperature mean. We excluded X of Y (Z%) of the summarized samples for this reason.
+Outlier exclusion based on 3.2 mean absolute deviations {Davies, 1993 #3138} from treatment and temperature mean. We excluded X of Y (Z%) of the summarized samples for this reason.
 
-Citation: Davies, P.L. and Gather, U. (1993). "The identification of multiple outliers" (with discussion). J. Amer. Statist. Assoc., 88, 782-801.
-
-nalysis. Outlier exclusion.
+All data analysis and statistics were performed using R version 3.2.3 {R Development Core Team, 2016 #4314}.
 
@@
 
+Incubation experiments can only go so far though...
+Morishita et al.
+{, 2014 #4316} measured GHG fluxes in the field at CPCRW and nearby forests, and found CO2 production to be correlated with both soil temperature and moisture, but CH4 uptake driven by temperature only.
+
 ### Acknowledgments
@@
 
-Table X. Summary of treatment effects on gas fluxes.
+Table X. Summary of temperature, WC effects on gas fluxes.
 
-Figure X. A faceted (Temperature ~ Treatment) plot showing fnormalized lux means and errors by incubation time.
- (Basically the figure already in the README diagnostics.)
-TFigur X. 
-Core water content across the course of the experiment. (Basically the figure already in the README diagnostics.)
+Figure X. A faceted (Temperature ~ Treatment) plot showing normalized flux means and errors by incubation time. (Basically the figure already in the README diagnostics.)
+
+Figure X. Core water content across the course of the experiment. (Basically the figure already in the README diagnostics.)
+
+Figure X. A pairs plot of variable correlations?

### 2016-02-02 R/4-fluxes.R
Added Peyton's new data file
Also shifted reading of mass data from script 2 to script 4

@@ -6,6 +6,8 @@ source("R/0-functions.R")
 SCRIPTNAME  	<- "4-fluxes.R"
 SUMMARYDATA   <- file.path(outputdir(scriptfolder = FALSE), "summarydata_clean.csv")
 
+DRYMASSDATA   <- "data/drymasses.csv"
+
 # ==============================================================================
 # Main 
 
@@ -15,6 +17,9 @@ printlog("Welcome to", SCRIPTNAME)
 
 summarydata <- read_csv(SUMMARYDATA)
 print_dims(summarydata)
+
+drymassdata <- read_csv(DRYMASSDATA, skip = 1)
+summarydata <- left_join(summarydata, drymassdata, by = "Core")
 
 # Create fluxdata and compute water content
 summarydata %>%

### 2016-02-02 manuscript/cpcrw_manuscript.Rmd
Methods - removals and exclusions

@@
 ---
+
+```{r, echo=FALSE}
+library(dplyr)
+```
 
@@
 #### Introduction
+
+Alaska forests under particular stress - browning {Ju, 2016 #4320}
 
@@
 
-**Statistical analysis**
+**Data and statistical analysis**
 
-Outlier exclusion based on 3.2 mean absolute deviations {Davies, 1993 #3138} from treatment and temperature mean. We excluded X of Y (Z%) of the summarized samples for this reason.
+We measured the rise in GHG concentrations from its minimum value (in the first 10 seconds of measurement after the analyzer was sealed to the soil core) to its maximum (in the first 45 seconds), considering this as the flux rate in ppm s^-1^ (CO2) or ppb s^-1^ (CH4).
+This value was calculated for each sample, i.e.
+each individual core measurement throughout the 100-day incubation.
+
+```{r, echo=FALSE}
+# compute outlier and exclusion numbers
+sdn <- readLines("../outputs/summarydata.csv") %>% length
+sdcn <- readLines("../outputs/summarydata_clean.csv") %>% length
+en <- readLines("../outputs/2.5-remove/excluded_data.csv") %>% length
+rn <- readLines("../outputs/2.5-remove/removed_data.csv") %>% length
+```
+
+Anomalous data were excluded based on their being more than 3.2 mean absolute deviations {Davies, 1993 #3138} from the treatment and temperature mean.
+We excluded `r en` of `r sdn` (`r round(en/sdn*100)`%) of the summarized samples for this reason.
+Other data (`r rn` of `r sdn`, `r round(rn/sdn*100)`%) were removed because of known instrument problems, e.g.
+the analyzer was left running after leaving a chamber.
+The final number of valid flux samples from the 100-day incubation was `r sdcn`.
+
+* joined with mass data
+* flux computation
+* stats
 

### 2016-02-02 manuscript_methods/Open_experiment.Rmd
Solidified outline, made ms document

---
title: "Running an open experiment: transparency and reproducibility in soil and ecosystem science"
author: "Ben Bond-Lamberty"
date: "February 2, 2016"
output: html_document
---

#### Introduction

Look at - Ruegg et al. paper on data life cycle!
Data paper - https://peerj.com/preprints/1448/
http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.1002295

Peyton's notes:

>Add to Introduction: 
>Reichman et al 2011 Science 331.pdf
>
>Wolkovich, E. M., Regetz, J. and O'Connor, M. I. (2012), Advances in global change research require open science by individual researchers. Glob Change Biol, 18: 2102–2110. doi:10.1111/j.1365-2486.2012.02693.x
>
>Add to reproducibility literature:
>http://journals.plos.org/plosbiology/article?id=10.1371%2Fjournal.pbio.1002333

#### Why more openness in science?

* Reproducibility

Thornton paper on archiving models

* Data loss

17% per year paper - Vines et al.

* Data quality

"All bugs are shallow" Raymond quote

* Increasing transparency of the process - increase public trust and access to science

* Value of collected databases - makes this easier
* BAAD TRY, FLUXNET, Dong-Gill, SRDB

#### Push and pull: funders, journals, and society

PLOS ONE, ESA, etc. policies

* NSF url
* DOE url
* NASA url
* NSERC, Wellcome trust, other?

#### Running a fully open experiment: one example

* GitHub
* Diagnostics page(s)
* Data: raw to summarized to final
* All statistical analyses
* "LRB" - notes 
* Manuscript(s)

#### Conclusions

This model not appropriate for everyone/thing!

Potential concerns

* Not appropriate for every kind of study
* Fear of "being scooped"
* Question of open from beginning or only after publication
* Privacy concerns

### 2016-02-02 manuscript_methods/Open_experiment.Rmd
Refining outline, adding citations and thoughts

@@
 title: "Running an open experiment: transparency and reproducibility in soil and ecosystem science"
-author: "Ben Bond-Lamberty"
+author: "Ben Bond-Lamberty, Peyton Smith, and Vanessa Bailey"
 date: "February 2, 2016"
@@
 
-Look at - Ruegg et al. paper on data life cycle!
-Data paper - https://peerj.com/preprints/1448/
-http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.1002295
-
-Peyton's notes:
-
->Add to Introduction: 
->Reichman et al 2011 Science 331.pdf
->
->Wolkovich, E.
-M., Regetz, J.
-and O'Connor, M.
-I.
-(2012), Advances in global change research require open science by individual researchers.
-Glob Change Biol, 18: 2102–2110.
-doi:10.1111/j.1365-2486.2012.02693.x
->
->Add to reproducibility literature:
->http://journals.plos.org/plosbiology/article?id=10.1371%2Fjournal.pbio.1002333
-
-#### Why more openness in science?
-
-* Reproducibility
-
-Thornton paper on archiving models
-
-* Data loss
-
-17% per year paper - Vines et al.
-
-* Data quality
-
-"All bugs are shallow" Raymond quote
-
-* Increasing transparency of the process - increase public trust and access to science
-
-* Value of collected databases - makes this easier
-* BAAD TRY, FLUXNET, Dong-Gill, SRDB
-
-#### Push and pull: funders, journals, and society
+General intro.
 
@@
 
-#### Running a fully open experiment: one example
+*Reproducibility*
+
+Surveying the 2000-2014 biomedical literature, Iqbal et al. {, 2016 #4334} found that none of 441 randomly-chosen studies provided raw data, and only one provided full protocols.
+
+"Reproducibility of experimental results is a fundamental criterion for credible scientific research" {Thornton, 2005 #2310}.
+More difficult for modeling work; this is a subset of the more general problem with respect to digital media, software, etc.
+(CITATION?).
+
+Scripted systems for data analysis!
+E.g.
+R {R Development Core Team, 2016 #4314}, Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record all infrmation about the data processing, analytical process and decisions, and statistical analysis.
+
+*Data loss*
+
+Vines et al.
+{, 2014 #4130} published a shocking result, based on surveying 516 articles from 2 to 22 years old: the odds of a data set being available fell by 17% each year, and the chances that the contact author's email address still worked declined by 7% per year.
+Reichman et al.
+{, 2011 #2862} estimated that less than 1% of ecological data collected is made available after publication, and noted, as an example, that most current and historical data relevant to the recent *Deepwater Horizon* oil spill are already inaccessible or lost {Reichman, 2011 #2862}.
+
+Scientists' use of strong and consistent data curation practices {Hart, 2015 #4292}{Rüegg, 2014 #3886} can mitigate the problem, but all our anecdotal experience and quantitative studies {Vines, 2014 #4130} suggest that in the long term, data cannot be reliably preserved by individual researchers.
+
+Data loss hits ecosystem, soil, and global change ecology particularly hard, as global to local climate changes make ecological data effectively irreproducible {Wolkovich, 2012 #3303}: we can never remeasure exactly the same system state.
+
+*Data and code quality*
+
+Providing open code does not magically produce bug-free code, or mistake free analyses {Easterbrook, 2014 #4137}.
+But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task (often, forever).
+
+Potential to increase pulic trust in research results.
+
+*Enabling synthetic and global-scale analyses*
+
+Many refs in {Wolkovich, 2012 #3303}.
+
+BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, SRDB {Bond-Lamberty, 2010 #2320}.
+
+#### An open experiment: one example
+
+This is example, we hope, of the impact of individual scientists {Wolkovich, 2012 #3303}.
 
@@
 
+Do we adequately credit researchers who contribute to global databases {Kattge, 2011 #3064} that subsequently produce high-impact papers {Díaz, 2016 #4318}?
+
+Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers efforts are adequately valued and cited {Kueffer, 2011 #4335}.
+
 This model not appropriate for everyone/thing!

### 2016-02-04 manuscript_methods/Open_experiment.Rmd
Added September test run data and email
See issue #31
Added figure and caption

@@
 
+VLB: Supplementary Information sections have become repositories of data and of non-core analyses. How does our open experiment replace (improve upon?) SI? Thoughts?
+
 Do we adequately credit researchers who contribute to global databases {Kattge, 2011 #3064} that subsequently produce high-impact papers {Díaz, 2016 #4318}?
@@
 
-This model not appropriate for everyone/thing!
+We don't present this model as perfect.
+First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ: for example, there was no unit testing to ensure module correctness, nor an automated build tool automating the workflow (i.e., the individual scripts all have to be run manually).
+
+Nor is this examine appropriate, or applicable, to all scientific research, even just in the fields of soil, ecosystem, and global change science.
 
@@
 
-* Not appropriate for every kind of study
 * Fear of "being scooped"
@@
 * Privacy concerns
+
+**References**
+
+**Figure 1.** The 'repository' containing all code and data of a recent soil incubation experiment.
+From top to bottom: raw data from a lab analyzer is uploaded; an initial script processes raw data into a standardized, summarized form; a second script looks for inconsistencies and outliers in the data; a third computes final data products, summaries, statistics, and figures; and a final one integrates these products to produce a submission-ready manuscript.
+(This schematic slightly simplifies the actual repository structure.) The repository is publicly available at https://github.com/bpbond/cpcrw_incubation, and throughout the experiment showed real-time updated diagnostics summarizing progress.
+
+![Repository and workflow](figureX.png)

### 2016-02-04 manuscript/cpcrw_manuscript.Rmd
Added methods re open experiment
Text and formula re flux computation

@@
 ```{r, echo=FALSE}
-library(dplyr)
+library(dplyr, warn.conflicts = FALSE)
 ```
@@
 * flux computation
+
+The GHG concentration measurements for each sample (i.e., each gas, core, and date/time) were used to calculate a linear rate of change (δc/δt), based on the concentration rise from a minimum (at 2-10 seconds after measurement began) to a maximum (at 10-45 seconds).
+Each core’s respiration flux (F) was then calculated following e.g.
+Steduto et al.
+{, 2002 #3383} as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where V is the core-specific system volume, M the core dry mass as determined at the end of the incubation, Pa atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), R the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and T the chamber air temperature (K) at time of measurement.
+The final respiration rate was expressed as mg C kg soil^-1^ day^-1^.
+TODO check all units
+
+Q10 computation
+
 * stats
@@
 All data analysis and statistics were performed using R version 3.2.3 {R Development Core Team, 2016 #4314}.
-
-**Data and code availability**
-
-Note open experiment...
+This experiment was run as an "open experiment" with all analysis code, data (raw to final), diagnostics, etc., available in real time at https://github.com/bpbond/cpcrw_incubation.
+The flux data backing the main results have also been archived at TODO.
 

### 2016-02-04 manuscript_methods/Open_experiment.Rmd
Beefed up intro, tweaked figure

@@
 ---
-title: "Running an open experiment: transparency and reproducibility in soil and ecosystem science"
+title: 'Running an open experiment: transparency and reproducibility in soil and ecosystem
+  science'
 author: "Ben Bond-Lamberty, Peyton Smith, and Vanessa Bailey"
@@
 
+#### Abstract
+
+Researchers in soil and ecosystem science, and almost every other field, as increasingly being pushed--by funders, journals, governments, and themselves--to ensure transparency and reproducibility of their work.
+A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, and enable powerful meta- and cross-disciplinary analyses.
+We describe a recent "open experiment", in which we documented every aspect of a soil incubation experiment online, making publicly available raw data, scripts, diagnostics, final analyses, and manuscripts in real time.
+Our use of tools such as version control, issue tracking, and open-source statistical software ensured data integrity and transparency.
+There are many avenues to improve scientific reproducibility and data availability, of which is this only one example; we acknowledge that this is not an approach suited for every experiment or situation.
+Nonetheless, we encourage the communities in our respective fields to consider its advantages, and lead rather than follow with respect to scientific reproducibility, transparency, and data availability.
+
 #### Introduction
@@
 
-PLOS ONE, ESA, etc. policies
+Journals are adopting increasingly stringent data access and deposition policies, e.g.
+Scientific Data (http://www.nature.com/sdata/data-policies), PLOS ONE (http://journals.plos.org/plosone/s/data-availability),  {Scientific Data, 2016 #4339}{PLOS ONE, 2014 #4338}, and Science (http://www.sciencemag.org/authors/science-editorial-policies#data-deposition).
+These policies generally share common assumptions and goals: maximizing access to scientific data; encouraging deposition into structured repositories as opposed to journal supplementary information; and specifying that it is not acceptable for authors to be solely responsible for ensuring data access.
 
-* NSF url
-* DOE url
-* NASA url
-* NSERC, Wellcome trust, other?
+Funders are moving in this direction as well, with organizations such as the U.S.
+National Science Foundation (http://www.nsf.gov/bfa/dias/policy/dmp.jsp) and the U.K.
+Wellcome Trust (http://www.wellcome.ac.uk/About-us/Policy/Spotlight-issues/Data-sharing/Data-management-and-sharing/index.htm) requiring explicit data management plans, unfettered reasonable access to primary data, and use of established repositories.
+
+Finally, growing numbers of scientists are pushing for open data on moral/political (why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results?) grounds, also on purely scientific ones: ensuring reproducibility {Stodden, 2011 #4340}.
+This latter issue--reproducibility--is at the heart of science {Kuhn, 1962 #4341} and has a number of facets worth considering briefly.
 
@@
 R {R Development Core Team, 2016 #4314}, Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record all infrmation about the data processing, analytical process and decisions, and statistical analysis.
+This and other best practices of scientific computing {Wilson, 2014 #4337} help produce reproducible, robust, and transparent science.
 
@@
 Data loss hits ecosystem, soil, and global change ecology particularly hard, as global to local climate changes make ecological data effectively irreproducible {Wolkovich, 2012 #3303}: we can never remeasure exactly the same system state.
+
+A subtler data loss issue is the "file drawer problem" {Rosenthal, 1979 #2291}.
 
@@
 
+TODO - maybe a second figure showing one of the diagnostic plots, partway through the incubation?
+
 **Figure 1.** The 'repository' containing all code and data of a recent soil incubation experiment.

### 2016-02-04 manuscript_methods/Open_experiment.Rmd
Added Nosek and Lewandowsky refs

@@
 General intro.
+
+Pushing scientists towards "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.
 
@@
 
+Transparency is no magic panacea ensuring reproducibility {Easterbrook, 2014 #4137}, and it raises genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
+We also recognize that our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}.
+
 VLB: Supplementary Information sections have become repositories of data and of non-core analyses. How does our open experiment replace (improve upon?) SI? Thoughts?

### 2016-02-08 R/4-fluxes.R
Normality test and data transformation
Now use broom package

@@ -92,10 +92,12 @@ fluxdata %>%
   mutate(incday = floor(inctime_days)) %>%
   group_by(Treatment, Temperature, incday) %>% 
   mutate(CO2_outlier = is_outlier(CO2_flux_µmol_g_s, devs = 3.0), 
-         # CH4 is so crazy variable we use a higher exclusion cutoff
+         # CH4 is so variable we use a higher exclusion cutoff
          CH4_outlier = is_outlier(CH4_flux_µmol_g_s, devs = 5.0)) %>%
   select(-incday) ->
   fluxdata
+
+fluxdata$incday <- NULL  # why doesn't the `select` above work?
 
 p <- ggplot(fluxdata, aes(inctime_days, CO2_flux_µmol_g_s, color = CO2_outlier))
 p <- p + geom_point() + facet_grid(Temperature~Treatment)

### 2016-02-08 R/5-analyze.R

@@ -6,6 +6,8 @@ source("R/0-functions.R")
 SCRIPTNAME  	<- "5-analyze.R"
 FLUXDATA   <- file.path(outputdir(scriptfolder = FALSE), "fluxdata.csv")
 
+library(broom)  # 0.4.0
+
 # ==============================================================================
 # Main 
 
@@ -17,38 +19,69 @@ fluxdata_orig <- read_csv(FLUXDATA)
 
 printlog("Transforming...")
 fluxdata_orig %>%
-  select(Core, Treatment, Temperature, incday, WC_gravimetric, CO2_flux_µmol_g_s, CH4_flux_µmol_g_s) %>%
+  select(Core, Treatment, Temperature, inctime_days, WC_gravimetric, CO2_flux_µmol_g_s, CH4_flux_µmol_g_s,
+         CO2_outlier, CH4_outlier) %>%
   melt(measure.vars = c("CO2_flux_µmol_g_s", "CH4_flux_µmol_g_s"),
        value.name = "flux_µmol_g_s") %>%
+  melt(measure.vars = c("CO2_outlier", "CH4_outlier"), 
+       variable.name = "outlier_name", value.name = "outlier") %>%
   mutate(Gas = substr(variable, 1, 3)) %>%
-  select(-variable) ->
+  select(-variable, -outlier_name) ->
   fluxdata
 
 fluxdata$Treatment <- factor(fluxdata$Treatment, levels = c("Field moisture", "Controlled drought", "Drought"))
 
 save_data(fluxdata, fname = "fluxdata_long")
 
+
 # -----------------------------------------------------------------------------
-# Does WC affect gas fluxes?
+# Examine data distribution
+# Log-transforming doesn't fix the lack of normality in our data,
+# but it's a major improvement; see graphs.
+
+fluxdata <- fluxdata %>% filter(flux_µmol_g_s > 0) 
+
+printlog("Graphing flux distributions and testing for normality...")
+p <- ggplot(fluxdata, aes(x = flux_µmol_g_s)) + geom_histogram(bins = 30)
+p <- p + facet_grid(~Gas, scales = "free") + ggtitle("Distribution of raw data")
+print(p)
+save_plot("distribution_raw")
+
+p <- ggplot(fluxdata, aes(x = log(flux_µmol_g_s))) + geom_histogram(bins = 30)
[…]

### 2016-02-08 R/5-analyze.R
Added some TODOs and model tweaks

@@ -7,6 +7,7 @@ SCRIPTNAME  	<- "5-analyze.R"
 FLUXDATA   <- file.path(outputdir(scriptfolder = FALSE), "fluxdata.csv")
 
 library(broom)  # 0.4.0
+library(reshape2) # 1.4.1
 
 # ==============================================================================
 # Main 
@@ -89,7 +90,7 @@ save_data(WC_effect)
 
 
 # -----------------------------------------------------------------------------
-# Main treatment effects
+# Effects of temperature and moisture on gas fluxes
 
 # Fitting a non-mixed-effects model for now
 # TODO: later, might want to explore having Core as a random effect
@@ -100,13 +101,20 @@ m_co2 <- lm(log(flux_µmol_g_s) ~ Temperature * WC_gravimetric,
             data = fluxdata, subset = Gas == "CO2")
 print(summary(m_co2))
 
+# TODO: a Q10 calculation. Use nls?
+# Q10 = R2/R1 ^ (10/(T2-T1))
+# fit_q10_model <- function( d ) {		# returns model or NA if nls errors out
+#   tryCatch( nls( Resp_mass ~ R20 * Q10 ^ ( ( Tair-20 )/10 ), data=d, start=c( R20=5, Q10=2 ) ),
+#             error=function( e ) NA )
+# }
+
 #lme(effort ~ Type, data = ergoStool, random = ~ 1 | Subject))
 # library(nlme)
 # m_co2_lme <- lme(log(flux_µmol_g_s) ~ Temperature * WC_gravimetric,
 #                  data = fluxdata, random = ~ 1 | Core)
 
 printlog("Fitting CH4 model...")
-m_ch4 <- lm(log(flux_µmol_g_s) ~ Temperature + WC_gravimetric,
+m_ch4 <- lm(log(flux_µmol_g_s) ~ Temperature * WC_gravimetric,
             data = fluxdata, subset = Gas == "CH4")
 print(summary(m_ch4))
 

### 2016-02-08 manuscript_openexperiment/Open_experiment.Rmd
Renamed methods ms folder

--- 2016-02-04 manuscript_methods/Open_experiment.Rmd
@@
 
-General intro.
+A wide variety of forces are pushing scientists to increase data access, both pre- and post-publication, and towards "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.
 
 Journals are adopting increasingly stringent data access and deposition policies, e.g.
-Scientific Data (http://www.nature.com/sdata/data-policies), PLOS ONE (http://journals.plos.org/plosone/s/data-availability),  {Scientific Data, 2016 #4339}{PLOS ONE, 2014 #4338}, and Science (http://www.sciencemag.org/authors/science-editorial-policies#data-deposition).
+Scientific Data (http://www.nature.com/sdata/data-policies), PLOS ONE (http://journals.plos.org/plosone/s/data-availability), and Science (http://www.sciencemag.org/authors/science-editorial-policies#data-deposition).
 These policies generally share common assumptions and goals: maximizing access to scientific data; encouraging deposition into structured repositories as opposed to journal supplementary information; and specifying that it is not acceptable for authors to be solely responsible for ensuring data access.
@@
 
-Finally, growing numbers of scientists are pushing for open data on moral/political (why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results?) grounds, also on purely scientific ones: ensuring reproducibility {Stodden, 2011 #4340}.
-This latter issue--reproducibility--is at the heart of science {Kuhn, 1962 #4341} and has a number of facets worth considering briefly.
+Finally, growing numbers of scientists are pushing for open science and data on moral/political grounds, as well as purely scientific ones, in particular ensuring reproducibility {Stodden, 2011 #4340}.
+We briefly discuss what we see as the primary arguments for data sharing and openness, before detailing a recent "open experiment" example.
 
-*Reproducibility*
+**Reproducibility**
 
-Surveying the 2000-2014 biomedical literature, Iqbal et al. {, 2016 #4334} found that none of 441 randomly-chosen studies provided raw data, and only one provided full protocols.
+Reproducibility of experimental results is at the heart of science {Kuhn, 1962 #4341}.
+Too often, however, it is impossible to attempt, because sufficient details and data are not publicly available.
+For example, surveying the 2000-2014 biomedical literature, Iqbal et al.
+{, 2016 #4334} found that none of 441 randomly-chosen studies provided raw data, and only one provided full protocols.
 
-"Reproducibility of experimental results is a fundamental criterion for credible scientific research" {Thornton, 2005 #2310}.
-More difficult for modeling work; this is a subset of the more general problem with respect to digital media, software, etc.
-(CITATION?).
+As code and data have become increasingly intertwined, the availability of the former has become a fundamental problem as well.
+Modeling groups have worked to improve reproducibility and archiving praactices {Thornton, 2005 #2310}.
+For scientists without programming skills, scripted systems for data analysis!
+E.g.
+R {R Development Core Team, 2016 #4314}, Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record information about the data processing, analytical process and decisions, and statistical analysis.
 
-Scripted systems for data analysis!
-E.g.
-R {R Development Core Team, 2016 #4314}, Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record all infrmation about the data processing, analytical process and decisions, and statistical analysis.
-This and other best practices of scientific computing {Wilson, 2014 #4337} help produce reproducible, robust, and transparent science.
-
-*Data loss*
+**Data loss**
 
 Vines et al.
-{, 2014 #4130} published a shocking result, based on surveying 516 articles from 2 to 22 years old: the odds of a data set being available fell by 17% each year, and the chances that the contact author's email address still worked declined by 7% per year.
-Reichman et al.
+{, 2014 #4130} published a shocking result, based on surveying 516 biology articles from 2 to 22 years old: the odds of a data set being available fell by 17% each year, and the chances that the contact author's email address still worked declined by 7% per year.
+Similarly, Reichman et al.
 {, 2011 #2862} estimated that less than 1% of ecological data collected is made available after publication, and noted, as an example, that most current and historical data relevant to the recent *Deepwater Horizon* oil spill are already inaccessible or lost {Reichman, 2011 #2862}.
-
-Scientists' use of strong and consistent data curation practices {Hart, 2015 #4292}{Rüegg, 2014 #3886} can mitigate the problem, but all our anecdotal experience and quantitative studies {Vines, 2014 #4130} suggest that in the long term, data cannot be reliably preserved by individual researchers.
+In global change ecology, Wolkovich et al.
+{, 2012 #3303} reported that only ~10% of raw data sets were eventually shared in their attempt to conduct a high-profile meta-analysis {Wolkovich, 2012 #3331}.
 
 Data loss hits ecosystem, soil, and global change ecology particularly hard, as global to local climate changes make ecological data effectively irreproducible {Wolkovich, 2012 #3303}: we can never remeasure exactly the same system state.
+A subtler data loss issue is the "file drawer problem" {Rosenthal, 1979 #2291}, where unpublished but potentially valuable data are lost fovever.
+Scientists' use of strong and consistent data curation practices {Hart, 2015 #4292}{Rüegg, 2014 #3886} can mitigate the problem, but all our anecdotal experience and quantitative studies {Vines, 2014 #4130} suggest that in the long term, data cannot be reliably preserved by individual researchers.
+This in turns means that the role of established, structured data and code repositories is critical.
 
-A subtler data loss issue is the "file drawer problem" {Rosenthal, 1979 #2291}.
+VLB: Supplementary Information sections have become repositories of data and of non-core analyses. How does our open experiment replace (improve upon?) SI? Thoughts?
 
-*Data and code quality*
+**Data and code quality**
 
@@
 
-Potential to increase pulic trust in research results.
+**Enabling synthetic and global-scale analyses**
 
-*Enabling synthetic and global-scale analyses*
+With greater data availability comes the potential to assemble large, internally consistent data sets--examples in biogeosciences and ecology include BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, and SRDB {Bond-Lamberty, 2010 #2320}.
+In turn, these data enable powerful meta-analyses and syntheses {Wolkovich, 2012 #3303}.
 
-Many refs in {Wolkovich, 2012 #3303}.
+**Institutional and social trust**
 
-BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, SRDB {Bond-Lamberty, 2010 #2320}.
+Why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results?
+
+Potential to increase public trust in research results.
+
+Governmental efforts {Holdren, 2013 #4344}.
 
@@
 
-This is example, we hope, of the impact of individual scientists {Wolkovich, 2012 #3303}.
+In early 2015, we planned a laboratory incubation experiment to characterize the chemical and biological properties of sub-Arctic, active layer soils subjected to changes in temperature and moisture.
+In this experiment, we would measure greenhouse gas fluxes from soil cores over 100 days, and measure the cores' physical, chemical, and biological characteristics under temperature and moisture changes.
+This required (i) a multidisciplinary team that was not located in one time zone; (ii) integrating a variety of different data; (iii) performing quality control and diagnostics rapidly, so if e.g.
+instrument problems arose we would lose only the minimum amount of time and data; (iv) tightly integrating data, statistical analyses, and manuscript results.
 
-* GitHub
-* Diagnostics page(s)
-* Data: raw to summarized to final
-* All statistical analyses
-* "LRB" - notes 
-* Manuscript(s)
+We used GitHub, a web-based Git repository hosting service, to...
+  "Git"" is a version control system...
+  A "repository"" is ...
+  A "web-based hosting service" is...
+
+The design of our repository is shown in **Figure 1**. Explain...
+
+This system had a number of desirable capabilities and characteristics:
+
+* The entire data processing and analytical system is online and documented.
+* The version control system let us make incremental changes, work out problems, look at histories (i.e., who made what change when).
+A issue tracker let us discuss problems, reference changes in the repository, create to-do lists, and assign responsibilities.
+* The public nature of the repository encouraged us to flesh out documentation, use clean and clear coding, and think about the longer-term ramifications of decisions.
+* Real-time diagnostics... (**Figure 2**)... Investing the effort to set up a software pipeline BEFOREHAND meant that we could reliably and easily identify and diagnose problems.
+* Everything necessary to reproduce study results.
+* Manuscripts as well!
+In fact, the "source code" for this manuscript, as well as its history, can be viewed at https://github.com/bpbond/cpcrw_incubation/blob/master/manuscript_methods/Open_experiment.Rmd.
+* Information webpage on the experiment and broader project.
 
@@
 
-VLB: Supplementary Information sections have become repositories of data and of non-core analyses. How does our open experiment replace (improve upon?) SI? Thoughts?
+Transparency is no magic panacea ensuring reproducibility {Easterbrook, 2014 #4137}, and it raises genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
+Scientists also frequently raise a variety of other concerns about data sharing, for example being "scooped" not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
+We cannot easily dismiss of these these: for example, does our community adequately credit researchers who contribute to global databases that subsequently produce high-impact papers?
+Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers efforts are adequately valued and cited {Kueffer, 2011 #4335}.
+We also recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}.
 
-Do we adequately credit researchers who contribute to global databases {Kattge, 2011 #3064} that subsequently produce high-impact papers {Díaz, 2016 #4318}?
+Nor is the model we present here--this public repository that was updated throughout the experiment and publication process--perfect.
+First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ: for example, there was no unit testing to ensure module correctness, nor an automated build tool automating the workflow (i.e., the individual scripts all have to be run manually).
+Second, it may not not be appropriate for applicable to all study types, even just in our professional fields of soil, ecosystem, and global change science.
+(And it certainly is not appropriate for sensitive human-subject data.) Third, Git and GitHub are designed for working with code, not data, and this causes some ...
 
-Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers efforts are adequately valued and cited {Kueffer, 2011 #4335}.
+Nonetheless... conclude.
 
-We don't present this model as perfect.
-First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ: for example, there was no unit testing to ensure module correctness, nor an automated build tool automating the workflow (i.e., the individual scripts all have to be run manually).
-
-Nor is this examine appropriate, or applicable, to all scientific research, even just in the fields of soil, ecosystem, and global change science.
-
-Potential concerns
-
-* Fear of "being scooped"
-* Question of open from beginning or only after publication
-* Privacy concerns
+This is example, we hope, of the impact of individual scientists' decisions and practices {Wolkovich, 2012 #3303}.
 
 **References**
-
-TODO - maybe a second figure showing one of the diagnostic plots, partway through the incubation?
 
@@
 
-![Repository and workflow](figureX.png)
+![Repository and workflow](figure1.png)
+
+**Figure 2.** Two of the many diagnostic figures that were produced, in real time, as the incubation experiment proceeded.
+They show (top) the number of replicate measurements made on each soil core, by date of measurement, and (bottom) measurements with a data mismatch problem, indicating a data entry or other error (in the figure no errors, which would be in red, are extant).
+A wide range of similar diagnostic graphics (see https://github.com/bpbond/cpcrw_incubation#current-diagnostics) and statistics let us quickly identify problems or interesting patterns in the data.
+
+![Sample diagnostics](figure2.png)

### 2016-02-08 manuscript/cpcrw_manuscript.Rmd
Added first table!
…and sketched out others and figures

@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, results='hide', message=FALSE, fig.keep='none'}
 library(dplyr, warn.conflicts = FALSE)
+setwd("../")
+source("R/5-analyze.R")
 ```
@@
 sdcn <- readLines("../outputs/summarydata_clean.csv") %>% length
-en <- readLines("../outputs/2.5-remove/excluded_data.csv") %>% length
-rn <- readLines("../outputs/2.5-remove/removed_data.csv") %>% length
+en <- readLines("../outputs/3-remove/excluded_data.csv") %>% length
+rn <- readLines("../outputs/3-remove/removed_data.csv") %>% length
 ```
 
-Anomalous data were excluded based on their being more than 3.2 mean absolute deviations {Davies, 1993 #3138} from the treatment and temperature mean.
+Anomalous data were excluded based on their being more than 3.2 mean absolute deviations {Davies, 1993 #3138} from the treatment mean on each day.
 We excluded `r en` of `r sdn` (`r round(en/sdn*100)`%) of the summarized samples for this reason.
@@
 
-### Acknowledgments
+#### Acknowledgments
 
@@
 
-Table X. Summary of temperature, WC effects on gas fluxes.
+#### References
 
-Figure X. A faceted (Temperature ~ Treatment) plot showing normalized flux means and errors by incubation time. (Basically the figure already in the README diagnostics.)
+**Table 1.** Summary of gravimetric water content effects on gas fluxes, by treatment and temperature.
 
-Figure X. Core water content across the course of the experiment. (Basically the figure already in the README diagnostics.)
+```{r kable, echo=FALSE}
+# Print out the `WC_effect` data frame in a nice-looking table
+WC_effect$Gas[WC_effect$Gas == "CO2"] <- "CO<sub>2</sub>"  # html
+WC_effect$Gas[WC_effect$Gas == "CH4"] <- "CH<sub>4</sub>"  # html
+knitr::kable(WC_effect, digits = 3, 
+             col.names = c("Treatment", "Temperature", "Gas", "R<sup>2</sup>", "Sigma", "P", "AIC"))
+```
 
-Figure X. A pairs plot of variable correlations?
+**Table 2.** Effect of temperature and WC on cumulative C fluxes for each gas.
+
+**Figure 1.** Core water content across the course of the experiment. (Basically the figure already in the README diagnostics.)
+
+**Figure 2.** A faceted (Temperature ~ Treatment) plot showing normalized flux means and errors by incubation time. (Basically the figure already in the README diagnostics.)
+
+**Figure 3.** Cumulative fluxes over the incubation.

### 2016-02-08 R/5-analyze.R

@@ -79,15 +79,36 @@ printlog("Summarizing WC effect...")
 fluxdata %>%
   filter(flux_µmol_g_s > 0) %>%
   group_by(Treatment, Temperature, Gas) %>%
+  filter(!outlier) %>%
   do(mod = lm(log(flux_µmol_g_s) ~ WC_gravimetric, data = .)) %>%
   glance(mod) %>%
-  select(Treatment, Temperature, Gas, adj.r.squared, sigma, p.value, AIC) %>%
-  mutate(signif = p.value < 0.05) ->
+  select(Treatment, Temperature, Gas, adj.r.squared, sigma, p.value, AIC) ->
+#  mutate(signif = p.value < 0.05) ->
   WC_effect
 
 print(WC_effect)
 save_data(WC_effect)
 
+# -----------------------------------------------------------------------------
+# Per-core models - look at R2, etc. variability
+
+# printlog("Per-core models...")
+# fluxdata %>%
+#   filter(flux_µmol_g_s > 0) %>%
+#   group_by(Core, Gas) %>%
+#   do(mod = lm(log(flux_µmol_g_s) ~ Temperature * WC_gravimetric, data = .)) %>%
+#   glance(mod) %>%
+#   select(Core, Gas, adj.r.squared, sigma, p.value, AIC) ->
+#   per_core_models
+# 
+# print(per_core_models)
+# save_data(per_core_models)
+# 
+# p <- ggplot(per_core_models, aes(adj.r.squared)) + geom_histogram(bins = 30)
+# p <- p + facet_grid(Gas ~ .) 
+# p <- p + ggtitle("Per-core temp, WC models")
+# print(p)
+# save_plot("per_core_models")
 
 # -----------------------------------------------------------------------------
 # Effects of temperature and moisture on gas fluxes

### 2016-02-08 manuscript/cpcrw_manuscript.Rmd
Added two figures (fluxes over time)

@@
 ---
-title: "Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils"
+title: Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils
 author: "Ben Bond-Lamberty, Peyton Smith, Vanessa Bailey"
-#date: "February 1, 2016"
 output: html_document
@@
 ```{r, echo=FALSE}
-# compute outlier and exclusion numbers
+# TODO: this isn't good--fragile if files move or change names
+# Want everything in-memory (or something)
+# Compute outlier and exclusion numbers
 sdn <- readLines("../outputs/summarydata.csv") %>% length
@@
 WC_effect$Gas[WC_effect$Gas == "CH4"] <- "CH<sub>4</sub>"  # html
-knitr::kable(WC_effect, digits = 3, 
-             col.names = c("Treatment", "Temperature", "Gas", "R<sup>2</sup>", "Sigma", "P", "AIC"))
+knitr::kable(WC_effect, digits = 3, col.names = c("Gas", "Treatment", "Temperature", "R<sup>2</sup>", "Sigma", "P", "AIC"))
 ```
@@
 
-**Figure 2.** A faceted (Temperature ~ Treatment) plot showing normalized flux means and errors by incubation time. (Basically the figure already in the README diagnostics.)
+**Figure 2.** A faceted (Temperature ~ Treatment) plot showing normalized CO2 flux means and errors by incubation time. (Basically the figure already in the README diagnostics.)
 
-**Figure 3.** Cumulative fluxes over the incubation.
+```{r, echo=FALSE}
+print(figureB)
+```
+
+**Figure 3.** A faceted (Temperature ~ Treatment) plot showing normalized CH4 flux means and errors by incubation time. (Basically the figure already in the README diagnostics.)
+
+```{r, echo=FALSE}
+print(figureC)
+```
+
+**Figure 4.** Cumulative fluxes over the incubation.

### 2016-02-08 R/5-analyze.R

@@ -36,6 +36,34 @@ save_data(fluxdata, fname = "fluxdata_long")
 
 
 # -----------------------------------------------------------------------------
+# Fluxes over time figure
+
+printlog("fluxes over time...")
+fluxdata %>%
+  mutate(incday = floor(inctime_days)) %>%
+  group_by(Gas, Temperature,Treatment, incday) %>%
+  summarise(flux = mean(flux_µmol_g_s), flux_sd = sd(flux_µmol_g_s)) ->
+  fluxdata_figsBC
+
+figsBC <- function(fd) {
+  panellabels <- 
+  ggplot(fd, aes(incday, flux)) + 
+    geom_point() + 
+    facet_grid(Temperature ~ Treatment) + 
+    geom_errorbar(aes(ymin = flux - flux_sd, ymax = flux + flux_sd)) +
+    xlab("Incubation day") + ylab(expression(µmol~g^{-1}~s^{-1}))
+}
+
+fluxdata_figsBC %>%
+  filter(Gas == "CO2") %>%
+  figsBC ->
+  figureB
+fluxdata_figsBC %>%
+  filter(Gas == "CH4") %>%
+  figsBC ->
+  figureC
+
+# -----------------------------------------------------------------------------
 # Examine data distribution
 # Log-transforming doesn't fix the lack of normality in our data,
 # but it's a major improvement; see graphs.
@@ -78,12 +106,12 @@ save_data(shapiro_trans)
 printlog("Summarizing WC effect...")
 fluxdata %>%
   filter(flux_µmol_g_s > 0) %>%
-  group_by(Treatment, Temperature, Gas) %>%
+  group_by(Gas, Treatment, Temperature) %>%
   filter(!outlier) %>%
   do(mod = lm(log(flux_µmol_g_s) ~ WC_gravimetric, data = .)) %>%
   glance(mod) %>%
-  select(Treatment, Temperature, Gas, adj.r.squared, sigma, p.value, AIC) ->
-#  mutate(signif = p.value < 0.05) ->
+  select(Gas, Treatment, Temperature, adj.r.squared, sigma, p.value, AIC) ->
+  #  mutate(signif = p.value < 0.05) ->
   WC_effect
 
 print(WC_effect)

### 2016-02-08 manuscript/cpcrw_manuscript.Rmd
Update cpcrw_manuscript.Rmd
added methods text

--- 2016-02-08 manuscript/cpcrw_manuscript.Rmd
@@
 The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density.
-Subsamples were collected and stored at -20 °C for potential enzyme activity and ATP analysis.
+Subsamples were collected and stored at -20 °C for dissolved organic carbon measurements.
 
@@
 Dry CH~4~ and CO~2~ concentrations were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group.
-Cores were weighed immediately after (PEYTON) gas measurements.
+Cores were weighed immediately after gas measurements.
 Ambient air was measured between treatment groups, and before starting measurements in a chamber.
+BEN: How was ambient air measurements used in this study?
+don't see any mention of it in remaining text, etc.
 
-The incubation experiment concluded on 10 December 2015. PEYTON: text re C, N, etc. analyses.
+The incubation experiment concluded on 9 December 2015, following the final CO~2~ and CH~4~ readings.
+Each soil core was maintained at the treatment-dependent temperature and moisture content (by weight) until removed for destructive sampling, December 14 - 18, 2015.
+Subsamples were collected and composited throughout each soil core for future dissolved organic carbon analysis (110 +/- 24 g dry weight equivalent) and dry-weight calculations (~ 28 g) upon destrucitve core sampling.
+The remaining core material was air-dried and then separated into particles > 2mm diameter and soil < 2mm diameter using a U.S.
+Standard Test Sieve, No.
+10 (Fisherbrand,  Pittsburg, PA, USA).
+Both the dry weight and the volume of soil ( < 2 mm) were used in calculations of gravimetric and volumetric soil moisture content (_Gardner, WH.
+1986 Chapter 21 in Klute, A., Ed.
+Methods of Soil Analysis.
+Part 1.
+Physical and mineralogical methods.
+2nd ed.
+Soil Science Society of American, Inc.
+Madison, Wisconsin_).
+Soil volume was caclulated as the total core volume - volume of particles greater than 2 mm diameter, determined using water displacement of particles > 2 mm diameter.
+Air-dried soil (< 2 mm diamter) and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia extension in February, 2016 for total C, N of dried soil samples and dissolved organic carbon from samples stored at -20 °C.
 

### 2016-02-09 R/5-analyze.R
Added a Figure A - water content changes

@@ -34,6 +34,13 @@ fluxdata$Treatment <- factor(fluxdata$Treatment, levels = c("Field moisture", "C
 
 save_data(fluxdata, fname = "fluxdata_long")
 
+# -----------------------------------------------------------------------------
+# Water content over time figure
+
+figureA <- ggplot(fluxdata_orig, aes(inctime_days, WC_gravimetric * 100, color=Treatment, group=Core)) 
+figureA <- figureA + geom_point() + geom_line()
+figureA <- figureA + facet_grid(~Temperature) 
+figureA <- figureA + xlab("Incubation day") + ylab("Gravimetric water content (%)")
 
 # -----------------------------------------------------------------------------
 # Fluxes over time figure

### 2016-02-09 manuscript/cpcrw_manuscript.Rmd
Tweaks to Peyton's methods text

--- 2016-02-08 manuscript/cpcrw_manuscript.Rmd
@@
 Dry CH~4~ and CO~2~ concentrations were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group.
-Cores were weighed immediately after (PEYTON) gas measurements.
+Cores were weighed immediately after gas measurements.
 Ambient air was measured between treatment groups, and before starting measurements in a chamber.
+BEN: How was ambient air measurements used in this study?
+don't see any mention of it in remaining text, etc.
+(TODO)
 
-The incubation experiment concluded on 10 December 2015. PEYTON: text re C, N, etc. analyses.
+The incubation experiment concluded on 9 December 2015, following the final CO~2~ and CH~4~ readings.
+Each soil core was maintained at the treatment-dependent temperature and moisture content (by mass) until removed for destructive sampling, December 14-18, 2015.
+Sub-samples were collected and composited throughout each soil core for dissolved organic carbon analysis (110 ± 24 g dry mass equivalent) and dry-mass calculations (~28 g each).
+The remaining core material was air-dried and separated into particles (>2 mm diameter) and soil (<2 mm) using a U.S.
+Standard Test Sieve No.
+10 (Fisherbrand,  Pittsburg, PA, USA).
+The dry mass and volume of soil were used in calculations of gravimetric and volumetric soil moisture content, respectively {Gardner, 1986 #4345}.
+Soil volume was caclulated as the total core volume minus the volume of particles >2 mm diameter, with the latter determined by water displacement.
+Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic C.
+TODO: get UGA analytical methods.
 
@@
 
+```{r, echo=FALSE}
+print(figureA)
+```
+
 **Figure 2.** A faceted (Temperature ~ Treatment) plot showing normalized CO2 flux means and errors by incubation time. (Basically the figure already in the README diagnostics.)

### 2016-02-09 R/4-fluxes.R
Fixed gravimetric WC calculation
Whoops.

@@ -29,7 +29,9 @@ summarydata <- left_join(summarydata, drymassdata, by = "Core")
 
 summarydata %>%
   filter(Treatment != "Ambient") %>%
-  mutate(WC_gravimetric = (Mass_g - SoilDryMass_g) / SoilDryMass_g) %>%
+  mutate(WC_gravimetric = 
+           (Mass_g - CoreSleeveMass_g - NonsoilMassLarge_g - SoilDryMass_g) / 
+           (SoilDryMass_g)) %>%
   select(Core, SoilDryMass_g, SoilVolume_cm3, WC_gravimetric,
          Treatment, Temperature, CO2_ppm_s, CH4_ppb_s, inctime_days) ->
   fluxdata

### 2016-02-09 R/6-plots.R
README update

@@ -1,78 +1,0 @@
-# NOTE THIS DOESN'T WORK WITH CURRENT DATASET YET
-# Ben Bond-Lamberty July 2015
-
-source("0-functions.R")
-
-SCRIPTNAME  	<- "4-plots.R"
-FLUXDATA      <- paste0(OUTPUT_DIR, "fluxdata.csv")  # output from script 3
-RAWDATA_SAMPLES      <- paste0(OUTPUT_DIR, "rawdata_samples.csv.gz")  # output from script 2
-
-# ==============================================================================
-# Main 
-
-sink(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), split=T) # open log
-
-printlog("Welcome to", SCRIPTNAME)
-
-printlog("Reading in flux data...")
-fluxdata <- read_csv(FLUXDATA)
-fluxdata$DATETIME <- ymd_hms(fluxdata$DATETIME, tz="America/Los_Angeles")
-print_dims(fluxdata)
-print(summary(fluxdata))
-
-fluxdata <- subset(fluxdata, !is.na(CORE))
-
-p <- ggplot(fluxdata, aes(DATETIME, CORE, color=STRUCTURE)) + geom_point()
-p <- p + ggtitle("QC - measurement time of day")
-print(p)
-save_plot("QC_time_of_day")
-
-sdata <- subset(fluxdata, CORE != "Ambient")
-
-# Histograms of the flux numbers
-p <- ggplot(sdata, aes(x = CO2_flux_mgC_hr)) + geom_histogram()
-p <- p + facet_grid(STRUCTURE~WETTING)
-print(p)
-save_plot("QC_CO2_flux_distribution")
-p <- ggplot(sdata, aes(x = CH4_flux_mgC_hr)) + geom_histogram()
-p <- p + facet_grid(STRUCTURE~WETTING)
-print(p)
-save_plot("QC_CH4_flux_distribution")
-
-p <- ggplot(sdata, aes(MOISTURE, CO2_flux_mgC_hr, color=WETTING)) + geom_boxplot()
-p <- p + facet_grid(STRUCTURE~INPUT)
-print(p)
-save_plot("QC_CO2")
-
-p <- ggplot(sdata, aes(MOISTURE, CH4_flux_mgC_hr, color=WETTING)) + geom_boxplot()
-p <- p + facet_grid(STRUCTURE~INPUT)
-print(p)
[…]

### 2016-02-09 R/2-summarize.R
Updated to 20 C throughout, reran diagnostics

@@ -32,7 +32,7 @@ qc_valvemap <- function(valvemap) {
     PROBLEM <- TRUE
   }
   nomass <- valvemap %>% 
-    filter(Core != "Ambient4" & Core != "Ambient22" & is.na(Mass_g))
+    filter(Core != "Ambient4" & Core != "Ambient20" & is.na(Mass_g))
   if(nrow(nomass)) {
     printlog("WARNING - some cores have no mass data")
     print(nomass)

### 2016-02-09 R/3.3-qc.R

@@ -65,7 +65,7 @@ p <- p + scale_size_discrete(guide = FALSE)
 p <- p + ggtitle("Orphan samples (no matching date/valve info)")
 save_diagnostic(p, "orphan_samples")
 
-p <- ggplot(subset(summarydata, Core != "Ambient4" & Core != "Ambient22"), 
+p <- ggplot(subset(summarydata, Core != "Ambient4" & Core != "Ambient20"), 
             aes(DATETIME, Core, size = is.na(Mass_g), color = !is.na(Mass_g)))
 p <- p + geom_point() 
 p <- p + scale_color_manual("Has mass data", values = c("TRUE" = "grey", "FALSE" = "red"))
@@ -98,7 +98,7 @@ save_data(treatment_summary, scriptfolder = FALSE)
 printlog(SEPARATOR)
 printlog("Computing core CVs...")
 core_cv <- summarydata %>% 
-  filter(Core != "Ambient4" & Core != "Ambient22") %>%
+  filter(Core != "Ambient4" & Core != "Ambient20") %>%
   group_by(Date, Core) %>% 
   summarise(CO2_ppm_s_cv = sd(CO2_ppm_s) / mean(CO2_ppm_s),
             min_CO2_time = mean(min_CO2_time),
@@ -202,7 +202,7 @@ printlog("Individual cores over time")
 # Reorder the 'Core' factor so that facets are ordered by Treatment and Temperature
 summarydata$TT <- paste(summarydata$Treatment, summarydata$Temperature)
 sd_no_A <- summarydata %>% 
-  filter(Core != "Ambient4" & Core != "Ambient22") %>% 
+  filter(Core != "Ambient4" & Core != "Ambient20") %>% 
   arrange(TT)
 sd_no_A$Core <- factor(sd_no_A$Core, levels = unique(as.character(sd_no_A$Core)))
 p <- qplot(inctime_days, CO2_ppm_s, data = sd_no_A, color = TT)
@@ -211,7 +211,7 @@ p <- p + facet_wrap(~Core, ncol = 6)
 print(p)
 save_plot("CO2_incday")
 
-p <- qplot(incday, CH4_ppb_s, data = sd_no_A, color = paste(Treatment, Temperature))
+p <- qplot(inctime_days, CH4_ppb_s, data = sd_no_A, color = paste(Treatment, Temperature))
[…]

### 2016-02-09 manuscript/cpcrw_manuscript.Rmd
Outlining results and discussion

@@
 
-* joined with mass data
-* flux computation
-
 The GHG concentration measurements for each sample (i.e., each gas, core, and date/time) were used to calculate a linear rate of change (δc/δt), based on the concentration rise from a minimum (at 2-10 seconds after measurement began) to a maximum (at 10-45 seconds).
@@
 {, 2002 #3383} as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where V is the core-specific system volume, M the core dry mass as determined at the end of the incubation, Pa atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), R the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and T the chamber air temperature (K) at time of measurement.
-The final respiration rate was expressed as mg C kg soil^-1^ day^-1^.
+The final respiration rate was expressed on a dry soil mass basis (mg C kg soil^-1^ day^-1^).
 TODO check all units
 
-Q10 computation
-
+* joined with mass and ancillary data
+* correlations
+* Q10 computation
 * stats
@@
 
-Results.
+* Soil WC changes over incubation (Figure 1)
+
+* Gas fluxes over incubation (Figures 2 and 3)
+* Statistical tests of WC and temperature on fluxes
+
+* Cumulative fluxes (Figure 4)
+* CO2:CH4 ratio
+* Statistical tests of WC and temperature on cumulative fluxes
 
@@
 
-Discussion.
+Compare absolute flux rates to other published studies from boreal mineral soils.
 
@@
 
+Note high CO2:CH4 ratios, compare with Treat et al. and other studies.
+
+Rises in boreal soil temperatures - who has quantified this? Changes in soil WC much less certain. Thermokarst etc., deepening of active layer - that's the larger implications of this work.
+
 #### Acknowledgments
 
-Acknowledgments.
+This research was supported by Office of Science of the U.S.
+Department of Energy as part of the Terrestrial Ecosystem Sciences Program.
+The Pacific Northwest National Laboratory is operated for DOE by Battelle Memorial Institute under contract DE-AC05-76RL01830.
 

### 2016-02-09 manuscript/cpcrw_manuscript.Rmd
Manuscript tweaks, plot fixes

@@
 title: Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils
-author: "Ben Bond-Lamberty, Peyton Smith, Vanessa Bailey"
+author: "Ben Bond-Lamberty^1^, Peyton Smith^2^, Vanessa Bailey^2^"
 output: html_document
 ---
+^1^Joint Global Change Research Institute, DOE Pacific Northwest National Laboratory, College Park, MD USA; 
+bondlamberty@pnnl.gov
+
+^2^Biological Sciences Division, Pacific Northwest National Laboratory, Richland, WA USA
 
@@
 #### Abstract
-
-Peyton potential text to use in abstract:
-
->The goal of this study was to characterize the chemical and biological properties of sub-Arctic, active layer soils subjected to changes in temperature and moisture.
-We also sought to determine the relative control temperature and moisture have on the fate of C and greenhouse gas emissions in active layer soils lying directly above permafrost.
-We measured carbon dioxide and methane emissions on soil at two different temperatures (4 and 20C) and moisture contents (field moist and drought-induced) throughout a 100-day incubation.
-We also determined soil structure, extracellular enzyme activity, total nitrogen, carbon, and the molecular composition of C pre and post-incubation cores.
-We hypothesized that the field moist 20˚C cores have a greater potential to lose C as both CO2 and CH4 than cores held at 4˚C due to increased aerobic and anaerobic microbial activity.
-We also hypothesize that drought-induced cores will have a higher potential to lose more ‘complex’ carbon compounds (aromatic-C) as these compounds are desorbed due to increased surface tension.
 
@@
 Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
-We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at difference temperature and moisture levels.We hypothesized that soils maintained at field moisture and high temperature would lose more C, both as CO~2~ and CH~4~, than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity.
+We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at difference temperature and moisture levels.
+We hypothesized that soils maintained at field moisture and high temperature would lose more C, both as CO~2~ and CH~4~, than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity.
 In addition, cores subject to drought (reduced soil moisture) would have a higher potential to lose more ‘complex’ carbon compounds (aromatic C) as these compounds are desorbed due to increased surface tension.
@@
 
-Anomalous data were excluded based on their being more than 3.2 mean absolute deviations {Davies, 1993 #3138} from the treatment mean on each day.
+Anomalous data were excluded based on their being more than 3.0 (for CO~2~) or 5.0 (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within each day, treatment, and temperature.
 We excluded `r en` of `r sdn` (`r round(en/sdn*100)`%) of the summarized samples for this reason.
@@
 All data analysis and statistics were performed using R version 3.2.3 {R Development Core Team, 2016 #4314}.
-This experiment was run as an "open experiment" with all analysis code, data (raw to final), diagnostics, etc., available in real time at https://github.com/bpbond/cpcrw_incubation.
-The flux data backing the main results have also been archived at TODO.
+This experiment was run as an "open experiment" with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
+The flux data backing the main results have been archived at TODO.
 
@@
 
+* Soil C, N, DOC means
+
 * Gas fluxes over incubation (Figures 2 and 3)
-* Statistical tests of WC and temperature on fluxes
+* Statistical tests of WC, T, C, N on individual core fluxes (Table 2)
 
 * Cumulative fluxes (Figure 4)
-* CO2:CH4 ratio
+* CO2:CH4 ratios
 * Statistical tests of WC and temperature on cumulative fluxes
@@
 
+------
+
 **Table 1.** Summary of gravimetric water content effects on gas fluxes, by treatment and temperature.
@@
 
-**Table 2.** Effect of temperature and WC on cumulative C fluxes for each gas.
+**Table 2.** Statistical tests of WC, T, C, N on individual core fluxes.
 
-**Figure 1.** Core water content across the course of the experiment. (Basically the figure already in the README diagnostics.)
+**Figure 1.** Core water content across the course of the experiment.
 
@@
 
-**Figure 2.** A faceted (Temperature ~ Treatment) plot showing normalized CO2 flux means and errors by incubation time. (Basically the figure already in the README diagnostics.)
+**Figure 2.** A faceted (Temperature ~ Treatment) plot showing normalized CO2 flux means and errors by incubation time.
 
@@
 
-**Figure 3.** A faceted (Temperature ~ Treatment) plot showing normalized CH4 flux means and errors by incubation time. (Basically the figure already in the README diagnostics.)
+**Figure 3.** A faceted (Temperature ~ Treatment) plot showing normalized CH4 flux means and errors by incubation time.
 
@@
 **Figure 4.** Cumulative fluxes over the incubation.
+
+```{r, echo=FALSE, warning=FALSE}
+print(figureD)   # TODO: wtf is going on here?
+```

### 2016-02-09 R/4-fluxes.R

@@ -192,15 +192,15 @@ fd_cumulative$Treatment <- factor(fd_cumulative$Treatment,
 fd_cumulative$Temperature <- as.factor(fd_cumulative$Temperature)
 
 # Add a few dummy rows so graph bars are spaced correctly
-dmy <- expand.grid(Treatment = unique(fd_cumulative$Treatment),
+dmy <- data.frame(Treatment = "Controlled drought",
                    Temperature = 4,
                    Gas = unique(fd_cumulative$Gas),
+                   cum_flux_mgC = NA,
+                   cum_flux_mgC_sd = NA,
                    stringsAsFactors = FALSE)
-dmy$cum_flux_mgC_sd <- NA
-dmy$cum_flux_mgC <- NA
-fd_cumulative <- rbind(fd_cumulative, dmy)
+fluxdata_cumulative <- rbind(fd_cumulative, dmy)
 
-p3 <- ggplot(fd_cumulative, aes(Temperature, cum_flux_mgC, fill = Treatment)) + 
+p3 <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Treatment)) + 
   geom_bar(stat = 'identity', position = position_dodge()) +
   geom_errorbar(aes(color = Treatment, 
                     ymin = cum_flux_mgC * 0.9, 
@@ -210,6 +210,7 @@ p3 <- ggplot(fd_cumulative, aes(Temperature, cum_flux_mgC, fill = Treatment)) +
   ylab(paste("Cumulative C (mg) over", floor(max(fluxdata$inctime_days)), "days")) +
   ggtitle("Cumulative C by gas, treatment, temperature")
 save_diagnostic(p3, "cumulative_gas")
+save_data(fluxdata_cumulative, scriptfolder = FALSE)
 
 # -----------------------------------------------------------------------------
 # Done!

### 2016-02-09 R/5-analyze.R

@@ -5,6 +5,7 @@ source("R/0-functions.R")
 
 SCRIPTNAME  	<- "5-analyze.R"
 FLUXDATA   <- file.path(outputdir(scriptfolder = FALSE), "fluxdata.csv")
+FLUXDATA_CUMULATIVE <- file.path(outputdir(scriptfolder = FALSE), "fluxdata_cumulative.csv")
 
 library(broom)  # 0.4.0
 library(reshape2) # 1.4.1
@@ -71,6 +72,23 @@ fluxdata_figsBC %>%
   figureC
 
 # -----------------------------------------------------------------------------
+# Cumulative flux figure
+
+fluxdata_cumulative <- read_csv(FLUXDATA_CUMULATIVE)
+fluxdata_cumulative$Treatment <- factor(fluxdata_cumulative$Treatment, 
+                                        levels = c("Field moisture", "Controlled drought", "Drought"))
+fluxdata_cumulative$Temperature <- as.factor(fluxdata_cumulative$Temperature)
+
+figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Treatment)) + 
+  geom_bar(stat = 'identity', position = position_dodge()) +
+  geom_errorbar(aes(color = Treatment, 
+                    ymin = cum_flux_mgC * 0.9, 
+                    ymax = cum_flux_mgC + cum_flux_mgC_sd), 
+                position = position_dodge(0.9), width = 0.4) +  
+  facet_grid(Gas ~ ., scales = "free") +
+  ylab(paste("Cumulative C (mg) over", floor(max(fluxdata$inctime_days)), "days"))
+
+# -----------------------------------------------------------------------------
 # Examine data distribution
 # Log-transforming doesn't fix the lack of normality in our data,
 # but it's a major improvement; see graphs.
@@ -116,7 +134,7 @@ fluxdata %>%
   group_by(Gas, Treatment, Temperature) %>%
   filter(!outlier) %>%
   do(mod = lm(log(flux_µmol_g_s) ~ WC_gravimetric, data = .)) %>%
-  glance(mod) %>%
+  broom::glance(mod) %>%
   select(Gas, Treatment, Temperature, adj.r.squared, sigma, p.value, AIC) ->
   #  mutate(signif = p.value < 0.05) ->
   WC_effect

### 2016-02-09 manuscript/cpcrw_manuscript.Rmd
Manuscript tweaks

@@
 Sub-samples were collected and composited throughout each soil core for dissolved organic carbon analysis (110 ± 24 g dry mass equivalent) and dry-mass calculations (~28 g each).
-The remaining core material was air-dried and separated into particles (>2 mm diameter) and soil (<2 mm) using a U.S.
+The remaining core material was air-dried and separated into particles (≥2 mm diameter) and soil (<2 mm) using a U.S.
 Standard Test Sieve No.
@@
 The dry mass and volume of soil were used in calculations of gravimetric and volumetric soil moisture content, respectively {Gardner, 1986 #4345}.
-Soil volume was caclulated as the total core volume minus the volume of particles >2 mm diameter, with the latter determined by water displacement.
+Soil volume was caclulated as the total core volume minus the volume of particles ≥2 mm diameter, with the latter determined by water displacement.
 Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic C.

### 2016-02-09 manuscript_openexperiment/Open_experiment.Rmd
Manuscript tweaks

@@
   science'
-author: "Ben Bond-Lamberty, Peyton Smith, and Vanessa Bailey"
+author: "Ben Bond-Lamberty^1^, Peyton Smith^2^, Vanessa Bailey^2^"
 date: "February 2, 2016"
@@
 ---
+^1^Joint Global Change Research Institute, DOE Pacific Northwest National Laboratory, College Park, MD USA; 
+bondlamberty@pnnl.gov
+
+^2^Biological Sciences Division, Pacific Northwest National Laboratory, Richland, WA USA
 

### 2016-02-09 manuscript_openexperiment/Open_experiment.Rmd
Text changes and additions for first draft
Also added first Word draft to repo, and tweaked figure 1.

--- 2016-02-08 manuscript_openexperiment/Open_experiment.Rmd
@@
   science'
-author: "Ben Bond-Lamberty, Peyton Smith, and Vanessa Bailey"
+author: "Ben Bond-Lamberty^1^, Peyton Smith^2^, Vanessa Bailey^2^"
 date: "February 2, 2016"
-output: html_document
+output: word_document
 ---
+^1^Joint Global Change Research Institute, DOE Pacific Northwest National Laboratory, College Park, MD USA; 
+bondlamberty@pnnl.gov
+
+^2^Biological Sciences Division, Pacific Northwest National Laboratory, Richland, WA USA
 
@@
 A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, and enable powerful meta- and cross-disciplinary analyses.
-We describe a recent "open experiment", in which we documented every aspect of a soil incubation experiment online, making publicly available raw data, scripts, diagnostics, final analyses, and manuscripts in real time.
-Our use of tools such as version control, issue tracking, and open-source statistical software ensured data integrity and transparency.
-There are many avenues to improve scientific reproducibility and data availability, of which is this only one example; we acknowledge that this is not an approach suited for every experiment or situation.
-Nonetheless, we encourage the communities in our respective fields to consider its advantages, and lead rather than follow with respect to scientific reproducibility, transparency, and data availability.
+Here we describe a recent "open experiment", in which we documented every aspect of a soil incubation experiment online.
+This involved making raw data, scripts, diagnostics, final analyses, and manuscripts publicly available in real time.
+We found that using tools such as version control, issue tracking, and open-source statistical software improved data integrity, accelerated our team's communications, and ensured transparency.
+There are many avenues to improve scientific reproducibility and data availability, of which is this only one example, and it is not an approach suited for every experiment or situation.
+Nonetheless, we encourage the communities in our respective fields to consider its advantages, and to lead rather than follow with respect to scientific reproducibility, transparency, and data availability.
 
@@
 
-A wide variety of forces are pushing scientists to increase data access, both pre- and post-publication, and towards "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.
+Science is becomingly increasingly collaborative and data-intensive {Adams, 2012 #4348}.
+At the same time, a wide variety of factors are encouraging, at times pushing, scientists to increase data access and use "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.
 
 Journals are adopting increasingly stringent data access and deposition policies, e.g.
-Scientific Data (http://www.nature.com/sdata/data-policies), PLOS ONE (http://journals.plos.org/plosone/s/data-availability), and Science (http://www.sciencemag.org/authors/science-editorial-policies#data-deposition).
+Scientific Data[^1], PLOS ONE[^2], and Science[^3].
 These policies generally share common assumptions and goals: maximizing access to scientific data; encouraging deposition into structured repositories as opposed to journal supplementary information; and specifying that it is not acceptable for authors to be solely responsible for ensuring data access.
+Funders are moving in this direction as well, with organizations such as the U.S.
+National Science Foundation[^4] and the U.K.
+Wellcome Trust[^5], along with many others, requiring explicit data management plans, unfettered reasonable access to primary data, and use of established repositories.
 
-Funders are moving in this direction as well, with organizations such as the U.S.
-National Science Foundation (http://www.nsf.gov/bfa/dias/policy/dmp.jsp) and the U.K.
-Wellcome Trust (http://www.wellcome.ac.uk/About-us/Policy/Spotlight-issues/Data-sharing/Data-management-and-sharing/index.htm) requiring explicit data management plans, unfettered reasonable access to primary data, and use of established repositories.
+[^1]: http://www.nature.com/sdata/data-policies
+[^2]: http://journals.plos.org/plosone/s/data-availability
+[^3]: http://www.sciencemag.org/authors/science-editorial-policies#data-deposition
+[^4]: http://www.nsf.gov/bfa/dias/policy/dmp.jsp
+[^5]: http://www.wellcome.ac.uk/About-us/Policy/Spotlight-issues/Data-sharing/Data-management-and-sharing/index.htm
 
-Finally, growing numbers of scientists are pushing for open science and data on moral/political grounds, as well as purely scientific ones, in particular ensuring reproducibility {Stodden, 2011 #4340}.
+Finally, growing numbers of scientists are pushing for open science and data on moral/political grounds, as well as purely scientific ones, in particular ensuring reproducibility {Stodden, 2011 #4340} and enabling larger synthetic activities.
+Such analyses {Wolkovich, 2012 #3303} are made possible by the assembly of large, internally consistent data sets.
+Examples in ecology, soil science, and biogeosciences include BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, and SRDB {Bond-Lamberty, 2010 #2320}.
+
 We briefly discuss what we see as the primary arguments for data sharing and openness, before detailing a recent "open experiment" example.
@@
 {, 2016 #4334} found that none of 441 randomly-chosen studies provided raw data, and only one provided full protocols.
+The issue of reproducibility in ecology (and its many related fields) has been raised and debated for years {Cassey, 2006 #4349} but ecological, soil science, and global change journals differ widely in their data deposition requirements.
 
@@
 Modeling groups have worked to improve reproducibility and archiving praactices {Thornton, 2005 #2310}.
-For scientists without programming skills, scripted systems for data analysis!
-E.g.
-R {R Development Core Team, 2016 #4314}, Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record information about the data processing, analytical process and decisions, and statistical analysis.
+Scientists in all fields are increasingly building and using software in their work, however, often without strong training in this area {Wilson, 2014 #4337}.
+In addition to open-source data analysis languages such as R {R Development Core Team, 2016 #4314}, scientific workflow systems such as Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record information about the data processing, analytical process and decisions, and statistical analysis.
+Providing open code does not magically produce bug-free code, or mistake free analyses {Easterbrook, 2014 #4137}.
+But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task (often, forever).
 
@@
 Similarly, Reichman et al.
-{, 2011 #2862} estimated that less than 1% of ecological data collected is made available after publication, and noted, as an example, that most current and historical data relevant to the recent *Deepwater Horizon* oil spill are already inaccessible or lost {Reichman, 2011 #2862}.
+{, 2011 #2862} estimated that less than 1% of ecological data collected is made available after publication, and noted, as an example, that much current and historical data relevant to the recent *Deepwater Horizon* oil spill are already inaccessible or lost {Reichman, 2011 #2862}.
 In global change ecology, Wolkovich et al.
-{, 2012 #3303} reported that only ~10% of raw data sets were eventually shared in their attempt to conduct a high-profile meta-analysis {Wolkovich, 2012 #3331}.
+{, 2012 #3303} reported that only ~10% of raw data sets were eventually shared in their attempt to conduct a meta-analysis {Wolkovich, 2012 #3331}.
 
@@
 
-VLB: Supplementary Information sections have become repositories of data and of non-core analyses. How does our open experiment replace (improve upon?) SI? Thoughts?
+TODO - VLB: Supplementary Information sections have become repositories of data and of non-core analyses.
+How does our open experiment replace (improve upon?) SI?
+Thoughts?
+- journals not ready, and don't want {Cassey, 2006 #4349}.
 
-**Data and code quality**
+**Institutional and social trust** (TODO)
 
-Providing open code does not magically produce bug-free code, or mistake free analyses {Easterbrook, 2014 #4137}.
-But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task (often, forever).
+Finally, there are longer-term issues of trust to consider with respect to reproducibility and openness.
 
-**Enabling synthetic and global-scale analyses**
+Governmental efforts to have the results of federally funded scientific research made available to the public, industry, and across the scientific community {Holdren, 2013 #4344}.
 
-With greater data availability comes the potential to assemble large, internally consistent data sets--examples in biogeosciences and ecology include BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, and SRDB {Bond-Lamberty, 2010 #2320}.
-In turn, these data enable powerful meta-analyses and syntheses {Wolkovich, 2012 #3303}.
+* Trust in correctness of science: Potential to increase public trust in research results (e.g. climate change)
+* Trust in utility of science: Why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results?
 
-**Institutional and social trust**
-
-Why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results?
-
-Potential to increase public trust in research results.
-
-Governmental efforts {Holdren, 2013 #4344}.
+TODO: mention that of course authors have some rights.
+But what is being denied, increasingly, is that "the intellectual property rights of publishing ecologists" {Cassey, 2006 #4349} take precedence over all other factors, at least as far as publicly-funded research is concerned.
 
@@
 
-We used GitHub, a web-based Git repository hosting service, to...
-  "Git"" is a version control system...
-  A "repository"" is ...
-  A "web-based hosting service" is...
+We used GitHub, a web-based Git repository hosting service, to store our data, scripts, and manuscripts.
+"Git" is a popular, free, and open source version control software: it tracks changes (when, what, by whom) made in a "repository", a collection of folders and files.
+Many scientific and other users of Git use the "GitHub" web service, as it offers excellent performance and a wide variety useful additional functinality, particularly for teams or collaborative projects.
 
-The design of our repository is shown in **Figure 1**. Explain...
+The design of our repository is shown in **Figure 1**, and the repository itself can be found at https://github.com/bpbond/cpcrw_incubation.
+It consists of a series of scripts that comprise a *pipeline*, start with raw data and ending with final analyses, figures, and manuscripts.
 
-This system had a number of desirable capabilities and characteristics:
+This system had a number of characteristics:
 
 * The entire data processing and analytical system is online and documented.
+It is written in R {R Development Core Team, 2016 #4314}, an open-source and widely used language and environment for statistical computing and graphics.
 * The version control system let us make incremental changes, work out problems, look at histories (i.e., who made what change when).
-A issue tracker let us discuss problems, reference changes in the repository, create to-do lists, and assign responsibilities.
+* A issue tracker let us discuss problems, reference changes in the repository, create to-do lists, and assign responsibilities.
+* Information webpage on the experiment and broader project.
+* Manuscripts as well!
+(Including this one.) Results are linked directly into the manuscripts, so that if a result changes, it is updated automatically in the text.
+Tools such as R Markdown[^7] have made this process much easier to build and maintain.
+
+[^7]: http://rmarkdown.rstudio.com
+
+Advantages:
+
 * The public nature of the repository encouraged us to flesh out documentation, use clean and clear coding, and think about the longer-term ramifications of decisions.
-* Real-time diagnostics... (**Figure 2**)... Investing the effort to set up a software pipeline BEFOREHAND meant that we could reliably and easily identify and diagnose problems.
-* Everything necessary to reproduce study results.
-* Manuscripts as well!
-In fact, the "source code" for this manuscript, as well as its history, can be viewed at https://github.com/bpbond/cpcrw_incubation/blob/master/manuscript_methods/Open_experiment.Rmd.
-* Information webpage on the experiment and broader project.
+* Real-time diagnostics (**Figure 2**) let our team, program managers, and interested other parties see at a glance the progress of the experiment.
+* The issue tracker helped our team--which was not physically located all in one place--to communicate, discuss, and track "issues" (i.e., problems or questions that arose).
+* Investing the effort to set up a software pipeline before, not after, the experiment was performed meant that we could reliably and easily identify and diagnose problems.
 
@@
 
-Transparency is no magic panacea ensuring reproducibility {Easterbrook, 2014 #4137}, and it raises genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
-Scientists also frequently raise a variety of other concerns about data sharing, for example being "scooped" not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
+Transparency is no magic panacea ensuring reproducibility {Easterbrook, 2014 #4137}, and may raise genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
+Scientists also frequently cite other concerns about data sharing, for example being "scooped" not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
 We cannot easily dismiss of these these: for example, does our community adequately credit researchers who contribute to global databases that subsequently produce high-impact papers?
 Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers efforts are adequately valued and cited {Kueffer, 2011 #4335}.
-We also recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}.
+We also recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}--that is, scientists do not receive 'proper' credit for datasets, relative to publications, even when the former may be much more broadly valueable.
 
-Nor is the model we present here--this public repository that was updated throughout the experiment and publication process--perfect.
+The model we present here--a public repository updated throughout the experiment, analysis, and publication process--is far from perfect.
 First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ: for example, there was no unit testing to ensure module correctness, nor an automated build tool automating the workflow (i.e., the individual scripts all have to be run manually).
 Second, it may not not be appropriate for applicable to all study types, even just in our professional fields of soil, ecosystem, and global change science.
-(And it certainly is not appropriate for sensitive human-subject data.) Third, Git and GitHub are designed for working with code, not data, and this causes some ...
+(And it certainly is not appropriate for sensitive human-subject data.) Third, Git and GitHub are designed for working with code, not data, although this situation continues to improve on a technical level[^6].
 
-Nonetheless... conclude.
+[^6]: https://github.com/blog/1885-better-word-highlighting-in-diffs
+
+Nonetheless (TODO: it's a lot better than nothing, and offers instrument-to-final product reproducibility).
+
+It doesn't have to be adopted wholesale: for example, the use of an issue tracker alone might be valuable for some groups and improve their science (though not reproducibility).
 
 This is example, we hope, of the impact of individual scientists' decisions and practices {Wolkovich, 2012 #3303}.
+
+We encourage the communities in our respective fields to consider its advantages, and to lead rather than follow with respect to scientific reproducibility, transparency, and data availability.
 

### 2016-02-09 R/4-fluxes.R
Gravimetric WC update
Data corrections due to sorting and data entry mistakes. See issue #38.

@@ -27,14 +27,33 @@ summarydata <- left_join(summarydata, drymassdata, by = "Core")
 # -----------------------------------------------------------------------------
 # Create fluxdata and compute water content
 
+printlog("Computing water contents...")
 summarydata %>%
   filter(Treatment != "Ambient") %>%
   mutate(WC_gravimetric = 
            (Mass_g - CoreSleeveMass_g - NonsoilMassLarge_g - SoilDryMass_g) / 
-           (SoilDryMass_g)) %>%
-  select(Core, SoilDryMass_g, SoilVolume_cm3, WC_gravimetric,
+           (SoilDryMass_g),
+         WC_volumetric = (Mass_g - CoreSleeveMass_g - NonsoilMassLarge_g - SoilDryMass_g) / 
+           (SoilVolume_cm3)) %>%
+  select(Core, SoilDryMass_g, SoilVolume_cm3, WC_gravimetric, WC_volumetric,
          Treatment, Temperature, CO2_ppm_s, CH4_ppb_s, inctime_days) ->
   fluxdata
+
+print(summary(fluxdata$WC_gravimetric))
+print(summary(fluxdata$WC_volumetric))
+
+p <- ggplot(fluxdata, aes(inctime_days, WC_gravimetric, color=paste(Treatment, Temperature), group=Core))
+p <- p + geom_line() + facet_wrap(~Core) + scale_color_discrete("")
+print(p)
+save_plot("WC_gravimetric")
+
+printlog("WC_gravimetric summary by core:")
+fluxdata %>% 
+  group_by(Core) %>% 
+  summarise_each(funs(min, mean, max), WC_gravimetric) %>% 
+  arrange(mean) %>%
+  as.data.frame %>%
+  print
 
 # -----------------------------------------------------------------------------
 # Flux computation

### 2016-02-09 R/5-analyze.R

@@ -142,6 +142,11 @@ fluxdata %>%
 print(WC_effect)
 save_data(WC_effect)
 
+# A quick visualization
+p <- qplot(WC_gravimetric, flux_µmol_g_s, data=fluxdata) 
+p <- p + facet_grid(Gas~Temperature, scales="free") + geom_smooth(method='lm')
+save_plot(p, pname = "WC_gravimetric_effect")
+
 # -----------------------------------------------------------------------------
 # Per-core models - look at R2, etc. variability
 

### 2016-02-10 manuscript/cpcrw_manuscript.Rmd
Gravimetric WC update
Data corrections due to sorting and data entry mistakes. See issue #38.
Added Tukey HSD test on cumulative fluxes

--- 2016-02-09 manuscript/cpcrw_manuscript.Rmd
@@
 ```{r, echo=FALSE, results='hide', message=FALSE, fig.keep='none'}
+pclean <- function(x) {  # pretty formatting of p-values
+  ifelse(x < 0.001, "< 0.001", paste("=", formatC(x, digits = 3, format = 'f'))) %>%
+    paste("P", .)
+}
+
 library(dplyr, warn.conflicts = FALSE)
@@
 Sub-samples were collected and composited throughout each soil core for dissolved organic carbon analysis (110 ± 24 g dry mass equivalent) and dry-mass calculations (~28 g each).
-The remaining core material was air-dried and separated into particles (>2 mm diameter) and soil (<2 mm) using a U.S.
+The remaining core material was air-dried and separated into particles (>2 mm diameter) and soil (≤2 mm) using a U.S.
 Standard Test Sieve No.
@@
 
-We measured the rise in GHG concentrations from its minimum value (in the first 10 seconds of measurement after the analyzer was sealed to the soil core) to its maximum (in the first 45 seconds), considering this as the flux rate in ppm s^-1^ (CO2) or ppb s^-1^ (CH4).
+We measured the rise in GHG concentrations from its minimum value (in the first 10 seconds of measurement after the analyzer was sealed to the soil core) to its maximum (in the first 45 seconds), considering this as the flux rate in ppm s^-1^ (CO~2~) or ppb s^-1^ (CH~4~).
 This value was calculated for each sample, i.e.
@@
 
+The effect of temperature and treatment (drought, controlled drought, or field moisture conditions) on cumulative gas fluxes was evaluted with a post-hoc Tukey Honest Significant Differences test.
+
 All data analysis and statistics were performed using R version 3.2.3 {R Development Core Team, 2016 #4314}.
@@
 * Cumulative fluxes (Figure 4)
+
+Cumulative CO~2~ evolved was highly affected by temperature (`r pclean(co2_hsd$Temperature["p adj"])`), and field moisture cores emitted significantly more CO~2~ than the other two treatments at both temperatures (P < 0.001 for both).
+There was no difference between the 20 °C drought and controlled drought treatments (`r pclean(co2_hsd$Treatment["Drought-Controlled drought", "p adj"])`).
+Neither temperature (`r pclean(ch4_hsd$Temperature[,"p adj"])`) nor treatment (mean `r pclean(mean(ch4_hsd$Treatment[,"p adj"]))`) was a significant factor in predicting cumulative CH~4~ fluxes.
+
 * CO2:CH4 ratios

### 2016-02-10 R/5-analyze.R

@@ -6,6 +6,7 @@ source("R/0-functions.R")
 SCRIPTNAME  	<- "5-analyze.R"
 FLUXDATA   <- file.path(outputdir(scriptfolder = FALSE), "fluxdata.csv")
 FLUXDATA_CUMULATIVE <- file.path(outputdir(scriptfolder = FALSE), "fluxdata_cumulative.csv")
+FD_CUMULATIVE_CORE <- file.path(outputdir(scriptfolder = FALSE), "fd_cumulative_core.csv")
 
 library(broom)  # 0.4.0
 library(reshape2) # 1.4.1
@@ -72,8 +73,29 @@ fluxdata_figsBC %>%
   figureC
 
 # -----------------------------------------------------------------------------
-# Cumulative flux figure
+# Cumulative flux figure and Tukey HSD tests
 
+printlog("Running Tukey HSD tests on cumulative emissions...")
+fd_cumulative_core <- read_csv(FD_CUMULATIVE_CORE) %>%
+  mutate(Temperature = as.factor(Temperature))
+co2_aov <- aov(cum_flux_mgC ~ Treatment + Temperature, 
+               data = subset(fd_cumulative_core, Gas == "CO2"))
+co2_hsd <- TukeyHSD(co2_aov)
+print(co2_hsd)
+ch4_aov <- aov(cum_flux_mgC ~ Treatment + Temperature, 
+               data = subset(fd_cumulative_core, Gas == "CH4"))
+ch4_hsd <- TukeyHSD(ch4_aov)
+print(ch4_hsd)
+
+# Convert the significance data to data frame, for easier (clearer)
+# access later by the manuscript code
+co2_hsd$Temperature <- as.data.frame(co2_hsd$Temperature)
+co2_hsd$Treatment <- as.data.frame(co2_hsd$Treatment)
+ch4_hsd$Temperature <- as.data.frame(ch4_hsd$Temperature)
+ch4_hsd$Treatment <- as.data.frame(ch4_hsd$Treatment)
+
+
+printlog("Making plot...")
 fluxdata_cumulative <- read_csv(FLUXDATA_CUMULATIVE)
 fluxdata_cumulative$Treatment <- factor(fluxdata_cumulative$Treatment, 
                                         levels = c("Field moisture", "Controlled drought", "Drought"))
@@ -87,6 +109,7 @@ figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Tre
                 position = position_dodge(0.9), width = 0.4) +  
[…]

### 2016-02-10 manuscript/cpcrw_manuscript.Rmd
Added CO2:CH4 ratio computation and ms text

@@
 * CO2:CH4 ratios
+
+```{r, echo=FALSE}
+minmag <- min(gas_ratio$CO2CH4_ratio) %>% log10 %>% floor %>%
+  c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine")[.]
+minratio <- min(gas_ratio$CO2CH4_ratio / 1e6)  %>% formatC(digits = 1, format = 'f')
+mintrt <- gas_ratio$Treatment[which.min(gas_ratio$CO2CH4_ratio)]
+mintemp <- gas_ratio$Temperature[which.min(gas_ratio$CO2CH4_ratio)]
+maxratio <- max(gas_ratio$CO2CH4_ratio / 1e6) %>% formatC(digits = 1, format = 'f')
+maxtrt <- gas_ratio$Treatment[which.max(gas_ratio$CO2CH4_ratio)]
+maxtemp <- gas_ratio$Temperature[which.max(gas_ratio$CO2CH4_ratio)]
+```
+
+The production of C from CO~2~ was over `r minmag` order of magnitudes higher than that from CH~4~, with CO~2~:CH~4~ ratios ranging from `r minratio` million in the `r mintemp` °C `r maxtrt` treatment, to `r maxratio` million in the `r maxtemp` °C `r maxtrt` treatment.
+
 * Statistical tests of WC and temperature on cumulative fluxes

### 2016-02-10 R/5-analyze.R

@@ -56,7 +56,7 @@ fluxdata %>%
 
 figsBC <- function(fd) {
   panellabels <- 
-  ggplot(fd, aes(incday, flux)) + 
+    ggplot(fd, aes(incday, flux)) + 
     geom_point() + 
     facet_grid(Temperature ~ Treatment) + 
     geom_errorbar(aes(ymin = flux - flux_sd, ymax = flux + flux_sd)) +
@@ -110,6 +110,17 @@ figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Tre
   facet_grid(Gas ~ ., scales = "free") +
   ylab(paste("Cumulative C (mg) over", floor(max(fluxdata$inctime_days)), "days"))
 
+
+# -----------------------------------------------------------------------------
+# CO2:CH4 emissions ratio
+
+printlog("Computing emissions ratios...")
+fluxdata_cumulative %>% 
+  group_by(Treatment, Temperature) %>% 
+  summarise(CO2CH4_ratio = max(cum_flux_mgC) / min(cum_flux_mgC)) %>%
+  na.omit %>%
+  print ->
+  gas_ratio
 
 # -----------------------------------------------------------------------------
 # Examine data distribution

### 2016-02-10 R/0-functions.R
Defining away key analytical constants

@@ -17,6 +17,12 @@ library(stringr)       # 1.0.0
 library(luzlogr)       # 0.1.2
 
 # -----------------------------------------------------------------------------
+# Parameters for key analytical choices
+# Defined here so they can be easily used in code AND manuscript
+CO2_EXCLUDE_DEVS <- 3.0
+CH4_EXCLUDE_DEVS <- 5.0
+
+# -----------------------------------------------------------------------------
 # Print dimensions of data frame
 print_dims <- function(d, dname=deparse(substitute(d))) {
   stopifnot(is.data.frame(d))

### 2016-02-10 R/4-fluxes.R

@@ -112,9 +112,9 @@ printlog("Identifying and plotting outliers...")
 fluxdata %>%
   mutate(incday = floor(inctime_days)) %>%
   group_by(Treatment, Temperature, incday) %>% 
-  mutate(CO2_outlier = is_outlier(CO2_flux_µmol_g_s, devs = 3.0), 
+  mutate(CO2_outlier = is_outlier(CO2_flux_µmol_g_s, devs = CO2_EXCLUDE_DEVS), 
          # CH4 is so variable we use a higher exclusion cutoff
-         CH4_outlier = is_outlier(CH4_flux_µmol_g_s, devs = 5.0)) %>%
+         CH4_outlier = is_outlier(CH4_flux_µmol_g_s, devs = CH4_EXCLUDE_DEVS)) %>%
   select(-incday) ->
   fluxdata
 

### 2016-02-10 manuscript/cpcrw_manuscript.Rmd
Misc changes
Changed WC presentation; named code chunks; cleaned up exclude/outlier
computation and text.

--- 2016-02-10 manuscript/cpcrw_manuscript.Rmd
@@
 
-```{r, echo=FALSE, results='hide', message=FALSE, fig.keep='none'}
+```{r setup, echo=FALSE, fig.keep='none', message=FALSE, results='hide'}
+library(dplyr, warn.conflicts = FALSE)
+
 pclean <- function(x) {  # pretty formatting of p-values
@@
 
-library(dplyr, warn.conflicts = FALSE)
+# Run the analysis script, which will leave a lot of figures, tables, and data in memory for us
 setwd("../")
@@
 These groups included two incubation temperatures (4 and 20 °C), each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
-The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 °C "drought" group.
+The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status would close matched those of the 4 °C "drought" cores, which we anticipated would dry more slowly than their 20 °C counterparts.
 The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density.
@@
 
-```{r, echo=FALSE}
+```{r compute_excludes, echo=FALSE}
+# Compute outlier and exclusion numbers
 # TODO: this isn't good--fragile if files move or change names
 # Want everything in-memory (or something)
-# Compute outlier and exclusion numbers
+rn <- readLines("../outputs/3-remove/removed_data.csv") %>% length
 sdn <- readLines("../outputs/summarydata.csv") %>% length
-sdcn <- readLines("../outputs/summarydata_clean.csv") %>% length
-en <- readLines("../outputs/3-remove/excluded_data.csv") %>% length
-rn <- readLines("../outputs/3-remove/removed_data.csv") %>% length
+fluxn <- nrow(fluxdata)
+en <- sum(fluxdata$outlier)
 ```
 
-Anomalous data were excluded based on their being more than 3.0 (for CO~2~) or 5.0 (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within each day, treatment, and temperature.
-We excluded `r en` of `r sdn` (`r round(en/sdn*100)`%) of the summarized samples for this reason.
-Other data (`r rn` of `r sdn`, `r round(rn/sdn*100)`%) were removed because of known instrument problems, e.g.
+Anomalous data were excluded based on their being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within each day, treatment, and temperature.
+We excluded `r en` of `r fluxn` (`r round(en/fluxn*100)`%) of the gas flux measurements for this reason.
+Other data (`r round(rn/sdn*100)`%) were removed because of known instrument problems, e.g.
 the analyzer was left running after leaving a chamber.
-The final number of valid flux samples from the 100-day incubation was `r sdcn`.
+The final number of valid flux samples from the 100-day incubation was `r sum(!fluxdata$outlier)`.
 
@@
 * Gas fluxes over incubation (Figures 2 and 3)
+
 * Statistical tests of WC, T, C, N on individual core fluxes (Table 2)
@@
 
-Cumulative CO~2~ evolved was highly affected by temperature (`r pclean(co2_hsd$Temperature["p adj"])`), and field moisture cores emitted significantly more CO~2~ than the other two treatments at both temperatures (P < 0.001 for both).
-There was no difference between the 20 °C drought and controlled drought treatments (`r pclean(co2_hsd$Treatment["Drought-Controlled drought", "p adj"])`).
+Cumulative CO~2~ evolved was highly affected by temperature (`r pclean(co2_hsd$Temperature["p adj"])`), and "field moisture" cores emitted significantly more CO~2~ than the other two treatments at both temperatures (P < 0.001 for both).
+There was no difference between the 20 °C "drought" and "controlled drought" treatments (`r pclean(co2_hsd$Treatment["Drought-Controlled drought", "p adj"])`).
 Neither temperature (`r pclean(ch4_hsd$Temperature[,"p adj"])`) nor treatment (mean `r pclean(mean(ch4_hsd$Treatment[,"p adj"]))`) was a significant factor in predicting cumulative CH~4~ fluxes.
@@
 * CO2:CH4 ratios
-* Statistical tests of WC and temperature on cumulative fluxes
+
+```{r compute_ratios, echo=FALSE}
+# Ratio data is stored in the `gas_ratio` d.f. computed by analysis script
+minmag <- min(gas_ratio$CO2CH4_ratio) %>% log10 %>% floor %>%
+  c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine")[.]
+minratio <- min(gas_ratio$CO2CH4_ratio / 1e6)  %>% formatC(digits = 1, format = 'f')
+mintrt <- gas_ratio$Treatment[which.min(gas_ratio$CO2CH4_ratio)]
+mintemp <- gas_ratio$Temperature[which.min(gas_ratio$CO2CH4_ratio)]
+maxratio <- max(gas_ratio$CO2CH4_ratio / 1e6) %>% formatC(digits = 1, format = 'f')
+maxtrt <- gas_ratio$Treatment[which.max(gas_ratio$CO2CH4_ratio)]
+maxtemp <- gas_ratio$Temperature[which.max(gas_ratio$CO2CH4_ratio)]
+```
+
+The production of C from CO~2~ was over `r minmag` order of magnitudes higher than that from CH~4~, with CO~2~:CH~4~ C ratios ranging from `r minratio` million in the `r mintemp` °C "`r maxtrt`" treatment, to `r maxratio` million in the `r maxtemp` °C "`r maxtrt`" treatment.
 
@@
 
-```{r kable, echo=FALSE}
+```{r table1, echo=FALSE}
 # Print out the `WC_effect` data frame in a nice-looking table
@@
 
-**Figure 1.** Core water content across the course of the experiment.
+**Figure 1.** Core water content across the course of the incubation experiment by temperature (left panel 4 °C, right panel 20 °C) and treatment.
 
-```{r, echo=FALSE}
-print(figureA)
+```{r figureA, echo=FALSE}
+figureA
 ```
@@
 
-```{r, echo=FALSE}
-print(figureB)
+```{r figureB, echo=FALSE}
+figureB
 ```
@@
 
-```{r, echo=FALSE}
-print(figureC)
+```{r figureC, echo=FALSE}
+figureC
 ```
@@
 
-```{r, echo=FALSE, warning=FALSE}
-print(figureD)   # TODO: wtf is going on here?
+```{r figureD, echo=FALSE, warning=FALSE}
+figureD
 ```

### 2016-02-10 R/5-analyze.R

@@ -39,10 +39,10 @@ save_data(fluxdata, fname = "fluxdata_long")
 # -----------------------------------------------------------------------------
 # Water content over time figure
 
-figureA <- ggplot(fluxdata_orig, aes(inctime_days, WC_gravimetric * 100, color=Treatment, group=Core)) 
+figureA <- ggplot(fluxdata_orig, aes(inctime_days, WC_gravimetric, color=Treatment, group=Core)) 
 figureA <- figureA + geom_point() + geom_line()
 figureA <- figureA + facet_grid(~Temperature) 
-figureA <- figureA + xlab("Incubation day") + ylab("Gravimetric water content (%)")
+figureA <- figureA + xlab("Incubation day") + ylab("Gravimetric water content (fraction dry mass)")
 
 # -----------------------------------------------------------------------------
 # Fluxes over time figure

### 2016-02-11 R/0-functions.R
Centralized file definitions and shifted entirely to `readr`

@@ -3,11 +3,8 @@
 # logged csv[.gz|zip] read/write; and a few other handy things.
 # Ben Bond-Lamberty March 2015
 
-INPUT_DIR     <- "data/"
-OUTPUT_DIR		<- "outputs/"
-#CHECKPOINTDATE	<- "2015-03-05" # comment out to not use checkpoint
-SEPARATOR		<- "-------------------"
-
+# -----------------------------------------------------------------------------
+# Key packages used across scripts
 library(ggplot2)       # 2.0.0
 theme_set(theme_bw())
 library(dplyr)         # 0.4.3
@@ -19,8 +16,23 @@ library(luzlogr)       # 0.1.2
 # -----------------------------------------------------------------------------
 # Parameters for key analytical choices
 # Defined here so they can be easily used in code AND manuscript
-CO2_EXCLUDE_DEVS <- 3.0
-CH4_EXCLUDE_DEVS <- 5.0
+CO2_EXCLUDE_DEVS <- 3.0    # CO2 outlier boundary, in mean absolute deviations
+CH4_EXCLUDE_DEVS <- 5.0    # CH4 outlier boundary, in mean absolute deviations
+
+# -----------------------------------------------------------------------------
+# Key files and directories shared between scripts
+OUTPUT_DIR		         <- "outputs/"
+DIAGNOSTICS_DIR        <- "qc_plots/"
+RAWDATA_FILE           <- file.path(OUTPUT_DIR, "rawdata.csv.gz")
+RAWDATA_SAMPLES_FILE   <- file.path(OUTPUT_DIR, "rawdata_samples.csv.gz")
+SUMMARYDATA_FILE       <- file.path(OUTPUT_DIR, "summarydata.csv")
+SUMMARYDATA_CLEAN_FILE <- file.path(OUTPUT_DIR, "summarydata_clean.csv")
+FLUXDATA_FILE          <- file.path(OUTPUT_DIR, "fluxdata.csv")
+FLUXDATA_CUM_FILE      <- file.path(OUTPUT_DIR, "fluxdata_cum.csv")
+FLUXDATA_CUM_CORE_FILE <- file.path(OUTPUT_DIR, "fluxdata_cum_core.csv")
+
+SEPARATOR		  <- "-------------------"
+
 
 # -----------------------------------------------------------------------------
 # Print dimensions of data frame
@@ -33,7 +45,7 @@ print_dims <- function(d, dname=deparse(substitute(d))) {
 # Return output directory (perhaps inside a script-specific folder)
[…]

### 2016-02-11 R/1-data.R

@@ -54,16 +54,13 @@ process_directory <- function(input_path, tempfile) {
   }
 }
 
-
 # ==============================================================================
 # Main 
 
 openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE)
 
 printlog("Welcome to", SCRIPTNAME)
-
 printlog("Data directory is", DATA_DIR)
-
 tf <- tempfile()
 printlog("Tempfile is", tf)
 
@@ -76,7 +73,7 @@ print_dims(rawdata)
 print(summary(rawdata))
 
 printlog("Writing output file...")
-save_data(rawdata, scriptfolder = FALSE, gzip = TRUE)
+save_data(rawdata, fn = RAWDATA_FILE, scriptfolder = FALSE)
 
 printlog("All done with", SCRIPTNAME)
 closelog()

### 2016-02-11 R/2-summarize.R

@@ -12,7 +12,6 @@ source("R/0-functions.R")
 SCRIPTNAME  	<- "2-summarize.R"
 PROBLEM       <- FALSE
 
-RAWDATA      <- file.path(OUTPUT_DIR, "rawdata.csv.gz")  # output from script 1
 VALVEMAP     <- "data/valvemap.csv"
 TREATMENTS   <- "data/treatments.csv"
 
@@ -48,7 +47,8 @@ openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # o
 printlog("Welcome to", SCRIPTNAME)
 
 printlog("Reading in raw data...")
-rawdata <- readr::read_csv(RAWDATA, col_types = "ccddddiiiddddddddc") %>%
+rawdata <- read_csv(RAWDATA_FILE, col_types = "ccddddiiiddddddddc") %>%
+#readr::read_csv(file.path(INPUT_DIR, RAWDATA_FILE), col_types = "ccddddiiiddddddddc") %>%
   # immediate discard columns we don't need
   select(DATE, TIME, MPVPosition, CH4_dry, CO2_dry, h2o_reported)
 print_dims(rawdata)
@@ -86,7 +86,7 @@ rawdata_samples <- rawdata %>%
 
 # Load and QC the valvemap data
 printlog("Loading valve map data...")
-valvemap <- read_csv(VALVEMAP, skip = 1)
+valvemap <- read_csv(VALVEMAP, skip = 1, col_types = "ccnncdc")
 printlog( "Converting date/time info to POSIXct..." )
 valvemap$StartDateTime <- mdy_hm(paste(valvemap$Date, valvemap$Time_set_start))
 valvemap <- arrange(valvemap, StartDateTime)
@@ -180,8 +180,8 @@ save_data(MPVPosition_checkdata)
 # -----------------------------------------------------------------------------
 # Done! 
 
-save_data(summarydata, scriptfolder = FALSE)
-save_data(rawdata_samples, scriptfolder = FALSE, gzip = TRUE)
+save_data(summarydata, fn = SUMMARYDATA_FILE, scriptfolder = FALSE)
+save_data(rawdata_samples, fn = RAWDATA_SAMPLES_FILE, scriptfolder = FALSE)
 
 printlog("All done with", SCRIPTNAME)
 closelog()

### 2016-02-11 R/3-remove.R

@@ -11,7 +11,6 @@ PROBLEM       <- FALSE
 CATEGORY_ERROR <- "Error"
 CATEGORY_OUTLIER <- "Outlier"
 
-SUMMARYDATA  <- file.path(OUTPUT_DIR, "summarydata.csv")  # output from script 2
 REMOVALS     <- file.path("data/", "removals.csv")
 
 # ==============================================================================
@@ -22,7 +21,7 @@ openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # o
 printlog("Welcome to", SCRIPTNAME)
 
 printlog("Reading in summary data...")
-summarydata <- read_csv(SUMMARYDATA) %>%
+summarydata <- read_csv(SUMMARYDATA_FILE) %>%
   mutate(DATETIME = ymd_hms(DATETIME))
 print_dims(summarydata)
 
@@ -79,7 +78,7 @@ for(i in seq_len(nrow(removals))) {
 
 printlog(SEPARATOR)
 printlog("All done with removals and exclusions")
-save_data(summarydata_clean, scriptfolder=FALSE)
+save_data(summarydata_clean, fn = SUMMARYDATA_CLEAN_FILE, scriptfolder = FALSE)
 
 printlog("Removed data:", nrow(removed_data), "of", nrow(summarydata), "=",
          round(nrow(removed_data) / nrow(summarydata) * 100, 1), "%")

### 2016-02-11 R/3.3-qc.R

@@ -8,10 +8,6 @@ source("R/0-functions.R")
 SCRIPTNAME  	<- "3-qc.R"
 PROBLEM       <- FALSE
 
-SUMMARYDATA     <- file.path(OUTPUT_DIR, "summarydata_clean.csv")  # output from script 3
-RAWDATA_SAMPLES <- file.path(OUTPUT_DIR, "rawdata_samples.csv.gz")  # output from script 1
-
-
 # Main ===========================================================================
 
 openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # open log
@@ -19,7 +15,7 @@ openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # o
 printlog("Welcome to", SCRIPTNAME)
 
 printlog("Reading in summary data...")
-summarydata <- read_csv(SUMMARYDATA)
+summarydata <- read_csv(SUMMARYDATA_CLEAN_FILE)
 print_dims(summarydata)
 
 summarydata$DATETIME <- ymd_hms(summarydata$DATETIME)
@@ -42,7 +38,7 @@ print(summarydata %>% group_by(Core) %>% summarise(n()) %>% as.data.frame())
 
 summarydata %>%   # save a list of samples and corresponding cores
   select(DATETIME, MPVPosition, Core) %>% 
-  save_data(fname="samplelist.csv")
+  save_data(fn = "samplelist.csv")
 
 printlog("Summaries for max_CH4 and max_CO2:")
 summary(summarydata$max_CO2)
@@ -116,7 +112,7 @@ core_cv <- core_cv[1:25,]    # only do this many detail plots
 MAX_COMBINED_PLOTS <- 9
 
 printlog("Reading in raw data...")
-rawdata <- readr::read_csv(RAWDATA_SAMPLES, col_types = "ccddddcid")
+rawdata <- readr::read_csv(RAWDATA_SAMPLES_FILE, col_types = "ccddddcid")
 rawdata$samplenum <- as.factor(rawdata$samplenum)
 summarydata$samplenum <- as.factor(summarydata$samplenum)
 

### 2016-02-11 R/3.5-investigate.R

@@ -10,11 +10,7 @@ source("R/0-functions.R")
 SCRIPTNAME  	<- "3.5-investigate.R"
 PROBLEM       <- FALSE
 
-SUMMARYDATA  <- file.path(OUTPUT_DIR, "summarydata_clean.csv")  # output from script 3
-RAWDATA      <- file.path(OUTPUT_DIR, "rawdata_samples.csv.gz")  # output from script 1
-TREATMENT_SUMMARYDATA  <- file.path(OUTPUT_DIR, "treatment_summary.csv")  # output from script 3
-
-library(digest)
+library(digest)  # 0.6.9
 
 # -----------------------------------------------------------------------------
 # make diagnostic plots etc. for a particular gas
@@ -90,9 +86,9 @@ openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # o
 printlog("Welcome to", SCRIPTNAME)
 
 printlog("Reading in raw data...")
-rawdata <- readr::read_csv(RAWDATA)
+rawdata <- readr::read_csv(RAWDATA_FILE)
 printlog("Reading in summary data...")
-summarydata <- read_csv(SUMMARYDATA)
+summarydata <- read_csv(SUMMARYDATA_FILE)
 summarydata$DATETIME <- ymd_hms(summarydata$DATETIME)
 printlog("Reading in treatment summaries...")
 tsummary <- read_csv(TREATMENT_SUMMARYDATA)

### 2016-02-11 R/4-fluxes.R

@@ -6,7 +6,6 @@ source("R/0-functions.R")
 SCRIPTNAME  	<- "4-fluxes.R"
 PROBLEM       <- FALSE
 
-SUMMARYDATA   <- file.path(outputdir(scriptfolder = FALSE), "summarydata_clean.csv")
 DRYMASSDATA   <- "data/drymasses.csv"
 
 library(reshape2)      # 1.4.1
@@ -18,7 +17,7 @@ openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # o
 
 printlog("Welcome to", SCRIPTNAME)
 
-summarydata <- read_csv(SUMMARYDATA)
+summarydata <- read_csv(SUMMARYDATA_CLEAN_FILE)
 print_dims(summarydata)
 
 drymassdata <- read_csv(DRYMASSDATA, skip = 1)
@@ -132,7 +131,7 @@ p <- p + ggtitle("Dry mass-corrected flux rates")
 print(p)
 save_diagnostic(p, "CH4_outliers")
 
-save_data(fluxdata, scriptfolder = FALSE)
+save_data(fluxdata, fn = FLUXDATA_FILE, scriptfolder = FALSE)
 
 
 # -----------------------------------------------------------------------------
@@ -195,7 +194,7 @@ fluxdata %>%
   summarise(cum_flux_mgC = last(value)) -> 
   fd_cumulative_core
 
-save_data(fd_cumulative_core, scriptfolder = FALSE)
+save_data(fd_cumulative_core, fn = FLUXDATA_CUM_CORE_FILE, scriptfolder = FALSE)
 
 fd_cumulative_core %>%   # already grouped by Treatment, Temperature, Gas
   summarise(cum_flux_mgC_sd = sd(cum_flux_mgC),
@@ -229,7 +228,7 @@ p3 <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Treatmen
   ylab(paste("Cumulative C (mg) over", floor(max(fluxdata$inctime_days)), "days")) +
   ggtitle("Cumulative C by gas, treatment, temperature")
 save_diagnostic(p3, "cumulative_gas")
-save_data(fluxdata_cumulative, scriptfolder = FALSE)
+save_data(fluxdata_cumulative, fn = FLUXDATA_CUM_FILE, scriptfolder = FALSE)
 
 # -----------------------------------------------------------------------------
 # Done!

### 2016-02-11 R/5-analyze.R

@@ -1,12 +1,9 @@
-# Computes fluxes. In progress.
-# Ben Bond-Lamberty December 2015
+# Final analysis script: statistical tests, final plots, etc.
+# Ben Bond-Lamberty January 2016
 
 source("R/0-functions.R")
 
 SCRIPTNAME  	<- "5-analyze.R"
-FLUXDATA   <- file.path(outputdir(scriptfolder = FALSE), "fluxdata.csv")
-FLUXDATA_CUMULATIVE <- file.path(outputdir(scriptfolder = FALSE), "fluxdata_cumulative.csv")
-FD_CUMULATIVE_CORE <- file.path(outputdir(scriptfolder = FALSE), "fd_cumulative_core.csv")
 
 library(broom)  # 0.4.0
 library(reshape2) # 1.4.1
@@ -18,7 +15,7 @@ openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # o
 
 printlog("Welcome to", SCRIPTNAME)
 
-fluxdata_orig <- read_csv(FLUXDATA)
+fluxdata_orig <- read_csv(FLUXDATA_FILE)
 
 printlog("Transforming...")
 fluxdata_orig %>%
@@ -34,7 +31,7 @@ fluxdata_orig %>%
 
 fluxdata$Treatment <- factor(fluxdata$Treatment, levels = c("Field moisture", "Controlled drought", "Drought"))
 
-save_data(fluxdata, fname = "fluxdata_long")
+save_data(fluxdata, fn = "fluxdata_long.csv")
 
 # -----------------------------------------------------------------------------
 # Water content over time figure
@@ -76,7 +73,7 @@ fluxdata_figsBC %>%
 # Cumulative flux figure and Tukey HSD tests
 
 printlog("Running Tukey HSD tests on cumulative emissions...")
-fd_cumulative_core <- read_csv(FD_CUMULATIVE_CORE) %>%
+fd_cumulative_core <- read_csv(FLUXDATA_CUM_CORE_FILE) %>%
   mutate(Temperature = as.factor(Temperature))
 co2_aov <- aov(cum_flux_mgC ~ Treatment + Temperature, 
                data = subset(fd_cumulative_core, Gas == "CO2"))
@@ -96,7 +93,7 @@ ch4_hsd$Treatment <- as.data.frame(ch4_hsd$Treatment)
 
 
 printlog("Making plot...")
[…]

### 2016-02-11 manuscript/cpcrw_manuscript.Rmd
WC and BD calculations and results text

@@
 title: Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils
-author: "Ben Bond-Lamberty^1^, Peyton Smith^2^, Vanessa Bailey^2^"
+author: "Ben Bond-Lamberty^1^, A. Peyton Smith^2^, Vanessa Bailey^2^"
 output: html_document
@@
 }
+
+# 'Pretty n' function to round a numeric value and print that # of digits
+pn <- function(x, n) formatC(round(x, n), digits = n, format = 'f')
 
@@
 
-* Soil WC changes over incubation (Figure 1)
+* Core BD
+
+```{r coredata, echo=FALSE}
+bd <- with(coredata, SoilDryMass_g / CoreVolume_cm3)
+schist <- with(coredata, NonsoilMassLarge_g / (NonsoilMassLarge_g + SoilDryMass_g)) * 100
+sdm <- coredata$SoilDryMass_g
+```
+
+The `r nrow(coredata)` experimental cores had a bulk density of `r pn(mean(bd), 2)` ± `r pn(sd(bd), 2)` (mean ± sd) g cm^-3^.
+Large (>2 mm) particles, primarily schist, comprised `r round(mean(schist), 0)`% ± `r round(sd(schist), 0)`% of the cores' total mass.
+Soil (≤2 mm) dry mass was `r round(mean(sdm), 0)` ± `r round(sd(sdm), 0)` g.
+
+* Soil WC changes over incubation
+
+```{r wc, echo=FALSE}
+# wc and wc_alltrt are computed by 5-analyze.R script
+wcbegin <- subset(wc_alltrt, stage == "Beginning")
+wcend <- subset(wc, stage == "End")
+wcendmean <- wcend$mean
+names(wcendmean) <- wcend$Treatment
+wcendsd <- wcend$sd
+names(wcendsd) <- wcend$Treatment
+```
+
+Gravimetric water content was `r pn(wcbegin$mean, 2)` ± `r pn(wcbegin$sd, 2)` (min `r pn(wcbegin$min, 2)`, max `r pn(wcbegin$max, 2)`) at the beginning of the incubation (**Figure 1**).
+"Field moisture" cores were on average unchanged (`r pn(wcendmean["Field moisture"], 2)` ± `r pn(wcendsd["Field moisture"], 2)`) at the end of the incubation, but both the drought treatments, which did not differ from each other in their effect on gravimetric water content, had declined to `r pn(wcendmean["Drought"], 2)` ±  `r pn(wcendsd["Drought"], 2)`.
 

### 2016-02-11 manuscript_openexperiment/Open_experiment.Rmd
WC and BD calculations and results text

@@
   science'
-author: "Ben Bond-Lamberty^1^, Peyton Smith^2^, Vanessa Bailey^2^"
+author: "Ben Bond-Lamberty^1^, A. Peyton Smith^2^, Vanessa Bailey^2^"
 date: "February 2, 2016"

### 2016-02-11 R/0-functions.R

@@ -30,6 +30,7 @@ SUMMARYDATA_CLEAN_FILE <- file.path(OUTPUT_DIR, "summarydata_clean.csv")
 FLUXDATA_FILE          <- file.path(OUTPUT_DIR, "fluxdata.csv")
 FLUXDATA_CUM_FILE      <- file.path(OUTPUT_DIR, "fluxdata_cum.csv")
 FLUXDATA_CUM_CORE_FILE <- file.path(OUTPUT_DIR, "fluxdata_cum_core.csv")
+COREDATA_FILE          <- "data/drymasses.csv"
 
 SEPARATOR		  <- "-------------------"
 

### 2016-02-11 R/4-fluxes.R

@@ -6,8 +6,6 @@ source("R/0-functions.R")
 SCRIPTNAME  	<- "4-fluxes.R"
 PROBLEM       <- FALSE
 
-DRYMASSDATA   <- "data/drymasses.csv"
-
 library(reshape2)      # 1.4.1
 
 # ==============================================================================
@@ -20,7 +18,7 @@ printlog("Welcome to", SCRIPTNAME)
 summarydata <- read_csv(SUMMARYDATA_CLEAN_FILE)
 print_dims(summarydata)
 
-drymassdata <- read_csv(DRYMASSDATA, skip = 1)
+drymassdata <- read_csv(COREDATA_FILE, skip = 1)
 summarydata <- left_join(summarydata, drymassdata, by = "Core")
 
 # -----------------------------------------------------------------------------

### 2016-02-11 R/5-analyze.R

@@ -15,6 +15,8 @@ openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # o
 
 printlog("Welcome to", SCRIPTNAME)
 
+coredata <- read_csv(COREDATA_FILE, skip = 1)
+
 fluxdata_orig <- read_csv(FLUXDATA_FILE)
 
 printlog("Transforming...")
@@ -34,12 +36,27 @@ fluxdata$Treatment <- factor(fluxdata$Treatment, levels = c("Field moisture", "C
 save_data(fluxdata, fn = "fluxdata_long.csv")
 
 # -----------------------------------------------------------------------------
-# Water content over time figure
+# Water content over time figure and data
 
 figureA <- ggplot(fluxdata_orig, aes(inctime_days, WC_gravimetric, color=Treatment, group=Core)) 
 figureA <- figureA + geom_point() + geom_line()
 figureA <- figureA + facet_grid(~Temperature) 
 figureA <- figureA + xlab("Incubation day") + ylab("Gravimetric water content (fraction dry mass)")
+
+fluxdata$incday <- floor(fluxdata$inctime_days)
+fluxdata$stage <- NA
+fluxdata$stage[fluxdata$incday == min(fluxdata$incday)] <- "Beginning"
+fluxdata$stage[fluxdata$incday == max(fluxdata$incday)] <- "End"
+fluxdata %>% 
+  filter(!is.na(stage)) %>%
+  group_by(stage, Treatment) %>% 
+  summarise_each(funs(mean, sd, max, min), WC_gravimetric) ->
+  wc
+fluxdata %>% 
+  filter(!is.na(stage)) %>%
+  group_by(stage) %>% 
+  summarise_each(funs(mean, sd, max, min), WC_gravimetric) ->
+  wc_alltrt
 
 # -----------------------------------------------------------------------------
 # Fluxes over time figure
@@ -84,13 +101,12 @@ ch4_aov <- aov(cum_flux_mgC ~ Treatment + Temperature,
 ch4_hsd <- TukeyHSD(ch4_aov)
 print(ch4_hsd)
 
-# Convert the significance data to data frame, for easier (clearer)
+# Convert the significance data to data frame, for easier (and clearer)
 # access later by the manuscript code
 co2_hsd$Temperature <- as.data.frame(co2_hsd$Temperature)
[…]

### 2016-02-11 manuscript/cpcrw_manuscript.Rmd
Added volumetric WC data and results

@@
 
-* Core BD
-
 ```{r coredata, echo=FALSE}
@@
 
-* Soil WC changes over incubation
-
 ```{r wc, echo=FALSE}
 # wc and wc_alltrt are computed by 5-analyze.R script
-wcbegin <- subset(wc_alltrt, stage == "Beginning")
-wcend <- subset(wc, stage == "End")
-wcendmean <- wcend$mean
-names(wcendmean) <- wcend$Treatment
-wcendsd <- wcend$sd
-names(wcendsd) <- wcend$Treatment
+gwcbegin <- subset(gwc_alltrt, stage == "Beginning")
+gwcend <- subset(gwc, stage == "End")
+gwcendmean <- gwcend$mean
+names(gwcendmean) <- gwcend$Treatment
+gwcendsd <- gwcend$sd
+names(gwcendsd) <- gwcend$Treatment
+vwcbegin <- subset(vwc_alltrt, stage == "Beginning")
+vwcend <- subset(vwc_alltrt, stage == "End")
 ```
 
-Gravimetric water content was `r pn(wcbegin$mean, 2)` ± `r pn(wcbegin$sd, 2)` (min `r pn(wcbegin$min, 2)`, max `r pn(wcbegin$max, 2)`) at the beginning of the incubation (**Figure 1**).
-"Field moisture" cores were on average unchanged (`r pn(wcendmean["Field moisture"], 2)` ± `r pn(wcendsd["Field moisture"], 2)`) at the end of the incubation, but both the drought treatments, which did not differ from each other in their effect on gravimetric water content, had declined to `r pn(wcendmean["Drought"], 2)` ±  `r pn(wcendsd["Drought"], 2)`.
+Gravimetric water content was `r pn(gwcbegin$mean, 2)` ± `r pn(gwcbegin$sd, 2)` (min `r pn(gwcbegin$min, 2)`, max `r pn(gwcbegin$max, 2)`) at the beginning of the incubation (**Figure 1**).
+"Field moisture" cores were on average unchanged (`r pn(gwcendmean["Field moisture"], 2)` ± `r pn(gwcendsd["Field moisture"], 2)`) at the end of the incubation, but both the drought treatments, which did not differ from each other in their effect on gravimetric water content, had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
+Volumetric water content values ranged from `r pn(vwcbegin$mean, 2)` ± `r pn(vwcbegin$sd, 2)` (min `r pn(vwcbegin$min, 2)`, max `r pn(vwcbegin$max, 2)`) at the beginning of the experiment to `r pn(vwcend$mean, 2)` ± `r pn(vwcend$sd, 2)` (min `r pn(vwcend$min, 2)`, max `r pn(vwcend$max, 2)`) at the end.
 

### 2016-02-11 R/0-functions.R

@@ -1,6 +1,5 @@
-# Common function definitions.
-# Most critically, this file provides easy ggplot and data saving; 
-# logged csv[.gz|zip] read/write; and a few other handy things.
+# Common function and file location definitions.
+# This script is source'd by every other script before executing.
 # Ben Bond-Lamberty March 2015
 
 # -----------------------------------------------------------------------------
@@ -37,7 +36,7 @@ SEPARATOR		  <- "-------------------"
 
 # -----------------------------------------------------------------------------
 # Print dimensions of data frame
-print_dims <- function(d, dname=deparse(substitute(d))) {
+print_dims <- function(d, dname = deparse(substitute(d))) {
   stopifnot(is.data.frame(d))
   printlog(dname, "rows =", nrow(d), "cols =", ncol(d))
 } # print_dims
@@ -49,7 +48,7 @@ print_dims <- function(d, dname=deparse(substitute(d))) {
 outputdir <- function(scriptfolder = TRUE) {
   output_dir <- OUTPUT_DIR
   if(scriptfolder) output_dir <- file.path(output_dir, sub(".R$", "", SCRIPTNAME))
-  if(!file.exists(output_dir)) dir.create(output_dir)
+  if(!file.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
   output_dir
 } # outputdir
 
@@ -102,7 +101,7 @@ is_outlier <- function(x, devs = 3.2) {
   # J. Amer. Statist. Assoc., 88, 782-801.
   x <- na.omit(x)
   lims <- median(x) + c(-1, 1) * devs * mad(x, constant = 1)
-  x < lims[ 1 ] | x > lims[2]
+  x < lims[1] | x > lims[2]
 } # is_outlier
 
 # -----------------------------------------------------------------------------
@@ -116,11 +115,6 @@ save_diagnostic <- function(p, pname, printit = TRUE, ...) {
   save_plot(pname, ...)
 }
 
-
-if(!file.exists(OUTPUT_DIR)) {
-  printlog("Creating", OUTPUT_DIR)
-  dir.create(OUTPUT_DIR)
-}
 
 # -----------------------------------------------------------------------------
 #CHECKPOINTDATE	<- "2015-03-05" # comment out to not use checkpoint

### 2016-02-11 R/5-analyze.R

@@ -18,10 +18,13 @@ printlog("Welcome to", SCRIPTNAME)
 coredata <- read_csv(COREDATA_FILE, skip = 1)
 
 fluxdata_orig <- read_csv(FLUXDATA_FILE)
+fluxdata_orig$Treatment <- factor(fluxdata_orig$Treatment, 
+                                  levels = c("Field moisture", "Controlled drought", "Drought"))
 
 printlog("Transforming...")
 fluxdata_orig %>%
-  select(Core, Treatment, Temperature, inctime_days, WC_gravimetric, CO2_flux_µmol_g_s, CH4_flux_µmol_g_s,
+  select(Core, Treatment, Temperature, inctime_days, WC_gravimetric, WC_volumetric,
+         CO2_flux_µmol_g_s, CH4_flux_µmol_g_s,
          CO2_outlier, CH4_outlier) %>%
   melt(measure.vars = c("CO2_flux_µmol_g_s", "CH4_flux_µmol_g_s"),
        value.name = "flux_µmol_g_s") %>%
@@ -30,8 +33,6 @@ fluxdata_orig %>%
   mutate(Gas = substr(variable, 1, 3)) %>%
   select(-variable, -outlier_name) ->
   fluxdata
-
-fluxdata$Treatment <- factor(fluxdata$Treatment, levels = c("Field moisture", "Controlled drought", "Drought"))
 
 save_data(fluxdata, fn = "fluxdata_long.csv")
 
@@ -51,12 +52,22 @@ fluxdata %>%
   filter(!is.na(stage)) %>%
   group_by(stage, Treatment) %>% 
   summarise_each(funs(mean, sd, max, min), WC_gravimetric) ->
-  wc
+  gwc
 fluxdata %>% 
   filter(!is.na(stage)) %>%
   group_by(stage) %>% 
   summarise_each(funs(mean, sd, max, min), WC_gravimetric) ->
-  wc_alltrt
+  gwc_alltrt
+fluxdata %>% 
+  filter(!is.na(stage)) %>%
+  group_by(stage, Treatment) %>% 
+  summarise_each(funs(mean, sd, max, min), WC_volumetric) ->
+  vwc
+fluxdata %>% 
+  filter(!is.na(stage)) %>%
+  group_by(stage) %>% 
+  summarise_each(funs(mean, sd, max, min), WC_volumetric) ->
+  vwc_alltrt
 
 # -----------------------------------------------------------------------------
 # Fluxes over time figure

### 2016-02-24 manuscript/cpcrw_manuscript.Rmd
Updated with comments from @apeyton

@@
 The anaerobic conditions common following permafrost thaw lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and interact with soil temperature effects in long-term incubations {Elberling, 2013 #3720}.
-Such laboratory studies have weaknesses, but their controllled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
+Such laboratory studies have weaknesses, but their controlled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
+TODO: add Kirschbaum 2006.
+Most studies have focused on surface soils or permafrost soils, however, neglecting deep soils that experience fluctuating temperature with season and moisture with potential permafrost thaw ..or something less awkward (TODO; CITE).
+Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
 
 The goal of this study was to examine how temperature and moisture control C loss and GHG (CO~2~ and CH~4~) emissions from soils sampled from the bottom of the annual active layer--i.e., directly above permafrost--in an Alaskan boreal forest.
-Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
 We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at difference temperature and moisture levels.
 We hypothesized that soils maintained at field moisture and high temperature would lose more C, both as CO~2~ and CH~4~, than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity.
-In addition, cores subject to drought (reduced soil moisture) would have a higher potential to lose more ‘complex’ carbon compounds (aromatic C) as these compounds are desorbed due to increased surface tension.
-TODO: maybe change/reword this last sentence.
 
@@
 
+APS comments to address re above - TODO:
+
+* It doesn’t look like there was an interactive effect of temp*moisture, but just checking if you included that in the stat model?
+* This is good IF the controlled drought was really following the same dry-down rate of the 4C drought treatment (see comment on Figure 2 below).
+This way a comparison between 20C drought and 4C drought is more appropriate and meaningful.
+* Totally surprised by this [last sentence above]. Assumed CH4 would be higher in cores sitting in water (i.e. field moist) – perhaps this CH4→ CO2 ‘effect’ is occurring in these cores as well!?!! 
+* Cumulative CO2 between cores held at 4˚ C and cores held at 20˚ C in the“field moisture” treatment did not (???!!!) differ (TRUE?!! Kinda looks that way and that is interesting and useful
+
 * CO2:CH4 ratios
@@
 
+Treat et al.
+2014 also showed that CH4 production from organic permafrost soils has a positive relationship with temperature (cites several studies: Lupasco et al.
+2012, etc.).
+perhaps mention (briefly) here - that these are not surface layer soils (where majority of microbial activity and C mineralization of labile C takes place) and that taking them out of depth (where they are not exposed to O2 as readily, etc.) for our laboratory incubation may differ from in situ CO2/CH4 production.
+
+Increased temperature alone may not drive C mineralization rates in forest-mineral soils (Giardina and Ryan, 2000), but its corresponding increase in soil moisture content can lead to changes in microbial community structure and enhance anaerobic activity that may amplify CO2, CH4 and N2O fluxes (Xue et al, 2016).
+
 Note high CO2:CH4 ratios, compare with Treat et al. and other studies.
@@
 
+"Barfed bullets" from APS:
+
+*	Seeta’s paper: Mineral soil: greenhouse treatment (warming treatment) = 0.646 ± 0.04 GREATER THAN control = 0.43 ± 0.05 C-CO2 per g soil per day in 14 day incubation ( α ≤ 0.1).
+Mineral soil also had greater active biomass and extractable-OM).
+(Sistla et al.
+2013 Nature).
+*	Xue et al.
+2016: active layer, contained organic horizon: soil moisture increased in warmed plots: ~50% moisture compared to ~45% (p < 0.05.
+C Flux (g/m2) from ecosystem respiration was greater in warmed plots (~500 g/m2) compared to control plots (~400) (p < 0.05).
+Authors remarked that soil moisture/climate was important factor in shifting functional composition of microbial community (GeoChip), esp.
+with anaerobic activity.
+We did not see differences in our cum C-CH4 loss with temperature or moisture, but we did see relationship between moisture and CH4 in both field moist treatments (4˚ and 20˚) and 20˚ drought.
+*	Treat et al.
+2015.
+Meta-analysis of anaerobic incubations.
+Mineral soils 20-100 cm depth had a mean CO2 production (ug CO2-C /g C/day) of 100.6 +/- 10.3 (33.2 +/- 3.7 per g DW)  (7840 per m2/d).
+All soils < 10C were 67 +/- 16 and > 20C was 172 +/- 17.
+Dry soils 68 +/- 8 and Fluctuating = 364 +/- 60 and inundated = 105 +/- 11.
+*	Schaufler et al.
+2010: 33.7 +/- 2.7C02-C (mg C/ m2/ hr) averaged across all temperature (5-20C) and moisture (20-80% WFPS) in forested, boreal site, with 33.7 +/- 2.7.
+CH4-C = -30.2 +/- 4.5.
+Optimum moisture content was 40% WFPS for this site – or highest CO2 evolved at this moisture content.
+Carbon dioxide emissions were linearly correlated with temperature (p = 0.000) and exponentially correlated with WFPS (p = 0.000).
+CH4 uptake (oxidation) was highest in the northern sites (included boreal site), and was also the driest soil moisture conditions.
+-44.8 +/- 5.5.
+ug CH4-C/m2/hr and decreased with increasing soil moisture.
+Effect of WFPS on CH4 oxidation, exponential, p  = 0.000.
+*	Wu et al.
+2010.
+Mongolian steppe 900-1400msl.
+Soils = chernozems, but there was a gleyosol.
+Soil moisture and temperature influenced CH4.
+Soil drought limited CH4 uptake at all 15C.
+CO2 in meadow sites – increased with temperature (< 30 mg C/m2/hr @ 5C to 90 at 15C) and with moisture (<30-50 in dry treatment from 5-15C, and 50-90 in high rain treatment at 5-15C).
+
+Why would we expect to see greater CO2 evolved in higher temps?
+
+*	Increased temp (to a certain degree) increases enzyme kinetics (i.e.
+Arrhenius equation) – rate of enzyme reactions increase, thus decomposition increases (and CO2 production increases) (Kirschbaun  1995, 2006 for SOM-temp, Wallenstein/Allison/Schimel for enzyme-temp, etc.
+so many others)
+*	Increased temperature enhances the metabolic rate of microorganisms – thus increasing the production of extracellular enzymes (state somewhere prior that extracellular enzymes catalyze fundamental biogeochemical reactions in SOC decomposition and nutrient cycling).
+Cusack et al.
+(2010) suggests temperature-sensitivity of decomposition may be indirectly related to temperature dependence of enzyme production AND turnover.
+*	Carbon use efficiency (CUE), a measure of how well a microorganism can assimilate C into biomass relative to how much C is lost as CO2,  has been shown to vary with temperature (CITE – Lopez-Moran 2007; Price and Sowers 2004 !) with decreases in CUE at higher temperatures leading to greater CO2 production during SOM decomposition (Steinweg et al.
+2008).
+Cite Balser and Wixon 2009 in regards to microbial-mediated effects on temp-sensitivity of C decomposition.
+*	Higher temperatures enhance desorption rates of OM with minerals (ten Hulscher and Cornelissen 1996; Kalbitz et al.
+2000 – others) making more OC available in solution (Xu & Saiers, 2010; Plante et al.
+2009), and thus accessible for adjacent, competent microorganisms (CITE MORE).
+If we want to get technical – warmer temperatures enhance desorption of complex OC to minerals, while simultaneously providing more surface area for simple OC compounds to interact with minerals (i.e.
+simple C adsorps and complex C desorps in warmer temps).
+
 #### Acknowledgments
 
+We are grateful to Jamie Hollingsworth for information about, and faciliating access to, the Caribou Poker Creeks Research Watershed Long-Term Ecological Research(LTER site.
 This research was supported by Office of Science of the U.S.
@@
 
+APS comment: So are we seeing more CH4 produced in cores that I watered? Something to note in discussion?  TODO
+
 **Table 2.** Statistical tests of WC, T, C, N on individual core fluxes.
@@
 ```
+
+APS comment: Did the controlled drought treatment match the 4C drought treatment in terms of rate of water loss (for both gravimetric AND volumetric)?
+Since the controlled drought treatment was meant to control for increased water loss with the higher temp, I want to make sure that it really was doing that.
+From this graph, it “looks” like it followed the same trend as the 20C drought treatment.
+TODO
 
@@
 ```
+
+APS comment: Was difference in cum CO2 between field moist cores and drought cores the same for both temperatures???
+Granted cum Co2 was greater overall in 20C, but interesting that the difference between treatments appear equal in both temperatures.
+This is a bit difference from Treat et al.
+2014 (granted they had permafrost/peat soils) – but Treat reported larger differences between wet/dry (aerobic/anaerobic) treatments at 20C but no differences at -0.5C, but explain this due to frozen water limiting microbial respiration.
+TODO

### 2016-02-24 manuscript_openexperiment/Open_experiment.Rmd
Incorporated comments from VLB

@@
 
+I really want to put something in about research collaborations becoming increasingly global, and the current state of the internet and associated collaborative tools create opportunities never-before-known to conduct research across (previously) geographic barriers.
+
 Journals are adopting increasingly stringent data access and deposition policies, e.g.
@@
 Examples in ecology, soil science, and biogeosciences include BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, and SRDB {Bond-Lamberty, 2010 #2320}.
+Online databases and collaborative tools have expanded the abilities of researchers to collaborate across large distances, both improving data access and facilitating research partnerships (TODO).
 
@@
 Reproducibility of experimental results is at the heart of science {Kuhn, 1962 #4341}.
-Too often, however, it is impossible to attempt, because sufficient details and data are not publicly available.
+Too often, however, sufficient details and data are not publicly available to even attempt to reproduce a study.
 For example, surveying the 2000-2014 biomedical literature, Iqbal et al.
@@
 As code and data have become increasingly intertwined, the availability of the former has become a fundamental problem as well.
-Modeling groups have worked to improve reproducibility and archiving praactices {Thornton, 2005 #2310}.
+Modeling groups have worked to improve reproducibility and archiving practices {Thornton, 2005 #2310}.
 Scientists in all fields are increasingly building and using software in their work, however, often without strong training in this area {Wilson, 2014 #4337}.
@@
 Vines et al.
-{, 2014 #4130} published a shocking result, based on surveying 516 biology articles from 2 to 22 years old: the odds of a data set being available fell by 17% each year, and the chances that the contact author's email address still worked declined by 7% per year.
+{, 2014 #4130} published a shocking finding, based on a survey of 516 biology articles from 2 to 22 years old: the odds of a data set being available fell by 17% each year, and the chances that the contact author's email address still worked declined by 7% per year.
 Similarly, Reichman et al.
@@
 
-Data loss hits ecosystem, soil, and global change ecology particularly hard, as global to local climate changes make ecological data effectively irreproducible {Wolkovich, 2012 #3303}: we can never remeasure exactly the same system state.
-A subtler data loss issue is the "file drawer problem" {Rosenthal, 1979 #2291}, where unpublished but potentially valuable data are lost fovever.
+Data loss hits ecosystem, soil, and global change ecology particularly hard, as climate changes make ecological data effectively irreproducible {Wolkovich, 2012 #3303}: we can never remeasure exactly the same system state.
+A subtler data loss issue is the "file drawer problem" {Rosenthal, 1979 #2291}, where unpublished but potentially valuable data are lost forever.
 Scientists' use of strong and consistent data curation practices {Hart, 2015 #4292}{Rüegg, 2014 #3886} can mitigate the problem, but all our anecdotal experience and quantitative studies {Vines, 2014 #4130} suggest that in the long term, data cannot be reliably preserved by individual researchers.
@@
 - journals not ready, and don't want {Cassey, 2006 #4349}.
+
+You don’t comment on data journals (e.g. Scientific Data), or other repositories that provide a doi for your data (the CZO has a great page on this: http://criticalzone.org/national/data/data-doi/)
+Thoughts on including?
+
+Also – the genomics world has a long history of requiring sequences to be deposited, but there are multiple repositories, and often they lack active curation.
+Therefore we get a hodgepodge of unchecked data of varying quality distributed across 3-5 repositories around the globe.
+These repositories may gain or lose favor, and some may even collapse, rendering the data they stored lost, or difficult to retrieve.
 
@@
 The design of our repository is shown in **Figure 1**, and the repository itself can be found at https://github.com/bpbond/cpcrw_incubation.
-It consists of a series of scripts that comprise a *pipeline*, start with raw data and ending with final analyses, figures, and manuscripts.
+It consists of a series of scripts that feed their results from one to the next, starting with raw data and ending with final analyses, figures, and manuscripts.
 
@@
 * The version control system let us make incremental changes, work out problems, look at histories (i.e., who made what change when).
-* A issue tracker let us discuss problems, reference changes in the repository, create to-do lists, and assign responsibilities.
+* An issue tracker let us discuss problems, reference changes in the repository, create to-do lists, and assign responsibilities.
 * Information webpage on the experiment and broader project.
@@
 * The public nature of the repository encouraged us to flesh out documentation, use clean and clear coding, and think about the longer-term ramifications of decisions.
-* Real-time diagnostics (**Figure 2**) let our team, program managers, and interested other parties see at a glance the progress of the experiment.
+* Real-time diagnostics (**Figure 2**) let the team, our DOE program managers, and other interested parties see at a glance the progress of the experiment.
 * The issue tracker helped our team--which was not physically located all in one place--to communicate, discuss, and track "issues" (i.e., problems or questions that arose).
 * Investing the effort to set up a software pipeline before, not after, the experiment was performed meant that we could reliably and easily identify and diagnose problems.
+* The real-time collaboration, and the public exposure of all stages inspired us to keep all activities moving quickly, leading to (Insert timeline??)
 
@@
 Scientists also frequently cite other concerns about data sharing, for example being "scooped" not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
-We cannot easily dismiss of these these: for example, does our community adequately credit researchers who contribute to global databases that subsequently produce high-impact papers?
+We cannot easily dismiss of these: for example, does our community adequately credit researchers who contribute to global databases that subsequently produce high-impact papers?
 Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers efforts are adequately valued and cited {Kueffer, 2011 #4335}.
-We also recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}--that is, scientists do not receive 'proper' credit for datasets, relative to publications, even when the former may be much more broadly valueable.
+We also recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}--that is, scientists do not receive 'proper' credit for datasets, relative to publications, even when the former may be much more broadly valuable.
 
@@
 First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ: for example, there was no unit testing to ensure module correctness, nor an automated build tool automating the workflow (i.e., the individual scripts all have to be run manually).
-Second, it may not not be appropriate for applicable to all study types, even just in our professional fields of soil, ecosystem, and global change science.
-(And it certainly is not appropriate for sensitive human-subject data.) Third, Git and GitHub are designed for working with code, not data, although this situation continues to improve on a technical level[^6].
+Second, this approach may not be appropriate for, nor applicable to, all study types, even just in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive human-subject data.
+Third, Git and GitHub are designed for working with code, not data, although this situation continues to improve on a technical level[^6].
 

### 2016-02-26 manuscript_openexperiment/Open_experiment.Rmd
Incorporated Peyton's comments

@@
 date: "February 2, 2016"
-output: word_document
+output: html_document
 ---
@@
 
-Researchers in soil and ecosystem science, and almost every other field, as increasingly being pushed--by funders, journals, governments, and themselves--to ensure transparency and reproducibility of their work.
+Researchers in soil and ecosystem science, and almost every other field, as increasingly being pushed--by funders, journals, governments, and themselves--to increase transparency and ensure reproducibility of their work.
 A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, and enable powerful meta- and cross-disciplinary analyses.
 Here we describe a recent "open experiment", in which we documented every aspect of a soil incubation experiment online.
-This involved making raw data, scripts, diagnostics, final analyses, and manuscripts publicly available in real time.
-We found that using tools such as version control, issue tracking, and open-source statistical software improved data integrity, accelerated our team's communications, and ensured transparency.
+This involved making all raw data, scripts, diagnostics, final analyses, and manuscripts publicly available in real time.
+We found that using tools such as version control, issue tracking, and open-source statistical software improved data integrity, accelerated our team's communication and productivity, and ensured transparency.
 There are many avenues to improve scientific reproducibility and data availability, of which is this only one example, and it is not an approach suited for every experiment or situation.
@@
 
-Science is becomingly increasingly collaborative and data-intensive {Adams, 2012 #4348}.
+Science is becomingly increasingly collaborative and data-intensive {Adams, 2012 #4348}; in conjunction with revolutions in Internet-based communication, this has created new reserach opportunities across former geographic and disciplinary barriers.
 At the same time, a wide variety of factors are encouraging, at times pushing, scientists to increase data access and use "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.
 
-I really want to put something in about research collaborations becoming increasingly global, and the current state of the internet and associated collaborative tools create opportunities never-before-known to conduct research across (previously) geographic barriers.
-
-Journals are adopting increasingly stringent data access and deposition policies, e.g.
+These factors include journals, which are adopting increasingly stringent data access and deposition policies, e.g.
 Scientific Data[^1], PLOS ONE[^2], and Science[^3].
@@
 
-Finally, growing numbers of scientists are pushing for open science and data on moral/political grounds, as well as purely scientific ones, in particular ensuring reproducibility {Stodden, 2011 #4340} and enabling larger synthetic activities.
+Finally, growing numbers of scientists are pushing for open science and data on moral/political grounds, with respect to taxpayer-funded research, as well as purely scientific ones, in particular ensuring reproducibility {Stodden, 2011 #4340} and enabling larger synthetic activities.
 Such analyses {Wolkovich, 2012 #3303} are made possible by the assembly of large, internally consistent data sets.
@@
 
-**Reproducibility**
+**Repeatability and reproducibility**
 
-Reproducibility of experimental results is at the heart of science {Kuhn, 1962 #4341}.
-Too often, however, sufficient details and data are not publicly available to even attempt to reproduce a study.
+Reproducibility of experimental results is at the heart of science and a requirement for results to be accepted as factual {Kuhn, 1962 #4341}.
+Too often, however, sufficient details and data are not publicly available to even attempt to repeat a study (i.e.
+perform it again in a comparable manner, while not expecting exactly the same result).
 For example, surveying the 2000-2014 biomedical literature, Iqbal et al.
 {, 2016 #4334} found that none of 441 randomly-chosen studies provided raw data, and only one provided full protocols.
-The issue of reproducibility in ecology (and its many related fields) has been raised and debated for years {Cassey, 2006 #4349} but ecological, soil science, and global change journals differ widely in their data deposition requirements.
+The issues of reproducibility and repeatability in ecology (and its many related fields) have been raised and debated for years {Cassey, 2006 #4349}{Ellison, 2010 #4363} but ecological, soil science, and global change journals differ widely in their data deposition requirements, when such requirements exist at all.
 
@@
 Modeling groups have worked to improve reproducibility and archiving practices {Thornton, 2005 #2310}.
-Scientists in all fields are increasingly building and using software in their work, however, often without strong training in this area {Wilson, 2014 #4337}.
+Scientists in all fields are increasingly building and using software in their work, though often without strong training in this area {Wilson, 2014 #4337}.
 In addition to open-source data analysis languages such as R {R Development Core Team, 2016 #4314}, scientific workflow systems such as Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record information about the data processing, analytical process and decisions, and statistical analysis.
-Providing open code does not magically produce bug-free code, or mistake free analyses {Easterbrook, 2014 #4137}.
-But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task (often, forever).
+Providing open code does not magically produce bug-free code, mistake-free analyses, or instantly better science {Easterbrook, 2014 #4137}.
+But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task (often until key details have been forgotten, if not forever).
 
@@
 Data loss hits ecosystem, soil, and global change ecology particularly hard, as climate changes make ecological data effectively irreproducible {Wolkovich, 2012 #3303}: we can never remeasure exactly the same system state.
+This results in a critical role for syntheses and meta-analyses {Ellison, 2010 #4363}{Cassey, 2006 #4349}, which depend on the existence of documented protocols and (ideally raw) data.
 A subtler data loss issue is the "file drawer problem" {Rosenthal, 1979 #2291}, where unpublished but potentially valuable data are lost forever.
 Scientists' use of strong and consistent data curation practices {Hart, 2015 #4292}{Rüegg, 2014 #3886} can mitigate the problem, but all our anecdotal experience and quantitative studies {Vines, 2014 #4130} suggest that in the long term, data cannot be reliably preserved by individual researchers.
+
 This in turns means that the role of established, structured data and code repositories is critical.
@@
 TODO - VLB: Supplementary Information sections have become repositories of data and of non-core analyses.
-How does our open experiment replace (improve upon?) SI?
-Thoughts?
-- journals not ready, and don't want {Cassey, 2006 #4349}.
+Yet, data provided in Supplemental Information sections while available for use, may be accessible only in inconvenient formats (e.g.
+pdf) or have restrictions for use in many journals (Molloy 2011, Neylon 2012).
+
+How does our open experiment replace (improve upon?) SI? Thoughts?  - journals not ready, and don't want {Cassey, 2006 #4349}.
 
@@
 * Trust in utility of science: Why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results?
+
+While open science is often touted as paving the way for enhancing public trust in science, there are still many social and political challenges to overcome (Grandy et al.
+2012).
+For example, the language used in academic publications is often difficult for someone outside the specific field to understand (whether the reader is a scientist or not) and so simply providing unfettered access of scientific publications to the public may not improve the public’s perception and trust in science.
+While there is a strong need for the public to better trust science, the reverse is also true.
+Trust by scientists and institutions in the public to accurately interpret results scientists and institutions is also another challenge in the quest for open science (CITE).
+List examples: BSE risks, causes and beef sales (90s), or  GMO perceptions vs science (described in Millstone & van Zwanenberg 2000 Nature Medicine), or public perception fueling debates on bird flu mutation research which resulted in 60 day ban on bird flu research…hmm, perhaps not the best examples.
 
@@
 * An issue tracker let us discuss problems, reference changes in the repository, create to-do lists, and assign responsibilities.
-* Information webpage on the experiment and broader project.
+* Informational webpage on the experiment and broader project.
 * Manuscripts as well!
@@
 Tools such as R Markdown[^7] have made this process much easier to build and maintain.
+* Log files provide an audit trail {Ellison, 2010 #4363} of what analytical steps were taken and, critically, the specific versions used for each unit of software, including the main R system itself.
+This is critical as software changes with time.
 
@@
 
+Disadvantages:
+•	Convincing co-authors who are uncomfortable with pre-published data being made publically available of the benefits of open science
+•	Having Peyton adhere to a level of professionalism in all comments and communications as they were made publically available ☺ #firstworldproblem
+•	Writing a manuscript in code was not a skill all co-authors possessed, nor was it the most easy or efficient way to collaborate on a paper (ex: reference manager?) ☺
+
 #### Conclusions
@@
 Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers efforts are adequately valued and cited {Kueffer, 2011 #4335}.
-We also recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}--that is, scientists do not receive 'proper' credit for datasets, relative to publications, even when the former may be much more broadly valuable.
+We also recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}.
+That is, scientists do not receive 'proper' credit for datasets relative to publications, even when the former may be much more broadly valuable, in almost every area: hiring, promotion and tenure decisions, impact metrics, etc.
 
 The model we present here--a public repository updated throughout the experiment, analysis, and publication process--is far from perfect.
-First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ: for example, there was no unit testing to ensure module correctness, nor an automated build tool automating the workflow (i.e., the individual scripts all have to be run manually).
+First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ: for example, there was no unit testing to ensure module correctness, nor a build tool automating the workflow (i.e., the individual scripts all have to be run manually).
 Second, this approach may not be appropriate for, nor applicable to, all study types, even just in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive human-subject data.
@@
 [^6]: https://github.com/blog/1885-better-word-highlighting-in-diffs
+
+There have been attempts in environmental science to create data repositories – for example, the Data Observation Network for Earth, DataONE (https://www.dataone.org/what-dataone), provides a repository for earth science data and makes it accessible to the both the scientific and general public with the mission to: “ensure the preservation, access, use and reuse of multi-scale, multi-discipline, and multi-national science data…[for use] by researchers, educators, and the public to better understand and conserve life on earth and the environment that sustains it”  and allows researchers to cite data, workflows or visualizations located in the DataONE archive (Reichman et al.
+2011).
+Nonetheless (TODO: it's a lot better than nothing, and offers instrument-to-final product reproducibility).
+
+Another nonetheless: A study from the UK group, Research Information Network (as mentioned in Molloy 2011), titled “To share or not to share: research data outputs”[^7], interviewed 100 scientists about open data and in their summary of findings, study participants reported that they would be more incentivized to partake in open data practices if there was more “evidence that there are benefits to be had from publishing datasets (e.g.
+through case studies)”[^8].
+So in essence, making our open science experience into a case study may help contribute to the growing movement of open science/may encourage other researchers to participate.
+
+[^7]: http://www.rin.ac.uk/our-work/data-management-and-curation/share-or-not-share-research-data-outputs
+[^8]: http://www.rin.ac.uk/system/files/attachments/To-share-data-outputs-report.pdf
 
@@
 This is example, we hope, of the impact of individual scientists' decisions and practices {Wolkovich, 2012 #3303}.
+
+APS: Something that we haven’t really addressed in this paper is what I consider as one of the most valuable advantages of having our experiment open: The exposure!
+More visibility for us as scientists and PNNL.
+I love that.
+I also think that (while it hasn’t yet), one of us will get on another publication/collaboration due to this dataset being made public.
+I’m not sure where that goes in this MS and if there is a citation to back up my claims – but I’ll look for one nevertheless.
 

### 2016-02-26 manuscript_openexperiment/Open_experiment.Rmd
Streamlining and rewriting
Still need to deal with “Trust” section

@@
 Researchers in soil and ecosystem science, and almost every other field, as increasingly being pushed--by funders, journals, governments, and themselves--to increase transparency and ensure reproducibility of their work.
-A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, and enable powerful meta- and cross-disciplinary analyses.
+A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, enable powerful meta- and cross-disciplinary analyses, and increase the efficiency of publicly-funded research.
 Here we describe a recent "open experiment", in which we documented every aspect of a soil incubation experiment online.
@@
 These factors include journals, which are adopting increasingly stringent data access and deposition policies, e.g.
-Scientific Data[^1], PLOS ONE[^2], and Science[^3].
+*Scientific Data*[^1], *PLOS ONE*[^2], and *Science*[^3].
 These policies generally share common assumptions and goals: maximizing access to scientific data; encouraging deposition into structured repositories as opposed to journal supplementary information; and specifying that it is not acceptable for authors to be solely responsible for ensuring data access.
@@
 Examples in ecology, soil science, and biogeosciences include BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, and SRDB {Bond-Lamberty, 2010 #2320}.
-Online databases and collaborative tools have expanded the abilities of researchers to collaborate across large distances, both improving data access and facilitating research partnerships (TODO).
+Online databases and collaborative tools have expanded the abilities of researchers to collaborate across large distances, both improving data access and facilitating multidisciplinary research partnerships {Markowitz, 2015 #4368}.
 
@@
 As code and data have become increasingly intertwined, the availability of the former has become a fundamental problem as well.
-Modeling groups have worked to improve reproducibility and archiving practices {Thornton, 2005 #2310}.
-Scientists in all fields are increasingly building and using software in their work, though often without strong training in this area {Wilson, 2014 #4337}.
+Specialized modeling groups have worked to improve reproducibility and archiving practices {Thornton, 2005 #2310}.
+More generally, scientists in all fields are increasingly building and using software in their work, though often without strong training in this area {Wilson, 2014 #4337}.
 In addition to open-source data analysis languages such as R {R Development Core Team, 2016 #4314}, scientific workflow systems such as Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record information about the data processing, analytical process and decisions, and statistical analysis.
@@
 This in turns means that the role of established, structured data and code repositories is critical.
+These provide a much-needed improvement on "Supplementary Information" sections accompanying journal articles, which have become *de facto* repositories of data, but often accessible only in inconvenient formats (e.g.
+PDF), have restrictions for use, and uncertain long-term availability {Molloy, 2011 #4366;Neylon, 2012 #4367}.
+Finally, most journals do not have expertise in data storage and management {Cassey, 2006 #4349}.
+In contrast, the best repositories provid easy data uploading, assignment of data digital object identifiers (DOIs), and long-term stability and availability.
+They do, however, eliminate all risks: for example, the genomics sciences have a long history of requiring sequences to be deposited, but there are multiple repositories of varying curatorial levels, data quality, and formats, causing longer-term problems {Lagesen, 2010 #4369}.
 
-TODO - VLB: Supplementary Information sections have become repositories of data and of non-core analyses.
-Yet, data provided in Supplemental Information sections while available for use, may be accessible only in inconvenient formats (e.g.
-pdf) or have restrictions for use in many journals (Molloy 2011, Neylon 2012).
+**Institutional and social trust**
 
-How does our open experiment replace (improve upon?) SI? Thoughts?  - journals not ready, and don't want {Cassey, 2006 #4349}.
+Finally, there are longer-term issues of trust to consider.
 
-You don’t comment on data journals (e.g. Scientific Data), or other repositories that provide a doi for your data (the CZO has a great page on this: http://criticalzone.org/national/data/data-doi/)
-Thoughts on including?
-
-Also – the genomics world has a long history of requiring sequences to be deposited, but there are multiple repositories, and often they lack active curation.
-Therefore we get a hodgepodge of unchecked data of varying quality distributed across 3-5 repositories around the globe.
-These repositories may gain or lose favor, and some may even collapse, rendering the data they stored lost, or difficult to retrieve.
-
-**Institutional and social trust** (TODO)
-
-Finally, there are longer-term issues of trust to consider with respect to reproducibility and openness.
+An explicit governemntal rationale for increasing transparency and open data in science
 
@@
 * Trust in correctness of science: Potential to increase public trust in research results (e.g. climate change)
-* Trust in utility of science: Why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results?
+* Trust in utility of science: Why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results? Why should politicians and other
 
-While open science is often touted as paving the way for enhancing public trust in science, there are still many social and political challenges to overcome (Grandy et al.
-2012).
+greater return on public investment in scientific research {Neylon, 2012 #4367}.
+
+While open science is often touted as paving the way for enhancing public trust in science, there are still many social and political challenges to overcome {Grand, 2012 #4370}.
 For example, the language used in academic publications is often difficult for someone outside the specific field to understand (whether the reader is a scientist or not) and so simply providing unfettered access of scientific publications to the public may not improve the public’s perception and trust in science.
@@
 * An issue tracker let us discuss problems, reference changes in the repository, create to-do lists, and assign responsibilities.
-* Informational webpage on the experiment and broader project.
-* Manuscripts as well!
-(Including this one.) Results are linked directly into the manuscripts, so that if a result changes, it is updated automatically in the text.
+* An informational webpage provided non-technical explanations of the experiment and broader project.
+* Manuscripts were directly tied into the data system (cf.
+**Figure 1**), with numerical results flowing directly into e.g.
+results sections.
+This meant that changes to the data (and thus statistical results, etc.) propagated automatically and consistently.
 Tools such as R Markdown[^7] have made this process much easier to build and maintain.
 * Log files provide an audit trail {Ellison, 2010 #4363} of what analytical steps were taken and, critically, the specific versions used for each unit of software, including the main R system itself.
-This is critical as software changes with time.
+This is critical as software changes with time, a potentially big problem in reproducing previous analyses.
 
-[^7]: http://rmarkdown.rstudio.com
+[^6]: http://rmarkdown.rstudio.com
 
-Advantages:
+We found a number of advantages to this system:
 
@@
 * Investing the effort to set up a software pipeline before, not after, the experiment was performed meant that we could reliably and easily identify and diagnose problems.
-* The real-time collaboration, and the public exposure of all stages inspired us to keep all activities moving quickly, leading to (Insert timeline??)
-
-Disadvantages:
-•	Convincing co-authors who are uncomfortable with pre-published data being made publically available of the benefits of open science
-•	Having Peyton adhere to a level of professionalism in all comments and communications as they were made publically available ☺ #firstworldproblem
-•	Writing a manuscript in code was not a skill all co-authors possessed, nor was it the most easy or efficient way to collaborate on a paper (ex: reference manager?) ☺
+* The real-time collaboration, and the public exposure of all stages inspired us to keep all activities moving quickly--in this case, it was less than ten months between sampling soil cores in the Alaskan forest and submitting two resulting manuscripts.
+* Our project and experiment received more exposure and publicity than it otherwise would have.
 
@@
 Scientists also frequently cite other concerns about data sharing, for example being "scooped" not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
-We cannot easily dismiss of these: for example, does our community adequately credit researchers who contribute to global databases that subsequently produce high-impact papers?
+The benefits are not always clear to researchers who might otherwise be open to following open-data practices {Molloy, 2011 #4366}.
+We cannot easily dismiss all of these concerns: for example, does our community adequately credit researchers who contribute to global databases that subsequently produce high-impact papers?
 Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers efforts are adequately valued and cited {Kueffer, 2011 #4335}.
+
 We also recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}.
 That is, scientists do not receive 'proper' credit for datasets relative to publications, even when the former may be much more broadly valuable, in almost every area: hiring, promotion and tenure decisions, impact metrics, etc.
+The advent of 'data descriptor' articles, in journals such as *Scientific Data*, improves but does not solve this problem.
 
@@
 Second, this approach may not be appropriate for, nor applicable to, all study types, even just in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive human-subject data.
-Third, Git and GitHub are designed for working with code, not data, although this situation continues to improve on a technical level[^6].
+Third, Git and GitHub are designed for working with code, not data, although this situation continues to improve on a technical level[^7].
 
-[^6]: https://github.com/blog/1885-better-word-highlighting-in-diffs
+[^7]: https://github.com/blog/1885-better-word-highlighting-in-diffs
 
@@
 Nonetheless (TODO: it's a lot better than nothing, and offers instrument-to-final product reproducibility).
-
-Another nonetheless: A study from the UK group, Research Information Network (as mentioned in Molloy 2011), titled “To share or not to share: research data outputs”[^7], interviewed 100 scientists about open data and in their summary of findings, study participants reported that they would be more incentivized to partake in open data practices if there was more “evidence that there are benefits to be had from publishing datasets (e.g.
-through case studies)”[^8].
-So in essence, making our open science experience into a case study may help contribute to the growing movement of open science/may encourage other researchers to participate.
-
-[^7]: http://www.rin.ac.uk/our-work/data-management-and-curation/share-or-not-share-research-data-outputs
-[^8]: http://www.rin.ac.uk/system/files/attachments/To-share-data-outputs-report.pdf
 
@@
 This is example, we hope, of the impact of individual scientists' decisions and practices {Wolkovich, 2012 #3303}.
-
-APS: Something that we haven’t really addressed in this paper is what I consider as one of the most valuable advantages of having our experiment open: The exposure!
-More visibility for us as scientists and PNNL.
-I love that.
-I also think that (while it hasn’t yet), one of us will get on another publication/collaboration due to this dataset being made public.
-I’m not sure where that goes in this MS and if there is a citation to back up my claims – but I’ll look for one nevertheless.
 

### 2016-02-27 manuscript_openexperiment/Open_experiment.Rmd
Finished revisions for second draft

@@
 
-Researchers in soil and ecosystem science, and almost every other field, as increasingly being pushed--by funders, journals, governments, and themselves--to increase transparency and ensure reproducibility of their work.
-A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, enable powerful meta- and cross-disciplinary analyses, and increase the efficiency of publicly-funded research.
-Here we describe a recent "open experiment", in which we documented every aspect of a soil incubation experiment online.
-This involved making all raw data, scripts, diagnostics, final analyses, and manuscripts publicly available in real time.
+Researchers in soil and ecosystem science, and almost every other field, as increasingly being pushed--by funders, journals, governments, and themselves--to increase transparency and reproducibility of their work.
+A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, enable powerful meta- and cross-disciplinary analyses, and increase trust in, and the efficiency of, publicly-funded research.
+Here we describe a recent "open experiment", in which we documented every aspect of a soil incubation experiment online, making all raw data, scripts, diagnostics, final analyses, and manuscripts publicly available in real time.
 We found that using tools such as version control, issue tracking, and open-source statistical software improved data integrity, accelerated our team's communication and productivity, and ensured transparency.
@@
 PDF), have restrictions for use, and uncertain long-term availability {Molloy, 2011 #4366;Neylon, 2012 #4367}.
-Finally, most journals do not have expertise in data storage and management {Cassey, 2006 #4349}.
-In contrast, the best repositories provid easy data uploading, assignment of data digital object identifiers (DOIs), and long-term stability and availability.
-They do, however, eliminate all risks: for example, the genomics sciences have a long history of requiring sequences to be deposited, but there are multiple repositories of varying curatorial levels, data quality, and formats, causing longer-term problems {Lagesen, 2010 #4369}.
+Most journals do not have expertise in data storage and management {Cassey, 2006 #4349}.
+In contrast, the best repositories provide easy data uploading, assignment of data digital object identifiers (DOIs), and long-term stability and availability.
+They do not, however, eliminate all risks: for example, the genomics sciences have a long history of requiring sequences to be deposited, but there are multiple repositories of varying curatorial levels, data quality, and formats, causing longer-term problems {Lagesen, 2010 #4369}.
 
@@
 
-Finally, there are longer-term issues of trust to consider.
+Finally, there are longer-term issues of trust to consider--in particular, the public's trust in science and science's trust in people {Grand, 2012 #4370}.
+Both of these have been weakened by public controversies over particular issues--most particularly relevant to our fields of research, climate change {IPCC, 2013 #3916}.
+In addition to such issues of trust in the *correctness* of science, there is that of trust in *utility* of science: why should the public fund scientists if the latter are not producing demonstrably replicable and reusable results?
+Why should politicians and other stakeholders not push for greater return on public investment in scientific research, and conclude that more data openness serves this cause {Neylon, 2012 #4367}?
 
-An explicit governemntal rationale for increasing transparency and open data in science
-
-Governmental efforts to have the results of federally funded scientific research made available to the public, industry, and across the scientific community {Holdren, 2013 #4344}.
-
-* Trust in correctness of science: Potential to increase public trust in research results (e.g. climate change)
-* Trust in utility of science: Why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results? Why should politicians and other
-
-greater return on public investment in scientific research {Neylon, 2012 #4367}.
-
-While open science is often touted as paving the way for enhancing public trust in science, there are still many social and political challenges to overcome {Grand, 2012 #4370}.
-For example, the language used in academic publications is often difficult for someone outside the specific field to understand (whether the reader is a scientist or not) and so simply providing unfettered access of scientific publications to the public may not improve the public’s perception and trust in science.
-While there is a strong need for the public to better trust science, the reverse is also true.
-Trust by scientists and institutions in the public to accurately interpret results scientists and institutions is also another challenge in the quest for open science (CITE).
-List examples: BSE risks, causes and beef sales (90s), or  GMO perceptions vs science (described in Millstone & van Zwanenberg 2000 Nature Medicine), or public perception fueling debates on bird flu mutation research which resulted in 60 day ban on bird flu research…hmm, perhaps not the best examples.
-
-TODO: mention that of course authors have some rights.
-But what is being denied, increasingly, is that "the intellectual property rights of publishing ecologists" {Cassey, 2006 #4349} take precedence over all other factors, at least as far as publicly-funded research is concerned.
+These reasons have led, at least in part, to governments  efforts to have the results of federally funded scientific research made available to the public, industry, and across the scientific community {Holdren, 2013 #4344}.
+Strong "open science" and "open data" movements argue that the completeness of information provided by open science is fundamentally beneficial, complementing and perhaps replacing older systems for establishing trust within science {Grand, 2012 #4370}, and between science and the public.
+These movements increasingly deny that "the intellectual property rights of publishing [scientists]" {Cassey, 2006 #4349} take precedence over all other factors, at least as far as publicly-funded research is concerned.
 
@@
 
-Transparency is no magic panacea ensuring reproducibility {Easterbrook, 2014 #4137}, and may raise genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
-Scientists also frequently cite other concerns about data sharing, for example being "scooped" not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
+Transparency is no magic panacea ensuring repeatability and reproducibility {Easterbrook, 2014 #4137}, and may raise genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
+Nor will open science by itself fundamentally enhance public trust in science, as there are still many social and political challenges to overcome {Grand, 2012 #4370}.
+
+Scientists also frequently cite other concerns: about data sharing, being "scooped", not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
 The benefits are not always clear to researchers who might otherwise be open to following open-data practices {Molloy, 2011 #4366}.
@@
 The model we present here--a public repository updated throughout the experiment, analysis, and publication process--is far from perfect.
-First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ: for example, there was no unit testing to ensure module correctness, nor a build tool automating the workflow (i.e., the individual scripts all have to be run manually).
+First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ, in particular with respect to automation.
 Second, this approach may not be appropriate for, nor applicable to, all study types, even just in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive human-subject data.
-Third, Git and GitHub are designed for working with code, not data, although this situation continues to improve on a technical level[^7].
+Third, Git and GitHub are designed for working with code, not data.
+Although this situation continues to improve on a technical level[^7], it still requires a fair amount of coding expertise to set up and run a experimental repository such as the one detailed here.
 
@@
 
-There have been attempts in environmental science to create data repositories – for example, the Data Observation Network for Earth, DataONE (https://www.dataone.org/what-dataone), provides a repository for earth science data and makes it accessible to the both the scientific and general public with the mission to: “ensure the preservation, access, use and reuse of multi-scale, multi-discipline, and multi-national science data…[for use] by researchers, educators, and the public to better understand and conserve life on earth and the environment that sustains it”  and allows researchers to cite data, workflows or visualizations located in the DataONE archive (Reichman et al.
-2011).
-Nonetheless (TODO: it's a lot better than nothing, and offers instrument-to-final product reproducibility).
-
-Nonetheless (TODO: it's a lot better than nothing, and offers instrument-to-final product reproducibility).
-
-It doesn't have to be adopted wholesale: for example, the use of an issue tracker alone might be valuable for some groups and improve their science (though not reproducibility).
-
-This is example, we hope, of the impact of individual scientists' decisions and practices {Wolkovich, 2012 #3303}.
-
+In summary, a variety of forces continue to push scientists towards more transparency in their methods, code, and data, with the goals of increasing reproducibility, enabling syntheses and meta-analyses, and improving trust in, and return from, publicly-funded science.
+While imperfect, the open experiment example we highlight here does offer instrument-to-final product reproducibility and a very high level of transparency.
+Elements of it (e.g., the use of issue-tracking software) might be usefully adopted in isolation, but we hope the entire experiment will be an example of individual scientists' decisions and practices {Wolkovich, 2012 #3303} having a larger impact.
 We encourage the communities in our respective fields to consider its advantages, and to lead rather than follow with respect to scientific reproducibility, transparency, and data availability.

### 2016-02-27 manuscript/cpcrw_manuscript.Rmd
Code and ms text summarizing DOC, C, N

@@
 Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
+
+TODO: look at Podrebarac (2016) - Soils isolated during incubation underestimate temperature sensitivity of respiration and its response to climate history
 
@@
 Soil (≤2 mm) dry mass was `r round(mean(sdm), 0)` ± `r round(sd(sdm), 0)` g.
+
+Sample DOC was `r pn(cndata_summary["DOC_mg_kg_mean"], 2)` ± `r pn(cndata_summary["DOC_mg_kg_sd"], 2)` mg kg^-1^.
+Carbon content was `r pn(cndata_summary["C_percent_mean"], 2)`% ± `r pn(cndata_summary["C_percent_sd"], 2)`%, while N content was `r pn(cndata_summary["N_percent_mean"], 2)`% ± `r pn(cndata_summary["N_percent_sd"], 2)`%.
+Neither temperature was treatment exerted any significant effect on these properties (P > 0.1 for all).
 

### 2016-02-27 R/0-functions.R

@@ -22,14 +22,18 @@ CH4_EXCLUDE_DEVS <- 5.0    # CH4 outlier boundary, in mean absolute deviations
 # Key files and directories shared between scripts
 OUTPUT_DIR		         <- "outputs/"
 DIAGNOSTICS_DIR        <- "qc_plots/"
+
 RAWDATA_FILE           <- file.path(OUTPUT_DIR, "rawdata.csv.gz")
 RAWDATA_SAMPLES_FILE   <- file.path(OUTPUT_DIR, "rawdata_samples.csv.gz")
+VALVEMAP               <- "data/valvemap.csv"
+TREATMENTS             <- "data/treatments.csv"
 SUMMARYDATA_FILE       <- file.path(OUTPUT_DIR, "summarydata.csv")
 SUMMARYDATA_CLEAN_FILE <- file.path(OUTPUT_DIR, "summarydata_clean.csv")
 FLUXDATA_FILE          <- file.path(OUTPUT_DIR, "fluxdata.csv")
 FLUXDATA_CUM_FILE      <- file.path(OUTPUT_DIR, "fluxdata_cum.csv")
 FLUXDATA_CUM_CORE_FILE <- file.path(OUTPUT_DIR, "fluxdata_cum_core.csv")
 COREDATA_FILE          <- "data/drymasses.csv"
+CNDATA_FILE            <- "data/uga-soil_samples.csv"
 
 SEPARATOR		  <- "-------------------"
 

### 2016-02-27 R/2-summarize.R

@@ -12,8 +12,6 @@ source("R/0-functions.R")
 SCRIPTNAME  	<- "2-summarize.R"
 PROBLEM       <- FALSE
 
-VALVEMAP     <- "data/valvemap.csv"
-TREATMENTS   <- "data/treatments.csv"
 
 # -----------------------------------------------------------------------------
 # QC the valvemap data

### 2016-02-27 R/5-analyze.R

@@ -37,7 +37,39 @@ fluxdata_orig %>%
 save_data(fluxdata, fn = "fluxdata_long.csv")
 
 # -----------------------------------------------------------------------------
+# Read C, N, DOC data; summarise; test for treatment effects; merge with fluxdata
+
+printlog(SEPARATOR)
+printlog("Reading", CNDATA_FILE)
+read_csv(CNDATA_FILE) %>%
+  select(DOC_mg_kg,	C_percent, N_percent, Core) %>%
+  left_join(read_csv(TREATMENTS, skip = 1), by = "Core") ->
+  cndata
+
+cndata %>%
+  group_by(Core) %>%
+  summarise_each(funs(mean(., na.rm = TRUE)), DOC_mg_kg, C_percent, N_percent) %>%
+  summarise_each(funs(mean, sd), DOC_mg_kg, C_percent, N_percent) %>%
+  print %>%
+  unlist(.[1,]) ->
+  cndata_summary
+
+lm(formula = DOC_mg_kg ~ Treatment * Temperature, data = cndata) %>%
+  anova %>% print
+lm(formula = C_percent ~ Treatment * Temperature, data = cndata) %>%
+  anova %>% print
+lm(formula = N_percent ~ Treatment * Temperature, data = cndata) %>%
+  anova %>% print
+
+# fluxdata %>%
+#   left_join(cndata, by = c("Core", "Treatment", "Temperature")) ->
+#   fluxdata
+
+# -----------------------------------------------------------------------------
 # Water content over time figure and data
+
+printlog(SEPARATOR)
+printlog("Water content over time...")
 
 figureA <- ggplot(fluxdata_orig, aes(inctime_days, WC_gravimetric, color=Treatment, group=Core)) 
 figureA <- figureA + geom_point() + geom_line()
@@ -72,7 +104,8 @@ fluxdata %>%
 # -----------------------------------------------------------------------------
 # Fluxes over time figure
 
-printlog("fluxes over time...")
+printlog(SEPARATOR)
+printlog("Fluxes over time...")
 fluxdata %>%
   mutate(incday = floor(inctime_days)) %>%
   group_by(Gas, Temperature,Treatment, incday) %>%
@@ -100,6 +133,7 @@ fluxdata_figsBC %>%
 # -----------------------------------------------------------------------------
 # Cumulative flux figure and Tukey HSD tests
 
+printlog(SEPARATOR)
[…]

### 2016-02-29 manuscript/cpcrw_manuscript.Rmd
Mixed-effects models for fluxes and water content
Changed to proper (linear mixed-effects models, treating core as a
random effect) models for testing effects on gas fluxes and water
content. Two new tables in ms summarize the flux effects, and methods
text added about this.

@@
 
+The effects of temperature, gravimetric water content, C percent, N percent, and DOC concentration on instantaneous gas fluxes were evaluated using a linear mixed-effects model fit by the R function *lme* in the R 'nlme' package, , version `r packageVersion("nlme")`.
+Because the dependent variable (CO~2~ or CH~4~ flux) was non-normally distributed, it was transformed using a natural-logarithm transformation.
+Core number was treated as a random effect in the model.
+We then used performed stepwise model selection by Akaike's information criterion (AIC) using the *stepAIC* function in the R 'MASS' package, version `r packageVersion("MASS")`.
+A linear mixed-effects model was also used to evaluate the effect of treatment on core water content.
 The effect of temperature and treatment (drought, controlled drought, or field moisture conditions) on cumulative gas fluxes was evaluted with a post-hoc Tukey Honest Significant Differences test.
 
-All data analysis and statistics were performed using R version 3.2.3 {R Development Core Team, 2016 #4314}.
+All data analysis and statistics were performed using R version `r R.version.string` {R Development Core Team, 2016 #4314}.
 This experiment was run as an "open experiment" with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
@@
 Gravimetric water content was `r pn(gwcbegin$mean, 2)` ± `r pn(gwcbegin$sd, 2)` (min `r pn(gwcbegin$min, 2)`, max `r pn(gwcbegin$max, 2)`) at the beginning of the incubation (**Figure 1**).
-"Field moisture" cores were on average unchanged (`r pn(gwcendmean["Field moisture"], 2)` ± `r pn(gwcendsd["Field moisture"], 2)`) at the end of the incubation, but both the drought treatments, which did not differ from each other in their effect on gravimetric water content, had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
+"Field moisture" cores were on average unchanged (`r pn(gwcendmean["Field moisture"], 2)` ± `r pn(gwcendsd["Field moisture"], 2)`) at the end of the incubation, but both the drought treatments, which did not differ from each other in their effect on gravimetric water content (P = `r pn(p_trt_wc, 3)`), had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
 Volumetric water content values ranged from `r pn(vwcbegin$mean, 2)` ± `r pn(vwcbegin$sd, 2)` (min `r pn(vwcbegin$min, 2)`, max `r pn(vwcbegin$max, 2)`) at the beginning of the experiment to `r pn(vwcend$mean, 2)` ± `r pn(vwcend$sd, 2)` (min `r pn(vwcend$min, 2)`, max `r pn(vwcend$max, 2)`) at the end.
@@
 
-**Table 1.** Summary of gravimetric water content effects on gas fluxes, by treatment and temperature.
+**Table 1.** Statistical tests of WC, T, C, N on individual core CO2 fluxes.
 
 ```{r table1, echo=FALSE}
-# Print out the `WC_effect` data frame in a nice-looking table
-WC_effect$Gas[WC_effect$Gas == "CO2"] <- "CO<sub>2</sub>"  # html
-WC_effect$Gas[WC_effect$Gas == "CH4"] <- "CH<sub>4</sub>"  # html
-knitr::kable(WC_effect, digits = 3, col.names = c("Gas", "Treatment", "Temperature", "R<sup>2</sup>", "Sigma", "P", "AIC"))
+knitr::kable(summary(step_co2_lme)$tTable, digits = 3, col.names = c("Value", "SE", "DF", "T", "P"))
 ```
 
-APS comment: So are we seeing more CH4 produced in cores that I watered? Something to note in discussion?  TODO
+**Table 2.** Statistical tests of WC, T, C, N on individual core CH4 fluxes.
 
-**Table 2.** Statistical tests of WC, T, C, N on individual core fluxes.
+```{r table2, echo=FALSE}
+knitr::kable(summary(step_ch4_lme)$tTable, digits = 3, col.names = c("Value", "SE", "DF", "T", "P"))
+```
 

### 2016-02-29 R/5-analyze.R

@@ -5,6 +5,9 @@ source("R/0-functions.R")
 
 SCRIPTNAME  	<- "5-analyze.R"
 
+library(nlme)          # 3.1-122
+#library(MASS)          # 7.3-45
+
 library(broom)  # 0.4.0
 library(reshape2) # 1.4.1
 
@@ -34,8 +37,6 @@ fluxdata_orig %>%
   select(-variable, -outlier_name) ->
   fluxdata
 
-save_data(fluxdata, fn = "fluxdata_long.csv")
-
 # -----------------------------------------------------------------------------
 # Read C, N, DOC data; summarise; test for treatment effects; merge with fluxdata
 
@@ -43,12 +44,16 @@ printlog(SEPARATOR)
 printlog("Reading", CNDATA_FILE)
 read_csv(CNDATA_FILE) %>%
   select(DOC_mg_kg,	C_percent, N_percent, Core) %>%
+  group_by(Core) %>%
+  # Summarise by core, averaging duplicates
+  summarise_each(funs(mean(., na.rm = TRUE)), DOC_mg_kg, C_percent, N_percent) ->
+  cndata_orig
+
+cndata_orig %>%
   left_join(read_csv(TREATMENTS, skip = 1), by = "Core") ->
   cndata
 
 cndata %>%
-  group_by(Core) %>%
-  summarise_each(funs(mean(., na.rm = TRUE)), DOC_mg_kg, C_percent, N_percent) %>%
   summarise_each(funs(mean, sd), DOC_mg_kg, C_percent, N_percent) %>%
   print %>%
   unlist(.[1,]) ->
@@ -61,9 +66,12 @@ lm(formula = C_percent ~ Treatment * Temperature, data = cndata) %>%
 lm(formula = N_percent ~ Treatment * Temperature, data = cndata) %>%
   anova %>% print
 
-# fluxdata %>%
-#   left_join(cndata, by = c("Core", "Treatment", "Temperature")) ->
-#   fluxdata
+fluxdata %>%
+  left_join(cndata_orig, by = "Core") ->
+  fluxdata
+
+save_data(fluxdata, fn = "fluxdata_long.csv")
+
 
 # -----------------------------------------------------------------------------
 # Water content over time figure and data
@@ -82,7 +90,7 @@ fluxdata$stage[fluxdata$incday == min(fluxdata$incday)] <- "Beginning"
[…]

### 2016-02-29 R/4-fluxes.R
Changed primary units to µgC-gas/gC soil
…in addition to µgC-gas/gDW soil. This follows Treat et al. (2015) and
many other studies.

@@ -29,17 +29,18 @@ summarydata %>%
   filter(Treatment != "Ambient") %>%
   mutate(WC_gravimetric = 
            (Mass_g - CoreSleeveMass_g - NonsoilMassLarge_g - SoilDryMass_g) / 
-           (SoilDryMass_g),
+           SoilDryMass_g,
          WC_volumetric = (Mass_g - CoreSleeveMass_g - NonsoilMassLarge_g - SoilDryMass_g) / 
-           (SoilVolume_cm3)) %>%
-  select(Core, SoilDryMass_g, SoilVolume_cm3, WC_gravimetric, WC_volumetric,
+           SoilVolume_cm3) %>%
+  select(Core, Mass_g, SoilDryMass_g, SoilVolume_cm3, 
+         WC_gravimetric, WC_volumetric,
          Treatment, Temperature, CO2_ppm_s, CH4_ppb_s, inctime_days) ->
   fluxdata
 
 print(summary(fluxdata$WC_gravimetric))
 print(summary(fluxdata$WC_volumetric))
 
-p <- ggplot(fluxdata, aes(inctime_days, WC_gravimetric, color=paste(Treatment, Temperature), group=Core))
+p <- ggplot(fluxdata, aes(inctime_days, WC_gravimetric, color = paste(Treatment, Temperature), group=Core))
 p <- p + geom_line() + facet_wrap(~Core) + scale_color_discrete("")
 print(p)
 save_plot("WC_gravimetric")
@@ -56,7 +57,7 @@ fluxdata %>%
 # Flux computation
 
 # At this point, `fluxdata` has slopes (CO2 ppm/s and CH4 ppb/s).
-# We want to convert this to mg C/g soil/s, using
+# We want to convert this to mg C/s, using
 # A = dC/dt * V/M * Pa/RT (cf. Steduto et al. 2002), where
 # 	A is the flux (µmol/g/s)
 #	  dC/dt is raw respiration as above (mole fraction/s)
@@ -79,23 +80,23 @@ Pa 			<- 101						# kPa				(Richland is ~120 m asl)
 R 			<- 8.3145e+3			# cm3 kPa K−1 mol−1
 Tair    <- 273.1 + fluxdata$Temperature     # C -> K
 
-# Calculate mass-corrected respiration, µmol/g soil/s
-fluxdata$CO2_flux_µmol_g_s <- 
[…]

### 2016-03-01 R/0-functions.R
Reworked outlier testing procedure
We now divide into OUTLIER_GROUPS (currently 10) groups (sequentially
across time) and look for outliers within those groups. This should I
think better allow for considerably core-to-core variability, rather
than looking only at fluxes on a given day.

@@ -15,8 +15,9 @@ library(luzlogr)       # 0.2.0
 # -----------------------------------------------------------------------------
 # Parameters for key analytical choices
 # Defined here so they can be easily used in code AND manuscript
-CO2_EXCLUDE_DEVS <- 3.0    # CO2 outlier boundary, in mean absolute deviations
-CH4_EXCLUDE_DEVS <- 5.0    # CH4 outlier boundary, in mean absolute deviations
+OUTLIER_GROUPS   <- 10     # Divide data into groups and identify outliers in each
+CO2_EXCLUDE_DEVS <- 5.0    # CO2 outlier boundary, in mean absolute deviations
+CH4_EXCLUDE_DEVS <- 10.0   # CH4 outlier boundary, in mean absolute deviations
 
 # -----------------------------------------------------------------------------
 # Key files and directories shared between scripts

### 2016-03-01 R/4-fluxes.R

@@ -105,15 +105,17 @@ fluxdata$CH4_flux_mgC_hr <- CH4_flux_µmol_g_s /
 # -----------------------------------------------------------------------------
 # Outlier identification - see 
 # For each gas, group by Treatment, Temperature, and incubation day
-
 printlog("Identifying and plotting outliers...")
 fluxdata %>%
-  mutate(incday = floor(inctime_days)) %>%
-  group_by(Treatment, Temperature, incday) %>% 
-  mutate(CO2_outlier = is_outlier(CO2_flux_mgC_hr, devs = CO2_EXCLUDE_DEVS), 
+  arrange(inctime_days) %>%
+  ungroup %>%
+  mutate(group = cut(1:nrow(fluxdata), OUTLIER_GROUPS)) %>%
+  group_by(Treatment, Temperature, group) %>% 
+         # Identify outliers by flux/mass, i.e. normalizing for mass and water
+  mutate(CO2_outlier = is_outlier(CO2_flux_mgC_hr / Mass_g, devs = CO2_EXCLUDE_DEVS), 
          # CH4 is so variable we use a higher exclusion cutoff
-         CH4_outlier = is_outlier(CH4_flux_mgC_hr, devs = CH4_EXCLUDE_DEVS)) %>%
-  select(-incday) ->
+         CH4_outlier = is_outlier(CH4_flux_mgC_hr / Mass_g, devs = CH4_EXCLUDE_DEVS)) %>%
+  select(-group) ->
   fluxdata
 
 fluxdata$incday <- NULL  # why doesn't the `select` above work?

### 2016-03-01 R/5-analyze.R

@@ -159,13 +159,13 @@ p_trt_wc <- summary(m_wc)$tTable["TreatmentControlled drought", "p-value"]
 printlog(SEPARATOR)
 printlog("Fluxes over time...")
 fluxdata %>%
+  filter(!outlier) %>%
   mutate(incday = floor(inctime_days)) %>%
   group_by(Gas, Temperature,Treatment, incday) %>%
   summarise(flux = mean(flux_µgC_g_day), flux_sd = sd(flux_µgC_g_day)) ->
   fluxdata_figsBC
 
 figsBC <- function(fd) {
-  panellabels <- 
     ggplot(fd, aes(incday, flux)) + 
     geom_point() + 
     facet_grid(Temperature ~ Treatment) + 
@@ -238,7 +238,10 @@ fluxdata_cumulative %>%
 # Log-transforming doesn't fix the lack of normality in our data,
 # but it's a major improvement; see graphs.
 
-fluxdata <- fluxdata %>% filter(flux_µgC_gC_day > 0) 
+# Remove outliers and zero fluxes from the data set
+fluxdata %>% 
+  filter(flux_µgC_gC_day > 0 & !outlier) ->
+  fluxdata
 
 printlog("Graphing flux distributions and testing for normality...")
 p <- ggplot(fluxdata, aes(x = flux_µgC_gC_day)) + geom_histogram(bins = 30)

### 2016-03-01 manuscript/cpcrw_manuscript.Rmd
Added tentative UGA text, and flux summary results

@@
 
-# Run the analysis script, which will leave a lot of figures, tables, and data in memory for us
+# Run the main analysis script, which will leave figures, tables, and data 
+# in memory for us. Need to temporarily change the working directory to do so.
+olddir <- getwd()
 setwd("../")
 source("R/5-analyze.R")
+setwd(olddir)
 ```
@@
 Such dynamics may however be disrupted by increased fire disturbance {Johnstone, 2010 #2465}.
-In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon and nitrogen pools {Weiss, 2015 #4317}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
+In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
 
@@
 Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic C.
-TODO: get UGA analytical methods.
+Samples were combusted in an oxygen atmosphere at 1350 °C, converting elemental C and N into gaseous form.
+Infrared cells were then used to to determine C content, and a thermal conductivity cell to determine N.
+TODO: confirm UGA analytical methods and exact instruments.
 
@@
 
-* Soil C, N, DOC means
-
-* Gas fluxes over incubation (Figures 2 and 3)
+After removal of known problematic data, and exclusion of statistical outliers, CO~2~ flux rates measured during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"]/1000, 3)` mg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"]/1000, 3)` mg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"]/1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"]/1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"]/1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"]/1000, 3)`).
+CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 3)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 3)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 3)`).
+Gas fluxes varied considerably over the course of the incubation (**Figures 2 and 3**).
 
 * Statistical tests of WC, T, C, N on individual core fluxes (Table 2)
-
-* Cumulative fluxes (Figure 4)
 
@@
 * Cumulative CO2 between cores held at 4˚ C and cores held at 20˚ C in the“field moisture” treatment did not (???!!!) differ (TRUE?!! Kinda looks that way and that is interesting and useful
-
-* CO2:CH4 ratios
 

### 2016-03-01 manuscript/cpcrw_manuscript.Rmd
Bumping up flux values to ensure all positive numbers
Following Treat et al. (2015). Do we want to transform CH4 fluxes to ng?

@@
 The effects of temperature, gravimetric water content, C percent, N percent, and DOC concentration on instantaneous gas fluxes were evaluated using a linear mixed-effects model fit by the R function *lme* in the R 'nlme' package, , version `r packageVersion("nlme")`.
-Because the dependent variable (CO~2~ or CH~4~ flux) was non-normally distributed, it was transformed using a natural-logarithm transformation.
+Because the dependent variable (CO~2~ or CH~4~ flux) was non-normally distributed, it was transformed using a natural-logarithm (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^ to ensure all positive fluxes) transformation.
 Core number was treated as a random effect in the model.
@@
 
-**Table 1.** Statistical tests of WC, T, C, N on individual core CO2 fluxes.
+**Table 1.** Linear mixed-effects model parameters, testing effects of gravimetric water content (unitless), temperature (°C), soil C (%), soil N (%),  and dissolved organic carbon (mg kg^-1^) on individual core CO~2~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^).
+Dependent variable has units of log(µg C g C^-1^ day^-1^).
+Columns include parameter value; standard error (SE); degrees of freedom (DF); T statistic; and P value.
 
@@
 
-**Table 2.** Statistical tests of WC, T, C, N on individual core CH4 fluxes.
+**Table 2.** Linear mixed-effects model parameters, testing effects of gravimetric water content (unitless), temperature (°C), soil C (%), soil N (%),  and dissolved organic carbon (mg kg^-1^) on log-transformed, individual core CH~4~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^).
+Dependent variable has units of log(µg C g C^-1^ day^-1^).
+Columns include parameter value; standard error (SE); degrees of freedom (DF); T statistic; and P value.
 

### 2016-03-01 R/0-functions.R

@@ -18,6 +18,7 @@ library(luzlogr)       # 0.2.0
 OUTLIER_GROUPS   <- 10     # Divide data into groups and identify outliers in each
 CO2_EXCLUDE_DEVS <- 5.0    # CO2 outlier boundary, in mean absolute deviations
 CH4_EXCLUDE_DEVS <- 10.0   # CH4 outlier boundary, in mean absolute deviations
+FLUX_ADDITION <- 0.1       # Added to fluxes before log transformation
 
 # -----------------------------------------------------------------------------
 # Key files and directories shared between scripts

### 2016-03-01 R/5-analyze.R

@@ -323,10 +323,13 @@ save_data(shapiro_trans)
 printlog(SEPARATOR)
 printlog("Fitting CO2 model...")
 
+# Add a small value to all flux values to ensure they're positive before log-transform
+fluxdata$flux_µgC_gC_day1 <- fluxdata$flux_µgC_gC_day + FLUX_ADDITION
+
 # C_percent and N_percent are *highly* correlated (r=0.98+)
 # Generally it seems that N_percent produces slightly better model fits,
 # so we use it here, not considering C_percent further
-m_co2_lme <- lme(log(flux_µgC_gC_day) ~ Temperature * WC_gravimetric + 
+m_co2_lme <- lme(log(flux_µgC_gC_day1) ~ Temperature * WC_gravimetric + 
                    (Temperature + WC_gravimetric) * (N_percent + DOC_mg_kg), # C_percent + 
                  data = fluxdata,
                  subset = Gas == "CO2" & !outlier,
@@ -343,7 +346,7 @@ step_co2_lme <- MASS::stepAIC(m_co2_lme, direction = "both")
 
 printlog(SEPARATOR)
 printlog("Fitting CH4 model...")
-m_ch4_lme <- lme(log(flux_µgC_gC_day) ~ Temperature * WC_gravimetric + 
+m_ch4_lme <- lme(log(flux_µgC_gC_day1) ~ Temperature * WC_gravimetric + 
                    (Temperature + WC_gravimetric) * (N_percent + DOC_mg_kg), # C_percent + 
                  data = fluxdata, 
                  subset = Gas == "CH4" & !outlier,

### 2016-03-01 manuscript/cpcrw_manuscript.Rmd
Added UGA methods
Based on 01-March-2016 email: We ran DOC on the water extract of the
soils by “Shimadzu 5000 TOC Analyzer”. The total carbon and nitrogen
were analyzed by an “Elementar Vario Max CNS” combustion analyzer.

@@
 Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic C.
-Samples were combusted in an oxygen atmosphere at 1350 °C, converting elemental C and N into gaseous form.
-Infrared cells were then used to to determine C content, and a thermal conductivity cell to determine N.
-TODO: confirm UGA analytical methods and exact instruments.
+Samples were combusted in an oxygen atmosphere at 1350 °C, and measured for gaseous C and N using a Elementar Vario Max CNS.
+DOC was measured using a Shimadzu 5000 TOC Analyzer.
 

### 2016-03-29 manuscript/cpcrw_manuscript.Rmd
Introduction tweaks; beginning to organize discussion

@@
 
-Alaska forests under particular stress - browning {Ju, 2016 #4320}
-
 High latitude ecosystems are being subjected to rapid changes in climate {IPCC, 2013 #3916} and fire occurrence and intensity {Kasischke, 2010 #2631}.
-This has a wide variety of effects on ecosystems and society, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}.
-As a result, rising temperatures and increasing fire will result in permafrost degradation {Pastick, 2015 #4311} and changes in soil temperature and hydrology that are expected to increase soil greenhouse gas (GHG) fluxes to the atmosphere, promoting a positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
+This has a wide variety of effects on ecosystems and society, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}{Ju, 2016 #4320}.
+As a result, rising temperatures and increasing fire will result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372} and changes in soil temperature and hydrology that are expected to increase soil greenhouse gas (GHG) fluxes to the atmosphere, promoting a positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
 
@@
 
-TODO: look at Podrebarac (2016) - Soils isolated during incubation underestimate temperature sensitivity of respiration and its response to climate history
-
 The goal of this study was to examine how temperature and moisture control C loss and GHG (CO~2~ and CH~4~) emissions from soils sampled from the bottom of the annual active layer--i.e., directly above permafrost--in an Alaskan boreal forest.
@@
 
-The field component of this research took place in Caribou-Poker Creeks Research Watershed (CPCRW), part of the Bonanza Creek LTER (http://www.lter.uaf.edu/default.cfm).
-Annual average air temperature was -2.5 °C, and annual average precipitation was 400 mm {Petrone, 2006 #4313}.
+The field component of this research took place in Caribou-Poker Creeks Research Watershed (CPCRW), part of the Bonanza Creek LTER (http://www.lter.uaf.edu/research/study-sites-cpcrw).
+CPCRW is located in the Yukon-Tanana Uplands northeast of Fairbanks, AK, a part of the boreal forest that has seen strong increases in air temperature and forest browning {Ju, 2016 #4320} over several decades.
+Annual average air temperature was -2.5 °C, and annual average precipitation 400 mm {Petrone, 2006 #4313}.
 The watershed's lowlands and north-facing slopes are dominated by black spruce (*Picea mariana*), feathermoss (*Pleurozium schreberi* and others), and *Sphagnum* spp.; the drier south slopes tend to be deciduous with a mixture of quaking aspen (*Populus tremuloides*), paper birch (*Betula neoalaskana*), and patches of alder (*Alnus crispa*).
@@
 
+The cumulative flux numbers above imply Q10 values of XXXX...
+
 APS comments to address re above - TODO:
@@
 
-Compare absolute flux rates to other published studies from boreal mineral soils.
-
-Incubation experiments can only go so far though...
-Morishita et al.
-{, 2014 #4316} measured GHG fluxes in the field at CPCRW and nearby forests, and found CO2 production to be correlated with both soil temperature and moisture, but CH4 uptake driven by temperature only.
-
-Treat et al.
-2014 also showed that CH4 production from organic permafrost soils has a positive relationship with temperature (cites several studies: Lupasco et al.
-2012, etc.).
-perhaps mention (briefly) here - that these are not surface layer soils (where majority of microbial activity and C mineralization of labile C takes place) and that taking them out of depth (where they are not exposed to O2 as readily, etc.) for our laboratory incubation may differ from in situ CO2/CH4 production.
-
-Increased temperature alone may not drive C mineralization rates in forest-mineral soils (Giardina and Ryan, 2000), but its corresponding increase in soil moisture content can lead to changes in microbial community structure and enhance anaerobic activity that may amplify CO2, CH4 and N2O fluxes (Xue et al, 2016).
-
-Note high CO2:CH4 ratios, compare with Treat et al. and other studies.
+**Importance and context**
 
@@
 
-"Barfed bullets" from APS:
+Eastern Alaska high latitude warming and forest stress {Ju, 2016 #4320}.
 
-*	Seeta’s paper: Mineral soil: greenhouse treatment (warming treatment) = 0.646 ± 0.04 GREATER THAN control = 0.43 ± 0.05 C-CO2 per g soil per day in 14 day incubation ( α ≤ 0.1).
-Mineral soil also had greater active biomass and extractable-OM).
-(Sistla et al.
-2013 Nature).
-*	Xue et al.
-2016: active layer, contained organic horizon: soil moisture increased in warmed plots: ~50% moisture compared to ~45% (p < 0.05.
-C Flux (g/m2) from ecosystem respiration was greater in warmed plots (~500 g/m2) compared to control plots (~400) (p < 0.05).
-Authors remarked that soil moisture/climate was important factor in shifting functional composition of microbial community (GeoChip), esp.
-with anaerobic activity.
-We did not see differences in our cum C-CH4 loss with temperature or moisture, but we did see relationship between moisture and CH4 in both field moist treatments (4˚ and 20˚) and 20˚ drought.
+**Previous results**
+
+Morishita et al.
+{, 2014 #4316} measured GHG fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both soil temperature and moisture, consistent with our results, TODO.
+Morishita et al.
+{, 2014 #4316} found however that CH~4~ uptake (they observed no emissions) was driven by temperature only; this is...
+
+Waldrop et al.
+{, 2010 #4380} incubated  active-layer and permafrost soils from near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost); flux rates were 0.01-0.10 µmol CH~4~ day^-1^ g^-1^ and 1-8 µg C-CO2 hr^-1^ g^-1^.
+
+Treat et al. 2014 also showed that CH4 production from organic permafrost soils has a positive relationship with temperature (cites several studies: Lupasco et al. 2012, etc.)
 *	Treat et al.
@@
 Dry soils 68 +/- 8 and Fluctuating = 364 +/- 60 and inundated = 105 +/- 11.
+
+*	Xue et al.
+2016: active layer, contained organic horizon: soil moisture increased in warmed plots: ~50% moisture compared to ~45% (p < 0.05.
+C Flux (g/m2) from ecosystem respiration was greater in warmed plots (~500 g/m2) compared to control plots (~400) (p < 0.05).
+Authors remarked that soil moisture/climate was important factor in shifting functional composition of microbial community (GeoChip), esp.
+with anaerobic activity.
+
 *	Schaufler et al.
@@
 Effect of WFPS on CH4 oxidation, exponential, p  = 0.000.
+
 *	Wu et al.
@@
 
+Seeta’s paper: Mineral soil: greenhouse treatment (warming treatment) = 0.646 ± 0.04 GREATER THAN control = 0.43 ± 0.05 C-CO2 per g soil per day in 14 day incubation ( α ≤ 0.1).
+Mineral soil also had greater active biomass and extractable-OM).
+(Sistla et al.
+2013 Nature).
+
+**Temperature versus moisture sensitivity, and CO~2~ versus CH~4~**
+
 Why would we expect to see greater CO2 evolved in higher temps?
 
-*	Increased temp (to a certain degree) increases enzyme kinetics (i.e.
-Arrhenius equation) – rate of enzyme reactions increase, thus decomposition increases (and CO2 production increases) (Kirschbaun  1995, 2006 for SOM-temp, Wallenstein/Allison/Schimel for enzyme-temp, etc.
-so many others)
+*	Increased temp (to a certain degree) increases enzyme kinetics (i.e. Arrhenius equation) (Kirschbaun  1995, 2006 for SOM-temp, Wallenstein/Allison/Schimel for enzyme-temp, etc. so many others)
 *	Increased temperature enhances the metabolic rate of microorganisms – thus increasing the production of extracellular enzymes (state somewhere prior that extracellular enzymes catalyze fundamental biogeochemical reactions in SOC decomposition and nutrient cycling).
@@
 2000 – others) making more OC available in solution (Xu & Saiers, 2010; Plante et al.
-2009), and thus accessible for adjacent, competent microorganisms (CITE MORE).
-If we want to get technical – warmer temperatures enhance desorption of complex OC to minerals, while simultaneously providing more surface area for simple OC compounds to interact with minerals (i.e.
-simple C adsorps and complex C desorps in warmer temps).
+2009), and thus accessible for adjacent, competent microorganisms
+
+Note high CO2:CH4 ratios, compare with Treat et al. and other studies.
+
+Increased temperature alone may not drive C mineralization rates in forest-mineral soils (Giardina and Ryan, 2000), but its corresponding increase in soil moisture content can lead to changes in microbial community structure and enhance anaerobic activity that may amplify CO2, CH4 and N2O fluxes (Xue et al, 2016).
+
+**Limitations and weaknesses**
+
+Incubation experiments can only go so far though...
+
+these are not surface layer soils (where majority of microbial activity and C mineralization of labile C takes place) and that taking them out of depth (where they are not exposed to O2 as readily, etc.) for our laboratory incubation may differ from in situ CO2/CH4 production.
+
+Check out:
+Soil microbial respiration rate and temperature sensitivity along a north-south forest transect in eastern China: Patterns and influencing factors
+http://onlinelibrary.wiley.com/doi/10.1002/2015JG003217/abstract
+
+TODO: look at Podrebarac (2016) - Soils isolated during incubation underestimate temperature sensitivity of respiration and its response to climate history
 

### 2016-03-30 manuscript/cpcrw_manuscript.Rmd
Added Q10 calculation, corresponding methods & results text

@@
 
-Anomalous data were excluded based on their being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within each day, treatment, and temperature.
+Anomalous data were excluded based on their being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within a 10-day period, treatment, and temperature.
 We excluded `r en` of `r fluxn` (`r round(en/fluxn*100)`%) of the gas flux measurements for this reason.
@@
 The GHG concentration measurements for each sample (i.e., each gas, core, and date/time) were used to calculate a linear rate of change (δc/δt), based on the concentration rise from a minimum (at 2-10 seconds after measurement began) to a maximum (at 10-45 seconds).
-Each core’s respiration flux (F) was then calculated following e.g.
+Each core’s respiration flux (*F*) was then calculated following e.g.
 Steduto et al.
-{, 2002 #3383} as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where V is the core-specific system volume, M the core dry mass as determined at the end of the incubation, Pa atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), R the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and T the chamber air temperature (K) at time of measurement.
+{, 2002 #3383} as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where *V* is the core-specific system volume, *M* the core dry mass as determined at the end of the incubation, *P~a~* atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), *R* the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and *T* the chamber air temperature (K) at time of measurement.
 The final respiration rate was expressed on a dry soil mass basis (mg C kg soil^-1^ day^-1^).
@@
 
+Temperature sensitivity (Q~10~) was calculated for each gas and treatment as $\frac{F_2}{F_1}^{\left( \frac{10}{T_2-T_1} \right)}$ where *F~1~* and *F~2~* are the cumulative gas fluxes (mg C) at temperatures *T~1~* and *T~2~* (°C), respectively.
+
 * joined with mass and ancillary data
 * correlations
-* Q10 computation
-* stats
 
@@
 
-The cumulative flux numbers above imply Q10 values of XXXX...
-
-APS comments to address re above - TODO:
-
-* It doesn’t look like there was an interactive effect of temp*moisture, but just checking if you included that in the stat model?
-* This is good IF the controlled drought was really following the same dry-down rate of the 4C drought treatment (see comment on Figure 2 below).
-This way a comparison between 20C drought and 4C drought is more appropriate and meaningful.
-* Totally surprised by this [last sentence above]. Assumed CH4 would be higher in cores sitting in water (i.e. field moist) – perhaps this CH4→ CO2 ‘effect’ is occurring in these cores as well!?!! 
-* Cumulative CO2 between cores held at 4˚ C and cores held at 20˚ C in the“field moisture” treatment did not (???!!!) differ (TRUE?!! Kinda looks that way and that is interesting and useful
+The cumulative flux numbers above result in CH~4~ temperature sensitivity (Q~10~) values of `r pn(q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(q10_cumulative["Drought", "CH4"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CO~2~ were `r pn(q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(q10_cumulative["Drought", "CO2"], 1)`.
 

### 2016-03-30 R/5-analyze.R

@@ -146,8 +146,10 @@ fluxdata %>%
 
 printlog(SEPARATOR)
 printlog("Treatment effect on water content...")
+lowtemp <- min(fluxdata$Temperature, na.rm = TRUE)
+hightemp <- max(fluxdata$Temperature, na.rm = TRUE)
 fluxdata %>%
-  filter(Temperature == 20) ->
+  filter(Temperature == hightemp) ->
   fluxdata20
 fluxdata20$Treatment <- factor(fluxdata20$Treatment, levels = c("Drought", "Controlled drought", "Field moisture"))
 m_wc <- lme(WC_gravimetric ~ Treatment,
@@ -228,6 +230,18 @@ figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Tre
   facet_grid(Gas ~ ., scales = "free") +
   ylab(paste("Cumulative C (mg) over", floor(max(fluxdata$inctime_days)), "days"))
 
+# -----------------------------------------------------------------------------
+# Q10 based on cumulative fluxes
+# TODO: I hate the hardcoded temperature values below!
+
+stopifnot(length(unique(fluxdata_cumulative$Temperature)) == 2)
+fluxdata_cumulative %>%
+  reshape2::dcast(Treatment + Gas ~ Temperature, 
+                  value.var = "cum_flux_mgC") %>%
+  mutate(Q10 = (`20` / `4`) ^ (10 / (hightemp - lowtemp))) %>%
+  reshape2::dcast(Treatment ~ Gas, value.var = "Q10") ->  
+  q10_cumulative
+rownames(q10_cumulative) <- q10_cumulative$Treatment
 
 # -----------------------------------------------------------------------------
 # CO2:CH4 emissions ratio
@@ -279,50 +293,6 @@ fluxdata %>%
   print ->
   shapiro_trans
 save_data(shapiro_trans)
-
-# # -----------------------------------------------------------------------------
-# # Does WC affect gas fluxes within temperature and treatment?
-# 
-# printlog(SEPARATOR)
-# printlog("Summarizing WC effect...")
-# fluxdata %>%
-#   filter(flux_µmol_g_s > 0) %>%
-#   group_by(Gas, Treatment, Temperature) %>%
-#   filter(!outlier) %>%
-#   do(mod = lm(log(flux_µmol_g_s) ~ WC_gravimetric, data = .)) %>%
-#   broom::glance(mod) %>%
[…]

### 2016-03-30 manuscript/cpcrw_manuscript.Rmd
Methods text additions, revisions, and rearrangements

@@
 The final respiration rate was expressed on a dry soil mass basis (mg C kg soil^-1^ day^-1^).
-TODO check all units
 
-Temperature sensitivity (Q~10~) was calculated for each gas and treatment as $\frac{F_2}{F_1}^{\left( \frac{10}{T_2-T_1} \right)}$ where *F~1~* and *F~2~* are the cumulative gas fluxes (mg C) at temperatures *T~1~* and *T~2~* (°C), respectively.
-
-* joined with mass and ancillary data
-* correlations
-
-The effects of temperature, gravimetric water content, C percent, N percent, and DOC concentration on instantaneous gas fluxes were evaluated using a linear mixed-effects model fit by the R function *lme* in the R 'nlme' package, , version `r packageVersion("nlme")`.
+The effects of temperature, gravimetric water content, C percent, N percent, and DOC concentration on instantaneous gas fluxes were evaluated using a linear mixed-effects model fit by the R function *lme* in the R 'nlme' package, version `r packageVersion("nlme")`.
 Because the dependent variable (CO~2~ or CH~4~ flux) was non-normally distributed, it was transformed using a natural-logarithm (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^ to ensure all positive fluxes) transformation.
@@
 A linear mixed-effects model was also used to evaluate the effect of treatment on core water content.
+
+Cumulative respiration for each core and gas was calculated by linearly interpolating flux rates between measurement dates and summing respired C over the entire incubation.
 The effect of temperature and treatment (drought, controlled drought, or field moisture conditions) on cumulative gas fluxes was evaluted with a post-hoc Tukey Honest Significant Differences test.
+Temperature sensitivity (Q~10~) was calculated for each gas and treatment as $\frac{F_2}{F_1}^{\left( \frac{10}{T_2-T_1} \right)}$ where *F~1~* and *F~2~* are the cumulative gas fluxes (mg C) at temperatures *T~1~* and *T~2~* (°C), respectively.
 
@@
 
-We are grateful to Jamie Hollingsworth for information about, and faciliating access to, the Caribou Poker Creeks Research Watershed Long-Term Ecological Research(LTER site.
-This research was supported by Office of Science of the U.S.
+We are grateful to Jamie Hollingsworth for information about, and faciliating access to, the Caribou Poker Creeks Research Watershed Long-Term Ecological Research site.
+This research was supported by the Office of Science of the U.S.
 Department of Energy as part of the Terrestrial Ecosystem Sciences Program.
 The Pacific Northwest National Laboratory is operated for DOE by Battelle Memorial Institute under contract DE-AC05-76RL01830.
+
+#### Author contributions
+
+B.B.-L., A.P.S., and V.L.B.
+designed this experiment.
+B.B.-L.
+and A.P.S.
+performed field sampling, and A.P.S.
+led the laboratory incubation and analyses.
+B.B.-L.
+wrote the manuscript, with contributions from all authors.
 

### 2016-03-31 R/2-summarize.R
New `Remove_this` field in valve map, and relevant code

@@ -84,7 +84,7 @@ rawdata_samples <- rawdata %>%
 
 # Load and QC the valvemap data
 printlog("Loading valve map data...")
-valvemap <- read_csv(VALVEMAP, skip = 1, col_types = "ccnncdc")
+valvemap <- read_csv(VALVEMAP, skip = 1, col_types = "ccnncdcc")
 printlog( "Converting date/time info to POSIXct..." )
 valvemap$StartDateTime <- mdy_hm(paste(valvemap$Date, valvemap$Time_set_start_UTC))
 valvemap <- arrange(valvemap, StartDateTime)

### 2016-03-31 R/3-remove.R

@@ -9,7 +9,6 @@ SCRIPTNAME  	<- "3-remove.R"
 PROBLEM       <- FALSE
 
 CATEGORY_ERROR <- "Error"
-CATEGORY_OUTLIER <- "Outlier"
 
 REMOVALS     <- file.path("data/", "removals.csv")
 
@@ -28,15 +27,17 @@ print_dims(summarydata)
 # -----------------------------------------------------------------------------
 # Removals - outliers or errors that should be automatically removed
 
+printlog("Removing pre-flagged 'Remove_this' data...")
+removed_data <- subset(summarydata, !is.na(Remove_this))
+summarydata <- subset(summarydata, is.na(Remove_this))
+printlog("Summary data observations removed:", nrow(removed_data))
+
 printlog("Reading in removal data...")
 removals <- read_csv(REMOVALS, skip = 1)
 
 summarydata_clean <- summarydata
-removed_data <- data.frame()
-excluded_data <- data.frame()
 for(i in seq_len(nrow(removals))) {
   removed <- data.frame()
-  excluded <- data.frame()
   rem <- removals[i,]
   printlog(SEPARATOR)
   printlog("Removal", i)
@@ -52,9 +53,6 @@ for(i in seq_len(nrow(removals))) {
     if(rem$Category == CATEGORY_ERROR) {
       printlog("Removed entire sample", rem$samplenum)
       removed <- subset(summarydata_clean, samplenum == rem$samplenum)
-    } else if(rem$Category == CATEGORY_OUTLIER) {
-      printlog("Excluded entire sample", rem$samplenum)
-      excluded <- subset(summarydata_clean, samplenum == rem$samplenum)
     } else {
       stop("Unknown category!")
     }
@@ -63,17 +61,10 @@ for(i in seq_len(nrow(removals))) {
     stop("Method not implemented yet!")
   }
 
-    
   printlog("Summary data observations removed:", nrow(removed))
   removed_data <- rbind(removed_data, removed)
-  printlog("Summary data observations excluded:", nrow(excluded))
-  excluded_data <- rbind(excluded_data, excluded)
 }
 
-# -----------------------------------------------------------------------------
-# Exclusions - based on statistical test
-
-# -----------------------------------------------------------------------------
 # Done! 
 
 printlog(SEPARATOR)
[…]

### 2016-03-31 manuscript/cpcrw_manuscript.Rmd
Fixed outlier reporting

@@
 fluxn <- nrow(fluxdata)
-en <- sum(fluxdata$outlier)
 ```
 
-Anomalous data were excluded based on their being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within a 10-day period, treatment, and temperature.
-We excluded `r en` of `r fluxn` (`r round(en/fluxn*100)`%) of the gas flux measurements for this reason.
-Other data (`r round(rn/sdn*100)`%) were removed because of known instrument problems, e.g.
+Anomalous data were excluded based on their gas fluxes being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within a 10-day period, for a given treatment and temperature.
+We excluded `r outlier_count` of `r fluxn` (`r pn(outlier_count/fluxn*100, 1)`%) measurements for this reason.
+Other data (`r pn(rn/sdn*100, 1)`%) were removed because of known instrument problems, e.g.
 the analyzer was left running after leaving a chamber.

### 2016-03-31 R/0-functions.R

@@ -37,6 +37,7 @@ TREATMENTS             <- "data/treatments.csv"
 SUMMARYDATA_FILE       <- file.path(OUTPUT_DIR, "summarydata.csv")
 SUMMARYDATA_CLEAN_FILE <- file.path(OUTPUT_DIR, "summarydata_clean.csv")
 FLUXDATA_FILE          <- file.path(OUTPUT_DIR, "fluxdata.csv")
+FLUXDATA_FINAL_FILE    <- file.path(OUTPUT_DIR, "fluxdata_final.csv")
 FLUXDATA_CUM_FILE      <- file.path(OUTPUT_DIR, "fluxdata_cum.csv")
 FLUXDATA_CUM_CORE_FILE <- file.path(OUTPUT_DIR, "fluxdata_cum_core.csv")
 COREDATA_FILE          <- "data/drymasses.csv"

### 2016-03-31 R/4-fluxes.R

@@ -122,7 +122,6 @@ fluxdata$grp <- NULL  # why doesn't the `select` above work?
 p <- ggplot(fluxdata, aes(inctime_days, CO2_flux_mgC_hr, color = CO2_outlier))
 p <- p + geom_point() + facet_grid(Temperature ~ Treatment)
 print(p)
-#save_plot("CO2_outliers")
 save_diagnostic(p, "CO2_outliers")
 p <- ggplot(fluxdata, aes(inctime_days, CH4_flux_mgC_hr, color = CH4_outlier))
 p <- p + geom_point() + facet_grid(Temperature ~ Treatment)

### 2016-03-31 R/5-analyze.R

@@ -91,12 +91,15 @@ fluxdata %>%
          flux_µgC_gC_day = flux_µgC_g_day * C_percent) ->
   fluxdata
 
+save_data(fluxdata, fn = FLUXDATA_FINAL_FILE, scriptfolder = FALSE)
+
+# -----------------------------------------------------------------------------
+# Summarize flux information for each access by markdown code later
+
 fluxdata %>%
   group_by(Gas) %>%
   summarise_each(funs(min, max, mean, sd), flux_µgC_g_day, flux_µgC_gC_day) ->
   fluxsummary
-
-save_data(fluxdata, fn = "fluxdata_long.csv")
 
 nms <- fluxsummary$Gas
 fluxsummary <- as.data.frame(t(fluxsummary[-1]))
@@ -261,6 +264,7 @@ fluxdata_cumulative %>%
 # but it's a major improvement; see graphs.
 
 # Remove outliers and zero fluxes from the data set
+outlier_count <- sum(fluxdata$outlier)
 fluxdata %>% 
   filter(flux_µgC_gC_day > 0 & !outlier) ->
   fluxdata

### 2016-03-31 manuscript/cpcrw_manuscript.Rmd
Added another outlier test based on core CV
See issue #44

@@
 Anomalous data were excluded based on their gas fluxes being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within a 10-day period, for a given treatment and temperature.
-We excluded `r outlier_count` of `r fluxn` (`r pn(outlier_count/fluxn*100, 1)`%) measurements for this reason.
+We excluded `r mad_outlier_count` of `r pre_outlier_count` (`r pn(mad_outlier_count/pre_outlier_count*100, 1)`%) measurements for this reason.
+If the coefficient of variability of fluxes from any core on a single day exceeded `r pn(MAX_CO2_CV*100, 0)`%, the entire core was also excluded for that day (`r cv_outlier_count` data points, `r pn(cv_outlier_count/pre_outlier_count*100, 1)`%).
 Other data (`r pn(rn/sdn*100, 1)`%) were removed because of known instrument problems, e.g.

### 2016-03-31 R/0-functions.R

@@ -23,6 +23,7 @@ library(luzlogr)       # 0.2.0
 OUTLIER_GROUPS   <- 10     # Divide data into date groups and identify outliers in each
 CO2_EXCLUDE_DEVS <- 5.0    # CO2 outlier boundary, in mean absolute deviations
 CH4_EXCLUDE_DEVS <- 10.0   # CH4 outlier boundary, in mean absolute deviations
+MAX_CO2_CV       <- 1.4    # Max allowable CO2 CV for a given core on a given day
 FLUX_ADDITION <- 0.1       # Added to fluxes before log transformation
 
 # -----------------------------------------------------------------------------

### 2016-03-31 R/3.3-qc.R

@@ -97,11 +97,15 @@ core_cv <- summarydata %>%
   filter(Core != "Ambient4" & Core != "Ambient20") %>%
   group_by(Date, Core) %>% 
   summarise(CO2_ppm_s_cv = sd(CO2_ppm_s) / mean(CO2_ppm_s),
+            CH4_ppb_s_cv = sd(CH4_ppb_s) / mean(CH4_ppb_s),
             min_CO2_time = mean(min_CO2_time),
             max_CO2_time = mean(max_CO2_time),
             samplenums = paste(unique(samplenum), collapse = " ")) %>% 
   ungroup() %>% 
   arrange(desc(CO2_ppm_s_cv))
+
+save_plot("CO2_core_cv", p = qplot(CO2_ppm_s_cv, data = core_cv, bins = 40))
+save_plot("CH4_core_cv", p = qplot(CH4_ppb_s_cv, data = core_cv, bins = 40))
 
 p <-  qplot(1:nrow(core_cv), CO2_ppm_s_cv, data = core_cv)
 p <- p + ggtitle("Distribution of CO2 CV by core and date")

### 2016-03-31 R/4-fluxes.R

@@ -109,7 +109,10 @@ printlog("Identifying and plotting outliers...")
 fluxdata %>%
   arrange(inctime_days) %>%
   ungroup %>%
-  mutate(grp = cut(1:nrow(fluxdata), OUTLIER_GROUPS)) %>%
+  mutate(grp = cut(1:nrow(fluxdata), OUTLIER_GROUPS),
+         incday = floor(inctime_days)) %>%
+
+  # Outliers based on mean absolute deviation of flux from similar cores (in time, trt, temp)
   group_by(Treatment, Temperature, grp) %>% 
          # Identify outliers by flux/mass, i.e. normalizing for mass and water
   mutate(CO2_outlier = is_outlier(CO2_flux_mgC_hr / Mass_g, devs = CO2_EXCLUDE_DEVS), 
@@ -117,16 +120,28 @@ fluxdata %>%
          CH4_outlier = is_outlier(CH4_flux_mgC_hr / Mass_g, devs = CH4_EXCLUDE_DEVS)) ->
   fluxdata
 
-fluxdata$grp <- NULL  # why doesn't the `select` above work?
+  # Outliers based on core CV on a given day
+fluxdata %>%
+  group_by(incday, Core) %>%
+  summarise(CO2cv = sd(CO2_flux_mgC_hr, na.rm = TRUE) / mean(CO2_flux_mgC_hr, na.rm = TRUE),
+            CV_outlier = CO2cv > MAX_CO2_CV) %>%
+  left_join(fluxdata, ., by = c("incday", "Core")) ->
+  fluxdata
+
+fluxdata$grp <- fluxdata$incday <- NULL  # why doesn't the `select` above work?
 
 p <- ggplot(fluxdata, aes(inctime_days, CO2_flux_mgC_hr, color = CO2_outlier))
-p <- p + geom_point() + facet_grid(Temperature ~ Treatment)
+p <- p + geom_jitter() + facet_grid(Temperature ~ Treatment)
 print(p)
 save_diagnostic(p, "CO2_outliers")
 p <- ggplot(fluxdata, aes(inctime_days, CH4_flux_mgC_hr, color = CH4_outlier))
-p <- p + geom_point() + facet_grid(Temperature ~ Treatment)
+p <- p + geom_jitter() + facet_grid(Temperature ~ Treatment)
 print(p)
 save_diagnostic(p, "CH4_outliers")
[…]

### 2016-03-31 R/5-analyze.R

@@ -25,23 +25,38 @@ fluxdata_orig <- read_csv(FLUXDATA_FILE)
 fluxdata_orig$Treatment <- factor(fluxdata_orig$Treatment, 
                                   levels = c("Field moisture", "Controlled drought", "Drought"))
 
+# -----------------------------------------------------------------------------
+# Transform into long format and handle outliers
+
+co2_mad_outlier_count <- sum(fluxdata_orig$CO2_outlier, na.rm = TRUE)
+ch4_mad_outlier_count <- sum(fluxdata_orig$CH4_outlier, na.rm = TRUE)
+mad_outlier_count <- co2_mad_outlier_count + ch4_mad_outlier_count
+# The CV test applies to both CO2 and CH4, so # is multiplied by 2
+cv_outlier_count <- sum(fluxdata_orig$CV_outlier, na.rm = TRUE) * 2
+
 printlog("Transforming...")
 fluxdata_orig %>%
   select(samplenum, Core, Treatment, Temperature, inctime_days, 
          Mass_g, SoilDryMass_g, WC_gravimetric, WC_volumetric,
          CO2_flux_mgC_hr, CH4_flux_mgC_hr,
-         CO2_outlier, CH4_outlier) %>%
+         CO2_outlier, CH4_outlier, CV_outlier) %>%
   melt(measure.vars = c("CO2_flux_mgC_hr", "CH4_flux_mgC_hr"),
        value.name = "flux_mgC_hr") %>%
   mutate(Gas = substr(variable, 1, 3)) %>%
   select(-variable) ->
   fluxdata
 
+pre_outlier_count <- nrow(fluxdata)
+
 # Combine the gas-specific outlier fields into a single outlier flag
-# I originally was using reshape2::melt but this was duplicating rows
-fluxdata$outlier <- fluxdata$CO2_outlier
+fluxdata$outlier <- fluxdata$CO2_outlier | fluxdata$CV_outlier
 fluxdata$outlier[fluxdata$Gas == "CH4"] <- fluxdata$CH4_outlier[fluxdata$Gas == "CH4"]
-fluxdata$CO2_outlier <- fluxdata$CH4_outlier <- NULL
+fluxdata$CO2_outlier <- fluxdata$CH4_outlier <- fluxdata$CV_outlier <- NULL
+
+# Remove outliers and zero fluxes from the data set
+fluxdata %>% 
+  filter(flux_mgC_hr > 0 & !outlier) ->
+  fluxdata
 
[…]

### 2016-03-31 R/0-functions.R
Documentation note re last commit

@@ -24,6 +24,7 @@ OUTLIER_GROUPS   <- 10     # Divide data into date groups and identify outliers 
 CO2_EXCLUDE_DEVS <- 5.0    # CO2 outlier boundary, in mean absolute deviations
 CH4_EXCLUDE_DEVS <- 10.0   # CH4 outlier boundary, in mean absolute deviations
 MAX_CO2_CV       <- 1.4    # Max allowable CO2 CV for a given core on a given day
+                           # See https://github.com/bpbond/cpcrw_incubation/issues/44
 FLUX_ADDITION <- 0.1       # Added to fluxes before log transformation
 
 # -----------------------------------------------------------------------------

### 2016-04-08 manuscript/cpcrw_manuscript.Rmd
Intro and results: new text, summarize lme models

@@
 The anaerobic conditions common following permafrost thaw lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and interact with soil temperature effects in long-term incubations {Elberling, 2013 #3720}.
-Such laboratory studies have weaknesses, but their controlled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
-TODO: add Kirschbaum 2006.
-Most studies have focused on surface soils or permafrost soils, however, neglecting deep soils that experience fluctuating temperature with season and moisture with potential permafrost thaw ..or something less awkward (TODO; CITE).
+Such laboratory studies have weaknesses {Podrebarac, 2016 #4365}, but their controlled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
+Most studies have focused on surface soils or permafrost soils, however, neglecting deep active-layer soils that experience fluctuating temperature and, especially, moisture conditions.
 Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
@@
 We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at difference temperature and moisture levels.
-We hypothesized that soils maintained at field moisture and high temperature would lose more C, both as CO~2~ and CH~4~, than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity.
+We hypothesized that (i) CO~2~ would be the dominant pathway for C loss in these largely aerobic soils; (ii) soils maintained at field moisture and high temperature would lose more C-CO~2~ than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity; and (iii) core CH~4~ fluxes would be sensitive only to temperature.
 
@@
 Cores were weighed immediately after gas measurements.
-Ambient air was measured between treatment groups, and before starting measurements in a chamber.
-BEN: How was ambient air measurements used in this study?
-don't see any mention of it in remaining text, etc.
-(TODO)
+Ambient air was measured between treatment groups, and before starting measurements in a chamber, as a check on ambient CO~2~ conditions and instrument stability.
 
@@
 This experiment was run as an "open experiment" with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
-The flux data backing the main results have been archived at TODO.
+The summarized flux data backing the main results have been archived at [DOI TO BE FILLED IN UPON ACCEPTANCE].
 
@@
 Soil (≤2 mm) dry mass was `r round(mean(sdm), 0)` ± `r round(sd(sdm), 0)` g.
-
 Sample DOC was `r pn(cndata_summary["DOC_mg_kg_mean"], 2)` ± `r pn(cndata_summary["DOC_mg_kg_sd"], 2)` mg kg^-1^.
@@
 
-After removal of known problematic data, and exclusion of statistical outliers, CO~2~ flux rates measured during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"]/1000, 3)` mg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"]/1000, 3)` mg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"]/1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"]/1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"]/1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"]/1000, 3)`).
-CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 3)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 3)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 3)`).
+After removal of known problematic data, and exclusion of statistical outliers, CO~2~ flux rates measured during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 2)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 2)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 2)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 3)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 2)`).
+CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 2)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 2)`).
+
 Gas fluxes varied considerably over the course of the incubation (**Figures 2 and 3**).
-
-* Statistical tests of WC, T, C, N on individual core fluxes (Table 2)
-
-Cumulative CO~2~ evolved was highly affected by temperature (`r pclean(co2_hsd$Temperature["p adj"])`), and "field moisture" cores emitted significantly more CO~2~ than the other two treatments at both temperatures (P < 0.001 for both).
-There was no difference between the 20 °C "drought" and "controlled drought" treatments (`r pclean(co2_hsd$Treatment["Drought-Controlled drought", "p adj"])`).
-Neither temperature (`r pclean(ch4_hsd$Temperature[,"p adj"])`) nor treatment (mean `r pclean(mean(ch4_hsd$Treatment[,"p adj"]))`) was a significant factor in predicting cumulative CH~4~ fluxes.
-
-The cumulative flux numbers above result in CH~4~ temperature sensitivity (Q~10~) values of `r pn(q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(q10_cumulative["Drought", "CH4"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CO~2~ were `r pn(q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(q10_cumulative["Drought", "CO2"], 1)`.
+In the linear mixed-effects model, CO~2~ was strongly influenced by incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 1**).
+(Note that percent C and percent N were highly correlated (r ≥ 0.98) for these cores.
+Because percent N was a slightlyl stronger predictor, it was retained in the model while percent C was excluded.) The interaction between water content and soil N was also highly significant (P < 0.001), with high-N cores having little relationship between water content and CO~2~ flux (data not shown).
+Methane fluxes were most strongly correlated with percent N, while water content exhibited significant interactions with percent N and DOC (**Table 2**).
+Neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes, however.
 
@@
 
-The production of C from CO~2~ was over `r minmag` order of magnitudes higher than that from CH~4~, with CO~2~:CH~4~ C ratios ranging from `r minratio` million in the `r mintemp` °C "`r maxtrt`" treatment, to `r maxratio` million in the `r maxtemp` °C "`r maxtrt`" treatment.
+The cumulative production of C from CO~2~ was over `r minmag` order of magnitudes higher than that from CH~4~, with CO~2~:CH~4~ C ratios ranging from `r minratio` million in the `r mintemp` °C "`r maxtrt`" treatment, to `r maxratio` million in the `r maxtemp` °C "`r maxtrt`" treatment.
+Cumulative CO~2~ evolved was highly affected by temperature (`r pclean(co2_hsd$Temperature["p adj"])`), and "field moisture" cores emitted significantly more CO~2~ than the other two treatments at both temperatures (P < 0.001 for both).
+There was no difference between the 20 °C "drought" and "controlled drought" treatments (`r pclean(co2_hsd$Treatment["Drought-Controlled drought", "p adj"])`).
+Neither temperature (`r pclean(ch4_hsd$Temperature[,"p adj"])`) nor treatment (mean `r pclean(mean(ch4_hsd$Treatment[,"p adj"]))`) was a significant factor in predicting cumulative CH~4~ fluxes.
+
+The cumulative flux numbers above result in CH~4~ temperature sensitivity (Q~10~) values of `r pn(q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(q10_cumulative["Drought", "CH4"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CO~2~ were `r pn(q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(q10_cumulative["Drought", "CO2"], 1)`.
 
@@
 
+Several studies have measured microbial respiration and GHG fluxes from soils very close to our study site.
 Morishita et al.
-{, 2014 #4316} measured GHG fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both soil temperature and moisture, consistent with our results, TODO.
+{, 2014 #4316} measured gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both soil temperature and moisture, consistent with our results, TODO.
 Morishita et al.
@@
 Waldrop et al.
-{, 2010 #4380} incubated  active-layer and permafrost soils from near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost); flux rates were 0.01-0.10 µmol CH~4~ day^-1^ g^-1^ and 1-8 µg C-CO2 hr^-1^ g^-1^.
+{, 2010 #4380} incubated  active-layer and permafrost soils from near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost); flux rates were 0.01-0.10 µmol CH~4~ day^-1^ g^-1^ () and 1-8 µg C-CO2 hr^-1^ g^-1^.
+
+More generally,
 

### 2016-04-08 manuscript/cpcrw_manuscript.Rmd
Added WFPS calculation and Q10s normalized by it
See issue #30

@@
 ```{r coredata, echo=FALSE}
-bd <- with(coredata, SoilDryMass_g / CoreVolume_cm3)
+bd <- with(coredata, SoilDryMass_g / SoilVolume_cm3)
 schist <- with(coredata, NonsoilMassLarge_g / (NonsoilMassLarge_g + SoilDryMass_g)) * 100
@@
 Volumetric water content values ranged from `r pn(vwcbegin$mean, 2)` ± `r pn(vwcbegin$sd, 2)` (min `r pn(vwcbegin$min, 2)`, max `r pn(vwcbegin$max, 2)`) at the beginning of the experiment to `r pn(vwcend$mean, 2)` ± `r pn(vwcend$sd, 2)` (min `r pn(vwcend$min, 2)`, max `r pn(vwcend$max, 2)`) at the end.
+Water filled pore space, assuming a particle density of `r PARTICLE_DENSITY` g cm^-3^, was `r pn(min(fluxdata$WFPS_percent), 0)`-`r pn(max(fluxdata$WFPS_percent), 0)`%.
 
@@
 
-The cumulative flux numbers above result in CH~4~ temperature sensitivity (Q~10~) values of `r pn(q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(q10_cumulative["Drought", "CH4"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CO~2~ were `r pn(q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(q10_cumulative["Drought", "CO2"], 1)`.
+The cumulative flux numbers above result in CH~4~ temperature sensitivity (Q~10~) values of `r pn(Q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(Q10_cumulative["Drought", "CH4"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CO~2~ were `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)`.
+Computing Q~10~s based on fluxes normalized by water filled pore space decreased these values slightly: to `r pn(wfpsQ10_cumulative["Field moisture", "CH4"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CH4"], 1)` for CH~4~, respectively, and `r pn(wfpsQ10_cumulative["Field moisture", "CO2"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CO2"], 1)` for CO~2~.
 
@@
 Rises in boreal soil temperatures - who has quantified this? Changes in soil WC much less certain. Thermokarst etc., deepening of active layer - that's the larger implications of this work.
+
+Modeling results highly dependent on water content of soils after permafrost degradation.
 

### 2016-04-08 R/0-functions.R

@@ -26,6 +26,7 @@ CH4_EXCLUDE_DEVS <- 10.0   # CH4 outlier boundary, in mean absolute deviations
 MAX_CO2_CV       <- 1.4    # Max allowable CO2 CV for a given core on a given day
                            # See https://github.com/bpbond/cpcrw_incubation/issues/44
 FLUX_ADDITION <- 0.1       # Added to fluxes before log transformation
+PARTICLE_DENSITY <- 2.65   # g/cm3
 
 # -----------------------------------------------------------------------------
 # Key files and directories shared between scripts

### 2016-04-08 R/4-fluxes.R

@@ -27,18 +27,19 @@ summarydata <- left_join(summarydata, drymassdata, by = "Core")
 printlog("Computing water contents...")
 summarydata %>%
   filter(Treatment != "Ambient") %>%
-  mutate(WC_gravimetric = 
-           (Mass_g - CoreSleeveMass_g - NonsoilMassLarge_g - SoilDryMass_g) / 
-           SoilDryMass_g,
-         WC_volumetric = (Mass_g - CoreSleeveMass_g - NonsoilMassLarge_g - SoilDryMass_g) / 
-           SoilVolume_cm3) %>%
+  mutate(WaterMass_g = Mass_g - CoreSleeveMass_g - NonsoilMassLarge_g - SoilDryMass_g,
+         WC_gravimetric = WaterMass_g / SoilDryMass_g,
+         WC_volumetric = WaterMass_g / SoilVolume_cm3,
+         TotalPorosity = (1 - SoilDryMass_g / SoilVolume_cm3 / PARTICLE_DENSITY),
+         WFPS_percent =  (TotalPorosity - WC_volumetric) * 100) %>%
   select(samplenum, Core, Mass_g, SoilDryMass_g, SoilVolume_cm3, 
-         WC_gravimetric, WC_volumetric,
+         WC_gravimetric, WC_volumetric, WFPS_percent,
          Treatment, Temperature, CO2_ppm_s, CH4_ppb_s, inctime_days) ->
   fluxdata
 
 print(summary(fluxdata$WC_gravimetric))
 print(summary(fluxdata$WC_volumetric))
+print(summary(fluxdata$WFPS_percent))
 
 p <- ggplot(fluxdata, aes(inctime_days, WC_gravimetric, color = paste(Treatment, Temperature), group=Core))
 p <- p + geom_line() + facet_wrap(~Core) + scale_color_discrete("")
@@ -197,22 +198,23 @@ save_plot("cumulative_CH4")
 
 printlog("Calculating flux summary datasets...")
 fluxdata %>%
-  melt(id.vars = c("Treatment", "Temperature", "Core", "inctime_days"), 
+  melt(id.vars = c("Treatment", "Temperature", "Core", "inctime_days", "WFPS_percent"), 
        measure.vars = c("cumCO2_flux_mgC", "cumCH4_flux_mgC"),
        variable.name = "Gas") %>%
   mutate(Gas = substr(Gas, start = 4, stop = 6)) %>%   # rename
[…]

### 2016-04-08 R/5-analyze.R

@@ -37,7 +37,7 @@ cv_outlier_count <- sum(fluxdata_orig$CV_outlier, na.rm = TRUE) * 2
 printlog("Transforming...")
 fluxdata_orig %>%
   select(samplenum, Core, Treatment, Temperature, inctime_days, 
-         Mass_g, SoilDryMass_g, WC_gravimetric, WC_volumetric,
+         Mass_g, SoilDryMass_g, WC_gravimetric, WC_volumetric, WFPS_percent,
          CO2_flux_mgC_hr, CH4_flux_mgC_hr,
          CO2_outlier, CH4_outlier, CV_outlier) %>%
   melt(measure.vars = c("CO2_flux_mgC_hr", "CH4_flux_mgC_hr"),
@@ -157,7 +157,16 @@ fluxdata %>%
   group_by(stage) %>% 
   summarise_each(funs(mean, sd, max, min), WC_volumetric) ->
   vwc_alltrt
-
+fluxdata %>% 
+  filter(!is.na(stage)) %>%
+  group_by(stage, Treatment) %>% 
+  summarise_each(funs(mean, sd, max, min), WFPS_percent) ->
+  wfps
+fluxdata %>% 
+  filter(!is.na(stage)) %>%
+  group_by(stage) %>% 
+  summarise_each(funs(mean, sd, max, min), WFPS_percent) ->
+  wfps_alltrt
 
 # -----------------------------------------------------------------------------
 # Did 'Treatment' affect water content?
@@ -246,7 +255,7 @@ figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Tre
                     ymax = cum_flux_mgC + cum_flux_mgC_sd), 
                 position = position_dodge(0.9), width = 0.4) +  
   facet_grid(Gas ~ ., scales = "free") +
-  ylab(paste("Cumulative C (mg) over", floor(max(fluxdata$inctime_days)), "days"))
+  ylab(paste("Cumulative C (mg)"))
 
 # -----------------------------------------------------------------------------
 # Q10 based on cumulative fluxes
@@ -258,8 +267,17 @@ fluxdata_cumulative %>%
                   value.var = "cum_flux_mgC") %>%
   mutate(Q10 = (`20` / `4`) ^ (10 / (hightemp - lowtemp))) %>%
   reshape2::dcast(Treatment ~ Gas, value.var = "Q10") ->  
-  q10_cumulative
[…]

### 2016-04-15 manuscript_openexperiment/Open_experiment.Rmd
Tweak
Various tweaks - close to final draft

@@
 date: "February 2, 2016"
-output: html_document
+output: word_document
 ---
@@
 
-Science is becomingly increasingly collaborative and data-intensive {Adams, 2012 #4348}; in conjunction with revolutions in Internet-based communication, this has created new reserach opportunities across former geographic and disciplinary barriers.
-At the same time, a wide variety of factors are encouraging, at times pushing, scientists to increase data access and use "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.
+Science is becomingly increasingly collaborative and data-intensive {Adams, 2012 #4348}; in conjunction with revolutions in Internet-based communication, this has created new research opportunities across former geographic and disciplinary barriers.
+At the same time, many factors are encouraging, at times pushing, scientists to increase data access and use "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.
 
@@
 These policies generally share common assumptions and goals: maximizing access to scientific data; encouraging deposition into structured repositories as opposed to journal supplementary information; and specifying that it is not acceptable for authors to be solely responsible for ensuring data access.
-Funders are moving in this direction as well, with organizations such as the U.S.
+Funding agencies are moving in this direction as well, with organizations such as the U.S.
 National Science Foundation[^4] and the U.K.
@@
 
-Finally, growing numbers of scientists are pushing for open science and data on moral/political grounds, with respect to taxpayer-funded research, as well as purely scientific ones, in particular ensuring reproducibility {Stodden, 2011 #4340} and enabling larger synthetic activities.
+Finally, growing numbers of scientists are pushing for open science and data on moral and political grounds, as well as purely scientific ones, arguing that it is not acceptable to sequester taxpayer-funded research behind private publishers' paywalls {Neylon, 2012 #4367}.
+A second focus revolves around ensuring reproducibility {Stodden, 2011 #4340} and enabling larger synthetic activities.
 Such analyses {Wolkovich, 2012 #3303} are made possible by the assembly of large, internally consistent data sets.
@@
 
-We briefly discuss what we see as the primary arguments for data sharing and openness, before detailing a recent "open experiment" example.
+Here we briefly discuss what we see as the primary arguments for data sharing and openness, and then describe a recent "open experiment" example.
 
@@
 
-As code and data have become increasingly intertwined, the availability of the former has become a fundamental problem as well.
+As data and code have become increasingly intertwined, the availability of the latter has become a fundamental problem as well.
 Specialized modeling groups have worked to improve reproducibility and archiving practices {Thornton, 2005 #2310}.
@@
 Providing open code does not magically produce bug-free code, mistake-free analyses, or instantly better science {Easterbrook, 2014 #4137}.
-But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task (often until key details have been forgotten, if not forever).
+But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task--often until key details have been forgotten, if not forever.
 
@@
 In global change ecology, Wolkovich et al.
-{, 2012 #3303} reported that only ~10% of raw data sets were eventually shared in their attempt to conduct a meta-analysis {Wolkovich, 2012 #3331}.
+{, 2012 #3303} reported that they were able to acquire only ~10% of other researchers' raw data sets during a meta-analysis {Wolkovich, 2012 #3331}.
 
@@
 This in turns means that the role of established, structured data and code repositories is critical.
-These provide a much-needed improvement on "Supplementary Information" sections accompanying journal articles, which have become *de facto* repositories of data, but often accessible only in inconvenient formats (e.g.
-PDF), have restrictions for use, and uncertain long-term availability {Molloy, 2011 #4366;Neylon, 2012 #4367}.
+These provide a much-needed improvement on "Supplementary Information" sections accompanying journal articles, which have become *de facto* repositories of data, but often have inconvenient formats (e.g.
+PDF), poorly searchability, use restrictions, and uncertain long-term availability {Molloy, 2011 #4366;Neylon, 2012 #4367}.
 Most journals do not have expertise in data storage and management {Cassey, 2006 #4349}.
-In contrast, the best repositories provide easy data uploading, assignment of data digital object identifiers (DOIs), and long-term stability and availability.
+In contrast, the best repositories provide easy data uploading, immediate assignment of data digital object identifiers (DOIs), and long-term stability and availability.
 They do not, however, eliminate all risks: for example, the genomics sciences have a long history of requiring sequences to be deposited, but there are multiple repositories of varying curatorial levels, data quality, and formats, causing longer-term problems {Lagesen, 2010 #4369}.
@@
 
-These reasons have led, at least in part, to governments  efforts to have the results of federally funded scientific research made available to the public, industry, and across the scientific community {Holdren, 2013 #4344}.
+These reasons have led, at least in part, to governmental efforts to have the results of federally funded scientific research made available to the public, industry, and across the scientific community {Holdren, 2013 #4344}.
 Strong "open science" and "open data" movements argue that the completeness of information provided by open science is fundamentally beneficial, complementing and perhaps replacing older systems for establishing trust within science {Grand, 2012 #4370}, and between science and the public.
@@
 In early 2015, we planned a laboratory incubation experiment to characterize the chemical and biological properties of sub-Arctic, active layer soils subjected to changes in temperature and moisture.
-In this experiment, we would measure greenhouse gas fluxes from soil cores over 100 days, and measure the cores' physical, chemical, and biological characteristics under temperature and moisture changes.
+In this experiment, we would measure greenhouse gas fluxes from soil cores over 100 days, and measure the cores' physical, chemical, and biological characteristics under temperature and moisture changes {Bond-Lamberty, 2016 #4387}.
 This required (i) a multidisciplinary team that was not located in one time zone; (ii) integrating a variety of different data; (iii) performing quality control and diagnostics rapidly, so if e.g.
@@
 
-We used GitHub, a web-based Git repository hosting service, to store our data, scripts, and manuscripts.
-"Git" is a popular, free, and open source version control software: it tracks changes (when, what, by whom) made in a "repository", a collection of folders and files.
-Many scientific and other users of Git use the "GitHub" web service, as it offers excellent performance and a wide variety useful additional functinality, particularly for teams or collaborative projects.
+We used GitHub, a web-based *Git repository* hosting service, to store our data, scripts, and manuscripts.
+"Git" is a popular, free, and open source version control software: it tracks all changes (when, what, by whom) made in a "repository", a collection of folders and files.
+Many scientific and other users of Git use the "GitHub" or similar web services, as they offer a wide variety useful additional functionality, particularly for teams or collaborative projects.
 
@@
 * Log files provide an audit trail {Ellison, 2010 #4363} of what analytical steps were taken and, critically, the specific versions used for each unit of software, including the main R system itself.
-This is critical as software changes with time, a potentially big problem in reproducing previous analyses.
+This is critical as software changes with time, a potentially large problem in reproducing previous analyses.
 
@@
 * Investing the effort to set up a software pipeline before, not after, the experiment was performed meant that we could reliably and easily identify and diagnose problems.
-* The real-time collaboration, and the public exposure of all stages inspired us to keep all activities moving quickly--in this case, it was less than ten months between sampling soil cores in the Alaskan forest and submitting two resulting manuscripts.
+* The real-time collaboration, and the public exposure of all stages inspired us to keep all activities moving quickly.
+In this case, it was less than nine months between sampling soil cores in the Alaskan forest and submitting two resulting manuscripts.
 * Our project and experiment received more exposure and publicity than it otherwise would have.
@@
 We cannot easily dismiss all of these concerns: for example, does our community adequately credit researchers who contribute to global databases that subsequently produce high-impact papers?
-Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers efforts are adequately valued and cited {Kueffer, 2011 #4335}.
+Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers' efforts are adequately valued and cited {Kueffer, 2011 #4335}.
 
@@
 Third, Git and GitHub are designed for working with code, not data.
-Although this situation continues to improve on a technical level[^7], it still requires a fair amount of coding expertise to set up and run a experimental repository such as the one detailed here.
+Although this situation continues to improve on a technical level[^7], it still requires a fair amount of coding expertise to set up and run an experimental repository such as the one detailed here.
 
@@
 
-**References**
+#### Acknowledgments
 
-**Figure 1.** The 'repository' containing all code and data of a recent soil incubation experiment.
+This research was supported by the Office of Science of the U.S.
+Department of Energy as part of the Terrestrial Ecosystem Sciences Program.
+The Pacific Northwest National Laboratory is operated for DOE by Battelle Memorial Institute under contract DE-AC05-76RL01830.
+
+#### References
+
+**Figure 1.** The 'repository' containing all code and data of a recent soil incubation experiment {Bond-Lamberty, 2016 #4387}.
 From top to bottom: raw data from a lab analyzer is uploaded; an initial script processes raw data into a standardized, summarized form; a second script looks for inconsistencies and outliers in the data; a third computes final data products, summaries, statistics, and figures; and a final one integrates these products to produce a submission-ready manuscript.
@@
 
-**Figure 2.** Two of the many diagnostic figures that were produced, in real time, as the incubation experiment proceeded.
+**Figure 2.** Two of the many diagnostic figures that were produced as the incubation experiment proceeded.
 They show (top) the number of replicate measurements made on each soil core, by date of measurement, and (bottom) measurements with a data mismatch problem, indicating a data entry or other error (in the figure no errors, which would be in red, are extant).
 A wide range of similar diagnostic graphics (see https://github.com/bpbond/cpcrw_incubation#current-diagnostics) and statistics let us quickly identify problems or interesting patterns in the data.
+All diagnostics were publicly available in real time.
 

### 2016-04-15 manuscript/cpcrw_manuscript.Rmd
Revisions to introduction, methods, results
Also moved MAX_MINCONC_TIME and MAX_MAXCONC_TIME definitions to
`0-functions.R`.

@@
 This has a wide variety of effects on ecosystems and society, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}{Ju, 2016 #4320}.
-As a result, rising temperatures and increasing fire will result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372} and changes in soil temperature and hydrology that are expected to increase soil greenhouse gas (GHG) fluxes to the atmosphere, promoting a positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
+In particular, rising temperatures and increasing fire will result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372} and changes in soil temperature and hydrology that are expected to increase soil greenhouse gas (GHG) fluxes to the atmosphere, promoting a positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
 
@@
 GHG responses to wetting and thawing dynamics are much less certain, with substantial variability between studies {Kim, 2012 #2863}.
-The anaerobic conditions common following permafrost thaw lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and interact with soil temperature effects in long-term incubations {Elberling, 2013 #3720}.
-Such laboratory studies have weaknesses {Podrebarac, 2016 #4365}, but their controlled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
-Most studies have focused on surface soils or permafrost soils, however, neglecting deep active-layer soils that experience fluctuating temperature and, especially, moisture conditions.
-Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
+The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and interact with soil temperature effects in long-term incubations {Elberling, 2013 #3720}.
+Decadal warming, and perhaps drying, trends in Alaska {Bieniek, 2014 #4389} will work to counteract these effects, with uncertain consequences for the carbon cycle and climate system.
 
 The goal of this study was to examine how temperature and moisture control C loss and GHG (CO~2~ and CH~4~) emissions from soils sampled from the bottom of the annual active layer--i.e., directly above permafrost--in an Alaskan boreal forest.
+Most previous studies have focused on surface soils or permafrost soils, neglecting deep active-layer soils that experience fluctuating temperature and, especially, moisture conditions.
+Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
 We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at difference temperature and moisture levels.
-We hypothesized that (i) CO~2~ would be the dominant pathway for C loss in these largely aerobic soils; (ii) soils maintained at field moisture and high temperature would lose more C-CO~2~ than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity; and (iii) core CH~4~ fluxes would be sensitive only to temperature.
+We hypothesized that (i) CO~2~ would be the dominant pathway for C loss in these largely aerobic soils; (ii) soils maintained at field moisture and high temperature would lose more C-CO~2~ than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity; and (iii) core CH~4~ fluxes would be sensitive only to temperature, as no anaerobic conditions were planned for the cores.
 
@@
 These groups included two incubation temperatures (4 and 20 °C), each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
-The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status would close matched those of the 4 °C "drought" cores, which we anticipated would dry more slowly than their 20 °C counterparts.
+The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status would close match those of the 4 °C "drought" cores, which we anticipated would dry more slowly than their 20 °C counterparts.
 The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density.
@@
 
-We measured the rise in GHG concentrations from its minimum value (in the first 10 seconds of measurement after the analyzer was sealed to the soil core) to its maximum (in the first 45 seconds), considering this as the flux rate in ppm s^-1^ (CO~2~) or ppb s^-1^ (CH~4~).
+At each GHG measurement, we measured the rise in gas concentrations, considering this as the flux rate in ppm s^-1^ (CO~2~) or ppb s^-1^ (CH~4~).
+The GHG concentration measurements for each sample (i.e., each gas, core, and date/time) were used to calculate a linear rate of change (δc/δt), based on the concentration rise from a minimum (up to `r MAX_MINCONC_TIME` seconds after measurement began) to a maximum (at `r MAX_MINCONC_TIME`-`r MAX_MAXCONC_TIME` seconds).
 This value was calculated for each sample, i.e.
@@
 # TODO: this isn't good--fragile if files move or change names
-# Want everything in-memory (or something)
 rn <- readLines("../outputs/3-remove/removed_data.csv") %>% length
@@
 
-The GHG concentration measurements for each sample (i.e., each gas, core, and date/time) were used to calculate a linear rate of change (δc/δt), based on the concentration rise from a minimum (at 2-10 seconds after measurement began) to a maximum (at 10-45 seconds).
-Each core’s respiration flux (*F*) was then calculated following e.g.
+Each core’s respiration flux (*F*) was calculated following e.g.
 Steduto et al.
 {, 2002 #3383} as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where *V* is the core-specific system volume, *M* the core dry mass as determined at the end of the incubation, *P~a~* atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), *R* the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and *T* the chamber air temperature (K) at time of measurement.
-The final respiration rate was expressed on a dry soil mass basis (mg C kg soil^-1^ day^-1^).
+The final respiration rate was expressed on a dry soil mass basis (µg C g soil^-1^ day^-1^).
 
 The effects of temperature, gravimetric water content, C percent, N percent, and DOC concentration on instantaneous gas fluxes were evaluated using a linear mixed-effects model fit by the R function *lme* in the R 'nlme' package, version `r packageVersion("nlme")`.
-Because the dependent variable (CO~2~ or CH~4~ flux) was non-normally distributed, it was transformed using a natural-logarithm (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^ to ensure all positive fluxes) transformation.
+Because the dependent variable (CO~2~ or CH~4~ flux) was non-normally distributed, it was transformed using a natural-logarithm (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^ to ensure all positive fluxes, following Treat et al.
+2015) transformation.
 Core number was treated as a random effect in the model.
@@
 All data analysis and statistics were performed using R version `r R.version.string` {R Development Core Team, 2016 #4314}.
-This experiment was run as an "open experiment" with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
+This experiment was run as an "open experiment" {Bond-Lamberty, 2016 #4388} with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
 The summarized flux data backing the main results have been archived at [DOI TO BE FILLED IN UPON ACCEPTANCE].
@@
 
-Gas fluxes varied considerably over the course of the incubation (**Figures 2 and 3**).
+These means conceal considerable variablility over the course of the incubation (**Figures 2 and 3**).
 In the linear mixed-effects model, CO~2~ was strongly influenced by incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 1**).
 (Note that percent C and percent N were highly correlated (r ≥ 0.98) for these cores.
-Because percent N was a slightlyl stronger predictor, it was retained in the model while percent C was excluded.) The interaction between water content and soil N was also highly significant (P < 0.001), with high-N cores having little relationship between water content and CO~2~ flux (data not shown).
+Because percent N was a slightlyl stronger predictor, it was retained in the model while percent C was excluded.) The interaction between water content and soil N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), with high-N cores having little relationship between water content and CO~2~ flux (data not shown).
 Methane fluxes were most strongly correlated with percent N, while water content exhibited significant interactions with percent N and DOC (**Table 2**).
@@
 
-The cumulative flux numbers above result in CH~4~ temperature sensitivity (Q~10~) values of `r pn(Q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(Q10_cumulative["Drought", "CH4"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CO~2~ were `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)`.
-Computing Q~10~s based on fluxes normalized by water filled pore space decreased these values slightly: to `r pn(wfpsQ10_cumulative["Field moisture", "CH4"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CH4"], 1)` for CH~4~, respectively, and `r pn(wfpsQ10_cumulative["Field moisture", "CO2"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CO2"], 1)` for CO~2~.
+The cumulative flux numbers above result in CO~2~ temperature sensitivity (Q~10~) values of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CH~4~ were `r pn(Q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(Q10_cumulative["Drought", "CH4"], 1)`.
+Computing Q~10~s based on fluxes normalized by water filled pore space changed these values only slightly: to `r pn(wfpsQ10_cumulative["Field moisture", "CO2"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CO2"], 1)` for CO~2~, respectively, and `r pn(wfpsQ10_cumulative["Field moisture", "CH4"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CH4"], 1)` for CH~4~.
 
@@
 Morishita et al.
-{, 2014 #4316} measured gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both soil temperature and moisture, consistent with our results, TODO.
-Morishita et al.
-{, 2014 #4316} found however that CH~4~ uptake (they observed no emissions) was driven by temperature only; this is...
-
+{, 2014 #4316} measured gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture, consistent with our results They found however that CH~4~ uptake (no emissions were observed) was driven by temperature only.
 Waldrop et al.
-{, 2010 #4380} incubated  active-layer and permafrost soils from near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost); flux rates were 0.01-0.10 µmol CH~4~ day^-1^ g^-1^ () and 1-8 µg C-CO2 hr^-1^ g^-1^.
+{, 2010 #4380} incubated active-layer and permafrost soils from near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C.
+Flux rates were 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), differing by orders of magnitude between sites but roughly similar to our observed emissions, and ~1-5 µg C-CO2 hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed from our cores.
 
@@
 Incubation experiments can only go so far though...
+Such laboratory studies have weaknesses {Podrebarac, 2016 #4365}, but their controlled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
 

### 2016-04-15 R/0-functions.R

@@ -20,6 +20,8 @@ library(luzlogr)       # 0.2.0
 # -----------------------------------------------------------------------------
 # Parameters for key analytical choices
 # Defined here so they can be easily used in code AND manuscript
+MAX_MINCONC_TIME <- 10     # the minimum concentration has to occur in first 10 s
+MAX_MAXCONC_TIME <- 45     # the maximum concentration has to occur in first 45 s
 OUTLIER_GROUPS   <- 10     # Divide data into date groups and identify outliers in each
 CO2_EXCLUDE_DEVS <- 5.0    # CO2 outlier boundary, in mean absolute deviations
 CH4_EXCLUDE_DEVS <- 10.0   # CH4 outlier boundary, in mean absolute deviations

### 2016-04-15 R/2-summarize.R

@@ -106,9 +106,6 @@ matchfun <- function(DATETIME, MPVPosition) {
 }
 
 printlog( "Computing summary statistics for each sample..." )
-# The window in which we look for min and max concentrations
-MAX_MINCONC_TIME <- 10  # the minimum concentration has to occur in first 10 s
-MAX_MAXCONC_TIME <- 45  # the maximum concentration has to occur in first 45 s
 
 # We want to apply different criteria here, so three different pipelines
 # to compute the min and max gas concentrations

### 2016-04-15 manuscript/cpcrw_manuscript.Rmd
Incubation ms - discussion additions, some results

@@
 ---
-title: Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils
+title: Temperature and moisture effects on greenhouse gas emissions from deep active-layer
+  boreal soils
 author: "Ben Bond-Lamberty^1^, A. Peyton Smith^2^, Vanessa Bailey^2^"
@@
 
-Abstract.
+High latitude ecosystems are being subjected to rapid climatic changes, with rising air temperatures and fires expected to drive permafrost degradation and soil carbon (C) cycle changes.
+How these soils' greenhouse gas emissions will respond to changes in temperature, moisture, and vegetation is highly uncertain.
+We used a laboratory experiment to examine how temperature and moisture control CO~2~ and CH~4~ emissions from soils sampled from the bottom of the annual active layer in an Alaskan boreal forest.
+Gas emissions were measured from thirty cores, subjected to two temperatures and either field moisture conditions or drought, over a 100-day incubation.
+TODO...
 
@@
 The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and interact with soil temperature effects in long-term incubations {Elberling, 2013 #3720}.
-Decadal warming, and perhaps drying, trends in Alaska {Bieniek, 2014 #4389} will work to counteract these effects, with uncertain consequences for the carbon cycle and climate system.
+More uniformly across the landscape, however, decadal warming, and perhaps drying, trends in Alaska {Bieniek, 2014 #4389} will work to counteract these effects, with uncertain consequences for the carbon cycle and climate system.
 
@@
 Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
-We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at difference temperature and moisture levels.
+We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at different temperatures, subjecting some cores to drying treatments.
 We hypothesized that (i) CO~2~ would be the dominant pathway for C loss in these largely aerobic soils; (ii) soils maintained at field moisture and high temperature would lose more C-CO~2~ than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity; and (iii) core CH~4~ fluxes would be sensitive only to temperature, as no anaerobic conditions were planned for the cores.
@@
 Starting on 31 August 2015 we measured the cores' mass and GHG (CH~4~ and CO~2~) evolution four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation.
-Throughout the incubation, cores were mounted on 0.05-MPa ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action.
-This was done to moderate the drying of the 20 °C "controlled drought" group.
-The 4 and 20 °C "field moisture" groups received additional wetting from the top of the cores to maintain their water status.
+Throughout the incubation, cores were mounted on porous ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action.
+The "drought" cores were mounted on dry plates, but not allowed to drop below 5% water content.
+When necessary, cores received additional wetting from the top to maintain their water status at the desired level.
 
@@
 This experiment was run as an "open experiment" {Bond-Lamberty, 2016 #4388} with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
-The summarized flux data backing the main results have been archived at [DOI TO BE FILLED IN UPON ACCEPTANCE].
+The summarized flux data backing the main results have been archived at [DOI to be filled in on acceptance].
 
@@
 
-After removal of known problematic data, and exclusion of statistical outliers, CO~2~ flux rates measured during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 2)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 2)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 2)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 3)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 2)`).
+After removal of known problematic data, and exclusion of statistical outliers, CO~2~ flux rates measured during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 1)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 1)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 1)`).
 CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 2)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 2)`).
@@
 Because percent N was a slightlyl stronger predictor, it was retained in the model while percent C was excluded.) The interaction between water content and soil N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), with high-N cores having little relationship between water content and CO~2~ flux (data not shown).
+"Drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
+
 Methane fluxes were most strongly correlated with percent N, while water content exhibited significant interactions with percent N and DOC (**Table 2**).
@@
 
-Rises in boreal soil temperatures - who has quantified this? Changes in soil WC much less certain. Thermokarst etc., deepening of active layer - that's the larger implications of this work.
-
-Modeling results highly dependent on water content of soils after permafrost degradation.
-
-Eastern Alaska high latitude warming and forest stress {Ju, 2016 #4320}.
+Rises in boreal air temperatures, and unpredictable precipitation changes, will warm and dry many soils, increase vegetation stress {Ju, 2016 #4320}, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}.
+The resulting changes in the soil carbon cycle may already be visible {Bond-Lamberty, 2012 #3480} but are highly uncertain going forward, increasing the need for controlled laboratory studies of GHG emission dynamics from high-latitude soils.
 
@@
 Waldrop et al.
-{, 2010 #4380} incubated active-layer and permafrost soils from near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C.
-Flux rates were 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), differing by orders of magnitude between sites but roughly similar to our observed emissions, and ~1-5 µg C-CO2 hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed from our cores.
+{, 2010 #4380} incubated active-layer and permafrost soils from near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C; these values are higher than we observed, consistent with the lower temperature range {Hamdi, 2013 #3664}.
+Waldrop et al.
+{, 2010 #4380} observed flux rates of 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), differing by orders of magnitude between sites (but roughly similar to our observed CH~4~ emissions), and ~1-5 µg C-CO2 hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed from our cores.
+In an incubation of active-layer Alaskan permafrost peats, Treat et al.
+{, 2014 #4213} found CO~2~ and CH~4~ emissions to be strongly correlated with temperature and moisture.
+Finally, during the first 100 days of a year-long incubation of Fairbanks-area 0-10 cm mineral soils, Neff and Hooper {, 2002 #4390} observed fluxes of ~`r pn(0.2 * 1000 * 12/44, 0)`-`r pn(1.5 * 1000 * 12/44, 0)` µg C-CO~2~ g C^-1^ day^-1^, in line with the results here.
 
-More generally,
+More generally, in a pan-Arctic synthesis of anaerobic soil incubations, Treat et al.
+{, 2015 #4217} reported mean CO~2~ rates of 47 (all mineral soils) and 101 (for 20-100 cm soils) µg C-CO~2~ g C^-1^ day^-1^, somewhat lower than our aerobic incubation results.
+Weiss et al.
+{, 2015 #4317} found CO~2~ production from Yedoma permafrost samples to be correlated with both percent C and N, consistent with our results (**Table 1**).
+TODO.
 
-Treat et al. 2014 also showed that CH4 production from organic permafrost soils has a positive relationship with temperature (cites several studies: Lupasco et al. 2012, etc.)
-*	Treat et al.
-2015.
-Meta-analysis of anaerobic incubations.
-Mineral soils 20-100 cm depth had a mean CO2 production (ug CO2-C /g C/day) of 100.6 +/- 10.3 (33.2 +/- 3.7 per g DW)  (7840 per m2/d).
-All soils < 10C were 67 +/- 16 and > 20C was 172 +/- 17.
-Dry soils 68 +/- 8 and Fluctuating = 364 +/- 60 and inundated = 105 +/- 11.
+**Temperature versus moisture sensitivity**
 
-*	Xue et al.
-2016: active layer, contained organic horizon: soil moisture increased in warmed plots: ~50% moisture compared to ~45% (p < 0.05.
-C Flux (g/m2) from ecosystem respiration was greater in warmed plots (~500 g/m2) compared to control plots (~400) (p < 0.05).
-Authors remarked that soil moisture/climate was important factor in shifting functional composition of microbial community (GeoChip), esp.
-with anaerobic activity.
+Warming usually increases soil GHG fluxes, for example at depth in a long-term Arctic tundra experiment {Sistla, 2013 #3666}, as increased temperatures increment enzyme kinetics, enhance the production of extracellular enzymes, and enhance desorptions rates of organic matter from minerals.
+A key question, for both experimentalists and modelers {Falloon, 2011 #3045}, is to what degree such soils' emissions could by constrained by their moisture status--in particular, limited by low soil moisture driven by increases in high-latitude temperatures, vapor pressure deficit, and potentially precipitation changes.
 
-*	Schaufler et al.
-2010: 33.7 +/- 2.7C02-C (mg C/ m2/ hr) averaged across all temperature (5-20C) and moisture (20-80% WFPS) in forested, boreal site, with 33.7 +/- 2.7.
-CH4-C = -30.2 +/- 4.5.
-Optimum moisture content was 40% WFPS for this site – or highest CO2 evolved at this moisture content.
-Carbon dioxide emissions were linearly correlated with temperature (p = 0.000) and exponentially correlated with WFPS (p = 0.000).
-CH4 uptake (oxidation) was highest in the northern sites (included boreal site), and was also the driest soil moisture conditions.
--44.8 +/- 5.5.
-ug CH4-C/m2/hr and decreased with increasing soil moisture.
-Effect of WFPS on CH4 oxidation, exponential, p  = 0.000.
+Our results suggest that moisture limitation could exert a large effect on CO~2~ production for deep active-layer soils (**Figure 4**): "drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
+This effect was highly significant, and suggests that moisture limitations could exert a significant constraint on deep active-layer soils as they slowly warm.
+Such moisture constraints are thought to be already exerting effects on vegetation and soil fluxes at large scales {Ju, 2016 #4320}{Bond-Lamberty, 2012 #3480}, but our understanding of the interactive effects involved is poor.
 
-*	Wu et al.
-2010.
-Mongolian steppe 900-1400msl.
-Soils = chernozems, but there was a gleyosol.
-Soil moisture and temperature influenced CH4.
-Soil drought limited CH4 uptake at all 15C.
-CO2 in meadow sites – increased with temperature (< 30 mg C/m2/hr @ 5C to 90 at 15C) and with moisture (<30-50 in dry treatment from 5-15C, and 50-90 in high rain treatment at 5-15C).
+In contrast, the temperature sensitivities observed in this experiment are low (all less than 2.0, even when controlling for changes in soil moisture).
+Such values are not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
+It is important to note, however, that even if increased temperature alone may not always drive C mineralization rates in forest-mineral soils {Giardina, 2000 #2159}, but it is frequently linked with increases in soil moisture content can lead to changes in microbial community structure, amplifying GHG fluxes (Xue et al, 2016).
+TODO: PEYTON - WHAT IS THIS CITATION?
 
-Seeta’s paper: Mineral soil: greenhouse treatment (warming treatment) = 0.646 ± 0.04 GREATER THAN control = 0.43 ± 0.05 C-CO2 per g soil per day in 14 day incubation ( α ≤ 0.1).
-Mineral soil also had greater active biomass and extractable-OM).
-(Sistla et al.
-2013 Nature).
+**Methane**
 
-**Temperature versus moisture sensitivity, and CO~2~ versus CH~4~**
+We observed very low, but consistently positive, CH~4~ production from these unsaturated mineral soils.
+This is contrast to many field studies, e.g.
+Schaufler et al.
+{, 2010 #4391} who found consistent CH~4~ uptake (oxidation) in dry conditions at a boreal site.
 
-Why would we expect to see greater CO2 evolved in higher temps?
-
-*	Increased temp (to a certain degree) increases enzyme kinetics (i.e. Arrhenius equation) (Kirschbaun  1995, 2006 for SOM-temp, Wallenstein/Allison/Schimel for enzyme-temp, etc. so many others)
-*	Increased temperature enhances the metabolic rate of microorganisms – thus increasing the production of extracellular enzymes (state somewhere prior that extracellular enzymes catalyze fundamental biogeochemical reactions in SOC decomposition and nutrient cycling).
-Cusack et al.
-(2010) suggests temperature-sensitivity of decomposition may be indirectly related to temperature dependence of enzyme production AND turnover.
-*	Carbon use efficiency (CUE), a measure of how well a microorganism can assimilate C into biomass relative to how much C is lost as CO2,  has been shown to vary with temperature (CITE – Lopez-Moran 2007; Price and Sowers 2004 !) with decreases in CUE at higher temperatures leading to greater CO2 production during SOM decomposition (Steinweg et al.
-2008).
-Cite Balser and Wixon 2009 in regards to microbial-mediated effects on temp-sensitivity of C decomposition.
-*	Higher temperatures enhance desorption rates of OM with minerals (ten Hulscher and Cornelissen 1996; Kalbitz et al.
-2000 – others) making more OC available in solution (Xu & Saiers, 2010; Plante et al.
-2009), and thus accessible for adjacent, competent microorganisms
-
-Note high CO2:CH4 ratios, compare with Treat et al. and other studies.
-
-Increased temperature alone may not drive C mineralization rates in forest-mineral soils (Giardina and Ryan, 2000), but its corresponding increase in soil moisture content can lead to changes in microbial community structure and enhance anaerobic activity that may amplify CO2, CH4 and N2O fluxes (Xue et al, 2016).
+Treat et al. found a median CO~2~:CH~4~ production ratios of 387 for boreal sites, far lower than (but consistent with) our observed ratios of several million.
 
@@
 Incubation experiments can only go so far though...
-Such laboratory studies have weaknesses {Podrebarac, 2016 #4365}, but their controlled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
+
+Laboratory soil incubations have weaknesses, for example in what can be inferred about temperature sensitivity {Podrebarac, 2016 #4365}{Hamdi, 2013 #3664}, but their controlled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
 
@@
 
-Check out:
-Soil microbial respiration rate and temperature sensitivity along a north-south forest transect in eastern China: Patterns and influencing factors
-http://onlinelibrary.wiley.com/doi/10.1002/2015JG003217/abstract
+Necessity of integrated lab-field-modeling experiments...
 
-TODO: look at Podrebarac (2016) - Soils isolated during incubation underestimate temperature sensitivity of respiration and its response to climate history
+[//]: (comment comment comment)
 
@@
 
-APS comment: Did the controlled drought treatment match the 4C drought treatment in terms of rate of water loss (for both gravimetric AND volumetric)?
-Since the controlled drought treatment was meant to control for increased water loss with the higher temp, I want to make sure that it really was doing that.
-From this graph, it “looks” like it followed the same trend as the 20C drought treatment.
-TODO
-
 **Figure 2.** A faceted (Temperature ~ Treatment) plot showing normalized CO2 flux means and errors by incubation time.
@@
 ```
-
-APS comment: Was difference in cum CO2 between field moist cores and drought cores the same for both temperatures???
-Granted cum Co2 was greater overall in 20C, but interesting that the difference between treatments appear equal in both temperatures.
-This is a bit difference from Treat et al.
-2014 (granted they had permafrost/peat soils) – but Treat reported larger differences between wet/dry (aerobic/anaerobic) treatments at 20C but no differences at -0.5C, but explain this due to frozen water limiting microbial respiration.
-TODO

### 2016-04-15 R/5-analyze.R

@@ -258,6 +258,16 @@ figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Tre
   ylab(paste("Cumulative C (mg)"))
 
 # -----------------------------------------------------------------------------
+# Calculate drought reduction (%)
+
+fluxdata_cumulative %>% 
+  filter(Treatment %in% c("Drought", "Field moisture"), Gas == "CO2") %>% 
+  dcast(Temperature ~ Treatment, value.var = "cum_flux_mgC") %>% 
+  mutate(Reduction = (`Field moisture` - Drought) / `Field moisture` * 100) ->
+  drought_effect
+rownames(drought_effect) <- drought_effect$Temperature
+
+# -----------------------------------------------------------------------------
 # Q10 based on cumulative fluxes
 # TODO: I hate the hardcoded temperature values below!
 

### 2016-04-18 manuscript_openexperiment/Open_experiment.Rmd
Incorporate VLB feedback on open experiment draft

@@
 
-Researchers in soil and ecosystem science, and almost every other field, as increasingly being pushed--by funders, journals, governments, and themselves--to increase transparency and reproducibility of their work.
+Researchers in soil and ecosystem science, and almost every other field, are being pushed--by funders, journals, governments, and themselves--to increase transparency and reproducibility of their work.
 A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, enable powerful meta- and cross-disciplinary analyses, and increase trust in, and the efficiency of, publicly-funded research.
@@
 But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task--often until key details have been forgotten, if not forever.
+This then presents opportunities for real-time peer review of both code and data.
 
@@
 Data loss hits ecosystem, soil, and global change ecology particularly hard, as climate changes make ecological data effectively irreproducible {Wolkovich, 2012 #3303}: we can never remeasure exactly the same system state.
-This results in a critical role for syntheses and meta-analyses {Ellison, 2010 #4363}{Cassey, 2006 #4349}, which depend on the existence of documented protocols and (ideally raw) data.
+This results in a critical need for syntheses and meta-analyses {Ellison, 2010 #4363}{Cassey, 2006 #4349}, which depend on the existence of documented protocols and (ideally raw) data.
 A subtler data loss issue is the "file drawer problem" {Rosenthal, 1979 #2291}, where unpublished but potentially valuable data are lost forever.
@@
 These provide a much-needed improvement on "Supplementary Information" sections accompanying journal articles, which have become *de facto* repositories of data, but often have inconvenient formats (e.g.
-PDF), poorly searchability, use restrictions, and uncertain long-term availability {Molloy, 2011 #4366;Neylon, 2012 #4367}.
-Most journals do not have expertise in data storage and management {Cassey, 2006 #4349}.
+PDF), use restrictions, and uncertain long-term availability {Molloy, 2011 #4366;Neylon, 2012 #4367}.
+Most journals do not have the necessary and forward-thinking expertise in data storage and management {Cassey, 2006 #4349}.
 In contrast, the best repositories provide easy data uploading, immediate assignment of data digital object identifiers (DOIs), and long-term stability and availability.
@@
 
-Transparency is no magic panacea ensuring repeatability and reproducibility {Easterbrook, 2014 #4137}, and may raise genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
+By itself transparency does not guarantee repeatability and reproducibility {Easterbrook, 2014 #4137}, and it may raise genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
 Nor will open science by itself fundamentally enhance public trust in science, as there are still many social and political challenges to overcome {Grand, 2012 #4370}.

### 2016-04-22 manuscript/cpcrw_manuscript.Rmd
Abstract written; intro tweaked; typos fixed

@@
 
+```{r wc, echo=FALSE}
+# Compute soil moisture data for use in abstract and results below
+# wc and wc_alltrt are computed by 5-analyze.R script
+gwcbegin <- subset(gwc_alltrt, stage == "Beginning")
+gwcend <- subset(gwc, stage == "End")
+gwcendmean <- gwcend$mean
+names(gwcendmean) <- gwcend$Treatment
+gwcendsd <- gwcend$sd
+names(gwcendsd) <- gwcend$Treatment
+vwcbegin <- subset(vwc_alltrt, stage == "Beginning")
+vwcend <- subset(vwc_alltrt, stage == "End")
+```
+
+```{r compute_ratios, echo=FALSE}
+# Compute ratios between CO2 and CH4 for use in abstract and results below
+# Ratio data is stored in the `gas_ratio` d.f. computed by analysis script
+minmag <- min(gas_ratio$CO2CH4_ratio) %>% log10 %>% floor %>%
+  c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine")[.]
+minratio <- min(gas_ratio$CO2CH4_ratio / 1e6)  %>% formatC(digits = 1, format = 'f')
+mintrt <- gas_ratio$Treatment[which.min(gas_ratio$CO2CH4_ratio)]
+mintemp <- gas_ratio$Temperature[which.min(gas_ratio$CO2CH4_ratio)]
+maxratio <- max(gas_ratio$CO2CH4_ratio / 1e6) %>% formatC(digits = 1, format = 'f')
+maxtrt <- gas_ratio$Treatment[which.max(gas_ratio$CO2CH4_ratio)]
+maxtemp <- gas_ratio$Temperature[which.max(gas_ratio$CO2CH4_ratio)]
+```
+
 #### Abstract
 
-High latitude ecosystems are being subjected to rapid climatic changes, with rising air temperatures and fires expected to drive permafrost degradation and soil carbon (C) cycle changes.
-How these soils' greenhouse gas emissions will respond to changes in temperature, moisture, and vegetation is highly uncertain.
-We used a laboratory experiment to examine how temperature and moisture control CO~2~ and CH~4~ emissions from soils sampled from the bottom of the annual active layer in an Alaskan boreal forest.
-Gas emissions were measured from thirty cores, subjected to two temperatures and either field moisture conditions or drought, over a 100-day incubation.
-TODO...
+Rapid climatic changes, rising air temperatures, and increased fires are expected to drive permafrost degradation and soil carbon (C) cycle changes in many high-latitude ecosystems.
+How these soils will respond to changes in their temperature, moisture, and overlying vegetation is highly uncertain, but critical to understand given the large soil C stocks in these regions.
+We used a laboratory experiment to examine how temperature and moisture control CO~2~ and CH~4~ emissions from mineral soils sampled from the bottom of the annual active layer, i.e.
+directly above permafrost, in an Alaskan boreal forest.
+Gas emissions from thirty cores, subjected to two temperatures and either field moisture conditions or experimental drought, were tracked over a 100-day incubation; we also measured a variety of physical and chemical characteristics of the cores.
+Gravimetric water content was `r pn(gwcbegin$mean, 2)` ± `r pn(gwcbegin$sd, 2)` (unitless) at the beginning of the incubation; cores kept at field moisture were unchanged at the end, but drought cores had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
+Carbon dioxide fluxes were strongly influenced by incubation chamber temperature, core water content, and percent soil N.
+Methane fluxes were most strongly correlated with percent N, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes, however.
+The cumulative production of C from CO~2~ was over `r minmag` order of magnitudes higher than that from CH~4~.
+The CO~2~ temperature sensitivity (i.e.
+Q~10~) was `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively.
+Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results help us understand the potential future dynamics of such soils with future climate and permafrost changes.
 
@@
 
-High latitude ecosystems are being subjected to rapid changes in climate {IPCC, 2013 #3916} and fire occurrence and intensity {Kasischke, 2010 #2631}.
-This has a wide variety of effects on ecosystems and society, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}{Ju, 2016 #4320}.
-In particular, rising temperatures and increasing fire will result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372} and changes in soil temperature and hydrology that are expected to increase soil greenhouse gas (GHG) fluxes to the atmosphere, promoting a positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
+High latitude ecosystems are being subjected to rapid changes in climate {IPCC, 2013 #3916} and fire occurrence and intensity {Kasischke, 2010 #2631}, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}{Ju, 2016 #4320}.
+This will have a wide variety of effects on both ecosystems and society: in particular, rising temperatures and increasing fire will likely result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372} and changes in soil temperature, with subsequent hydrology changes that will influence soil greenhouse gas (GHG) fluxes to the atmosphere.
+Such fluxes are a large component of the global C cycle and could thus promote a positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
 
-The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such a feedback remain highly uncertain {Schuur, 2015 #4319}.
+The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such any such feedback remain highly uncertain {Schuur, 2015 #4319}.
 While northern soils hold enormous quantities {Tarnocai, 2009 #2231} of potentially mineralizable soil organic carbon (SOC), vegetation succession dynamics promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}.
@@
 Temperature and moisture typically have strong and often interactive influences on soil GHG emissions.
-A long literature of laboratory incubations, field observations, and meta-analyses has documented increased CO2 fluxes with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
+A long literature of laboratory incubations, field observations, and meta-analyses has documented increased CO~2~ fluxes with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
 GHG responses to wetting and thawing dynamics are much less certain, with substantial variability between studies {Kim, 2012 #2863}.
-The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and interact with soil temperature effects in long-term incubations {Elberling, 2013 #3720}.
+The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}; such interactions are critical to examine in the course of long-term incubation experiments {Elberling, 2013 #3720}.
 More uniformly across the landscape, however, decadal warming, and perhaps drying, trends in Alaska {Bieniek, 2014 #4389} will work to counteract these effects, with uncertain consequences for the carbon cycle and climate system.
 
-The goal of this study was to examine how temperature and moisture control C loss and GHG (CO~2~ and CH~4~) emissions from soils sampled from the bottom of the annual active layer--i.e., directly above permafrost--in an Alaskan boreal forest.
-Most previous studies have focused on surface soils or permafrost soils, neglecting deep active-layer soils that experience fluctuating temperature and, especially, moisture conditions.
-Such deep but unfrozen soils were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
+The goal of this study was to examine how temperature and moisture control C loss and GHG (CO~2~ and CH~4~) emissions from soils sampled from the bottom of the annual active layer–i.e., directly above permafrost–in an Alaskan boreal forest.
+Most previous studies have focused on surface soils or permafrost soils, neglecting deep active-layer soils that were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
 We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at different temperatures, subjecting some cores to drying treatments.
@@
 At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (N=6 in each group).
-These groups included two incubation temperatures (4 and 20 °C), each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
+These included two incubation temperatures (4 and 20 °C), each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
 The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status would close match those of the 4 °C "drought" cores, which we anticipated would dry more slowly than their 20 °C counterparts.
@@
 Cumulative respiration for each core and gas was calculated by linearly interpolating flux rates between measurement dates and summing respired C over the entire incubation.
-The effect of temperature and treatment (drought, controlled drought, or field moisture conditions) on cumulative gas fluxes was evaluted with a post-hoc Tukey Honest Significant Differences test.
+The effect of temperature and treatment (drought, controlled drought, or field moisture conditions) on cumulative gas fluxes was evaluated with a post-hoc Tukey Honest Significant Differences test.
 Temperature sensitivity (Q~10~) was calculated for each gas and treatment as $\frac{F_2}{F_1}^{\left( \frac{10}{T_2-T_1} \right)}$ where *F~1~* and *F~2~* are the cumulative gas fluxes (mg C) at temperatures *T~1~* and *T~2~* (°C), respectively.
@@
 
-```{r wc, echo=FALSE}
-# wc and wc_alltrt are computed by 5-analyze.R script
-gwcbegin <- subset(gwc_alltrt, stage == "Beginning")
-gwcend <- subset(gwc, stage == "End")
-gwcendmean <- gwcend$mean
-names(gwcendmean) <- gwcend$Treatment
-gwcendsd <- gwcend$sd
-names(gwcendsd) <- gwcend$Treatment
-vwcbegin <- subset(vwc_alltrt, stage == "Beginning")
-vwcend <- subset(vwc_alltrt, stage == "End")
-```
-
 Gravimetric water content was `r pn(gwcbegin$mean, 2)` ± `r pn(gwcbegin$sd, 2)` (min `r pn(gwcbegin$min, 2)`, max `r pn(gwcbegin$max, 2)`) at the beginning of the incubation (**Figure 1**).
@@
 
-After removal of known problematic data, and exclusion of statistical outliers, CO~2~ flux rates measured during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 1)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 1)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 1)`).
+After removal of known problematic data, and exclusion of statistical outliers, CO~2~ flux rates measured during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 1)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 1)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 1)`) over the 100 days.
 CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 2)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 2)`).
@@
 (Note that percent C and percent N were highly correlated (r ≥ 0.98) for these cores.
-Because percent N was a slightlyl stronger predictor, it was retained in the model while percent C was excluded.) The interaction between water content and soil N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), with high-N cores having little relationship between water content and CO~2~ flux (data not shown).
-"Drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
+Because percent N was a slightly stronger predictor, it was retained in the model while percent C was excluded.) The interaction between water content and percent N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), with high-N cores having little relationship between water content and CO~2~ flux (data not shown).
 
@@
 
-```{r compute_ratios, echo=FALSE}
-# Ratio data is stored in the `gas_ratio` d.f. computed by analysis script
-minmag <- min(gas_ratio$CO2CH4_ratio) %>% log10 %>% floor %>%
-  c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine")[.]
-minratio <- min(gas_ratio$CO2CH4_ratio / 1e6)  %>% formatC(digits = 1, format = 'f')
-mintrt <- gas_ratio$Treatment[which.min(gas_ratio$CO2CH4_ratio)]
-mintemp <- gas_ratio$Temperature[which.min(gas_ratio$CO2CH4_ratio)]
-maxratio <- max(gas_ratio$CO2CH4_ratio / 1e6) %>% formatC(digits = 1, format = 'f')
-maxtrt <- gas_ratio$Treatment[which.max(gas_ratio$CO2CH4_ratio)]
-maxtemp <- gas_ratio$Temperature[which.max(gas_ratio$CO2CH4_ratio)]
-```
-
 The cumulative production of C from CO~2~ was over `r minmag` order of magnitudes higher than that from CH~4~, with CO~2~:CH~4~ C ratios ranging from `r minratio` million in the `r mintemp` °C "`r maxtrt`" treatment, to `r maxratio` million in the `r maxtemp` °C "`r maxtrt`" treatment.
@@
 There was no difference between the 20 °C "drought" and "controlled drought" treatments (`r pclean(co2_hsd$Treatment["Drought-Controlled drought", "p adj"])`).
+"Drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
 Neither temperature (`r pclean(ch4_hsd$Temperature[,"p adj"])`) nor treatment (mean `r pclean(mean(ch4_hsd$Treatment[,"p adj"]))`) was a significant factor in predicting cumulative CH~4~ fluxes.
@@
 Warming usually increases soil GHG fluxes, for example at depth in a long-term Arctic tundra experiment {Sistla, 2013 #3666}, as increased temperatures increment enzyme kinetics, enhance the production of extracellular enzymes, and enhance desorptions rates of organic matter from minerals.
-A key question, for both experimentalists and modelers {Falloon, 2011 #3045}, is to what degree such soils' emissions could by constrained by their moisture status--in particular, limited by low soil moisture driven by increases in high-latitude temperatures, vapor pressure deficit, and potentially precipitation changes.
+A key question, for both experimentalists and modelers {Falloon, 2011 #3045}, is to what degree such soils' emissions could by constrained by their moisture status–in particular, limited by low soil moisture driven by increases in high-latitude temperatures, vapor pressure deficit, and potentially precipitation changes.
 

### 2016-04-22 manuscript/cpcrw_manuscript.Rmd
New draft- discussion fleshed out, figure legends

@@
 ^2^Biological Sciences Division, Pacific Northwest National Laboratory, Richland, WA USA
+
+[//]: This is the Rmarkdown file
 
@@
 This will have a wide variety of effects on both ecosystems and society: in particular, rising temperatures and increasing fire will likely result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372} and changes in soil temperature, with subsequent hydrology changes that will influence soil greenhouse gas (GHG) fluxes to the atmosphere.
-Such fluxes are a large component of the global C cycle and could thus promote a positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
+Such fluxes are a large component of the global C cycle and, because of the high C stocks of northern soils {Tarnocai, 2009 #2231}, could result in a significant and positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
 
@@
 Temperature and moisture typically have strong and often interactive influences on soil GHG emissions.
-A long literature of laboratory incubations, field observations, and meta-analyses has documented increased CO~2~ fluxes with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
+Laboratory incubations, field observations, and meta-analyses have documented increased CO~2~, and under some conditions CH~4~ {Olefeldt, 2013 #3456}, fluxes with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
 GHG responses to wetting and thawing dynamics are much less certain, with substantial variability between studies {Kim, 2012 #2863}.
-The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}; such interactions are critical to examine in the course of long-term incubation experiments {Elberling, 2013 #3720}.
+The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and such interactions are critical to examine in the course of long-term incubation experiments {Elberling, 2013 #3720}.
 More uniformly across the landscape, however, decadal warming, and perhaps drying, trends in Alaska {Bieniek, 2014 #4389} will work to counteract these effects, with uncertain consequences for the carbon cycle and climate system.
@@
 
-**Importance and context**
-
-Rises in boreal air temperatures, and unpredictable precipitation changes, will warm and dry many soils, increase vegetation stress {Ju, 2016 #4320}, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}.
-The resulting changes in the soil carbon cycle may already be visible {Bond-Lamberty, 2012 #3480} but are highly uncertain going forward, increasing the need for controlled laboratory studies of GHG emission dynamics from high-latitude soils.
-
-**Previous results**
-
 Several studies have measured microbial respiration and GHG fluxes from soils very close to our study site.
@@
 {, 2015 #4317} found CO~2~ production from Yedoma permafrost samples to be correlated with both percent C and N, consistent with our results (**Table 1**).
-TODO.
+The response of soil biota to stresses such as drought tend to differ between soil types, but be broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}.
 
-**Temperature versus moisture sensitivity**
+*Temperature versus moisture sensitivity*
 
@@
 
-In contrast, the temperature sensitivities observed in this experiment are low (all less than 2.0, even when controlling for changes in soil moisture).
+In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture).
 Such values are not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
-It is important to note, however, that even if increased temperature alone may not always drive C mineralization rates in forest-mineral soils {Giardina, 2000 #2159}, but it is frequently linked with increases in soil moisture content can lead to changes in microbial community structure, amplifying GHG fluxes (Xue et al, 2016).
-TODO: PEYTON - WHAT IS THIS CITATION?
+It is important to note, however, that even if increased temperature alone may not always drive C mineralization rates in forest-mineral soils {Giardina, 2000 #2159}, but it is frequently linked with increases in soil moisture content can lead to changes in microbial community structure and GHG fluxes {Xue, 2016 #4361}.
 
-**Methane**
+*Methane*
 
@@
 {, 2010 #4391} who found consistent CH~4~ uptake (oxidation) in dry conditions at a boreal site.
+Our results are however broadly consistent with data from 65 studies summarized by Olefeldt et al.
+{Olefeldt, 2013 #3456}, who found that CH~4~ emissions were more sensitive to soil temperature in wetter ecosystems; it would have been a surprise if the little methanogenic activity in our upland, well-drained soils was temperature-sensitive at all.
 
-Treat et al. found a median CO~2~:CH~4~ production ratios of 387 for boreal sites, far lower than (but consistent with) our observed ratios of several million.
+Methane was thus a far smaller C flux than CO~2~ from these soils.
+This is true more generally: for example, Treat et al.
+found a median CO~2~:CH~4~ production ratios of 387 for boreal sites, far lower than (but consistent with) our observed ratios of several million.
+MORE.
 
-**Limitations and weaknesses**
+*Limitations and weaknesses*
 
-Incubation experiments can only go so far though...
+There were weaknesses in our approach and experimental design that should be considered.
+Laboratory experiments offer precise control, but lack the *in situ* nature of field manipulations {Sistla, 2013 #3666}, raising uncertainties to what degree their results can be extrapolated.
+They also have more specific weaknesses, for example in what can be inferred about temperature sensitivity {Podrebarac, 2016 #4365}{Hamdi, 2013 #3664}.
+Nonetheless, the controlled environments of incubations provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
 
-Laboratory soil incubations have weaknesses, for example in what can be inferred about temperature sensitivity {Podrebarac, 2016 #4365}{Hamdi, 2013 #3664}, but their controlled environments provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
+We focused on a experimental drought, rather than flooding, because of the well-drained nature of the field site: it is unlikely that the mid-slope forest we sampled in will ever suffer from thermokarst or excessive soil moisture, but too-dry conditions are a serious possibility in this low-precipitation ecosystem {Barber, 2000 #1320}.
+In addition, the soils here are not surface layer soils (where the majority of microbial activity and C mineralization of labile C takes place); taking them out of depth (where they are less exposed to O~2~, for example) may significantly change the abiotic conditions to which the microbial community is adapted.
+MORE?
 
-these are not surface layer soils (where majority of microbial activity and C mineralization of labile C takes place) and that taking them out of depth (where they are not exposed to O2 as readily, etc.) for our laboratory incubation may differ from in situ CO2/CH4 production.
+*Conclusions*
 
-Necessity of integrated lab-field-modeling experiments...
-
-[//]: (comment comment comment)
+Rises in boreal air temperatures, and unpredictable precipitation changes, will warm and dry many soils, increase vegetation stress {Ju, 2016 #4320}{Barber, 2000 #1320}, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}.
+The resulting changes in the soil carbon cycle may already be visible {Bond-Lamberty, 2012 #3480} but are highly uncertain going forward, increasing the need for controlled laboratory studies, ideally integrated with field and modeling experiments, of GHG emission dynamics from high-latitude soils.
 
@@
 
-**Table 1.** Linear mixed-effects model parameters, testing effects of gravimetric water content (unitless), temperature (°C), soil C (%), soil N (%),  and dissolved organic carbon (mg kg^-1^) on individual core CO~2~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^).
+**Table 1.** Linear mixed-effects model parameters, testing effects of temperature (°C), gravimetric water content (unitless), soil C (%), soil N (%),  and dissolved organic carbon (mg kg^-1^) on individual core CO~2~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^); a colon (":") indicates an interaction.
 Dependent variable has units of log(µg C g C^-1^ day^-1^).
@@
 
-**Table 2.** Linear mixed-effects model parameters, testing effects of gravimetric water content (unitless), temperature (°C), soil C (%), soil N (%),  and dissolved organic carbon (mg kg^-1^) on log-transformed, individual core CH~4~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^).
+**Table 2.** Linear mixed-effects model parameters, testing effects of temperature (°C), gravimetric water content (unitless), soil N (%),and dissolved organic carbon (DOC, mg kg^-1^) on log-transformed, individual core CH~4~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^); a colon (":") indicates an interaction.
 Dependent variable has units of log(µg C g C^-1^ day^-1^).
@@
 
-**Figure 2.** A faceted (Temperature ~ Treatment) plot showing normalized CO2 flux means and errors by incubation time.
+**Figure 2.** Mass-normalized CO~2~ fluxes over the 100-day incubation, by temperature (4 and 20 °C, rows) and treatment (field moisture, drought, and controlled drought; columns).
+Error bars show core-to-core standard deviation.
 
@@
 
-**Figure 3.** A faceted (Temperature ~ Treatment) plot showing normalized CH4 flux means and errors by incubation time.
+**Figure 3.** Mass-normalized CH~4~ fluxes over the 100-day incubation, by temperature (4 and 20 °C, rows) and treatment (field moisture, drought, and controlled drought; columns).
+Error bars show core-to-core standard deviation.
 
@@
 
-**Figure 4.** Cumulative fluxes over the incubation.
+**Figure 4.** Cumulative C fluxes (mg) over the incubation, by gas (CO~2~ and CH~4~, rows) and treatment (columns). TODO: these need to be normalized.
 

### 2016-04-25 manuscript/cpcrw_manuscript.Rmd
Fixed "<0.001" problem in table p-values
Thanks @apeyton

@@
 ---
-title: Temperature and moisture effects on greenhouse gas emissions from deep active-layer
-  boreal soils
+title: Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils
 author: "Ben Bond-Lamberty^1^, A. Peyton Smith^2^, Vanessa Bailey^2^"
@@
 
-pclean <- function(x) {  # pretty formatting of p-values
-  ifelse(x < 0.001, "< 0.001", paste("=", formatC(x, digits = 3, format = 'f'))) %>%
-    paste("P", .)
+pclean <- function(x, digits = 3, printP = TRUE) {  # pretty formatting of p-values
+  ltstring <- paste0("< 0.", substr("000000000", 1, digits - 1), "1")
+  valstring <- ifelse(x < 10^-digits, ltstring, formatC(x, digits = digits, format = 'f'))
+  if(printP) {
+    ifelse(x < 0.001, valstring, paste("=", valstring)) %>% paste("P", .)
+  } else {
+    valstring
+  }
 }
@@
 ```{r table1, echo=FALSE}
-knitr::kable(summary(step_co2_lme)$tTable, digits = 3, col.names = c("Value", "SE", "DF", "T", "P"))
+t1 <- summary(step_co2_lme)$tTable %>% round(3) %>% as.data.frame()
+t1$`p-value` <- pclean(t1$`p-value`, printP = FALSE)
+knitr::kable(t1, digits = 3, col.names = c("Value", "SE", "DF", "T", "P"))
 ```
@@
 ```{r table2, echo=FALSE}
-knitr::kable(summary(step_ch4_lme)$tTable, digits = 3, col.names = c("Value", "SE", "DF", "T", "P"))
+t2 <- summary(step_co2_lme)$tTable %>% round(3) %>% as.data.frame()
+t2$`p-value` <- pclean(t2$`p-value`, printP = FALSE)
+knitr::kable(t2, digits = 3, col.names = c("Value", "SE", "DF", "T", "P"))
 ```

### 2016-04-25 manuscript/cpcrw_manuscript.Rmd
Fixed a couple mistakes in last commit

@@
 pclean <- function(x, digits = 3, printP = TRUE) {  # pretty formatting of p-values
-  ltstring <- paste0("< 0.", substr("000000000", 1, digits - 1), "1")
+  ltstring <- paste0("< 0.", paste(rep("0", digits - 1), collapse = ""), "1")
   valstring <- ifelse(x < 10^-digits, ltstring, formatC(x, digits = digits, format = 'f'))
   if(printP) {
-    ifelse(x < 0.001, valstring, paste("=", valstring)) %>% paste("P", .)
+    ifelse(x < 10^-digits, valstring, paste("=", valstring)) %>% paste("P", .)
   } else {

### 2016-04-25 manuscript/cpcrw_manuscript.Rmd
Fixed figure 3 warning, and more tweaks to pclean

--- 2016-04-25 manuscript/cpcrw_manuscript.Rmd
@@
 
-pclean <- function(x, digits = 3, printP = TRUE) {  # pretty formatting of p-values
-  ltstring <- paste0("< 0.", substr("000000000", 1, digits - 1), "1")
-  valstring <- ifelse(x < 10^-digits, ltstring, formatC(x, digits = digits, format = 'f'))
+# 'Pretty n' function to round a numeric value and print that # of digits
+pn <- function(x, n) formatC(round(x, n), digits = n, format = 'f')
+
+# 'Clean p value' function to pretty-print p value(s), specifically
+pclean <- function(x, digits = 3, printP = TRUE) {
+  ltstring <- paste0("< 0.", paste(rep("0", digits - 1), collapse = ""), "1")
+  valstring <- ifelse(x < 10^-digits, ltstring, pn(x, digits))
   if(printP) {
-    ifelse(x < 0.001, valstring, paste("=", valstring)) %>% paste("P", .)
+    ifelse(x < 10^-digits, valstring, paste("=", valstring)) %>% paste("P", .)
   } else {
@@
 }
-
-# 'Pretty n' function to round a numeric value and print that # of digits
-pn <- function(x, n) formatC(round(x, n), digits = n, format = 'f')
 

### 2016-04-25 R/5-analyze.R

@@ -203,7 +203,8 @@ fluxdata %>%
   fluxdata_figsBC
 
 figsBC <- function(fd) {
-    ggplot(fd, aes(incday, flux)) + 
+  fd <- filter(fd, !is.na(incday), !is.na(flux_sd), !is.na(flux))
+  ggplot(fd, aes(incday, flux)) + 
     geom_point() + 
     facet_grid(Temperature ~ Treatment) + 
     geom_errorbar(aes(ymin = flux - flux_sd, ymax = flux + flux_sd)) +

### 2016-04-25 R/5-analyze.R
Fixed colors to be colorblind-friendly - thanks @apeyton

@@ -128,10 +128,11 @@ save_data(fluxsummary)
 printlog(SEPARATOR)
 printlog("Water content over time...")
 
-figureA <- ggplot(fluxdata_orig, aes(inctime_days, WC_gravimetric, color=Treatment, group=Core)) 
-figureA <- figureA + geom_point() + geom_line()
-figureA <- figureA + facet_grid(~Temperature) 
-figureA <- figureA + xlab("Incubation day") + ylab("Gravimetric water content (fraction dry mass)")
+figureA <- ggplot(fluxdata_orig, aes(inctime_days, WC_gravimetric, color=Treatment, group=Core)) +
+  geom_point() + geom_line() + 
+  facet_grid(~Temperature) + 
+  xlab("Incubation day") + ylab("Gravimetric water content (fraction dry mass)") +
+  scale_color_brewer(palette = "Set1")
 
 fluxdata$incday <- floor(fluxdata$inctime_days)
 fluxdata$stage <- NA
@@ -256,7 +257,8 @@ figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Tre
                     ymax = cum_flux_mgC + cum_flux_mgC_sd), 
                 position = position_dodge(0.9), width = 0.4) +  
   facet_grid(Gas ~ ., scales = "free") +
-  ylab(paste("Cumulative C (mg)"))
+  ylab(paste("Cumulative C (mg)")) +
+  scale_fill_brewer(palette = "Set1") + scale_color_brewer(palette = "Set1")
 
 # -----------------------------------------------------------------------------
 # Calculate drought reduction (%)

### 2016-04-26 manuscript/cpcrw_manuscript.Rmd
Changed cumulative flux calculation and tests to per-gC basis
Doesn’t really change much, but absolutely necessary to do.

--- 2016-04-25 manuscript/cpcrw_manuscript.Rmd
@@
 
-pclean <- function(x, digits = 3, printP = TRUE) {  # pretty formatting of p-values
-  ltstring <- paste0("< 0.", substr("000000000", 1, digits - 1), "1")
-  valstring <- ifelse(x < 10^-digits, ltstring, formatC(x, digits = digits, format = 'f'))
+# 'Pretty n' function to round a numeric value and print that # of digits
+pn <- function(x, n) formatC(round(unlist(x), n), digits = n, format = 'f')
+
+# 'Clean p value' function to pretty-print p value(s), specifically
+pclean <- function(x, digits = 3, printP = TRUE) {
+  x <- as.vector(x)
+  ltstring <- paste0("< 0.", paste(rep("0", digits - 1), collapse = ""), "1")
+  valstring <- ifelse(x < 10^-digits, ltstring, pn(x, digits))
   if(printP) {
-    ifelse(x < 0.001, valstring, paste("=", valstring)) %>% paste("P", .)
+    ifelse(x < 10^-digits, valstring, paste("=", valstring)) %>% paste("P", .)
   } else {

### 2016-04-26 R/4-fluxes.R

@@ -167,26 +167,25 @@ fluxdata %>%
   group_by(Core) %>%
   mutate(delta_hrs = (inctime_days - lag(inctime_days)) * 24,
          CO2_flux_mgC = CO2_flux_mgC_hr_interp * delta_hrs,
-         cumCO2_flux_mgC = c(0, cumsum(CO2_flux_mgC[-1])),
+         cumCO2_flux_mgC_gC = c(0, cumsum(CO2_flux_mgC[-1])) / SoilDryMass_g,
          CH4_flux_mgC = CH4_flux_mgC_hr_interp * delta_hrs,
-         cumCH4_flux_mgC = c(0, cumsum(CH4_flux_mgC[-1]))) %>%
+         cumCH4_flux_mgC_gC = c(0, cumsum(CH4_flux_mgC[-1]))  / SoilDryMass_g) %>%
   select(-CO2_flux_mgC_hr_interp, -CH4_flux_mgC_hr_interp,
          -CO2_flux_mgC_hr_clean, -CH4_flux_mgC_hr_clean) ->
   fluxdata
-
 
 # -----------------------------------------------------------------------------
 # A few diagnostic plots
 
 printlog("Cumulative flux diagnostic plots...")
-p1 <- ggplot(fluxdata, aes(inctime_days, cumCO2_flux_mgC, group = Core)) + 
+p1 <- ggplot(fluxdata, aes(inctime_days, cumCO2_flux_mgC_gC, group = Core)) + 
   geom_line() + 
   facet_grid(Temperature ~ Treatment) +
   ggtitle("Cumulative CO2 by core and treatment")
 print(p1)
 save_plot("cumulative_CO2")
 
-p2 <- ggplot(fluxdata, aes(inctime_days, cumCH4_flux_mgC, group = Core)) + 
+p2 <- ggplot(fluxdata, aes(inctime_days, cumCH4_flux_mgC_gC, group = Core)) + 
   geom_line() + 
   facet_grid(Temperature ~ Treatment) + 
   ggtitle("Cumulative CH4 by core and treatment")
@@ -199,20 +198,20 @@ save_plot("cumulative_CH4")
 printlog("Calculating flux summary datasets...")
 fluxdata %>%
   melt(id.vars = c("Treatment", "Temperature", "Core", "inctime_days", "WFPS_percent"), 
-       measure.vars = c("cumCO2_flux_mgC", "cumCH4_flux_mgC"),
[…]

### 2016-04-26 R/5-analyze.R

@@ -228,11 +228,11 @@ printlog(SEPARATOR)
 printlog("Running Tukey HSD tests on cumulative emissions...")
 fd_cumulative_core <- read_csv(FLUXDATA_CUM_CORE_FILE) %>%
   mutate(Temperature = as.factor(Temperature))
-co2_aov <- aov(cum_flux_mgC ~ Treatment + Temperature, 
+co2_aov <- aov(cum_flux_mgC_gC ~ Treatment + Temperature, 
                data = subset(fd_cumulative_core, Gas == "CO2"))
 co2_hsd <- TukeyHSD(co2_aov)
 print(co2_hsd)
-ch4_aov <- aov(cum_flux_mgC ~ Treatment + Temperature, 
+ch4_aov <- aov(cum_flux_mgC_gC ~ Treatment + Temperature, 
                data = subset(fd_cumulative_core, Gas == "CH4"))
 ch4_hsd <- TukeyHSD(ch4_aov)
 print(ch4_hsd)
@@ -250,22 +250,24 @@ fluxdata_cumulative$Treatment <- factor(fluxdata_cumulative$Treatment,
                                         levels = c("Field moisture", "Controlled drought", "Drought"))
 fluxdata_cumulative$Temperature <- as.factor(fluxdata_cumulative$Temperature)
 
-figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC, fill = Treatment)) + 
+figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC_gC, fill = Treatment)) + 
   geom_bar(stat = 'identity', position = position_dodge()) +
   geom_errorbar(aes(color = Treatment, 
-                    ymin = cum_flux_mgC * 0.9, 
-                    ymax = cum_flux_mgC + cum_flux_mgC_sd), 
+                    ymin = cum_flux_mgC_gC * 0.9, 
+                    ymax = cum_flux_mgC_gC + cum_flux_mgC_gC_sd), 
                 position = position_dodge(0.9), width = 0.4) +  
   facet_grid(Gas ~ ., scales = "free") +
-  ylab(paste("Cumulative C (mg)")) +
+  ylab(expression(Cumulative~C~(mg~g~C^{-1}))) +
   scale_fill_brewer(palette = "Set1") + scale_color_brewer(palette = "Set1")
+
+#ylab(expression(µg~C~g~C^{-1}~day^{-1}))
 
[…]

### 2016-04-27 manuscript/cpcrw_manuscript.Rmd
Incorporated VLB and APS comments
Text clarified in a number of spots; discussion beefed up (though a
number of areas need to be beefed up still); figure 4 now has Tukey’s
HSD labels.

@@
 ---
-^1^Joint Global Change Research Institute, DOE Pacific Northwest National Laboratory, College Park, MD USA; 
-bondlamberty@pnnl.gov
+^1^Joint Global Change Research Institute, DOE Pacific Northwest National Laboratory, College Park, MD USA
 
 ^2^Biological Sciences Division, Pacific Northwest National Laboratory, Richland, WA USA
+
+*Correspondence to*: Ben Bond-Lamberty (bondlamberty@pnnl.gov)
 
@@
 
-Rapid climatic changes, rising air temperatures, and increased fires are expected to drive permafrost degradation and soil carbon (C) cycle changes in many high-latitude ecosystems.
+Rapid climatic changes, rising air temperatures, and increased fires are expected to drive permafrost degradation and alter soil carbon (C) cycling in many high-latitude ecosystems.
 How these soils will respond to changes in their temperature, moisture, and overlying vegetation is highly uncertain, but critical to understand given the large soil C stocks in these regions.
@@
 Gas emissions from thirty cores, subjected to two temperatures and either field moisture conditions or experimental drought, were tracked over a 100-day incubation; we also measured a variety of physical and chemical characteristics of the cores.
-Gravimetric water content was `r pn(gwcbegin$mean, 2)` ± `r pn(gwcbegin$sd, 2)` (unitless) at the beginning of the incubation; cores kept at field moisture were unchanged at the end, but drought cores had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
-Carbon dioxide fluxes were strongly influenced by incubation chamber temperature, core water content, and percent soil N.
-Methane fluxes were most strongly correlated with percent N, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes, however.
-The cumulative production of C from CO~2~ was over `r minmag` order of magnitudes higher than that from CH~4~.
+Gravimetric water content was `r pn(gwcbegin$mean, 2)` ± `r pn(gwcbegin$sd, 2)` (unitless) at the beginning of the incubation; cores at field moisture were unchanged at the end, but drought cores had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
+Carbon dioxide fluxes were strongly influenced by incubation chamber temperature, core water content, and percent soil nitrogen.
+Methane fluxes were most strongly correlated with percent nitrogen, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
+The cumulative production of C from CO~2~ was over `r minmag` orders of magnitudes higher than that from CH~4~.
 The CO~2~ temperature sensitivity (i.e.
 Q~10~) was `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively.
-Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results help us understand the potential future dynamics of such soils with future climate and permafrost changes.
+These results suggest that deep active-layer soils may be much more sensitive to changes in moisture than to temperature, a critical factor as discontinuous permafrost melts in interior Alaska.
+Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results help us understand the future dynamics, and feedback potential, of such soils.
 
-#### Introduction
+#### 1 Introduction
 
-High latitude ecosystems are being subjected to rapid changes in climate {IPCC, 2013 #3916} and fire occurrence and intensity {Kasischke, 2010 #2631}, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}{Ju, 2016 #4320}.
-This will have a wide variety of effects on both ecosystems and society: in particular, rising temperatures and increasing fire will likely result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372} and changes in soil temperature, with subsequent hydrology changes that will influence soil greenhouse gas (GHG) fluxes to the atmosphere.
-Such fluxes are a large component of the global C cycle and, because of the high C stocks of northern soils {Tarnocai, 2009 #2231}, could result in a significant and positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}.
+High latitude ecosystems are being subjected to rapid changes in climate {IPCC, 2013 #3916} and increases in fire frequency and intensity {Kasischke, 2010 #2631}, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}{Ju, 2016 #4320}.
+This will have a wide variety of ecosystem effects {Alexander, 2016 #4414}: in particular, rising temperatures and increasing fire will likely result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372}{Genet, 2013 #4412}{Helbig, 2016 #4413} and changes in soil temperature, with subsequent hydrology changes that will influence soil greenhouse gas (GHG) fluxes to the atmosphere.
+Such fluxes are a large component of the global C cycle and, because of the high C stocks of northern soils {Tarnocai, 2009 #2231}, could result in a significant and positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}{Schaefer, 2014 #4117}.
 
 The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such any such feedback remain highly uncertain {Schuur, 2015 #4319}.
-While northern soils hold enormous quantities {Tarnocai, 2009 #2231} of potentially mineralizable soil organic carbon (SOC), vegetation succession dynamics promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}.
-Such dynamics may however be disrupted by increased fire disturbance {Johnstone, 2010 #2465}.
+While northern soils hold enormous quantities {Tarnocai, 2009 #2231} of potentially mineralizable soil organic carbon (SOC), vegetation and succession dynamics (for example, thermal insulation by mosses) promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}{Turetsky, 2012 #3405}.
+Such dynamics may however be disrupted by increased fire disturbance, particularly with more-intense fires {Johnstone, 2010 #2465}{Genet, 2013 #4412}.
 In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
@@
 Temperature and moisture typically have strong and often interactive influences on soil GHG emissions.
-Laboratory incubations, field observations, and meta-analyses have documented increased CO~2~, and under some conditions CH~4~ {Olefeldt, 2013 #3456}, fluxes with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
-GHG responses to wetting and thawing dynamics are much less certain, with substantial variability between studies {Kim, 2012 #2863}.
+Laboratory incubations, field observations, and meta-analyses have documented increased fluxes of CO~2~, and under some conditions of CH~4~ {Olefeldt, 2013 #3456}, with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
+Greenhouse gas (GHG) responses to wetting and thawing dynamics are much less certain, with substantial variability between studies {Kim, 2012 #2863}.
 The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and such interactions are critical to examine in the course of long-term incubation experiments {Elberling, 2013 #3720}.
-More uniformly across the landscape, however, decadal warming, and perhaps drying, trends in Alaska {Bieniek, 2014 #4389} will work to counteract these effects, with uncertain consequences for the carbon cycle and climate system.
+More uniformly across the landscape, however, decadal warming, and perhaps drying, trends in Alaska {Bieniek, 2014 #4389} may counteract these effects.
+A critical question, then, is how the structure, chemistry, and microbial communities of current active-layer soils will respond to almost-certain temperature and moisture changes in the future {Xue, 2016 #4361}.
 
-The goal of this study was to examine how temperature and moisture control C loss and GHG (CO~2~ and CH~4~) emissions from soils sampled from the bottom of the annual active layer–i.e., directly above permafrost–in an Alaskan boreal forest.
+The goal of this study was to examine how temperature and moisture control GHG (CO~2~ and CH~4~) emissions from soils sampled from the bottom of the annual active layer–i.e., directly above permafrost–in an Alaskan boreal forest.
 Most previous studies have focused on surface soils or permafrost soils, neglecting deep active-layer soils that were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
-We also aimed to characterize the chemical and biological properties of these soils as they were subjected to a 100-day incubation at different temperatures, subjecting some cores to drying treatments.
-We hypothesized that (i) CO~2~ would be the dominant pathway for C loss in these largely aerobic soils; (ii) soils maintained at field moisture and high temperature would lose more C-CO~2~ than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity; and (iii) core CH~4~ fluxes would be sensitive only to temperature, as no anaerobic conditions were planned for the cores.
+We also aimed to characterize the chemical and structural properties of these soils following a 100-day incubation at different temperatures, subjecting some cores to drying treatments.
+We hypothesized that (i) CO~2~ would be the dominant pathway for C loss in these largely aerobic soils; (ii) soils maintained at field moisture and high temperature would lose more C-CO~2~ than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity; and (iii) core CH~4~ fluxes would be sensitive only to temperature, as no anaerobic conditions were imposed on the cores.
 
-#### Methods
+#### 2 Methods
 
-**Field sampling**
+**2.1 Field sampling**
 
@@
 Annual average air temperature was -2.5 °C, and annual average precipitation 400 mm {Petrone, 2006 #4313}.
-The watershed's lowlands and north-facing slopes are dominated by black spruce (*Picea mariana*), feathermoss (*Pleurozium schreberi* and others), and *Sphagnum* spp.; the drier south slopes tend to be deciduous with a mixture of quaking aspen (*Populus tremuloides*), paper birch (*Betula neoalaskana*), and patches of alder (*Alnus crispa*).
+The watershed's lowlands and north-facing slopes are dominated by black spruce (*Picea mariana* (Mill.) BSP), feathermoss (*Pleurozium schreberi* and others), and *Sphagnum* spp.; the drier south slopes tend to be deciduous with a mixture of trembling aspen (*Populus tremuloides* Michx.), paper birch (*Betula neoalaskana*), and patches of alder (*Alnus crispa*).
 
@@
 
-**Laboratory incubation**
+**2.2 Laboratory incubation**
 
@@
 At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (N=6 in each group).
-These included two incubation temperatures (4 and 20 °C), each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
+These included two incubation temperatures of 4 and 20 °C, following the protocol of a number of previous boreal incubation studies {cf.
+Table 1 in \Treat, 2015 #4217}.
+Within each temperature there were two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
 The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status would close match those of the 4 °C "drought" cores, which we anticipated would dry more slowly than their 20 °C counterparts.
 The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density.
-Subsamples were collected and stored at -20 °C for dissolved organic carbon measurements.
+Subsamples were collected and stored at -20 °C for dissolved organic carbon measurements or air-dried for soil C and N (see below).
 
-On 18 August 2015 cores were placed into one of two growth chambers (Conviron CMP6050 Control Systems, Winnipeg, Canada) maintained at 4 and 20 °C temperatures and 70% relative humidity and allowed to equilibrate for two weeks.
-Starting on 31 August 2015 we measured the cores' mass and GHG (CH~4~ and CO~2~) evolution four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation.
-Throughout the incubation, cores were mounted on porous ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action.
+On 18 August 2015 cores were placed into one of two growth chambers (Conviron Control Systems BDW80, Winnipeg, Canada) maintained at 4 and 20 °C temperatures and 70% relative humidity and allowed to equilibrate for two weeks.
+Starting on 31 August 2015 we measured the cores' mass and GHG (CH~4~ and CO~2~) emissions four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation.
+Throughout the incubation, cores with a 200 µm mesh screen fit to the base were mounted on porous ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action.
 The "drought" cores were mounted on dry plates, but not allowed to drop below 5% water content.
@@
 The dry mass and volume of soil were used in calculations of gravimetric and volumetric soil moisture content, respectively {Gardner, 1986 #4345}.
-Soil volume was caclulated as the total core volume minus the volume of particles >2 mm diameter, with the latter determined by water displacement.
+Soil volume was calculated as the total core volume minus the volume of particles >2 mm diameter, with the latter determined by water displacement.
 Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic C.
@@
 
-**Data and statistical analysis**
+**2.3 Data and statistical analysis**
 
@@
 
-Anomalous data were excluded based on their gas fluxes being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within a 10-day period, for a given treatment and temperature.
-We excluded `r mad_outlier_count` of `r pre_outlier_count` (`r pn(mad_outlier_count/pre_outlier_count*100, 1)`%) measurements for this reason.
-If the coefficient of variability of fluxes from any core on a single day exceeded `r pn(MAX_CO2_CV*100, 0)`%, the entire core was also excluded for that day (`r cv_outlier_count` data points, `r pn(cv_outlier_count/pre_outlier_count*100, 1)`%).
-Other data (`r pn(rn/sdn*100, 1)`%) were removed because of known instrument problems, e.g.
-the analyzer was left running after leaving a chamber.
-The final number of valid flux samples from the 100-day incubation was `r sum(!fluxdata$outlier)`.
-
 Each core’s respiration flux (*F*) was calculated following e.g.
@@
 
-The effects of temperature, gravimetric water content, C percent, N percent, and DOC concentration on instantaneous gas fluxes were evaluated using a linear mixed-effects model fit by the R function *lme* in the R 'nlme' package, version `r packageVersion("nlme")`.
+Anomalous data were excluded based on their gas fluxes being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within a 10-day period, for a given treatment and temperature.
+We excluded `r mad_outlier_count` of `r pre_outlier_count` (`r pn(mad_outlier_count/pre_outlier_count*100, 1)`%) measurements for this reason.
+If the coefficient of variability (CV) of fluxes from any core on a single day exceeded `r pn(MAX_CO2_CV*100, 0)`%, a value chosen based on the distribution of CVs across all cores, the entire core was excluded for that day (`r cv_outlier_count` data points, `r pn(cv_outlier_count/pre_outlier_count*100, 1)`%).
+Other data (`r pn(rn/sdn*100, 1)`%) were removed because of known instrument problems, e.g.
+the analyzer was left running after leaving a chamber.
+The final number of valid flux samples from the 100-day incubation was `r sum(!fluxdata$outlier)`.
+
+The effects of temperature, gravimetric water content, percent C, percent N, and DOC concentration on instantaneous gas fluxes were evaluated using a linear mixed-effects model fit by the R function *lme* in the R 'nlme' package, version `r packageVersion("nlme")`.
 Because the dependent variable (CO~2~ or CH~4~ flux) was non-normally distributed, it was transformed using a natural-logarithm (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^ to ensure all positive fluxes, following Treat et al.
 2015) transformation.
-Core number was treated as a random effect in the model.
+Soil core was treated as a random effect in the model.
 We then used performed stepwise model selection by Akaike's information criterion (AIC) using the *stepAIC* function in the R 'MASS' package, version `r packageVersion("MASS")`.
@@
 
-#### Results
+#### 3 Results
 
 ```{r coredata, echo=FALSE}
+# Summarize a few bulk density metrics
 bd <- with(coredata, SoilDryMass_g / SoilVolume_cm3)
@@
 Carbon content was `r pn(cndata_summary["C_percent_mean"], 2)`% ± `r pn(cndata_summary["C_percent_sd"], 2)`%, while N content was `r pn(cndata_summary["N_percent_mean"], 2)`% ± `r pn(cndata_summary["N_percent_sd"], 2)`%.
-Neither temperature was treatment exerted any significant effect on these properties (P > 0.1 for all).
+Neither temperature nor moisture treatment exerted any significant effect on these highly variable properties (P > 0.1 for all).
 
@@
 "Field moisture" cores were on average unchanged (`r pn(gwcendmean["Field moisture"], 2)` ± `r pn(gwcendsd["Field moisture"], 2)`) at the end of the incubation, but both the drought treatments, which did not differ from each other in their effect on gravimetric water content (P = `r pn(p_trt_wc, 3)`), had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
-Volumetric water content values ranged from `r pn(vwcbegin$mean, 2)` ± `r pn(vwcbegin$sd, 2)` (min `r pn(vwcbegin$min, 2)`, max `r pn(vwcbegin$max, 2)`) at the beginning of the experiment to `r pn(vwcend$mean, 2)` ± `r pn(vwcend$sd, 2)` (min `r pn(vwcend$min, 2)`, max `r pn(vwcend$max, 2)`) at the end.
+Volumetric water content values ranged from `r pn(vwcbegin$mean, 2)` ± `r pn(vwcbegin$sd, 2)` (min `r pn(vwcbegin$min, 2)`, max `r pn(vwcbegin$max, 2)`) at the beginning of the experiment to `r pn(vwcend$mean, 2)` ± `r pn(vwcend$sd, 2)` (min `r pn(vwcend$min, 2)`, max `r pn(vwcend$max, 2)`) at the end across all cores.
 Water filled pore space, assuming a particle density of `r PARTICLE_DENSITY` g cm^-3^, was `r pn(min(fluxdata$WFPS_percent), 0)`-`r pn(max(fluxdata$WFPS_percent), 0)`%.
 
-After removal of known problematic data, and exclusion of statistical outliers, CO~2~ flux rates measured during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 1)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 1)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 1)`) over the 100 days.
+Carbon dioxide fluxes during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 1)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 1)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 1)`) over the 100 days.
 CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 2)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 2)`).
 
-These means conceal considerable variablility over the course of the incubation (**Figures 2 and 3**).
+These means conceal considerable variability over the course of the incubation (**Figures 2 and 3**).
 In the linear mixed-effects model, CO~2~ was strongly influenced by incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 1**).
@@
 
-The cumulative production of C from CO~2~ was over `r minmag` order of magnitudes higher than that from CH~4~, with CO~2~:CH~4~ C ratios ranging from `r minratio` million in the `r mintemp` °C "`r maxtrt`" treatment, to `r maxratio` million in the `r maxtemp` °C "`r maxtrt`" treatment.
-Cumulative CO~2~ evolved was highly affected by temperature (`r pclean(co2_hsd$Temperature["p adj"])`), and "field moisture" cores emitted significantly more CO~2~ than the other two treatments at both temperatures (P < 0.001 for both).
-There was no difference between the 20 °C "drought" and "controlled drought" treatments (`r pclean(co2_hsd$Treatment["Drought-Controlled drought", "p adj"])`).
+The cumulative production of C from CO~2~ (**Figure 4**) was over `r minmag` order of magnitudes higher than that from CH~4~, with CO~2~:CH~4~ C ratios ranging from `r minratio` million in the `r mintemp` °C "`r maxtrt`" treatment, to `r maxratio` million in the `r maxtemp` °C "`r maxtrt`" treatment.
+Cumulative CO~2~ evolved was highly affected by temperature (`r pclean(co2_hsd$Temperature["p adj"])`), and "field moisture" cores emitted significantly more CO~2~ than the other two moisture treatments at both temperatures (P < 0.001 for both, with no significant interactive effect).
+There was no difference between fluxes from the 20 °C "drought" and "controlled drought" treatments (`r pclean(co2_hsd$Treatment["Drought-Controlled drought", "p adj"])`).
 "Drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
-Neither temperature (`r pclean(ch4_hsd$Temperature[,"p adj"])`) nor treatment (mean `r pclean(mean(ch4_hsd$Treatment[,"p adj"]))`) was a significant factor in predicting cumulative CH~4~ fluxes.
+Neither temperature (`r pclean(ch4_hsd$Temperature[,"p adj"])`) nor moisture treatment (mean `r pclean(mean(ch4_hsd$Treatment[,"p adj"]))`) was a significant factor in predicting cumulative CH~4~ fluxes.
 
 The cumulative flux numbers above result in CO~2~ temperature sensitivity (Q~10~) values of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CH~4~ were `r pn(Q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(Q10_cumulative["Drought", "CH4"], 1)`.
-Computing Q~10~s based on fluxes normalized by water filled pore space changed these values only slightly: to `r pn(wfpsQ10_cumulative["Field moisture", "CO2"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CO2"], 1)` for CO~2~, respectively, and `r pn(wfpsQ10_cumulative["Field moisture", "CH4"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CH4"], 1)` for CH~4~.
+Computing Q~10~ values based on fluxes normalized by water filled pore space changed these values only slightly: to `r pn(wfpsQ10_cumulative["Field moisture", "CO2"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CO2"], 1)` for CO~2~, respectively, and `r pn(wfpsQ10_cumulative["Field moisture", "CH4"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CH4"], 1)` for CH~4~.
 
-#### Discussion
+#### 4 Discussion
 
+We found that CO~2~, but not CH~4~, fluxes from these oxic active-layer mineral soils were highly sensitive to temperature and drought effects.
 Several studies have measured microbial respiration and GHG fluxes from soils very close to our study site.
@@
 Waldrop et al.
-{, 2010 #4380} incubated active-layer and permafrost soils from near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C; these values are higher than we observed, consistent with the lower temperature range {Hamdi, 2013 #3664}.
+{, 2010 #4380} incubated active-layer and permafrost soils from *Picea mariana* sites near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C; these values are higher than we observed, consistent with the lower temperature range {Hamdi, 2013 #3664}.
 Waldrop et al.
@@
 Weiss et al.
-{, 2015 #4317} found CO~2~ production from Yedoma permafrost samples to be correlated with both percent C and N, consistent with our results (**Table 1**).
-The response of soil biota to stresses such as drought tend to differ between soil types, but be broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}.
+{, 2015 #4317} found CO~2~ production from Siberian Yedoma permafrost samples to be correlated with both percent C and N, consistent with our results (**Table 1**); The response of soil biota to stresses such as drought tend to differ between soil types and organisms, but be broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}, making such comparisons between e.g.
+Alaska and Siberia incubations useful, in spite of their significant ecological and climatic differences.
 
-*Temperature versus moisture sensitivity*
+*4.1 Temperature versus moisture sensitivity*
 
-Warming usually increases soil GHG fluxes, for example at depth in a long-term Arctic tundra experiment {Sistla, 2013 #3666}, as increased temperatures increment enzyme kinetics, enhance the production of extracellular enzymes, and enhance desorptions rates of organic matter from minerals.
-A key question, for both experimentalists and modelers {Falloon, 2011 #3045}, is to what degree such soils' emissions could by constrained by their moisture status–in particular, limited by low soil moisture driven by increases in high-latitude temperatures, vapor pressure deficit, and potentially precipitation changes.
+Warming usually increases soil GHG fluxes, for example at depth in a long-term Arctic tundra experiment {Sistla, 2013 #3666}, as increased temperatures enhance the production of extracellular enzymes, increase enzyme activities, and enhance desorption rates of organic matter from minerals.
+A key question, for both experimentalists and modelers {Falloon, 2011 #3045}, is to what degree such soils' emissions could by constrained by their moisture status, that is itself driven by increases in high-latitude temperatures, vapor pressure deficit, and potentially precipitation changes.
 
@@
 
-In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture).
-Such values are not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
-It is important to note, however, that even if increased temperature alone may not always drive C mineralization rates in forest-mineral soils {Giardina, 2000 #2159}, but it is frequently linked with increases in soil moisture content can lead to changes in microbial community structure and GHG fluxes {Xue, 2016 #4361}.
-
-*Methane*
+In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture), but not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
+It is important to note, however, that while increased temperature does not always drive C mineralization rates in forest mineral soils {Giardina, 2000 #2159}, it is linked with increases in soil moisture content can lead to changes in microbial community structure and GHG fluxes {Xue, 2016 #4361}.
+More surprisingly, at least initially, was that Q~10~ values were lower in the drought treatment cores, although one would expect a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 than at 20.
+TODO: THOUGHTS ON THIS FROM A MICROBIAL PERSPECTIVE?
 
@@
 Our results are however broadly consistent with data from 65 studies summarized by Olefeldt et al.
-{Olefeldt, 2013 #3456}, who found that CH~4~ emissions were more sensitive to soil temperature in wetter ecosystems; it would have been a surprise if the little methanogenic activity in our upland, well-drained soils was temperature-sensitive at all.
+{, 2013 #3456}, who found that CH~4~ emissions were more sensitive to soil temperature in wetter ecosystems; it would have been a surprise if the little methanogenic activity in our upland, well-drained soils was temperature-sensitive at all.
+Thus we see little opportunity for CH~4~ to be a significant contributor, now or in the future, to such soils' C fluxes and climate feedback risk.
 
-Methane was thus a far smaller C flux than CO~2~ from these soils.
+Methane was thus a far smaller C flux than CO~2~ from these soils, in particular at higher temperatures (as CO~2~ was responsive to temperature, but CH~4~ was not).
 This is true more generally: for example, Treat et al.
-found a median CO~2~:CH~4~ production ratios of 387 for boreal sites, far lower than (but consistent with) our observed ratios of several million.
-MORE.
+found a median CO~2~:CH~4~ production ratio of 387 for boreal sites, far lower than (but consistent with) our observed ratios of several million, and global flux estimates of CH~4~ are orders of magnitude lower than those of CO~2~ {Kirschke, 2013 #3796}.
 
-*Limitations and weaknesses*
+*4.2 Soil nitrogen*
+
+*4.3 Limitations and weaknesses*
 
@@
 
-*Conclusions*
+*5 Conclusions*
 
-Rises in boreal air temperatures, and unpredictable precipitation changes, will warm and dry many soils, increase vegetation stress {Ju, 2016 #4320}{Barber, 2000 #1320}, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}.
-The resulting changes in the soil carbon cycle may already be visible {Bond-Lamberty, 2012 #3480} but are highly uncertain going forward, increasing the need for controlled laboratory studies, ideally integrated with field and modeling experiments, of GHG emission dynamics from high-latitude soils.
+Rises in boreal air temperatures, and unpredictable precipitation changes, will warm and dry many soils, increase vegetation stress {Ju, 2016 #4320}{Barber, 2000 #1320}, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}, with uncertain consequences for soil dynamics and GHG fluxes.
+In this laboratory experiment, we found that CO~2~ fluxes were strongly influenced by temperature, water content, and soil nitrogen, while CH~4~ fluxes were much smaller and not sensitive to temperature or water content in these well-drained mineral soils.
+This suggests that understanding how soil moisture might change with permafrost degradation, how soil biota will respond to these changes, and how models should treat soil organic matter decomposition with respect to multiple and interacting drivers are all critical areas of research going forward.
+Further controlled field and laboratory studies, ideally tightly integrated with modeling experiments, are critical to understand GHG emission dynamics from high-latitude soils.
 
@@
 Error bars show core-to-core standard deviation.
+The "controlled drought" treatment, for 20 °C only, was meant to dry cores at roughly the same rate as the drought cores at 4 °C.
 
@@
 
-**Figure 4.** Cumulative C fluxes (mg) over the incubation, by gas (CO~2~ and CH~4~, rows) and treatment (columns). TODO: these need to be normalized.
+**Figure 4.** Cumulative C fluxes (mg g C^-1^) over the incubation, by gas (CH~4~ and CO~2~, top and bottom panels respectively) and treatment (columns).
+Letters within a panel indicate significant differences based on Tukey's HSD.
 

### 2016-04-27 R/5-analyze.R

@@ -260,7 +260,25 @@ figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC_gC, fill = 
   ylab(expression(Cumulative~C~(mg~g~C^{-1}))) +
   scale_fill_brewer(palette = "Set1") + scale_color_brewer(palette = "Set1")
 
-#ylab(expression(µg~C~g~C^{-1}~day^{-1}))
+# Make the labels for this figure. Kind of a PITA
+library(agricolae)  # v1.2.3
+fluxdata_cumulative %>% 
+  group_by(Gas, Treatment, Temperature) %>% 
+  summarise(cum_flux_mgC_gC = max(cum_flux_mgC_gC + cum_flux_mgC_gC_sd) * 1.1) %>%
+  mutate(trt = paste(Gas, Treatment, Temperature, sep = ":")) %>%
+  group_by(Gas) %>%
+  mutate(cum_flux_mgC_gC = max(cum_flux_mgC_gC, na.rm = TRUE) * 1.1) ->
+  labeldata
+out_co2 <- HSD.test(co2_aov, trt = c("Treatment", "Temperature"))
+out_co2$groups$trt <- paste("CO2", trimws(out_co2$groups$trt), sep = ":")
+out_ch4 <- HSD.test(ch4_aov, trt = c("Treatment", "Temperature"))
+out_ch4$groups$trt <- paste("CH4", trimws(out_ch4$groups$trt), sep = ":")
+rbind(out_co2$groups, out_ch4$groups) %>%
+  left_join(labeldata, by = "trt") ->
+  labeldata
+figureD <- figureD + geom_text(data = labeldata, 
+                               aes(label = M), 
+                               position = position_dodge(width = 1))
 
 # -----------------------------------------------------------------------------
 # Calculate drought reduction (%)

### 2016-05-13 manuscript/cpcrw_manuscript.Rmd
Many minor tweaks to intro, methods, results

@@
 Gravimetric water content was `r pn(gwcbegin$mean, 2)` ± `r pn(gwcbegin$sd, 2)` (unitless) at the beginning of the incubation; cores at field moisture were unchanged at the end, but drought cores had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
-Carbon dioxide fluxes were strongly influenced by incubation chamber temperature, core water content, and percent soil nitrogen.
+Carbon dioxide fluxes were strongly influenced by incubation chamber temperature, core water content, and percent soil nitrogen, and had a temperature sensitivity (i.e.
+Q~10~) of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively..
 Methane fluxes were most strongly correlated with percent nitrogen, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
 The cumulative production of C from CO~2~ was over `r minmag` orders of magnitudes higher than that from CH~4~.
-The CO~2~ temperature sensitivity (i.e.
-Q~10~) was `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively.
 These results suggest that deep active-layer soils may be much more sensitive to changes in moisture than to temperature, a critical factor as discontinuous permafrost melts in interior Alaska.
-Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results help us understand the future dynamics, and feedback potential, of such soils.
+Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results provide insight into their future dynamics and feedback potential with future climate change.
 
@@
 The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and such interactions are critical to examine in the course of long-term incubation experiments {Elberling, 2013 #3720}.
-More uniformly across the landscape, however, decadal warming, and perhaps drying, trends in Alaska {Bieniek, 2014 #4389} may counteract these effects.
+Decadal warming and drying trends in Alaska {Bieniek, 2014 #4389} may counteract these effects, however.
 A critical question, then, is how the structure, chemistry, and microbial communities of current active-layer soils will respond to almost-certain temperature and moisture changes in the future {Xue, 2016 #4361}.
@@
 CPCRW is located in the Yukon-Tanana Uplands northeast of Fairbanks, AK, a part of the boreal forest that has seen strong increases in air temperature and forest browning {Ju, 2016 #4320} over several decades.
-Annual average air temperature was -2.5 °C, and annual average precipitation 400 mm {Petrone, 2006 #4313}.
+Annual average air temperature is -2.5 °C, and annual average precipitation 400 mm {Petrone, 2006 #4313}.
 The watershed's lowlands and north-facing slopes are dominated by black spruce (*Picea mariana* (Mill.) BSP), feathermoss (*Pleurozium schreberi* and others), and *Sphagnum* spp.; the drier south slopes tend to be deciduous with a mixture of trembling aspen (*Populus tremuloides* Michx.), paper birch (*Betula neoalaskana*), and patches of alder (*Alnus crispa*).
@@
 Thirty-nine soil cores, each 30 cm high by 7.5 cm wide, were taken using a soil recovery augur (AMS Inc., American Falls, ID) on 3-5 August 2015.
-We sampled from the bottom (within 0-2 cm) of the active layer, which averaged 80 cm depth.
+We sampled from the bottom (within 0-2 cm of permafrost) of the active layer, which averaged 80 cm depth.
 Sample points were randomly located in the transition zone described above, and separated by 2-5 m.
-Cores were kept cool in the field before being packed in dry ice and shipped to the lab in Richland, WA within 48-72 hours of collection.
+Cores were kept cool in the field before being packed in dry ice and shipped to Richland, WA within 48-72 hours of collection.
 
@@
 Starting on 31 August 2015 we measured the cores' mass and GHG (CH~4~ and CO~2~) emissions four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation.
-Throughout the incubation, cores with a 200 µm mesh screen fit to the base were mounted on porous ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action.
+Throughout the incubation, cores had a 200 µm mesh screen fit to the base and were mounted on porous ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action.
 The "drought" cores were mounted on dry plates, but not allowed to drop below 5% water content.
@@
 
-At each GHG measurement, we measured the rise in gas concentrations, considering this as the flux rate in ppm s^-1^ (CO~2~) or ppb s^-1^ (CH~4~).
-The GHG concentration measurements for each sample (i.e., each gas, core, and date/time) were used to calculate a linear rate of change (δc/δt), based on the concentration rise from a minimum (up to `r MAX_MINCONC_TIME` seconds after measurement began) to a maximum (at `r MAX_MINCONC_TIME`-`r MAX_MAXCONC_TIME` seconds).
-This value was calculated for each sample, i.e.
-each individual core measurement throughout the 100-day incubation.
-
 ```{r compute_excludes, echo=FALSE}
@@
 
-Each core’s respiration flux (*F*) was calculated following e.g.
-Steduto et al.
-{, 2002 #3383} as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where *V* is the core-specific system volume, *M* the core dry mass as determined at the end of the incubation, *P~a~* atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), *R* the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and *T* the chamber air temperature (K) at time of measurement.
+For each measurement of each sample throughout the 100-day incubation (i.e., each gas, core, and date/time), we used the rise in gas concentrations to calculate a flux rate in ppm s^-1^ (CO~2~) or ppb s^-1^ (CH~4~), a linear rate of change (δc/δt) based on the concentration rise from a minimum (up to `r MAX_MINCONC_TIME` seconds after measurement began) to a maximum (at `r MAX_MINCONC_TIME`-`r MAX_MAXCONC_TIME` seconds).
+Each core’s respiration flux (*F*) was then calculated as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where *V* is the core-specific system volume, *M* the core dry mass as determined at the end of the incubation, *P~a~* atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), *R* the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and *T* the chamber air temperature (K) at time of measurement.
 The final respiration rate was expressed on a dry soil mass basis (µg C g soil^-1^ day^-1^).
@@
 The effect of temperature and treatment (drought, controlled drought, or field moisture conditions) on cumulative gas fluxes was evaluated with a post-hoc Tukey Honest Significant Differences test.
-Temperature sensitivity (Q~10~) was calculated for each gas and treatment as $\frac{F_2}{F_1}^{\left( \frac{10}{T_2-T_1} \right)}$ where *F~1~* and *F~2~* are the cumulative gas fluxes (mg C) at temperatures *T~1~* and *T~2~* (°C), respectively.
+Temperature sensitivity (Q~10~) was calculated for each gas and treatment as $\frac{F_2}{F_1}^{\left( \frac{10}{T_2-T_1} \right)}$ where *F~1~* and *F~2~* are the cumulative gas fluxes (mg C g C^-1^) at temperatures *T~1~* and *T~2~* (°C), respectively.
 
@@
 This experiment was run as an "open experiment" {Bond-Lamberty, 2016 #4388} with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
-The summarized flux data backing the main results have been archived at [DOI to be filled in on acceptance].
+The summarized flux data backing the main results have been archived at [*DOI to be filled in on acceptance*].
 
@@
 Volumetric water content values ranged from `r pn(vwcbegin$mean, 2)` ± `r pn(vwcbegin$sd, 2)` (min `r pn(vwcbegin$min, 2)`, max `r pn(vwcbegin$max, 2)`) at the beginning of the experiment to `r pn(vwcend$mean, 2)` ± `r pn(vwcend$sd, 2)` (min `r pn(vwcend$min, 2)`, max `r pn(vwcend$max, 2)`) at the end across all cores.
-Water filled pore space, assuming a particle density of `r PARTICLE_DENSITY` g cm^-3^, was `r pn(min(fluxdata$WFPS_percent), 0)`-`r pn(max(fluxdata$WFPS_percent), 0)`%.
+Water filled pore space, assuming a particle density of `r PARTICLE_DENSITY` g cm^-3^, was `r pn(min(fluxdata$WFPS_percent), 0)`-`r pn(max(fluxdata$WFPS_percent), 0)`% over all cores, moisture treatments, and temperatures.
 
@@
 Methane fluxes were most strongly correlated with percent N, while water content exhibited significant interactions with percent N and DOC (**Table 2**).
-Neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes, however.
+Neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
 
@@
 Waldrop et al.
-{, 2010 #4380} incubated active-layer and permafrost soils from *Picea mariana* sites near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C; these values are higher than we observed, consistent with the lower temperature range {Hamdi, 2013 #3664}.
+{, 2010 #4380} incubated active-layer and permafrost soils from *Picea mariana* sites near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C; these values are higher than we observed, consistent with the lower temperature range {Hamdi, 2013 #3664} and fundamental biokinetics {Davidson, 2006 #1669}.
 Waldrop et al.
-{, 2010 #4380} observed flux rates of 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), differing by orders of magnitude between sites (but roughly similar to our observed CH~4~ emissions), and ~1-5 µg C-CO2 hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed from our cores.
+{, 2010 #4380} also observed flux rates of 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), differing by orders of magnitude between sites (but roughly similar to our observed CH~4~ emissions), and ~1-5 µg C-CO2 hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed from our cores.
 In an incubation of active-layer Alaskan permafrost peats, Treat et al.

### 2016-05-17 manuscript_openexperiment/Open_experiment.Rmd
Close-to-final edits - add a couple refs, etc.

@@
 
-Researchers in soil and ecosystem science, and almost every other field, are being pushed--by funders, journals, governments, and themselves--to increase transparency and reproducibility of their work.
+Researchers in soil and ecosystem science, and almost every other field, are being pushed--by funders, journals, governments, and their peers--to increase transparency and reproducibility of their work.
 A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, enable powerful meta- and cross-disciplinary analyses, and increase trust in, and the efficiency of, publicly-funded research.
-Here we describe a recent "open experiment", in which we documented every aspect of a soil incubation experiment online, making all raw data, scripts, diagnostics, final analyses, and manuscripts publicly available in real time.
+Many scientists however lack experience in, and may be unsure of the benefits of, making their data and fully-reproducible analyses publicly available.
+Here we describe a recent "open experiment", in which we documented every aspect of a soil incubation online, making all raw data, scripts, diagnostics, final analyses, and manuscripts available in real time.
 We found that using tools such as version control, issue tracking, and open-source statistical software improved data integrity, accelerated our team's communication and productivity, and ensured transparency.
@@
 
-#### Introduction
+#### 1. Introduction
 
 Science is becomingly increasingly collaborative and data-intensive {Adams, 2012 #4348}; in conjunction with revolutions in Internet-based communication, this has created new research opportunities across former geographic and disciplinary barriers.
-At the same time, many factors are encouraging, at times pushing, scientists to increase data access and use "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.
+At the same time, many factors are pushing scientists to increase data access and use "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.
 
-These factors include journals, which are adopting increasingly stringent data access and deposition policies, e.g.
-*Scientific Data*[^1], *PLOS ONE*[^2], and *Science*[^3].
-These policies generally share common assumptions and goals: maximizing access to scientific data; encouraging deposition into structured repositories as opposed to journal supplementary information; and specifying that it is not acceptable for authors to be solely responsible for ensuring data access.
+Scientific journals are adopting increasingly stringent data access and deposition policies, e.g.
+those of *Scientific Data*[^1], *PLOS ONE*[^2], and *Science*[^3].
+These policies generally share common assumptions and goals: maximizing access to  data; encouraging deposition into structured repositories as opposed to journal supplementary information; and specifying that it is not acceptable for authors to be solely responsible for ensuring data access.
 Funding agencies are moving in this direction as well, with organizations such as the U.S.
@@
 A second focus revolves around ensuring reproducibility {Stodden, 2011 #4340} and enabling larger synthetic activities.
-Such analyses {Wolkovich, 2012 #3303} are made possible by the assembly of large, internally consistent data sets.
-Examples in ecology, soil science, and biogeosciences include BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, and SRDB {Bond-Lamberty, 2010 #2320}.
-Online databases and collaborative tools have expanded the abilities of researchers to collaborate across large distances, both improving data access and facilitating multidisciplinary research partnerships {Markowitz, 2015 #4368}.
+Such analyses {Wolkovich, 2012 #3303} are made possible by the assembly of large, internally consistent data sets; examples in ecology, soil science, and biogeosciences of such databases include BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, and SRDB {Bond-Lamberty, 2010 #2320}.
+Online databases and collaborative tools have also expanded the abilities of researchers to collaborate across large distances, both improving data access and facilitating multidisciplinary research partnerships {Markowitz, 2015 #4368}.
 
@@
 
-**Repeatability and reproducibility**
+**1.1. Repeatability and reproducibility**
 
 Reproducibility of experimental results is at the heart of science and a requirement for results to be accepted as factual {Kuhn, 1962 #4341}.
-Too often, however, sufficient details and data are not publicly available to even attempt to repeat a study (i.e.
+Too often, however, sufficient details and data are not publicly available to even *repeat* a study (i.e.
 perform it again in a comparable manner, while not expecting exactly the same result).
@@
 But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task--often until key details have been forgotten, if not forever.
-This then presents opportunities for real-time peer review of both code and data.
+This also allows for real-time peer review of both code and data.
 
-**Data loss**
+**1.2. Data loss**
 
@@
 Similarly, Reichman et al.
-{, 2011 #2862} estimated that less than 1% of ecological data collected is made available after publication, and noted, as an example, that much current and historical data relevant to the recent *Deepwater Horizon* oil spill are already inaccessible or lost {Reichman, 2011 #2862}.
+{, 2011 #2862} estimated that less than 1% of ecological data collected is made available after publication, and noted, as an example, that much current and historical data relevant to the 2010 *Deepwater Horizon* oil spill are already inaccessible or lost {Reichman, 2011 #2862}.
 In global change ecology, Wolkovich et al.
-{, 2012 #3303} reported that they were able to acquire only ~10% of other researchers' raw data sets during a meta-analysis {Wolkovich, 2012 #3331}.
+{, 2012 #3303} reported that they were able to acquire only ~10% of other researchers' raw data sets in preparation for a meta-analysis {Wolkovich, 2012 #3331}.
 
@@
 A subtler data loss issue is the "file drawer problem" {Rosenthal, 1979 #2291}, where unpublished but potentially valuable data are lost forever.
-Scientists' use of strong and consistent data curation practices {Hart, 2015 #4292}{Rüegg, 2014 #3886} can mitigate the problem, but all our anecdotal experience and quantitative studies {Vines, 2014 #4130} suggest that in the long term, data cannot be reliably preserved by individual researchers.
+Scientists' use of strong and consistent data curation practices {Hart, 2015 #4292}{Rüegg, 2014 #3886} can mitigate the problem, but both our anecdotal experience and quantitative studies {Vines, 2014 #4130} suggest that in the long term, data cannot be reliably preserved by individual researchers.
 
@@
 PDF), use restrictions, and uncertain long-term availability {Molloy, 2011 #4366;Neylon, 2012 #4367}.
-Most journals do not have the necessary and forward-thinking expertise in data storage and management {Cassey, 2006 #4349}.
+Most journals neither desire nor have the necessary expertise in data storage and management {Cassey, 2006 #4349}.
 In contrast, the best repositories provide easy data uploading, immediate assignment of data digital object identifiers (DOIs), and long-term stability and availability.
-They do not, however, eliminate all risks: for example, the genomics sciences have a long history of requiring sequences to be deposited, but there are multiple repositories of varying curatorial levels, data quality, and formats, causing longer-term problems {Lagesen, 2010 #4369}.
+They do not, however, eliminate all risks: for example, the genomics sciences have a long history of requiring sequences to be deposited, but there are multiple repositories of varying curatorial levels, data quality, and formats {Lagesen, 2010 #4369}.
 
-**Institutional and social trust**
+**1.3. Institutional and social trust**
 
@@
 
-#### An open experiment: one example
+#### 2. An open experiment: one example
 
@@
 In this case, it was less than nine months between sampling soil cores in the Alaskan forest and submitting two resulting manuscripts.
+* Late changes to the analysis pipeline (for example, when we identified an incorrect calculation) did not result in time-consuming and error-prone pasting of new data values into our manuscript.
 * Our project and experiment received more exposure and publicity than it otherwise would have.
 
-#### Conclusions
+#### 3. Caveats and conclusions
 
@@
 Scientists also frequently cite other concerns: about data sharing, being "scooped", not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
-The benefits are not always clear to researchers who might otherwise be open to following open-data practices {Molloy, 2011 #4366}.
+The benefits are not always clear to researchers who might otherwise be open to following open-data practices {Molloy, 2011 #4366}{Hampton, 2015 #4420}.
 We cannot easily dismiss all of these concerns: for example, does our community adequately credit researchers who contribute to global databases that subsequently produce high-impact papers?
@@
 First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ, in particular with respect to automation.
-Second, this approach may not be appropriate for, nor applicable to, all study types, even just in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive human-subject data.
+Second, this approach may not be appropriate for, nor applicable to, all study types, even in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive human-subject data.
 Third, Git and GitHub are designed for working with code, not data.
 Although this situation continues to improve on a technical level[^7], it still requires a fair amount of coding expertise to set up and run an experimental repository such as the one detailed here.
+We applaud efforts to lower the technical and information hurdles to open science {Hampton, 2015 #4420}.
 
@@
 In summary, a variety of forces continue to push scientists towards more transparency in their methods, code, and data, with the goals of increasing reproducibility, enabling syntheses and meta-analyses, and improving trust in, and return from, publicly-funded science.
-While imperfect, the open experiment example we highlight here does offer instrument-to-final product reproducibility and a very high level of transparency.
-Elements of it (e.g., the use of issue-tracking software) might be usefully adopted in isolation, but we hope the entire experiment will be an example of individual scientists' decisions and practices {Wolkovich, 2012 #3303} having a larger impact.
+The open experiment example we highlight here offers instrument-to-final product reproducibility and a very high level of transparency, although it is only one of a number of possible models {Hampton, 2015 #4420}.
+Elements of this case study (e.g., the use of issue-tracking software) might be usefully adopted in isolation, but we hope the entire experiment will be an example of individual scientists' decisions and practices {Wolkovich, 2012 #3303} having a larger impact.
 We encourage the communities in our respective fields to consider its advantages, and to lead rather than follow with respect to scientific reproducibility, transparency, and data availability.
@@
 The Pacific Northwest National Laboratory is operated for DOE by Battelle Memorial Institute under contract DE-AC05-76RL01830.
+
+#### Author contribution
+
+B.B.-L. designed the repository and script system described here, and wrote the manuscript with contributions from all authors.
 

### 2016-05-18 manuscript_openexperiment/Open_experiment.Rmd
Initial ERL submission

@@
 
-Science is becomingly increasingly collaborative and data-intensive {Adams, 2012 #4348}; in conjunction with revolutions in Internet-based communication, this has created new research opportunities across former geographic and disciplinary barriers.
+Science is becoming increasingly collaborative and data-intensive {Adams, 2012 #4348}; in conjunction with revolutions in Internet-based communication, this has created new research opportunities across former geographic and disciplinary barriers.
 At the same time, many factors are pushing scientists to increase data access and use "best practices" in dealing with data and code {Hart, 2015 #4292;Wilson, 2014 #4337}.

### 2016-05-19 manuscript/cpcrw_manuscript.Rmd
Added soil N section to discussion
And other tweaks throughout.

@@
+---
+output: word_document
+---
+[//]: This Rmarkdown file generates the Biogeosciences manuscript
+
 ---
@@
 *Correspondence to*: Ben Bond-Lamberty (bondlamberty@pnnl.gov)
-
-[//]: This is the Rmarkdown file
 
@@
 Carbon dioxide fluxes were strongly influenced by incubation chamber temperature, core water content, and percent soil nitrogen, and had a temperature sensitivity (i.e.
-Q~10~) of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively..
-Methane fluxes were most strongly correlated with percent nitrogen, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
+Q~10~) of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively.
+Methane emissions were most strongly correlated with percent nitrogen, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
 The cumulative production of C from CO~2~ was over `r minmag` orders of magnitudes higher than that from CH~4~.
@@
 Such dynamics may however be disrupted by increased fire disturbance, particularly with more-intense fires {Johnstone, 2010 #2465}{Genet, 2013 #4412}.
-In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
+In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}{Karhu, 2014 #4124}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
 
@@
 At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (N=6 in each group).
-These included two incubation temperatures of 4 and 20 °C, following the protocol of a number of previous boreal incubation studies {cf.
-Table 1 in \Treat, 2015 #4217}.
+These included two incubation temperatures of 4 and 20 °C, following the protocol of a number of previous boreal incubation studies {Treat, 2015 #4217}.
 Within each temperature there were two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume.
 The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status would close match those of the 4 °C "drought" cores, which we anticipated would dry more slowly than their 20 °C counterparts.
-The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density.
+The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and N, and bulk density.
 Subsamples were collected and stored at -20 °C for dissolved organic carbon measurements or air-dried for soil C and N (see below).
@@
 Soil core was treated as a random effect in the model.
-We then used performed stepwise model selection by Akaike's information criterion (AIC) using the *stepAIC* function in the R 'MASS' package, version `r packageVersion("MASS")`.
+We then performed stepwise model selection by Akaike's information criterion (AIC) using the *stepAIC* function in the R 'MASS' package, version `r packageVersion("MASS")`.
 A linear mixed-effects model was also used to evaluate the effect of treatment on core water content.
@@
 
-All data analysis and statistics were performed using R version `r R.version.string` {R Development Core Team, 2016 #4314}.
-This experiment was run as an "open experiment" {Bond-Lamberty, 2016 #4388} with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
+All data analysis and statistics were performed using `r R.version.string` {R Development Core Team, 2016 #4314}.
+This experiment was run as an 'open experiment' {Bond-Lamberty, 2016 #4388} with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
 The summarized flux data backing the main results have been archived at [*DOI to be filled in on acceptance*].
@@
 Carbon content was `r pn(cndata_summary["C_percent_mean"], 2)`% ± `r pn(cndata_summary["C_percent_sd"], 2)`%, while N content was `r pn(cndata_summary["N_percent_mean"], 2)`% ± `r pn(cndata_summary["N_percent_sd"], 2)`%.
+Mean C:N was `r pn(cndata_summary["C_percent_mean"]  / cndata_summary["N_percent_mean"], 1)`.
 Neither temperature nor moisture treatment exerted any significant effect on these highly variable properties (P > 0.1 for all).
@@
 In the linear mixed-effects model, CO~2~ was strongly influenced by incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 1**).
-(Note that percent C and percent N were highly correlated (r ≥ 0.98) for these cores.
-Because percent N was a slightly stronger predictor, it was retained in the model while percent C was excluded.) The interaction between water content and percent N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), with high-N cores having little relationship between water content and CO~2~ flux (data not shown).
+Percent C and percent N were highly correlated (r = `r pn(cor(cndata$C_percent, cndata$N_percent), 2)`) for these cores.
+Because percent N was a slightly stronger predictor, it was retained in the model while percent C was excluded; cf.
+Colman and Schimel {, 2014 #4172}.
+The interaction between water content and percent N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), with high-N cores having little relationship between water content and CO~2~ flux (data not shown).
 
@@
 
-We found that CO~2~, but not CH~4~, fluxes from these oxic active-layer mineral soils were highly sensitive to temperature and drought effects.
+Rises in boreal air temperatures, and unpredictable precipitation changes, will warm and dry many soils, increase vegetation stress {Ju, 2016 #4320}{Barber, 2000 #1320}, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}, with uncertain consequences for soil dynamics and GHG fluxes.
+In this laboratory experiment we found that CO~2~, but not CH~4~, fluxes from these oxic active-layer mineral soils were sensitive to temperature and, in particular, moisture.
+
 Several studies have measured microbial respiration and GHG fluxes from soils very close to our study site.
 Morishita et al.
-{, 2014 #4316} measured gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture, consistent with our results They found however that CH~4~ uptake (no emissions were observed) was driven by temperature only.
+{, 2014 #4316} quantified gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture, consistent with our results They found however that CH~4~ uptake (no emissions were observed) was driven by temperature only.
 Waldrop et al.
@@
 {, 2015 #4217} reported mean CO~2~ rates of 47 (all mineral soils) and 101 (for 20-100 cm soils) µg C-CO~2~ g C^-1^ day^-1^, somewhat lower than our aerobic incubation results.
-Weiss et al.
-{, 2015 #4317} found CO~2~ production from Siberian Yedoma permafrost samples to be correlated with both percent C and N, consistent with our results (**Table 1**); The response of soil biota to stresses such as drought tend to differ between soil types and organisms, but be broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}, making such comparisons between e.g.
-Alaska and Siberia incubations useful, in spite of their significant ecological and climatic differences.
+The response of soil biota to stresses such as drought tends to differ between soil types and organisms, but be broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}, making such comparisons between useful, in spite of significant ecological and climatic differences of studies collected by meta-analyses such as Treat et al.
+{, 2015 #4217}.
 
@@
 In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture), but not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
+In contrast, observed surface CO~2~ fluxes at this range exhibited a Q~10~ of 5.1 ± 1.4 over a temperature range of 3.5-15 °C (personal communication, C.
+Anderson).
 It is important to note, however, that while increased temperature does not always drive C mineralization rates in forest mineral soils {Giardina, 2000 #2159}, it is linked with increases in soil moisture content can lead to changes in microbial community structure and GHG fluxes {Xue, 2016 #4361}.
-More surprisingly, at least initially, was that Q~10~ values were lower in the drought treatment cores, although one would expect a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 than at 20.
-TODO: THOUGHTS ON THIS FROM A MICROBIAL PERSPECTIVE?
+More surprisingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 than at 20 °C.
+**TODO: THOUGHTS ON THIS FROM A MICROBIAL PERSPECTIVE?**
 
-We observed very low, but consistently positive, CH~4~ production from these unsaturated mineral soils.
-This is contrast to many field studies, e.g.
-Schaufler et al.
-{, 2010 #4391} who found consistent CH~4~ uptake (oxidation) in dry conditions at a boreal site.
-Our results are however broadly consistent with data from 65 studies summarized by Olefeldt et al.
+We observed very low but positive CH~4~ production from these upland mineral soils.
+This is contrast to many field studies that have observed CH~4~ uptake (oxidation) in dry boreal sites {Matson, 2009 #2259}{Schaufler, 2010 #4391}.
+Anoxic microsites in soil can however provide enough CH~4~ production to balance low-level consumption in otherwise aerobic soils {Kammann, 2009 #4424}.
+In addition, our results are broadly consistent with data from 65 studies summarized by Olefeldt et al.
 {, 2013 #3456}, who found that CH~4~ emissions were more sensitive to soil temperature in wetter ecosystems; it would have been a surprise if the little methanogenic activity in our upland, well-drained soils was temperature-sensitive at all.
-Thus we see little opportunity for CH~4~ to be a significant contributor, now or in the future, to such soils' C fluxes and climate feedback risk.
-
-Methane was thus a far smaller C flux than CO~2~ from these soils, in particular at higher temperatures (as CO~2~ was responsive to temperature, but CH~4~ was not).
+Methane was also a far smaller C flux than CO~2~ from these soils, in particular at higher temperatures (as CO~2~ was responsive to temperature, but CH~4~ was not).
 This is true more generally: for example, Treat et al.
-found a median CO~2~:CH~4~ production ratio of 387 for boreal sites, far lower than (but consistent with) our observed ratios of several million, and global flux estimates of CH~4~ are orders of magnitude lower than those of CO~2~ {Kirschke, 2013 #3796}.
+found a median CO~2~:CH~4~ production ratio of 387 for boreal sites, far lower than (but consistent with) our observed ratios of several million.
+Thus we see little opportunity for CH~4~ to be a significant contributor to these upland soils' C fluxes and climate feedback risk, even accounting for the stronger radiative forcing of this gas.
 
 *4.2 Soil nitrogen*
+
+Somewhat unexpectedly, percent soil N was very significantly and positively correlated with both CO~2~ and CH~4~ fluxes (**Tables 1 and 2**).
+N interacts with microbial respiration via a number of complex, interactive, and still unclear mechanisms {Luo, 2006 #2443}, including reductions in belowground plant allocation, shifts in energy source or population of the saprotrophic community {Saiya-Cork, 2002 #3595} that leave it less capable of decomposing recalcitrant compounds , and perhaps abiotic stabliization mechanisms {Janssens, 2010 #2518}.
+Meta-analyses have generally shown negative to neutral effects of N deposition on microbial biomass {Treseder, 2008 #4421} and respiration {Ramirez, 2012 #4422}, and total soil respiration across ecosystems and biomes {Janssens, 2010 #2518}{Zhou, 2014 #3845}.
+These effect are likely due to several one or more mechanisms involving soil pH, ligninase enzymes, and phenol oxidase activity {Luo, 2006 #2443}.
+
+These conclusions have generally come from examining the effect of anthropogenic N deposition or experimental N amendments.
+Individual studies, including this one, have however observed positive correlations between ambient soil N and microbial respiration.
+For example, Weiss et al.
+{, 2015 #4317} found CO~2~ production from Siberian Yedoma permafrost samples to be correlated with both percent C and N, consistent with our active-layer results (**Table 1**).
+In an incubation of 84 North American soils, Colman and Schimel found that percent C and percent N were highly (R^2^ = 0.84) correlated and that the former was a significant predictor of microbial respiration--similar to our findings (C to N correlation of `r pn(cor(cndata$C_percent, cndata$N_percent), 2)`), except that percent N out-predicted percent C in this study.
+*In situ* respiration rates have also been shown to be negatively correlated with C:N at large spatial scales {Allaire, 2012 #4423}.
+Percent C and N both varied widely in our soil cores, even though they were collected within tens of meters of each other, suggesting that active-layer SOC response to temperature and moisture may also be highly spatially variable, even in a visually uniform boreal forest.
+Spatially explicit analyses of soil properties, temperatures {Bond-Lamberty, 2005 #1072}, and respiration {Allaire, 2012 #4423} are likely necessary to accurately constrain and predict soil fluxes in this ecosystem.
 
@@
 
-We focused on a experimental drought, rather than flooding, because of the well-drained nature of the field site: it is unlikely that the mid-slope forest we sampled in will ever suffer from thermokarst or excessive soil moisture, but too-dry conditions are a serious possibility in this low-precipitation ecosystem {Barber, 2000 #1320}.
+We focused on an experimental drought, rather than flooding, because of the well-drained nature of the field site: it is unlikely that the mid-slope forest we sampled in will ever suffer from thermokarst or excessive soil moisture, but too-dry conditions are a serious possibility in this low-precipitation ecosystem {Barber, 2000 #1320}.
 In addition, the soils here are not surface layer soils (where the majority of microbial activity and C mineralization of labile C takes place); taking them out of depth (where they are less exposed to O~2~, for example) may significantly change the abiotic conditions to which the microbial community is adapted.
-MORE?
+**TODO:MORE?**
 
@@
 
-Rises in boreal air temperatures, and unpredictable precipitation changes, will warm and dry many soils, increase vegetation stress {Ju, 2016 #4320}{Barber, 2000 #1320}, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}, with uncertain consequences for soil dynamics and GHG fluxes.
-In this laboratory experiment, we found that CO~2~ fluxes were strongly influenced by temperature, water content, and soil nitrogen, while CH~4~ fluxes were much smaller and not sensitive to temperature or water content in these well-drained mineral soils.
-This suggests that understanding how soil moisture might change with permafrost degradation, how soil biota will respond to these changes, and how models should treat soil organic matter decomposition with respect to multiple and interacting drivers are all critical areas of research going forward.
+In this laboratory experiment, we found that CO~2~ fluxes were strongly influenced by temperature and water content, and correlated with soil C and N, while CH~4~ fluxes were much smaller and not sensitive to temperature or water content in these well-drained mineral soils.
+This suggests that understanding how soil moisture might change with spatially variable permafrost degradation, how soil biota will respond to these changes, and how models should treat soil organic matter decomposition with respect to multiple and interacting drivers are all critical areas of research going forward.
 Further controlled field and laboratory studies, ideally tightly integrated with modeling experiments, are critical to understand GHG emission dynamics from high-latitude soils.

### 2016-05-26 manuscript/cpcrw_manuscript.Rmd
Incorporated VLB comments on May 19 draft

@@
 
-We sampled soils from a southeast slope (65.1620 °N, 147.4874 °W) at CPCRW, in a 60 m transition zone between lowland *Picea mariana* Mill BSP.
-and upland *Betula neoalaskana*, with significant white spruce (*Picea glauca*) presence.
+We sampled soils from a southeast slope (65.1620 °N, 147.4874 °W) at CPCRW, in a 60 m transition zone between lowland *Picea mariana* and upland *Betula neoalaskana*, with significant white spruce (*Picea glauca*) presence as well.
 Stand density in this transition zone was 4060 ± 2310 trees ha^-1^, with basal area of 27.9 ± 7.0 m^2^ ha^-1^.
@@
 Waldrop et al.
-{, 2010 #4380} also observed flux rates of 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), differing by orders of magnitude between sites (but roughly similar to our observed CH~4~ emissions), and ~1-5 µg C-CO2 hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed from our cores.
+{, 2010 #4380} also observed flux rates of 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), differing by orders of magnitude between sites (but roughly similar to our observed CH~4~ emissions), and ~1-5 µg C-CO~2~ hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed from our cores.
 In an incubation of active-layer Alaskan permafrost peats, Treat et al.
@@
 More surprisingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 than at 20 °C.
-**TODO: THOUGHTS ON THIS FROM A MICROBIAL PERSPECTIVE?**
+There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}.
+Unlike surface soils, active layer soils, which store large quantities of soil C {Mueller, 2015 #4435}, are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils may be older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}.
+Thus, it is not surprising that these more stable C compounds would be metabolized by processes that have been reported to be less temperature-sensitive.
 
@@
 Somewhat unexpectedly, percent soil N was very significantly and positively correlated with both CO~2~ and CH~4~ fluxes (**Tables 1 and 2**).
-N interacts with microbial respiration via a number of complex, interactive, and still unclear mechanisms {Luo, 2006 #2443}, including reductions in belowground plant allocation, shifts in energy source or population of the saprotrophic community {Saiya-Cork, 2002 #3595} that leave it less capable of decomposing recalcitrant compounds , and perhaps abiotic stabliization mechanisms {Janssens, 2010 #2518}.
+N interacts with microbial respiration via a number of complex, interactive, and still unclear mechanisms {Luo, 2006 #2443}, including reductions in belowground plant allocation, shifts in energy source or population of the saprotrophic community {Saiya-Cork, 2002 #3595} that leave it less capable of decomposing recalcitrant compounds, and perhaps abiotic stabilization mechanisms {Janssens, 2010 #2518}.
 Meta-analyses have generally shown negative to neutral effects of N deposition on microbial biomass {Treseder, 2008 #4421} and respiration {Ramirez, 2012 #4422}, and total soil respiration across ecosystems and biomes {Janssens, 2010 #2518}{Zhou, 2014 #3845}.
@@
 In addition, the soils here are not surface layer soils (where the majority of microbial activity and C mineralization of labile C takes place); taking them out of depth (where they are less exposed to O~2~, for example) may significantly change the abiotic conditions to which the microbial community is adapted.
-**TODO:MORE?**
+However, focusing on the active layer provides crucial information about the potential loss of C from these soils, a risk that needs to be well understood as permafrost degradation leads to expansions in the depth of the active layer across the Arctic.
 
@@
 
-We are grateful to Jamie Hollingsworth for information about, and faciliating access to, the Caribou Poker Creeks Research Watershed Long-Term Ecological Research site.
+We are grateful to Jamie Hollingsworth for information about, and facilitating access to, the Caribou Poker Creeks Research Watershed Long-Term Ecological Research site.
 This research was supported by the Office of Science of the U.S.

### 2016-06-01 manuscript/cpcrw_manuscript.Rmd
Final tweaks and Biogeosciences submitted version

@@
 Gravimetric water content was `r pn(gwcbegin$mean, 2)` ± `r pn(gwcbegin$sd, 2)` (unitless) at the beginning of the incubation; cores at field moisture were unchanged at the end, but drought cores had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
-Carbon dioxide fluxes were strongly influenced by incubation chamber temperature, core water content, and percent soil nitrogen, and had a temperature sensitivity (i.e.
-Q~10~) of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively.
+Carbon dioxide fluxes were strongly influenced by incubation chamber temperature, core water content, and percent soil nitrogen, and had a temperature sensitivity (Q~10~) of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively.
 Methane emissions were most strongly correlated with percent nitrogen, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
@@
 This experiment was run as an 'open experiment' {Bond-Lamberty, 2016 #4388} with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
-The summarized flux data backing the main results have been archived at [*DOI to be filled in on acceptance*].
+The summarized flux data backing the main results have been archived at [*DOI to be filled in*].
 
@@
 
-*4.1 Temperature versus moisture sensitivity*
+**4.1 Temperature versus moisture sensitivity**
 
@@
 In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture), but not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
-In contrast, observed surface CO~2~ fluxes at this range exhibited a Q~10~ of 5.1 ± 1.4 over a temperature range of 3.5-15 °C (personal communication, C.
-Anderson).
-It is important to note, however, that while increased temperature does not always drive C mineralization rates in forest mineral soils {Giardina, 2000 #2159}, it is linked with increases in soil moisture content can lead to changes in microbial community structure and GHG fluxes {Xue, 2016 #4361}.
+Observed surface CO~2~ fluxes at this range exhibited a Q~10~ of 5.1 ± 1.4 over a temperature range of 3.5-15 °C (personal communication, C.
+Anderson); these surface fluxes were measured over multiple months and include root respiration, however, confounding any direct comparison.
+It is also important to note that while increased temperature does not always drive C mineralization rates in forest mineral soils {Giardina, 2000 #2159}, it is linked with increases in soil moisture content can lead to changes in microbial community structure and GHG fluxes {Xue, 2016 #4361}.
+
 More surprisingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 than at 20 °C.
@@
 
-*4.2 Soil nitrogen*
+**4.2 Soil nitrogen**
 
@@
 
-*4.3 Limitations and weaknesses*
+**4.3 Limitations and weaknesses**
 
@@
 
-*5 Conclusions*
+#### 5 Conclusions
 

### 2016-06-28 manuscript_openexperiment/Response_to_reviewers.Rmd
Added response to reviewers doc.

---
title: "Response_to_reviewers"
author: "Ben Bond-Lamberty"
date: "June 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2016-06-28 manuscript_openexperiment/Response_to_reviewers.Rmd
Responses to Reviewer 2.

@@
 ```
+
+**Referee 2**
+
+*This is a rather unusual manuscript, reporting not on science but on a new way that science might be conducted.
+I was prepared to dislike the paper, but I found it rather interesting and easy to read.
+It might even convince me to join a disaggregate group of scientists to pursue a study of interest.
+If there is a special section of ERL that publishes such papers, I would encourage you to include it.*
+
+Thank you.
+
+*Only a couple of specialized comments*
+*1.       The authors should standardize whether web citations are given in the text or in footnotes.  I prefer the latter to keep the text clean.*
+
+We agree, and have moved all web citations--except the one referencing the main repository, the subject of this manuscript--to footnotes.
+(We will obviously follow journal requirements if they differ from this, however.)
+
+*2.       Figure 2, particularly the lower panel, needs a lot more explanation before it will be easy for readers to understand.*
+
+We have expanded and hopefully clarified the Figure 2 caption.
+
+**Referee 3**
+
+*The manuscript is a perspective and case study analysis on the value of running an "open" experiment.
+The idea behind this is to lay out the protocol, the treatments, the results and even the manuscript drafts all open for public viewing and commenting in real time, as data come in and results are developed.
+It's a fresh and radical approach to doing science and could strongly benefit rebuilding trust in our field, as the authors document and adequately reference literature on the topic.*
+
+*As a journal editor, I have been pushing in every article to provide a clear data access statement and to publish data in open and public repositories.
+It's nice to have a manuscript to point to on good reason for doing so with a clear example.*
+
+Thank you.
+
+*I don't have any recommendations for major changes to this short piece.
+However, I think the caveats section (section 3) could benefit from more discussion on barriers to adopting this approach, not just from the incentive side, but from the skills needed, time invested, and challenges in using systems like github to run experiments.
+It would be helpful to discuss what was hard in getting the open experiment to work, what took the most time, what tools were the most difficult or least developed or accessible to scientists.
+If anything most of us are well convinced of the arguments the authors make on why to share data and code and incentives are showing up in multiple places.
+The challenge is finding the time, motivation, and technical skills to do so.
+So perhaps some discussion of how the authors dealt with those would help convince the otherwise skittish researcher that this isn’t as hard as it looks to do!*
+
+TODO.

### 2016-06-28 manuscript_openexperiment/Open_experiment.Rmd
Responses to Reviewer 2.

@@
 author: "Ben Bond-Lamberty^1^, A. Peyton Smith^2^, Vanessa Bailey^2^"
-date: "February 2, 2016"
-output: word_document
+date: "July 1, 2016"
+output:
+  html_document: default
+  word_document: default
 ---
@@
 More generally, scientists in all fields are increasingly building and using software in their work, though often without strong training in this area {Wilson, 2014 #4337}.
-In addition to open-source data analysis languages such as R {R Development Core Team, 2016 #4314}, scientific workflow systems such as Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record information about the data processing, analytical process and decisions, and statistical analysis.
+In addition to open-source data analysis languages such as R {R Development Core Team, 2016 #4314}, scientific workflow systems such as Kepler[^6] or Taverna[^7] record information about the data processing, analytical process and decisions, and statistical analysis.
 Providing open code does not magically produce bug-free code, mistake-free analyses, or instantly better science {Easterbrook, 2014 #4137}.
@@
 This also allows for real-time peer review of both code and data.
+
+[^6]: https://kepler-project.org
+[^7]: http://www.taverna.org.uk
 
@@
 This meant that changes to the data (and thus statistical results, etc.) propagated automatically and consistently.
-Tools such as R Markdown[^7] have made this process much easier to build and maintain.
+Tools such as R Markdown[^8] have made this process much easier to build and maintain.
 * Log files provide an audit trail {Ellison, 2010 #4363} of what analytical steps were taken and, critically, the specific versions used for each unit of software, including the main R system itself.
@@
 
-[^6]: http://rmarkdown.rstudio.com
+[^8]: http://rmarkdown.rstudio.com
 
@@
 Third, Git and GitHub are designed for working with code, not data.
-Although this situation continues to improve on a technical level[^7], it still requires a fair amount of coding expertise to set up and run an experimental repository such as the one detailed here.
+Although this situation continues to improve on a technical level[^9], it still requires a fair amount of coding expertise to set up and run an experimental repository such as the one detailed here.
 We applaud efforts to lower the technical and information hurdles to open science {Hampton, 2015 #4420}.
 
-[^7]: https://github.com/blog/1885-better-word-highlighting-in-diffs
+[^9]: https://github.com/blog/1885-better-word-highlighting-in-diffs
 
@@
 **Figure 2.** Two of the many diagnostic figures that were produced as the incubation experiment proceeded.
-They show (top) the number of replicate measurements made on each soil core, by date of measurement, and (bottom) measurements with a data mismatch problem, indicating a data entry or other error (in the figure no errors, which would be in red, are extant).
-A wide range of similar diagnostic graphics (see https://github.com/bpbond/cpcrw_incubation#current-diagnostics) and statistics let us quickly identify problems or interesting patterns in the data.
+Top: the number of replicate measurements made on each soil core (y), by date of measurement (x); the 'gap' at 2015-11-11/12 shows a potential problem in our data-processing pipeline, related to the instrument's clock crossing midnight.
+Bottom: data matches between multiplexer valve number (y) and soil core, by date of measurement (x); grey dots indicate that every valve has been correctly matched with a soil core, while a red dot would indicate a data entry problem or other error.
+A wide range of similar diagnostic graphics and statistics let us quickly identify problems or interesting patterns in the data.
 All diagnostics were publicly available in real time.

### 2016-06-28 manuscript_openexperiment/Response_to_reviewers.Rmd
Responses to Reviewer 2; significant new text in ms re barriers and caveats.

@@
 ```
+
+We thank all referees for their thoughtful comments and careful assessment. Responses to their comments (italicized) are below, and significant changes in the manuscript text are marked in red.
 
@@
 
-TODO.
+We have taken much of the material in the former "Conclusions" section and split into a new section, "Barriers and caveats", which discusses some of these issues (skills, time, barriers for an 'average' scientist) in much greater depth.

### 2016-06-28 manuscript_openexperiment/Open_experiment.Rmd
Responses to Reviewer 2; significant new text in ms re barriers and caveats.

@@
 output:
+  word_document: default
   html_document: default
-  word_document: default
 ---
@@
 
-#### 3. Caveats and conclusions
+#### 3. Barriers and caveats
 
-By itself transparency does not guarantee repeatability and reproducibility {Easterbrook, 2014 #4137}, and it may raise genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
+There were thus significant advantages to implementing and using an automated analytical pipeline in an open repository.
+It is worth considering in depth, however, the costs of such an approach, the investments required, and the barriers we encountered (or that many other scientists would).
+
+* More upfront time is required in an open repository system, as the basic building blocks are assembled into an automated pipeline that runs from raw data to diagnostics to final products.
+While frustrating, this is generally a good thing however: automation makes later analytical stages much faster (see above), while openness provides a strong incentive to write clean, clear code from the beginning, for example, as opposed to deferring documentation.
+* To our knowledge, there is no 'template' for an experimental repository such as the one used here.
+This results in unnecessary time re-inventing wheels, in particular with respect to repository design decisions.
+* Effective data management takes some programming skills.
+We believe that basic programming, like basic statistics, should be a part of almost every scientist's training, and do not view this as a huge problem going forward.
+For scientists used to spreadsheet-oriented data processing, however, it requires new skills and practices.
+* Version control (here, the Git software) is critical for tracking provenance, ensuring robustness, and efficiently sharing changes.
+It introduces significant complexity, however, particularly for any use beyond the most simple operations.
+We believe version control software remains too difficult for most scientists to use effectively[^9].
+* Git and GitHub are fundamentally designed for working with code, not data or long documents.
+This situation continues to improve on a technical level[^10], but some operations (e.g., storing very large files; tracking small changes to columnar data; commenting on words or phrases) remain awkward or even impossible.
+
+[^9]: https://xkcd.com/1597/
+[^10]: E.g., https://github.com/blog/1885-better-word-highlighting-in-diffs
+
+In summary, while tools and platforms such as Git, GitHub, RStudio, and RMarkdown have drastically improved the feasibility and accessibility of running an 'open experiment', they remain significant hurdles for many scientists.
+We applaud efforts to lower the technical and information hurdles to open science {Hampton, 2015 #4420}.
+
+A number of caveats are also in order.
+The model we present here--a public repository updated throughout the experiment, analysis, and publication process--is far from perfect.
+For example, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ, in particular with respect to automation.
+In addition, this approach may not be appropriate for, nor applicable to, all study types, even in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive human-subject data.
+
+More generally, by itself transparency does not guarantee repeatability and reproducibility {Easterbrook, 2014 #4137}, and it may raise genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
 Nor will open science by itself fundamentally enhance public trust in science, as there are still many social and political challenges to overcome {Grand, 2012 #4370}.
@@
 
-We also recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}.
+Finally, we recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}.
 That is, scientists do not receive 'proper' credit for datasets relative to publications, even when the former may be much more broadly valuable, in almost every area: hiring, promotion and tenure decisions, impact metrics, etc.
@@
 
-The model we present here--a public repository updated throughout the experiment, analysis, and publication process--is far from perfect.
-First, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ, in particular with respect to automation.
-Second, this approach may not be appropriate for, nor applicable to, all study types, even in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive human-subject data.
-Third, Git and GitHub are designed for working with code, not data.
-Although this situation continues to improve on a technical level[^9], it still requires a fair amount of coding expertise to set up and run an experimental repository such as the one detailed here.
-We applaud efforts to lower the technical and information hurdles to open science {Hampton, 2015 #4420}.
+#### 4. Conclusions
 
-[^9]: https://github.com/blog/1885-better-word-highlighting-in-diffs
-
-In summary, a variety of forces continue to push scientists towards more transparency in their methods, code, and data, with the goals of increasing reproducibility, enabling syntheses and meta-analyses, and improving trust in, and return from, publicly-funded science.
+A variety of forces continue to push scientists towards more transparency in their methods, code, and data, with the goals of increasing reproducibility, enabling syntheses and meta-analyses, and improving trust in, and return from, publicly-funded science.
 The open experiment example we highlight here offers instrument-to-final product reproducibility and a very high level of transparency, although it is only one of a number of possible models {Hampton, 2015 #4420}.
-Elements of this case study (e.g., the use of issue-tracking software) might be usefully adopted in isolation, but we hope the entire experiment will be an example of individual scientists' decisions and practices {Wolkovich, 2012 #3303} having a larger impact.
+Elements of this case study (e.g., the use of issue-tracking or version control software) might be usefully adopted in isolation, but we hope the entire experiment will be an example of individual scientists' decisions and practices {Wolkovich, 2012 #3303} having a larger impact.
 We encourage the communities in our respective fields to consider its advantages, and to lead rather than follow with respect to scientific reproducibility, transparency, and data availability.

### 2016-06-29 manuscript_openexperiment/Open_experiment.Rmd
Incorporated VLB comments and resubmitted

@@
 Vines et al.
-{, 2014 #4130} published a shocking finding, based on a survey of 516 biology articles from 2 to 22 years old: the odds of a data set being available fell by 17% each year, and the chances that the contact author's email address still worked declined by 7% per year.
+{, 2014 #4130} published a shocking finding, based on a survey of 516 biology articles from 2 to 22 years old: the odds of a data set being available post-publication fell by 17% each year, and the chances that the contact author's email address still worked declined by 7% per year.
 Similarly, Reichman et al.
@@
 There were thus significant advantages to implementing and using an automated analytical pipeline in an open repository.
-It is worth considering in depth, however, the costs of such an approach, the investments required, and the barriers we encountered (or that many other scientists would).
+It is worth considering in depth, however, the costs of such an approach, the investments required, and the barriers encountered.
 
 * More upfront time is required in an open repository system, as the basic building blocks are assembled into an automated pipeline that runs from raw data to diagnostics to final products.
-While frustrating, this is generally a good thing however: automation makes later analytical stages much faster (see above), while openness provides a strong incentive to write clean, clear code from the beginning, for example, as opposed to deferring documentation.
+While frustrating, this is generally good: automation makes later analytical stages much faster (see above), while openness provides a strong incentive to write clean, clear code from the beginning, for example, as opposed to deferring documentation.
 * To our knowledge, there is no 'template' for an experimental repository such as the one used here.
-This results in unnecessary time re-inventing wheels, in particular with respect to repository design decisions.
-* Effective data management takes some programming skills.
-We believe that basic programming, like basic statistics, should be a part of almost every scientist's training, and do not view this as a huge problem going forward.
-For scientists used to spreadsheet-oriented data processing, however, it requires new skills and practices.
+This results in unnecessary time spent "re-inventing wheels", in particular with respect to repository design decisions.
+Clear, flexible, and powerful templates will help scientists take advantage of repository-based and open science approaches.
+* Effective data management in a repository system takes some programming skills.
+Basic programming is becoming an increasingly important part of most scientists' training, just as basic statistics has long been a critical skill.
+For scientists accustomed to spreadsheet-based data processing, investment in new skills and practices, or new partnerships with skilled programmers, will facilitate more open analyses.
 * Version control (here, the Git software) is critical for tracking provenance, ensuring robustness, and efficiently sharing changes.
 It introduces significant complexity, however, particularly for any use beyond the most simple operations.
-We believe version control software remains too difficult for most scientists to use effectively[^9].
+In our experience, version control software remains too difficult for most scientists to use effectively[^9].
+We urge software developers to work closely with diverse science communities to develop more user-friendly tools.
 * Git and GitHub are fundamentally designed for working with code, not data or long documents.
@@
 For example, there are a number of "best practices" of scientific computing {Wilson, 2014 #4337} that we did not employ, in particular with respect to automation.
-In addition, this approach may not be appropriate for, nor applicable to, all study types, even in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive human-subject data.
+In addition, this approach may not be appropriate for, nor applicable to, all study types, even in our professional fields of soil, ecosystem, and global change science; it certainly is not appropriate for sensitive data, e.g.
+human-subject research.
 
-More generally, by itself transparency does not guarantee repeatability and reproducibility {Easterbrook, 2014 #4137}, and it may raise genuine concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
+More generally, by itself transparency does not guarantee repeatability and reproducibility {Easterbrook, 2014 #4137}, and it may raise new concerns about, for example, protecting scientists from harassment {Lewandowsky, 2016 #4342}.
 Nor will open science by itself fundamentally enhance public trust in science, as there are still many social and political challenges to overcome {Grand, 2012 #4370}.
 
-Scientists also frequently cite other concerns: about data sharing, being "scooped", not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
+Scientists also frequently cite other concerns: about data sharing, being 'scooped', not receiving sufficient credit, and time constraints {Wolkovich, 2012 #3303}.
 The benefits are not always clear to researchers who might otherwise be open to following open-data practices {Molloy, 2011 #4366}{Hampton, 2015 #4420}.
@@
 Finally, we recognize that, in general, our professional and career incentives do not yet align well with "an open research culture" {Nosek, 2015 #4343}.
-That is, scientists do not receive 'proper' credit for datasets relative to publications, even when the former may be much more broadly valuable, in almost every area: hiring, promotion and tenure decisions, impact metrics, etc.
+That is, scientists do not receive appropriate credit for datasets relative to publications, as hiring, promotion and tenure decisions all tend to reward publications, not datasets.
 The advent of 'data descriptor' articles, in journals such as *Scientific Data*, improves but does not solve this problem.

### 2016-09-06 manuscript/response_to_reviewer_comments.txt
Started response to reviewers document

# Response to referee comments

We thank both referees for their thoughtful comments.

## Referee #1

1-1. The authors have measured CO2 and CH4 emissions from mineral soils sampled from an Alaskan boreal forest within a discontinuous permafrost zone. These soils were incubated for 100 days at two temperatures and moistures in order to better understand the effect of these variables, and others like soil chemistry, on microbial respiration rates. Temperature sensitivity of respiration was higher in the drought treatment than field moisture conditions, suggesting that respiration rates in these deep active layer thickness soils may be more sensitive to changes in moisture than to temperature. Overall this is a valuable data set. This manuscript contributes to the growing body of Arctic permafrost and active layer incubation literature that will help inform temperature sensitivities of the relatively large mineral C pool at depth, and the interactions of soil temperature with soil moisture, nutrients, and carbon quality that also influence the rate of C remineralization. There has been a lot of focus in incubation studies on the effect of inundation on CO2 and CH4 emissions, but drought is an important factor to consider as well.

1-2. I do think that the paper has within it data that is relevant to the scope of BG, but I think that a more in depth analysis of the data, and a better contextualization of the data within the Arctic literature is required. The methods and concepts are not novel, but this is an understudied system and very timely work. More is needed in order to reach substantial conclusions.

1-3. I think that the work would benefit from a more thorough comparison with boreal forest incubations across the Arctic.

1-4. N section would benefit from more Arctic-centric comparisons of N limitations and in particular of boreal forest N dynamics. Q10 can be temperature dependent, also depending on N limitation in the system.

1-5. This study raises interesting questions. In mineral soils, under woody vegetation that might be of low C quality, and slower C pool, one might expect higher temperature sensitivity. I think that these questions, even if not addressed directly by the data presented, should have been discussed more explicitly. Comparison with other Arctic woody plant systems would be instructive.

1-6. Studies have shown that moisture can have a weaker effect on temperature sensitivity early on during an incubation experiment, in the presence of more labile C. This relative to the effect on moisture on the Q-10 of cumulative respiration, reflecting slow turning over C – this could be an interesting analysis to include here, and would help to assess how

1-7. How do your results in terms of temperature and moisture sensitivity (especially under drought conditions) scale with Alaskan climate change predictions from modelers? How does it compare with deep soils incubations (mineral soils) from the Arctic, and from boreal ecosystems?

1-8. Line 31-34: I cannot find discussion of this point in the rest of the text, and while important, this statement is relatively vague and there are no cited references. Since it underpins the rational for studying deep, unfrozen Arctic soils, it would be helpful to expand on this more in the manuscript.

1-9. Lines 48-60: I think that this section would benefit from an introduction of the interactions between the specific ecosystem (upland boreal forest) you are studying, and its interaction with soil chemistry, since vegetation type is influential in terms of soil carbon quality and quantity. Woody plant biomass tends to have a higher C:N ratio relative to herbaceous dominated systems, and this tends to result in lower quality resources for microbial communities.

1-10. Lines 70-72: These are really important considerations, and it seems appropriate to discuss them more explicitly. How is the temperature and precipitation regime of the boreal forest of interior AK expected to change? There are also indirect effects of vegetation type on soil temperatures that could be discussed here.

1-11. Lines 72-74: While these are important questions, they are not really addressed in this study, and so either it might make sense to leave it out, or to discuss the particulars as they apply to this study, ie: the importance and questions related to C:N ratios.

1-12. Lines 77-80: I think a stronger argument for why deep active-layer soils can be made, and it would be helpful to clarify what are the “strong effects” of warming.

1-13. I cannot tell if C:N, %C and %N were measured at the end of the incubation. Could these results be collated in a table in the manuscript? Otherwise the methods section appears to be detailed and well written.

1-14. Line 232: In this section it would also be interesting to know the soil respiration decay rate per treatment over the course of the incubation experiment.
Line 238-240: Confusingly worded sentence.

1-15. I don’t think that the summary of nearby respiration studies add very much to the discussion section. Perhaps if the similarities and discrepancies were more integral to the central findings of the paper or integrated differently into the discussion they would seem more meaningful here. Perhaps comparing with other boreal incubations (eg: Lee et al., 2012; Lavoie et al., 2011) would help to provide some additional context.

1-16. Line 270: There is missing punctuation after the word “results”.

1-17. Line 286-293: Perhaps the new synthesis by Schadel et al., 2016, would also be a useful comparison here.

1-18. Line 293: That soil moisture may be as important a control on microbial respiration as temperature is an important finding in recent incubation studies, and the potential to define its interaction with temperature will help modelers of soil decomposition better constrain the physical parameters of microbial respiration rates. This feels buried in the manuscript, and I think that it would improve the paper if it were highlighted better throughout the text.

1-19. Line 311-317: This section could be better explained in the context of the discussion or omitted altogether. It seems less important to defend the plausibility of relatively low temperature sensitivity, but instead to try to explain it in the context of these soil characteristics. Could low temperature sensitivity be the result of low C quality in this deep soil environment?

1-20. Line 322-332: This section, which lays out the crux of the paper, the interaction between temperature and moisture sensitivity in driving microbial respiration is relatively vague. It would be good to describe the less-temperature-sensitive processes that would be important to consider for more stable-C metabolism. And how does moisture play a role here? Perhaps DOC becomes more limiting in the drought conditions?

1-21. Line 356: The Janssens et al., 2010, citation refers to a meta-analysis of temperate forest soils that are not nitrogen limited. There are studies focusing on Arctic N cycling that would be more appropriate, and many Arctic studies have shown that N availability can limit C mineralization rates. Is this site considered to be N limited in the deep active layer?

1-22. Line 367: Is this comparison, with North American soils, relevant to this study?

1-23. Line 383-384: Can you be more explicit in your meaning here? How do you mean that there is weakness in what can be inferred about temperature sensitivity from experiments?

## Referee #2

2-1. This paper studies carbon (C) release in the form of CO2 and CH4 from deep-active layer boreal soils using 30 soil cores in interior Alaska. After 100 days of incubation at two different temperatures and different moisture conditions, the authors find that incubation temperature, core water content and soil nitrogen strongly influenced CO2 release. This is a useful dataset that adds to the growing body of literature on C release from permafrost once thawed. There is a good amount of publications that deal with warmer temperatures in the Arctic and changing soil moisture condition but not many have simulated drought conditions. A probably unfortunate overlap in publication date with this paper is a recent meta-analysis published by Schädel et al. 2016, which showed that C release doubles when incubation temperature increases by 10◦C and that C release under aerobic conditions is three times higher than under anaerobic conditions.

2-2. It seems like this paper was published as a discussion paper before Schädel et al. 2016 was published and hence a discussion of the meta-analysis was not possible but should be addressed in the revisions.

2-3. The importance of the results would be more obvious if the discussion also contained an upscaling or circumpolar aspect of drought in the Arctic. It would be useful to have some discussion about the area that is expected to be most affected by drought. This is important as changes in temperature will affect most of the Arctic, whereas drought effects or dry soils will occur more locally.

2-4. 1) Throughout the manuscript, I have noticed that important papers from the permafrost literature are missing. This applies to C stocks in the permafrost area, Tarnocai et al. 2009 is a good paper but there are more recent and more accurate estimates of permafrost C stocks described in Hugelius et al. 2014 and Schuur et al. 2015 that should be cited. When it comes to the permafrost C feedback, Schuur et al. 2015 is currently the best and most up to date review. In addition, Koven et al. 2015 is a good one too. The discussion on incubation literature should include papers like Lavoie et al. 2011, Dutta et al. 2006, and Schädel et al. 2014.

2-5. 2) L. 31: Permafrost thaws and does not melt

2-6. 3) A better explanation is needed why deep-active layer soils are different to active layer or permafrost soils, I couldn’t find a strong argument for why they would behave differently. Also, deep-active layer soils are those that are the most impacted by inter annual variability in thaw depth and so they might switch between active layer in one year to permafrost in another, that’s worth some discussion as well.

2-7. 4) The statistics in this paper are generally good and I would like to compliment the authors on making the entire data set and analysis available online. I would still suggest that the manuscript would profit from some additional details on collinearity of the tested variables as well as model outputs such as AIC.

2-8. 5) Add a table with soil properties such as bulk density, %C etc.

2-9. 6) Why not include C/N as a variable in the statistical analysis? Schädel et al. 2014 showed that C/N is a good predictor of C release and can be used as a scaling factor. It would be interesting to see if C release from short-term incubations show the same result

2-10. 7) In the discussion, it would be good to also include the warming potential of CO2 and CH4 especially when making assumptions about the permafrost C feedback, it is briefly mentioned in line 348 but a more in depth discussion would be good

2-11. 8) the conclusions might be a bit strong given the data and previous results published

### 2016-09-27 manuscript/response_to_reviewer_comments.txt
Tweaks in preparation for responses
Responding to reviewer comments
Easy stuff first…

@@
 # Response to referee comments
+# Note that using LaTeX command below, as that's what Biogeosciences uses
+# http://www.biogeosciences.net/peer_review/commenting_on_a_paper.html
+# BBL September 2016
 
-We thank both referees for their thoughtful comments.
+We thank both referees for their thoughtful comments, which we have numbered below for convenient referencing. Line numbers refer to the discussion manuscript.
 
-## Referee #1
+{\bf Response to Referee #1}
 
-1-1.
-The authors have measured CO2 and CH4 emissions from mineral soils sampled from an Alaskan boreal forest within a discontinuous permafrost zone.
-These soils were incubated for 100 days at two temperatures and moistures in order to better understand the effect of these variables, and others like soil chemistry, on microbial respiration rates.
-Temperature sensitivity of respiration was higher in the drought treatment than field moisture conditions, suggesting that respiration rates in these deep active layer thickness soils may be more sensitive to changes in moisture than to temperature.
-Overall this is a valuable data set.
-This manuscript contributes to the growing body of Arctic permafrost and active layer incubation literature that will help inform temperature sensitivities of the relatively large mineral C pool at depth, and the interactions of soil temperature with soil moisture, nutrients, and carbon quality that also influence the rate of C remineralization.
+{\it 1-1.
+The authors have measured...
+This manuscript contributes to the growing body of Arctic permafrost and active layer incubation literature...
 There has been a lot of focus in incubation studies on the effect of inundation on CO2 and CH4 emissions, but drought is an important factor to consider as well.
+}
 
-1-2.
+Thank you.
+
+{\it 1-2.
 I do think that the paper has within it data that is relevant to the scope of BG, but I think that a more in depth analysis of the data, and a better contextualization of the data within the Arctic literature is required.
@@
 More is needed in order to reach substantial conclusions.
+}
 
-1-3. I think that the work would benefit from a more thorough comparison with boreal forest incubations across the Arctic.
+We agree that the methods and concepts are not particularly novel (although drought incubations are fairly unusual), and that this is an understudied system and timely–thank you.
 
-1-4.
+{\it 1-3. I think that the work would benefit from a more thorough comparison with boreal forest incubations across the Arctic. }
+
+TODO
+
+{\it 1-4.
 N section would benefit from more Arctic-centric comparisons of N limitations and in particular of boreal forest N dynamics.
 Q10 can be temperature dependent, also depending on N limitation in the system.
+}
 
-1-5.
+TODO
+
+{\it 1-5.
 This study raises interesting questions.
@@
 Comparison with other Arctic woody plant systems would be instructive.
+}
 
-1-6.
+TODO
+
+{\it 1-6.
 Studies have shown that moisture can have a weaker effect on temperature sensitivity early on during an incubation experiment, in the presence of more labile C.
-This relative to the effect on moisture on the Q-10 of cumulative respiration, reflecting slow turning over C – this could be an interesting analysis to include here, and would help to assess how
+This relative to the effect on moisture on the Q-10 of cumulative respiration, reflecting slow turning over C – this could be an interesting analysis to include here, and would help to assess how }
 
-1-7.
+TODO
+
+{\it 1-7.
 How do your results in terms of temperature and moisture sensitivity (especially under drought conditions) scale with Alaskan climate change predictions from modelers?
 How does it compare with deep soils incubations (mineral soils) from the Arctic, and from boreal ecosystems?
+}
 
-1-8.
+TODO
+
+{\it 1-8.
 Line 31-34: I cannot find discussion of this point in the rest of the text, and while important, this statement is relatively vague and there are no cited references.
 Since it underpins the rational for studying deep, unfrozen Arctic soils, it would be helpful to expand on this more in the manuscript.
+}
 
-1-9.
+TODO
+
+{\it 1-9.
 Lines 48-60: I think that this section would benefit from an introduction of the interactions between the specific ecosystem (upland boreal forest) you are studying, and its interaction with soil chemistry, since vegetation type is influential in terms of soil carbon quality and quantity.
 Woody plant biomass tends to have a higher C:N ratio relative to herbaceous dominated systems, and this tends to result in lower quality resources for microbial communities.
+}
 
-1-10.
+TODO
+
+{\it 1-10.
 Lines 70-72: These are really important considerations, and it seems appropriate to discuss them more explicitly.
@@
 There are also indirect effects of vegetation type on soil temperatures that could be discussed here.
+}
 
-1-11.
+TODO
+
+{\it 1-11.
 Lines 72-74: While these are important questions, they are not really addressed in this study, and so either it might make sense to leave it out, or to discuss the particulars as they apply to this study, ie: the importance and questions related to C:N ratios.
+}
 
-1-12. Lines 77-80: I think a stronger argument for why deep active-layer soils can be made, and it would be helpful to clarify what are the “strong effects” of warming.
+That's a good point–thank you. We will qualify this sentence, better tying it to the specific measurements made in this study (C, N, respiration rates, stress response).
 
-1-13.
+{\it 1-12. Lines 77-80: I think a stronger argument for why deep active-layer soils can be made, and it would be helpful to clarify what are the “strong effects” of warming. }
+
+TODO
+
+{\it 1-13.
 I cannot tell if C:N, %C and %N were measured at the end of the incubation.
@@
 Otherwise the methods section appears to be detailed and well written.
+}
 
-1-14. Line 232: In this section it would also be interesting to know the soil respiration decay rate per treatment over the course of the incubation experiment.
-Line 238-240: Confusingly worded sentence.
+C and N were measured only in the 'extra' group (l.
+128-129).
+This will be clarified in lines 162-, which we agree were a bit ambiguous.
+A table summarizing the moisture content, pH, soil carbon and N, and bulk density data is an excellent idea and will be added.
 
-1-15.
+{\it 1-14. Line 232: In this section it would also be interesting to know the soil respiration decay rate per treatment over the course of the incubation experiment. }
+
+This will be included in the new table (see comment 1-13 above).
+
+{\it 1-15. Line 238-240: Confusingly worded sentence. }
+
+Agreed.
+This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X% exhibited little relationship between water content and CO2 flux (data not shown)."
+
+{\it 1-16.
 I don’t think that the summary of nearby respiration studies add very much to the discussion section.
@@
 Perhaps comparing with other boreal incubations (eg: Lee et al., 2012; Lavoie et al., 2011) would help to provide some additional context.
+}
 
-1-16. Line 270: There is missing punctuation after the word “results”.
+TODO
 
-1-17. Line 286-293: Perhaps the new synthesis by Schadel et al., 2016, would also be a useful comparison here.
+{\it 1-17. Line 270: There is missing punctuation after the word “results”. }
 
-1-18.
+This will be fixed.
+
+{\it 1-18. Line 286-293: Perhaps the new synthesis by Schadel et al., 2016, would also be a useful comparison here. }
+
+This point was also made by Referee #2.
+The fact that we didn't cite the Schädel et al.
+(2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
+In our revision, we will significantly expand this paragraph, discussing and comparing to Schädel et al.
+(2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
+
+{\it 1-19.
 Line 293: That soil moisture may be as important a control on microbial respiration as temperature is an important finding in recent incubation studies, and the potential to define its interaction with temperature will help modelers of soil decomposition better constrain the physical parameters of microbial respiration rates.
 This feels buried in the manuscript, and I think that it would improve the paper if it were highlighted better throughout the text.
+}
 
-1-19.
+TODO
+
+{\it 1-20.
 Line 311-317: This section could be better explained in the context of the discussion or omitted altogether.
@@
 Could low temperature sensitivity be the result of low C quality in this deep soil environment?
+}
 
-1-20.
+TODO
+
+{\it 1-21.
 Line 322-332: This section, which lays out the crux of the paper, the interaction between temperature and moisture sensitivity in driving microbial respiration is relatively vague.
@@
 Perhaps DOC becomes more limiting in the drought conditions?
+}
 
-1-21.
+TODO
+
+{\it 1-22.
 Line 356: The Janssens et al., 2010, citation refers to a meta-analysis of temperate forest soils that are not nitrogen limited.
@@
 Is this site considered to be N limited in the deep active layer?
+}
 
-1-22. Line 367: Is this comparison, with North American soils, relevant to this study?
+TODO
 
-1-23. Line 383-384: Can you be more explicit in your meaning here? How do you mean that there is weakness in what can be inferred about temperature sensitivity from experiments?
+{\it 1-23. Line 367: Is this comparison, with North American soils, relevant to this study? }
 
-## Referee #2
+It's true that Colman and Schimel (2014) include only a few studies that could be termed boreal (from Maine, USA). We will either remove or de-emphasize/qualify this comparison.
 
-2-1.
-This paper studies carbon (C) release in the form of CO2 and CH4 from deep-active layer boreal soils using 30 soil cores in interior Alaska.
-After 100 days of incubation at two different temperatures and different moisture conditions, the authors find that incubation temperature, core water content and soil nitrogen strongly influenced CO2 release.
+{\it 1-24. Line 383-384: Can you be more explicit in your meaning here? How do you mean that there is weakness in what can be inferred about temperature sensitivity from experiments? }
+
+TODO
+
+{\bf Response to Referee #2}
+
+{\it 2-1.
+This paper studies...
 This is a useful dataset that adds to the growing body of literature on C release from permafrost once thawed.
-There is a good amount of publications that deal with warmer temperatures in the Arctic and changing soil moisture condition but not many have simulated drought conditions.
-A probably unfortunate overlap in publication date with this paper is a recent meta-analysis published by Schädel et al.
-2016, which showed that C release doubles when incubation temperature increases by 10◦C and that C release under aerobic conditions is three times higher than under anaerobic conditions.
+There is a good amount of publications that deal with warmer temperatures in the Arctic and changing soil moisture condition but not many have simulated drought conditions...
+}
 
-2-2.
+We appreciate the referee's assessment of the utility and interest of this dataset and analysis–thank you.
+
+{\it 2-2.
 It seems like this paper was published as a discussion paper before Schädel et al.
 2016 was published and hence a discussion of the meta-analysis was not possible but should be addressed in the revisions.
+}
 
-2-3.
+This point was also made by Referee #1.
+Yes, the fact that we didn't cite the Schädel et al.
+(2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
+In our revision, we will significantly expand the discussion on this point, comparing our results to Schädel et al.
+(2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
+
+{\it 2-3.
 The importance of the results would be more obvious if the discussion also contained an upscaling or circumpolar aspect of drought in the Arctic.
@@
 This is important as changes in temperature will affect most of the Arctic, whereas drought effects or dry soils will occur more locally.
+}
 
-2-4.
+TODO
+
+{\it 2-4.
 1) Throughout the manuscript, I have noticed that important papers from the permafrost literature are missing.
@@
 2014.
+}
 
-2-5. 2) L. 31: Permafrost thaws and does not melt
+TODO
 
-2-6.
+{\it 2-5. 2) L. 31: Permafrost thaws and does not melt }
+
+Thanks; this wording will be fixed.
+
+{\it 2-6.
 3) A better explanation is needed why deep-active layer soils are different to active layer or permafrost soils, I couldn’t find a strong argument for why they would behave differently.
 Also, deep-active layer soils are those that are the most impacted by inter annual variability in thaw depth and so they might switch between active layer in one year to permafrost in another, that’s worth some discussion as well.
+}
 
-2-7.
+TODO
+
+{\it 2-7.
 4) The statistics in this paper are generally good and I would like to compliment the authors on making the entire data set and analysis available online.
 I would still suggest that the manuscript would profit from some additional details on collinearity of the tested variables as well as model outputs such as AIC.
+}
 
-2-8. 5) Add a table with soil properties such as bulk density, %C etc.
+TODO
 
-2-9.
+{\it 2-8. 5) Add a table with soil properties such as bulk density, %C etc. }
+
+This useful suggestion was also made by Referee #1. We will do so.
+
+{\it 2-9.
 6) Why not include C/N as a variable in the statistical analysis?
@@
 2014 showed that C/N is a good predictor of C release and can be used as a scaling factor.
-It would be interesting to see if C release from short-term incubations show the same result
+It would be interesting to see if C release from short-term incubations show the same result }
 
-2-10.
-7) In the discussion, it would be good to also include the warming potential of CO2 and CH4 especially when making assumptions about the permafrost C feedback, it is briefly mentioned in line 348 but a more in depth discussion would be good
+TODO
 
-2-11. 8) the conclusions might be a bit strong given the data and previous results published
+{\it 2-10.
+7) In the discussion, it would be good to also include the warming potential of CO2 and CH4 especially when making assumptions about the permafrost C feedback, it is briefly mentioned in line 348 but a more in depth discussion would be good }
+
+That's a very good point–thank you–and will integrate well with an expanded comparison to the Schädel et al. (2016) paper (cf. 2-2 above) and other publications (2-4 above).
+
+{\it 2-11. 8) the conclusions might be a bit strong given the data and previous results published }
+
+TODO

### 2016-09-27 manuscript/response_to_reviewer_comments.txt
Tweaks during a conference call  :)

@@
 # Response to referee comments
-# Note that using LaTeX command below, as that's what Biogeosciences uses
+# Note LaTeX command below, as that's what Biogeosciences uses; see
 # http://www.biogeosciences.net/peer_review/commenting_on_a_paper.html
@@
 
-We thank both referees for their thoughtful comments, which we have numbered below for convenient referencing. Line numbers refer to the discussion manuscript.
+We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
+Per the Biogeosciences instructions, we have not prepared a revised manuscript, but instead respond below to all comments and questions.
 

### 2016-09-28 manuscript/response_to_reviewer_comments.txt
Done with Reviewer 2 (except one comment)

@@
 We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
-Per the Biogeosciences instructions, we have not prepared a revised manuscript, but instead respond below to all comments and questions.
+Per the {\it Biogeosciences} email instructions, we have not prepared a revised manuscript, but instead respond below to all comments and questions.
 
@@
 
-This point was also made by Referee #2.
+This point was also made by Referee #2 (comment 2-2).
 The fact that we didn't cite the Schädel et al.
@@
 
+We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
+Per the {\it Biogeosciences} email instructions, we have not prepared a revised manuscript, but instead respond below to all comments and questions.
+
 {\bf Response to Referee #2}
@@
 
-This point was also made by Referee #1.
+This point was also made by Referee #1 (comment 1-18).
 Yes, the fact that we didn't cite the Schädel et al.
@@
 
-TODO
+We will add this to the discussion, noting in particular, for example, observable anthropogenic influences on high-latitude precipitation (Wan et al.
+2015); drier and warmer conditions in boreal Eurasia (Buermann et al.
+2015); that growing season length in interior Alaska has increased by 45\% in the last century with no increase in precipitation (Wendler and Shulski 2009); and the complex and still-unresolved dynamics at play, meaning that, as the referee notes, drying and drought will probably be highly local in effect (Zhang et al.
+2012).
 
@@
 
-TODO
+We appreciate the referee drawing our attention to these omissions.
+Schuur et al.
+(2015) is already cited, Hugelius (2014) and Koven (2015; though we do cite his 2011 paper) are useful additions.
+The Lavoie paper is very useful with respect to N and microbial respiration, while Dutta et al.
+(2006), although it concerns Siberian soils, is also a good comparison.
+We had not included Schädel et al.
+(2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it would also be a good addition.
 
@@
 
-Thanks; this wording will be fixed.
+This wording will be fixed.
 
@@
 
-TODO
+This is a good point. We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils. For example:
+
+\begin{itemize}
+  \item Decomposition rates swing wildly between the frozen and unfrozen states of soil, due to the phase changes involved, microbial community dynamics, etc.
+Q10 respiratory coefficients, for example, are extremely high near the freeze-thaw boundary (Monson et al.
+2006).
+  \item Deep active-layer soils are subject to distinctive freeze-thaw and cryoturbation dynamics, which are likely to change their sensitivity to climate change and feedback potential.
+  \item The depth of thaw can change the source/sink status of boreal ecosystems (Goulden et al. 1998) and potentially entire biomes (Harden et al. 2012, Hayes et al. 2011).
+  \item Seasonally frozen soils pose particular problems for accurate modeling of high-latitude carbon dynamics (Nicolsky et al. 2007, Bao et al. 2016).
+  \item Deep soil carbon is likely to be a highly important contributor to future climate feedbacks.
+The modeling studies of Koven et al.
+(2011, 2015), for example, found that one-third of 21st century climate-induced carbon loss originated from seasonally frozen soil carbon.
+\end{itemize}
 
@@
 
-TODO
+Thank you. We appreciate the useful suggestions, and will provide these additional details in our revised manuscript.
 
-{\it 2-8. 5) Add a table with soil properties such as bulk density, %C etc. }
+{\it 2-8. 5) Add a table with soil properties such as bulk density, \%C etc. }
 
-This useful suggestion was also made by Referee #1. We will do so.
+This useful suggestion was also made by Referee #1 (comment 1-13). We will do so.
 
@@
 
-TODO
+We will add a sentence of caveats, noting in particular the useful but incremental nature of this study, given previous work and limitations of the experimental design.
+
+{\bf New references cited:}
+
+Bao, H., T.
+Koike, K.
+Yang, L.
+Wang, M.
+Shrestha and P.
+Lawford (2016).
+"Development of an enthalpy-based frozen soil model and its validation in a cold region in China." Journal of Geophysical Research-Atmospheres 121(10): 5259-5280.
+
+Buermann, W., B.
+Parida, M.
+Jung, G.
+M.
+MacDonald, C.
+J.
+Tucker and M.
+Reichstein (2014).
+"Recent shift in Eurasian boreal forest greening response may be associated with warmer and drier summers." Geophysical Research Letters in press.
+
+Goulden, M.
+L., S.
+C.
+Wofsy, J.
+W.
+Harden, S.
+E.
+Trumbore, P.
+M.
+Crill, S.
+T.
+Gower, T.
+Fries, B.
+C.
+Daube, S.-M.
+Fan, D.
+J.
+Sutton, A.
+M.
+Bazzaz and J.
+W.
+Munger (1998).
+"Sensitivity of boreal forest carbon balance to soil thaw." Science 279: 214-217.
+
+Harden, J.
+W., C.
+D.
+Koven, C.-L.
+Ping, G.
+Hugelius, A.
+D.
+McGuire, P.
+Camill, T.
+Jorgenson, P.
+Kuhry, G.
+J.
+Michaelson, J.
+A.
+O'Donnell, E.
+A.
+G.
+Schuur, C.
+Tarnocai, K.
+D.
+Johnson and G.
+Grosse (2012).
+"Field information links permafrost carbon to physical vulnerabilities of thawing." Geophysical Research Letters 39: L15704.
+
+Hayes, D.
+J., A.
+D.
+McGuire, D.
+W.
+Kicklighter, K.
+R.
+Gurney, T.
+J.
+Burnside and J.
+M.
+Melillo (2011).
+"Is the northern high-latitude land-based CO2 sink weakening?" Global Biochemical Cycles 25: GB3018.
+
+Koven, C.
+D., B.
+Ringeval, P.
+Friedlingstein, P.
+Ciais, P.
+Cadule, D.
+Khvorostyanov, G.
+Krinner and C.
+Tarnocai (2011).
+"Permafrost carbon-climate feedbacks accelerate global warming." Proceedings of the National Academy of Science 108(36): 14769-14774.
+
+Koven, C.
+D., D.
+M.
+Lawrence and W.
+J.
+Riley (2015).
+"Permafrost carbon−climate feedback is sensitive to deep soil carbon decomposability but not deep soil nitrogen dynamics." Proceedings of the National Academy of Science 112(12): 3752-3757.
+
+Monson, R.
+K., D.
+L.
+Lipson, S.
+P.
+Burns, A.
+A.
+Turnipseed, A.
+C.
+Delany, M.
+W.
+WIlliams and S.
+K.
+Schmidt (2006).
+"Winter forest soil respiration controlled by climate and microbial community composition." Nature 439: 711-714.
+
+Nicolsky, D. J., V. E. Romanovsky, V. A. Alexeev and D. M. Lawrence (2007). "Improved modeling of permafrost dynamics in a GCM land-surface scheme." Geophysical Research Letters 34: L08501.
+
+Wan, H., X. Zhang, F. W. Zwiers and S.-K. Min (2015). "Attributing northern high-latitude precipitation change over the period 1966–2005 to human influence." Climate Dynamics 45(7): 1713-1726.
+
+Wendler, G. and M. Shulski (2009). "A century of climate change for Fairbanks, Alaska." Arctic 62(3): 295-300.
+
+Zhang, X., J.
+He, J.
+Zhang, I.
+Polyakov, R.
+Gerdes, J.
+Inoue and P.
+Wu (2012).
+"Enhanced poleward moisture transport and amplified northern high-latitude wetting trend." Nature Climate Change 3: 47-51.

### 2016-09-28 manuscript/response_to_reviewer_comments.txt
Largely done with Reviewer 1 responses

@@
 
-We agree that the methods and concepts are not particularly novel (although drought incubations are fairly unusual), and that this is an understudied system and timely–thank you.
+We appreciate this comment, and agree that the methods and concepts are not particularly novel–although drought incubations are fairly unusual–and that this is an understudied system and timely.
 
@@
 
-TODO
+This point was raised, indirectly, by Referee #2 (comment 2-4).
+We will do so, for example comparing with studies such as the Schädel et al.
+(2016) meta-analysis (and references therein), Dutta et al.
+(2006), Lavoie et al.
+(2011), Karhu et al.
+(2010), and Czimczik et al.
+(2006, 2010).
 
@@
 
-TODO
+We will supplement this section with a better comparison to, for example, Lavoie et al.
+(2011), who incubated upland Alaskan boreal soils at two temperatures and two N addition levels; Sistla et al.
+(2012); and the combination meta-analysis and modeling study of Bouskill et al.
+(2014) that at N addition at high latitudes.
 
@@
 
-TODO
+This is similar to the referee's comment 1-9 (please see our response to that below), with the added factor of C quality.
+This idea, and studies focusing on it with respect to mineralization, has a long history (e.g.
+Nadelhoffer al.
+1991) but has been the subject of debate (e.g.
+Kleber 2010).
+We will address these questions more explicitly in our revision.
+HELP
 
@@
 
-TODO
+The first question is similar to a point raised by Referee #2 (comment 2-3).
+We will discuss this in more depth, noting in particular, for example, observable anthropogenic influences on high-latitude precipitation (Wan et al.
+2015); drier and warmer conditions in boreal Eurasia (Buermann et al.
+2015); that growing season length in interior Alaska has increased by 45\% in the last century with no increase in precipitation (Wendler and Shulski 2009); and the complex and still-unresolved dynamics at play, meaning that, as the referee notes, drying and drought will probably be highly local in effect (Zhang et al.
+2012).
+
+The second question largely repeats, we think, both referees' suggestions to better compare our results to previous work, in particular boreal and Arctic incubations; see our responses to 1-3 and 1-4 above.
 
@@
 
-TODO
+Referee #2 raised this point as well (comment 2-6), and it's a good point.
+We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
+For example:
+
+\begin{itemize}
+  \item Decomposition rates swing wildly between the frozen and unfrozen states of soil, due to the phase changes involved, microbial community dynamics, etc.
+Q10 respiratory coefficients, for example, are extremely high near the freeze-thaw boundary (Monson et al.
+2006).
+  \item Deep active-layer soils are subject to distinctive freeze-thaw and cryoturbation dynamics, which are likely to change their sensitivity to climate change and feedback potential.
+  \item The depth of thaw can change the source/sink status of boreal ecosystems (Goulden et al. 1998) and potentially entire biomes (Harden et al. 2012, Hayes et al. 2011).
+  \item Seasonally frozen soils pose particular problems for accurate modeling of high-latitude carbon dynamics (Nicolsky et al. 2007, Bao et al. 2016).
+  \item Deep soil carbon is likely to be a highly important contributor to future climate feedbacks.
+The modeling studies of Koven et al.
+(2011, 2015), for example, found that one-third of 21st century climate-induced carbon loss originated from seasonally frozen soil carbon.
+\end{itemize}
 
@@
 
-TODO
+Vegetation and ecosystem type is a significant factor that is not well explored here, we agree.
+We will add some points about this, noting for example the work of Högberg et al.
+(2007) on soil microbial community composition, the relationship of vegetation type and soil respiration (Raich and Schlesinger 1992, Bond-Lamberty and Thomson 2010), and how plant functional types affect ecosystem response more generally (Chapin et al.
+1996).
 
@@
 
-TODO
+We agree, although this largely echoes comments 1-7 and 1-9; please see our responses above.
 
@@
 
-TODO
+Agreed. See our response to comment 1-8 above.
 
 {\it 1-13.
-I cannot tell if C:N, %C and %N were measured at the end of the incubation.
+I cannot tell if C:N, \%C and \%N were measured at the end of the incubation.
 Could these results be collated in a table in the manuscript?
@@
 
-C and N were measured only in the 'extra' group (l.
-128-129).
-This will be clarified in lines 162-, which we agree were a bit ambiguous.
-A table summarizing the moisture content, pH, soil carbon and N, and bulk density data is an excellent idea and will be added.
+C and N were measured for all samples post-incubation, and in the 'extra' group (l.
+128-129) pre-incubation.
+This will be clarified in the methods, particularly lines 162-, which we agree were a bit ambiguous.
+A table summarizing the moisture content, pH, soil C and N, and bulk density data is an excellent idea and will be added.
 
@@
 Agreed.
-This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X% exhibited little relationship between water content and CO2 flux (data not shown)."
+This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)." TODO
 
@@
 
-TODO
+We agree that the comparison to other boreal respiration studies needs improvement, and this echoes Reviewer 2's comment 2-4.
+We will restructure and improve this summary and comparison section.
+The Lavoie (2011) paper is very useful with respect to N and microbial respiration, while Dutta et al.
+(2006), although it concerns Siberian soils, may also prove a useful comparison.
+We had not included Schädel et al.
+(2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but it would also be a good addition.
 
@@
 
-TODO
+Thanks for the useful suggestion; we will do so.
 
@@
 
-TODO
+That is an interesting and useful thought. In our revision, we will explore this idea... HELP
 
@@
 
-TODO
+HELP
 
@@
 
-TODO
+We agree that referring and comparing to studies such as Lavoie et al.
+(2011) and Bouskill et al.
+(2014), which focused specifically on high latitudes, would be a useful addition.
+We don't know of any studies examining the N limitation of deep soils at this site, but Allison et al.
+(2008, 2011) used fertilization and incubation experiments to examine how Alaskan boreal biota (plants and microbes) responded to N addition.
+We will discuss these results in our revision, along with other studies examining the relationship between N availability and C mineralization.
 
@@
 
-TODO
+We basically meant what the title of the Podrebarac et al.
+(2016) paper says: "Soils isolated during incubation underestimate temperature sensitivity of respiration and its response to climate history".
+I.e., incubation soils are isolated from their natural environment, and as a result we need to be cautious about extrapolating incubation results to {\it in situ} responses.
+We will clarify this in the text.
+
+{\bf New references cited:}
+
+Allison, S. D., C. I. Czimczik and K. K. Treseder (2008). "Microbial activity and soil respiration under nitrogen addition in Alaskan boreal forest." Global Change Biology 14(5): 1156-1168.
+
+Allison, S.
+D., T.
+B.
+Gartner, M.
+C.
+Mack, K.
+McGuire and K.
+K.
+Treseder (2011).
+"Nitrogen alters carbon dynamics during early succession in boreal forest." Soil Biology and Biochemistry 42: 1157-1164.
+
+Bond-Lamberty, B. and A. M. Thomson (2010). "A global database of soil respiration data." Biogeosciences 7: 1915-1926.
+
+Bouskill, N.
+J., W.
+J.
+Riley and J.
+Y.
+Tang (2014).
+"Meta-analysis of high-latitude nitrogen-addition and warming studies implies ecological mechanisms overlooked by land models." Biogeosciences 11: 6969-6983.
+
+Chapin III, F.
+S., M.
+S.
+Bret-Harte, S.
+E.
+Hobbie and H.
+Zhong (1996).
+"Plant functional types as predictors of transient responses of arctic vegetation to global change." Journal of Vegetation Science 7: 347-358.
+
+Czimczik, C. I., S. E. Trumbore, M. S. Carbone and G. C. Winston (2006). "Changing sources of soil respiration with time since fire in a boreal forest." Global Change Biology 12(6): 957-971.
+
+Czimczik, C. I. and J. M. Welker (2010). "Radiocarbon content of CO2 respired from high Arctic tundra in northwest Greenland." Arctic, Antarctic, and Alpine Research 42(3): 342-350.
+
+Dutta, K., E. A. G. Schuur, J. C. Neff and S. A. Zimov (2006). "Potential carbon release from permafrost soils of Northeastern Siberia." Global Change Biology 12: 2336-2351.
+
+Högberg, M. N., P. Högberg and D. D. Myrold (2007). "Is microbial community composition in boreal forest soils determined by pH, C-to-N ratio, the trees, or all three?" Oecologia 150(4): 590-601.
+
+Karhu, K., H.
+Fritze, K.
+Hämäläinen, P.
+Vanhala, H.
+Jungner, M.
+Oinonen, E.
+Sonninen, M.
+Tuomi, P.
+Spetz, V.
+Kitunen and J.
+Liski (2010).
+"Temperature sensitivity of soil carbon fractions in boreal forest soil." Ecology 91(2): 370-376.
+
+Kleber, M. (2010). "What is recalcitrant soil organic matter?" Environmental Chemistry 7(4): 320-332.
+
+Lavoie, M., M.
+C.
+Mack and E.
+A.
+G.
+Schuur (2011).
+"Effects of elevated nitrogen and temperature on carbon and nitrogen dynamics in Alaskan arctic and boreal soils." Journal of Geophysical Research-Biogeosciences 116: G03013.
+
+Nadelhoffer, K. J., A. E. Giblin, G. R. Shaver and J. A. Laundre (1991). "Effects of temperature and substrate quality on element mineralization in six arctic soils." Ecology 72(1): 242-253.
+
+Raich, J. W. and W. H. Schlesinger (1992). "The global carbon dioxide flux in soil respiration and its relationship to vegetation and climate." Tellus B 44(2): 81-99.
 

### 2016-09-30 R/5-analyze.R
Added analysis re changing T/SM effects with time

@@ -6,11 +6,12 @@ source("R/0-functions.R")
 
 SCRIPTNAME  	<- "5-analyze.R"
 
-library(nlme)          # 3.1-122
-#library(MASS)          # 7.3-45
+library(nlme)       # 3.1-128
+#library(MASS)       # 7.3-45
 
-library(broom)  # 0.4.0
-library(reshape2) # 1.4.1
+library(broom)      # 0.4.1
+library(reshape2)   # 1.4.1
+library(agricolae)  # v1.2.4
 
 # ==============================================================================
 # Main 
@@ -266,7 +267,7 @@ figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC_gC, fill = 
 save_plot("figureD", figureD)
 
 # Make the labels for this figure. Kind of a PITA
-library(agricolae)  # v1.2.3
+# Uses `HSD.test` in the `agricolae` package
 fluxdata_cumulative %>% 
   group_by(Gas, Treatment, Temperature) %>% 
   summarise(cum_flux_mgC_gC = max(cum_flux_mgC_gC + cum_flux_mgC_gC_sd) * 1.1) %>%
@@ -373,10 +374,17 @@ printlog("Fitting CO2 model...")
 # Add a small value to all flux values to ensure they're positive before log-transform
 fluxdata$flux_µgC_gC_day1 <- fluxdata$flux_µgC_gC_day + FLUX_ADDITION
 
+# Reviewer 1 had an interesting suggestion: see if moisture effects
+# change over time (specifically, the reviewer said, "Studies have shown that 
+# moisture can have a weaker effect on temperature sensitivity early on during 
+# an incubation experiment, in the presence of more labile C.")
+# Test this by looking at different thirds of the data
+fluxdata$third <- cut(fluxdata$incday, breaks = 3)
+
 # C_percent and N_percent are *highly* correlated (r=0.98+)
 # Generally it seems that N_percent produces slightly better model fits,
 # so we use it here, not considering C_percent further
-m_co2_lme <- lme(log(flux_µgC_gC_day1) ~ Temperature * WC_gravimetric + 
+m_co2_lme <- lme(log(flux_µgC_gC_day1) ~ Temperature * WC_gravimetric +
                    (Temperature + WC_gravimetric) * (N_percent + DOC_mg_kg), # C_percent + 
                  data = fluxdata,
                  subset = Gas == "CO2",
[…]

### 2016-09-30 R/5-analyze.R
Added analysis re C/N ratio per Reviewer 2

@@ -68,27 +68,33 @@ read_csv(CNDATA_FILE) %>%
   select(DOC_mg_kg,	C_percent, N_percent, Core) %>%
   group_by(Core) %>%
   # Summarise by core, averaging duplicates
-  summarise_each(funs(mean(., na.rm = TRUE)), DOC_mg_kg, C_percent, N_percent) ->
+  summarise_each(funs(mean(., na.rm = TRUE)), DOC_mg_kg, C_percent, N_percent) %>%
+  # Compute C:N ratio, following Reviewer 2's suggestion
+  mutate(CN = ifelse(N_percent > 0, C_percent / N_percent, NA)) ->
   cndata_orig
 
 # Compute correlations between the variables
-cndata_cor <- cor(cndata_orig[-1])
+cndata_cor <- cor(cndata_orig[-1], use = "complete.obs")
 
 cndata_orig %>%
   left_join(read_csv(TREATMENTS, skip = 1), by = "Core") ->
   cndata
 
 cndata %>%
-  summarise_each(funs(mean, sd), DOC_mg_kg, C_percent, N_percent) %>%
-  print %>%
-  unlist(.[1,]) ->
+  summarise_each(funs(mean, sd), DOC_mg_kg, C_percent, N_percent, CN) %>%
+  print ->
   cndata_summary
+# This isn't working via dplyr - compute manually
+cndata_summary$CN_mean <- mean(cndata$CN, na.rm = TRUE)
+cndata_summary$CN_sd <- sd(cndata$CN, na.rm = TRUE)
 
 lm(formula = DOC_mg_kg ~ Treatment * Temperature, data = cndata) %>%
   anova %>% print
 lm(formula = C_percent ~ Treatment * Temperature, data = cndata) %>%
   anova %>% print
 lm(formula = N_percent ~ Treatment * Temperature, data = cndata) %>%
+  anova %>% print
+lm(formula = CN ~ Treatment * Temperature, data = cndata) %>%
   anova %>% print
 
 fluxdata %>%
@@ -110,7 +116,7 @@ fluxdata %>%
 save_data(fluxdata, fn = FLUXDATA_FINAL_FILE, scriptfolder = FALSE)
 
 # -----------------------------------------------------------------------------
-# Summarize flux information for each access by markdown code later
+# Summarize flux information for easy access by markdown code later
 
 fluxdata %>%
   group_by(Gas) %>%
@@ -365,7 +371,7 @@ fluxdata %>%
 save_data(shapiro_trans)
 
[…]

### 2016-09-30 manuscript/response_to_reviewer_comments.txt
Tweaks before sending to APS and VLB
Finished last TODOs in reviewer responses

@@
 
-TODO
+This is an interesting suggestion, and an analysis we would be happy to include.
+As a preliminary step, before revising the manuscript, we have added code (see https://github.com/bpbond/cpcrw_incubation/commit/18884c836d855b94417f7ae2cef910c577bc40b) that examines if and how temperature and moisture sensitivities change over incubation time.
+Currently we do not any evidence for changes in CO2 moisture sensitivity with time, and weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time.
+We will discuss this issue, referencing previous work such as Reichstein et al.
+(2005).
 
@@
 
-Referee #2 raised this point as well (comment 2-6), and it's a good point.
+Referee #2 raised this point as well (comment 2-6), and it's a good one.
 We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
@@
 Agreed.
-This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)." TODO
+This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)."
 
@@
 
+Reichstein, M., J.-A.
+Subke, A.
+C.
+Angeli and J.
+D.
+Tenhunen (2005).
+"Does the temperature sensitivity of decomposition of soil organic matter depend upon water content, soil horizon, or incubation time?" Global Change Biology 11(10): 1754-1767.
+
 We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
@@
 We appreciate the referee drawing our attention to these omissions.
-Schuur et al.
+While Schuur et al.
 (2015) is already cited, Hugelius (2014) and Koven (2015; though we do cite his 2011 paper) are useful additions.
@@
 
-TODO
+This is an interesting suggestion, and an analysis we would be happy to include.
+As a preliminary step, before revising the manuscript, we have added code (see https://github.com/bpbond/cpcrw_incubation/commit/426a91e1bbd21200718b334d3295fbef40a1ea6) to compute C/N and examines its significance as a predictor.
+Currently C/N seems to be a poorer predictor than %N.
+We will discuss this issue, referencing previous work such as Schädel et al (2014).
 
@@
 
-That's a very good point–thank you–and will integrate well with an expanded comparison to the Schädel et al. (2016) paper (cf. 2-2 above) and other publications (2-4 above).
+That's a very good point–thank you–and will integrate well with an expanded comparison to the Schädel et al. (2016) paper (cf. comment 2-2 above) and other publications (comment 2-4 above).
 

### 2016-09-30 manuscript/response_to_reviewer_comments.txt
Finishing up response to reviewers

@@
 We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
-Per the {\it Biogeosciences} email instructions, we have not prepared a revised manuscript, but instead respond below to all comments and questions.
+Per the {\it Biogeosciences} email instructions, we have not yet prepared a revised manuscript, but instead respond below to all comments and questions.
 
@@
 
-This point was raised, indirectly, by Referee #2 (comment 2-4).
-We will do so, for example comparing with studies such as the Schädel et al.
+This point was raised by Referee #2 as well (comment 2-4).
+We will do so, comparing for example with studies such as the Schädel et al.
 (2016) meta-analysis (and references therein), Dutta et al.
@@
 Nadelhoffer al.
-1991) but has been the subject of debate (e.g.
-Kleber 2010).
-We will address these questions more explicitly in our revision.
-HELP
+1991); as the referee notes, from fundamental biokinetics (Davidson and Janssens 2006) we expect that decomposition Q10 should be inversely related to litter carbon quality.
+We will address these questions more explicitly in our revision, referring for example to incubation studies on this question (Fierer et al.
+2005).
 
@@
 This is an interesting suggestion, and an analysis we would be happy to include.
-As a preliminary step, before revising the manuscript, we have added code (see https://github.com/bpbond/cpcrw_incubation/commit/18884c836d855b94417f7ae2cef910c577bc40b) that examines if and how temperature and moisture sensitivities change over incubation time.
-Currently we do not any evidence for changes in CO2 moisture sensitivity with time, and weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time.
+As a preliminary step, before revising the manuscript, we have added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/18884c836d855b94417f7ae2cef910c577bc40b}) that examines if and how temperature and moisture sensitivities change over incubation time.
+Currently we do not observe any evidence for changes in CO2 moisture sensitivity with time, and only weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time.
 We will discuss this issue, referencing previous work such as Reichstein et al.
@@
 128-129) pre-incubation.
-This will be clarified in the methods, particularly lines 162-, which we agree were a bit ambiguous.
+This will be clarified in the methods, particularly lines 162-, which we agree were ambiguous.
 A table summarizing the moisture content, pH, soil C and N, and bulk density data is an excellent idea and will be added.
+Reviewer 2 also raised the idea of looking at C/N, and we have responded to it in detail (see comment 2-9).
 
@@
 
-We agree that the comparison to other boreal respiration studies needs improvement, and this echoes Reviewer 2's comment 2-4.
+We agree that the comparison to other boreal respiration studies needs improvement, and this echoes Referee #2's comment 2-4.
 We will restructure and improve this summary and comparison section.
@@
 
-That is an interesting and useful thought. In our revision, we will explore this idea... HELP
+We appreciate this useful advice and question.
+Certainly temperature sensitivity is, at least in theory, related to C quality (see our answer to 1-5 above), although low-quality litter (i.e.
+in a boreal coniferous forest) might be expected to be associated with {/it higher} Q10?
+In our revision, we will explore these ideas, considering how how factors such as vegetation, soil characteristics, and climate might explain the low observed temperature sensitivity.
+HELP
 
@@
 
+This is indeed interesting to consider: what mechanisms might produce a Q10 {/it increase} under drought conditions?
+This is opposite to what is usually observed (e.g.
+Jassal et al.
+2008), but the field is rife with contradictory results (von Lützow and Kögel-Knabner 2009).
+We will make this paragraph more specific in all these areas.
 HELP
@@
 
-{\bf New references cited:}
+\newpage
+\begin{footnotesize}
+{\bf References cited in response:}
 
@@
 
+Davidson, E. A. and I. A. Janssens (2006). "Temperature sensitivity of soil carbon decomposition and feedbacks to climate change." Nature 440: 165-173.
+
 Dutta, K., E. A. G. Schuur, J. C. Neff and S. A. Zimov (2006). "Potential carbon release from permafrost soils of Northeastern Siberia." Global Change Biology 12: 2336-2351.
 
+Fierer, N., J. M. Craine, K. McLauchlan and J. P. Schimel (2005). "Litter quality and the temperature sensitivity of decomposition." Ecology 86(2): 320-326.
+
 Högberg, M. N., P. Högberg and D. D. Myrold (2007). "Is microbial community composition in boreal forest soils determined by pH, C-to-N ratio, the trees, or all three?" Oecologia 150(4): 590-601.
+
+Jassal, R.
+S., T.
+A.
+Black, M.
+D.
+Novak, D.
+Gaumont-Guay and Z.
+Nesic (2008).
+"Effect of soil water stress on soil respiration and its temperature sensitivity in an 18-year-old temperate Douglas-fir stand." Global Change Biology 14(6): 1305-1318.
 
@@
 
+von Lützow, M. and I. Kögel-Knabner (2009). "Temperature sensitivity of soil organic matter decomposition—what do we know?" Biology and Fertility of Soils 46(1): 1-15.
+
+\end{footnotesize}
+
 We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
-Per the {\it Biogeosciences} email instructions, we have not prepared a revised manuscript, but instead respond below to all comments and questions.
+Per the {\it Biogeosciences} email instructions, we have not yet prepared a revised manuscript, but instead respond below to all comments and questions.
 
@@
 This is an interesting suggestion, and an analysis we would be happy to include.
-As a preliminary step, before revising the manuscript, we have added code (see https://github.com/bpbond/cpcrw_incubation/commit/426a91e1bbd21200718b334d3295fbef40a1ea6) to compute C/N and examines its significance as a predictor.
-Currently C/N seems to be a poorer predictor than %N.
+As a preliminary step, before revising the manuscript, we have added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/426a91e1bbd21200718b334d3295fbef40a1ea6}) to compute C/N and examines its significance as a predictor.
+Currently C/N seems to be a poorer predictor than \%N.
 We will discuss this issue, referencing previous work such as Schädel et al (2014).
@@
 "Enhanced poleward moisture transport and amplified northern high-latitude wetting trend." Nature Climate Change 3: 47-51.
+
+\end{footnotesize}

### 2016-09-30 manuscript/cpcrw_manuscript.Rmd
Finishing up response to reviewers

@@
 
-More surprisingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 than at 20 °C.
+More surprisingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 than at 20 °C (TODO-fix sentence).
 There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}.
-Unlike surface soils, active layer soils, which store large quantities of soil C {Mueller, 2015 #4435}, are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils may be older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}.
+Unlike surface soils, deep active layer soils, which store large quantities of soil C {Mueller, 2015 #4435}, are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils may be older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}.
 Thus, it is not surprising that these more stable C compounds would be metabolized by processes that have been reported to be less temperature-sensitive.

### 2016-10-02 manuscript/response_to_reviewer_comments.txt
Final tweaks to response to reviewers

@@
 We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
-Per the {\it Biogeosciences} email instructions, we have not yet prepared a revised manuscript, but instead respond below to all comments and questions.
+Per the {\it Biogeosciences} instructions, we have not yet prepared a revised manuscript, but respond below to all comments and questions.
 
@@
 
-{\it 1-1.
+{\it 1.1.
 The authors have measured...
@@
 
-{\it 1-2.
+{\it 1.2.
 I do think that the paper has within it data that is relevant to the scope of BG, but I think that a more in depth analysis of the data, and a better contextualization of the data within the Arctic literature is required.
@@
 
-{\it 1-3. I think that the work would benefit from a more thorough comparison with boreal forest incubations across the Arctic. }
+{\it 1.3. I think that the work would benefit from a more thorough comparison with boreal forest incubations across the Arctic. }
 
-This point was raised by Referee #2 as well (comment 2-4).
+This point was raised by Referee #2 as well (comment 2.4).
 We will do so, comparing for example with studies such as the Schädel et al.
@@
 
-{\it 1-4.
+{\it 1.4.
 N section would benefit from more Arctic-centric comparisons of N limitations and in particular of boreal forest N dynamics.
@@
 
-{\it 1-5.
+{\it 1.5.
 This study raises interesting questions.
@@
 
-This is similar to the referee's comment 1-9 (please see our response to that below), with the added factor of C quality.
+This is similar to the referee's comment 1.9 (please see our response to that below), with the added factor of C quality.
 This idea, and studies focusing on it with respect to mineralization, has a long history (e.g.
@@
 
-{\it 1-6.
+{\it 1.6.
 Studies have shown that moisture can have a weaker effect on temperature sensitivity early on during an incubation experiment, in the presence of more labile C.
@@
 This is an interesting suggestion, and an analysis we would be happy to include.
-As a preliminary step, before revising the manuscript, we have added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/18884c836d855b94417f7ae2cef910c577bc40b}) that examines if and how temperature and moisture sensitivities change over incubation time.
+As a preliminary step, before revising the manuscript, we have added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/18884c836d855b94417f7ae2cef910c577bc40b}) that examines if and how temperature and moisture sensitivities change over incubation time in this experiment.
 Currently we do not observe any evidence for changes in CO2 moisture sensitivity with time, and only weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time.
@@
 
-{\it 1-7.
+{\it 1.7.
 How do your results in terms of temperature and moisture sensitivity (especially under drought conditions) scale with Alaskan climate change predictions from modelers?
@@
 
-The first question is similar to a point raised by Referee #2 (comment 2-3).
-We will discuss this in more depth, noting in particular, for example, observable anthropogenic influences on high-latitude precipitation (Wan et al.
+The first question is similar to a point raised by Referee #2 (comment 2.3).
+We will discuss this in more depth, noting in particular, for example, the fact that there are observable anthropogenic influences on high-latitude precipitation (Wan et al.
 2015); drier and warmer conditions in boreal Eurasia (Buermann et al.
-2015); that growing season length in interior Alaska has increased by 45\% in the last century with no increase in precipitation (Wendler and Shulski 2009); and the complex and still-unresolved dynamics at play, meaning that, as the referee notes, drying and drought will probably be highly local in effect (Zhang et al.
+2015), and growing season length increases in interior Alaska with no increase in precipitation (Wendler and Shulski 2009); and the complex and still-unresolved dynamics at play, meaning that, as the referee notes, drying and drought will probably be highly local in effect (Zhang et al.
 2012).
 
-The second question largely repeats, we think, both referees' suggestions to better compare our results to previous work, in particular boreal and Arctic incubations; see our responses to 1-3 and 1-4 above.
+The second question largely repeats, we think, both referees' suggestions to better compare our results to previous work, in particular boreal and Arctic incubations; see our responses to 1.3 and 1.4 above.
 
-{\it 1-8.
+{\it 1.8.
 Line 31-34: I cannot find discussion of this point in the rest of the text, and while important, this statement is relatively vague and there are no cited references.
@@
 
-Referee #2 raised this point as well (comment 2-6), and it's a good one.
+Referee #2 raised this point as well (comment 2.6), and it's a good one.
 We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
@@
 \begin{itemize}
-  \item Decomposition rates swing wildly between the frozen and unfrozen states of soil, due to the phase changes involved, microbial community dynamics, etc.
+  \item Decomposition rates vary greatly between the frozen and unfrozen states of soil, due to the phase changes involved, microbial community dynamics, etc.
 Q10 respiratory coefficients, for example, are extremely high near the freeze-thaw boundary (Monson et al.
@@
 
-{\it 1-9.
+{\it 1.9.
 Lines 48-60: I think that this section would benefit from an introduction of the interactions between the specific ecosystem (upland boreal forest) you are studying, and its interaction with soil chemistry, since vegetation type is influential in terms of soil carbon quality and quantity.
@@
 
-{\it 1-10.
+{\it 1.10.
 Lines 70-72: These are really important considerations, and it seems appropriate to discuss them more explicitly.
@@
 
-We agree, although this largely echoes comments 1-7 and 1-9; please see our responses above.
+We agree, although this largely echoes comments 1.7 and 1.9; please see our responses above.
 
-{\it 1-11.
+{\it 1.11.
 Lines 72-74: While these are important questions, they are not really addressed in this study, and so either it might make sense to leave it out, or to discuss the particulars as they apply to this study, ie: the importance and questions related to C:N ratios.
@@
 
-{\it 1-12. Lines 77-80: I think a stronger argument for why deep active-layer soils can be made, and it would be helpful to clarify what are the “strong effects” of warming. }
+{\it 1.12. Lines 77-80: I think a stronger argument for why deep active-layer soils can be made, and it would be helpful to clarify what are the “strong effects” of warming. }
 
-Agreed. See our response to comment 1-8 above.
+Agreed. See our response to comment 1.8 above.
 
-{\it 1-13.
+{\it 1.13.
 I cannot tell if C:N, \%C and \%N were measured at the end of the incubation.
@@
 A table summarizing the moisture content, pH, soil C and N, and bulk density data is an excellent idea and will be added.
-Reviewer 2 also raised the idea of looking at C/N, and we have responded to it in detail (see comment 2-9).
+Reviewer 2 also raised the idea of looking at C/N, and we have responded to that suggestion in detail (see comment 2.9).
 
-{\it 1-14. Line 232: In this section it would also be interesting to know the soil respiration decay rate per treatment over the course of the incubation experiment. }
+{\it 1.14. Line 232: In this section it would also be interesting to know the soil respiration decay rate per treatment over the course of the incubation experiment. }
 
-This will be included in the new table (see comment 1-13 above).
+This will be included in the new table (see comment 1.13 above).
 
-{\it 1-15. Line 238-240: Confusingly worded sentence. }
+{\it 1.15. Line 238-240: Confusingly worded sentence. }
 
 Agreed.
-This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)."
+This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > _\% exhibited little relationship between water content and CO2 flux (data not shown)."
 
-{\it 1-16.
+{\it 1.16.
 I don’t think that the summary of nearby respiration studies add very much to the discussion section.
@@
 
-We agree that the comparison to other boreal respiration studies needs improvement, and this echoes Referee #2's comment 2-4.
+We agree that the comparison to other boreal respiration studies needs improvement, and this echoes Referee #2's comment 2.4.
 We will restructure and improve this summary and comparison section.
@@
 
-{\it 1-17. Line 270: There is missing punctuation after the word “results”. }
+{\it 1.17. Line 270: There is missing punctuation after the word “results”. }
 
@@
 
-{\it 1-18. Line 286-293: Perhaps the new synthesis by Schadel et al., 2016, would also be a useful comparison here. }
+{\it 1.18. Line 286-293: Perhaps the new synthesis by Schadel et al., 2016, would also be a useful comparison here. }
 
-This point was also made by Referee #2 (comment 2-2).
+This point was also made by Referee #2 (comment 2.2).
 The fact that we didn't cite the Schädel et al.
@@
 
-{\it 1-19.
+{\it 1.19.
 Line 293: That soil moisture may be as important a control on microbial respiration as temperature is an important finding in recent incubation studies, and the potential to define its interaction with temperature will help modelers of soil decomposition better constrain the physical parameters of microbial respiration rates.
@@
 
-{\it 1-20.
+{\it 1.20.
 Line 311-317: This section could be better explained in the context of the discussion or omitted altogether.
@@
 We appreciate this useful advice and question.
-Certainly temperature sensitivity is, at least in theory, related to C quality (see our answer to 1-5 above), although low-quality litter (i.e.
+Certainly temperature sensitivity is, at least in theory, related to C quality (see our answer to 1.5 above), although low-quality litter (i.e.
 in a boreal coniferous forest) might be expected to be associated with {/it higher} Q10?
@@
 
-{\it 1-21.
+{\it 1.21.
 Line 322-332: This section, which lays out the crux of the paper, the interaction between temperature and moisture sensitivity in driving microbial respiration is relatively vague.
@@
 
-This is indeed interesting to consider: what mechanisms might produce a Q10 {/it increase} under drought conditions?
+This is indeed interesting to consider: what mechanisms might produce a Q10 {\it increase} under drought conditions?
 This is opposite to what is usually observed (e.g.
@@
 
-{\it 1-22.
+{\it 1.22.
 Line 356: The Janssens et al., 2010, citation refers to a meta-analysis of temperate forest soils that are not nitrogen limited.
@@
 
-{\it 1-23. Line 367: Is this comparison, with North American soils, relevant to this study? }
+{\it 1.23. Line 367: Is this comparison, with North American soils, relevant to this study? }
 
@@
 
-{\it 1-24. Line 383-384: Can you be more explicit in your meaning here? How do you mean that there is weakness in what can be inferred about temperature sensitivity from experiments? }
+{\it 1.24. Line 383-384: Can you be more explicit in your meaning here? How do you mean that there is weakness in what can be inferred about temperature sensitivity from experiments? }
 
@@
 
-von Lützow, M. and I. Kögel-Knabner (2009). "Temperature sensitivity of soil organic matter decomposition—what do we know?" Biology and Fertility of Soils 46(1): 1-15.
+von Lützow, M. and I. Kögel-Knabner (2009). "Temperature sensitivity of soil organic matter decomposition--what do we know?" Biology and Fertility of Soils 46(1): 1-15.
 
@@
 We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
-Per the {\it Biogeosciences} email instructions, we have not yet prepared a revised manuscript, but instead respond below to all comments and questions.
+Per the {\it Biogeosciences} email instructions, we have not yet prepared a revised manuscript, but respond below to all comments and questions.
 
@@
 
-{\it 2-1.
+{\it 2.1.
 This paper studies...
@@
 
-{\it 2-2.
+{\it 2.2.
 It seems like this paper was published as a discussion paper before Schädel et al.
@@
 
-This point was also made by Referee #1 (comment 1-18).
+This point was also made by Referee #1 (comment 1.18).
 Yes, the fact that we didn't cite the Schädel et al.
@@
 
-{\it 2-3.
+{\it 2.3.
 The importance of the results would be more obvious if the discussion also contained an upscaling or circumpolar aspect of drought in the Arctic.
@@
 
-We will add this to the discussion, noting in particular, for example, observable anthropogenic influences on high-latitude precipitation (Wan et al.
+We will add this to the discussion, noting in particular the fact that there are observable anthropogenic influences on high-latitude precipitation (Wan et al.
 2015); drier and warmer conditions in boreal Eurasia (Buermann et al.
-2015); that growing season length in interior Alaska has increased by 45\% in the last century with no increase in precipitation (Wendler and Shulski 2009); and the complex and still-unresolved dynamics at play, meaning that, as the referee notes, drying and drought will probably be highly local in effect (Zhang et al.
+2015), and growing season length increases in interior Alaska with no increase in precipitation (Wendler and Shulski 2009); and the complex and still-unresolved dynamics at play, meaning that, as the referee notes, drying and drought will probably be highly local in effect (Zhang et al.
 2012).
 
-{\it 2-4.
+We will discuss this in more depth, noting in particular, for example,
+
+{\it 2.4.
 1) Throughout the manuscript, I have noticed that important papers from the permafrost literature are missing.
@@
 We had not included Schädel et al.
-(2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it would also be a good addition.
+(2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it would also be a reasonable addition.
 
-{\it 2-5. 2) L. 31: Permafrost thaws and does not melt }
+{\it 2.5. 2) L. 31: Permafrost thaws and does not melt }
 
@@
 
-{\it 2-6.
+{\it 2.6.
 3) A better explanation is needed why deep-active layer soils are different to active layer or permafrost soils, I couldn’t find a strong argument for why they would behave differently.
@@
 
-{\it 2-7.
+{\it 2.7.
 4) The statistics in this paper are generally good and I would like to compliment the authors on making the entire data set and analysis available online.
@@
 
-{\it 2-8. 5) Add a table with soil properties such as bulk density, \%C etc. }
+{\it 2.8. 5) Add a table with soil properties such as bulk density, \%C etc. }
 
-This useful suggestion was also made by Referee #1 (comment 1-13). We will do so.
+This useful suggestion was also made by Referee #1 (comment 1.13). We will do so.
 
-{\it 2-9.
+{\it 2.9.
 6) Why not include C/N as a variable in the statistical analysis?
@@
 This is an interesting suggestion, and an analysis we would be happy to include.
-As a preliminary step, before revising the manuscript, we have added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/426a91e1bbd21200718b334d3295fbef40a1ea6}) to compute C/N and examines its significance as a predictor.
+As a preliminary step, before revising the manuscript, we have added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/426a91e1bbd21200718b334d3295fbef40a1ea6}) to compute C/N and examine its significance as a predictor.
 Currently C/N seems to be a poorer predictor than \%N.
@@
 
-{\it 2-10.
+{\it 2.10.
 7) In the discussion, it would be good to also include the warming potential of CO2 and CH4 especially when making assumptions about the permafrost C feedback, it is briefly mentioned in line 348 but a more in depth discussion would be good }
 
-That's a very good point–thank you–and will integrate well with an expanded comparison to the Schädel et al. (2016) paper (cf. comment 2-2 above) and other publications (comment 2-4 above).
+That's a very good point–thank you–and will integrate well with an expanded comparison to the Schädel et al. (2016) paper (cf. comment 2.2 above) and other publications (comment 2.4 above).
 
-{\it 2-11. 8) the conclusions might be a bit strong given the data and previous results published }
+{\it 2.11. 8) the conclusions might be a bit strong given the data and previous results published }
 
-We will add a sentence of caveats, noting in particular the useful but incremental nature of this study, given previous work and limitations of the experimental design.
+We will add a sentence of caveats, noting in particular the useful but incremental nature of this study.
 

### 2016-10-02 manuscript/response_to_reviewer_comments.txt
Truly final tweaks to reviewer responses
Caught a few typos.

@@
 Agreed.
-This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > _\% exhibited little relationship between water content and CO2 flux (data not shown)."
+This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)."
 
@@
 In our revision, we will explore these ideas, considering how how factors such as vegetation, soil characteristics, and climate might explain the low observed temperature sensitivity.
-HELP
 
@@
 We will make this paragraph more specific in all these areas.
-HELP
 
@@
 2012).
-
-We will discuss this in more depth, noting in particular, for example,
 

### 2016-10-08 R/5-analyze.R
Added code to test effects of %C, %N, DOC, and C/N on _cumulative_ fluxes, following Reviewer 3's suggestion

@@ -237,8 +237,9 @@ save_plot("figureC", figureC)
 
 printlog(SEPARATOR)
 printlog("Running Tukey HSD tests on cumulative emissions...")
-fd_cumulative_core <- read_csv(FLUXDATA_CUM_CORE_FILE) %>%
-  mutate(Temperature = as.factor(Temperature))
+read_csv(FLUXDATA_CUM_CORE_FILE) %>%
+  mutate(Temperature = as.factor(Temperature)) ->
+  fd_cumulative_core
 co2_aov <- aov(cum_flux_mgC_gC ~ Treatment + Temperature, 
                data = subset(fd_cumulative_core, Gas == "CO2"))
 co2_hsd <- TukeyHSD(co2_aov)
@@ -291,6 +292,31 @@ rbind(out_co2$groups, out_ch4$groups) %>%
 figureD <- figureD + geom_text(data = labeldata, 
                                aes(label = M), 
                                position = position_dodge(width = 1))
+
+# -----------------------------------------------------------------------------
+# Test effects of %C, %N, etc. on cumulative fluxes (per Referee 3 suggestion)
+
+printlog("Merging cumulative fluxes with CN data...")
+fd_cumulative_core %>%
+  left_join(cndata_orig, by = "Core") ->
+  fd_cumulative_r3
+fd_cumulative_r3 <- fd_cumulative_r3[complete.cases(fd_cumulative_r3),]
+
+printlog("Fitting CO2 model for cumulative fluxes...")
+m_co2_cum <- lm(cum_flux_mgC_gC ~ Temperature + Treatment +
+                   N_percent + DOC_mg_kg + CN, # C_percent + 
+                 data = fd_cumulative_r3, 
+                 subset = Gas == "CO2")
+step_co2_cum <- MASS::stepAIC(m_co2_cum, direction = "both")
+print(summary(m_co2_cum))
+
+printlog("Fitting CH4 model for cumulative fluxes...")
+m_ch4_cum <- lm(cum_flux_mgC_gC ~ Temperature + Treatment +
+                   N_percent + DOC_mg_kg + CN, # C_percent + 
+                 data = fd_cumulative_r3, 
+                 subset = Gas == "CH4")
+step_ch4_cum <- MASS::stepAIC(m_ch4_cum, direction = "both")
+print(summary(step_ch4_cum))
 
 # -----------------------------------------------------------------------------
 # Calculate drought reduction (%)
[…]

### 2016-10-08 manuscript/response_to_reviewer_comments.txt
Updated to include response to Referee 3
See http://www.biogeosciences-discuss.net/bg-2016-234/#discussion

@@
-# Response to referee comments
-# Note LaTeX command below, as that's what Biogeosciences uses; see
-# http://www.biogeosciences.net/peer_review/commenting_on_a_paper.html
-# BBL September 2016
+% Response to referee comments
+% Note LaTeX command below, as that's what Biogeosciences uses; see
+% http://www.biogeosciences.net/peer_review/commenting_on_a_paper.html
+% BBL September 2016
 
-We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
+% October 2016: added Reviewer 3 comments (which somehow did not show up online)
+% and editor's decision letter at end
+
+% ======================================================================================
+We thank all referees for their thoughtful comments, which we have numbered for convenient referencing.
 Per the {\it Biogeosciences} instructions, we have not yet prepared a revised manuscript, but respond below to all comments and questions.
@@
 
-We thank both referees for their thoughtful comments, which we have numbered for convenient referencing.
-Per the {\it Biogeosciences} email instructions, we have not yet prepared a revised manuscript, but respond below to all comments and questions.
+% ======================================================================================
+We thank all referees for their thoughtful comments, which we have numbered for convenient referencing.
+Per the {\it Biogeosciences} instructions, we have not yet prepared a revised manuscript, but respond below to all comments and questions.
 
@@
 \end{footnotesize}
+
+% ======================================================================================
+We thank all referees for their thoughtful comments, which we have numbered for convenient referencing.
+Per the {\it Biogeosciences} instructions, we have not yet prepared a revised manuscript, but respond below to all comments and questions.
+
+{\bf Response to Referee #3 (comments were not online before editor's decision) }
+
+{\it 3.1.
+This manuscript presents...interesting and surprising correlations between CO2 production and soil nitrogen content.
+The main point of this paper, and I think that it’s an important one, is that many upland forest soils in the boreal region that have permafrost will dry, and the majority of C will be emitted as CO2, but the magnitude of C emission will decrease under drought conditions (imposed by freely draining and no water inputs) by ~ 50-70\%.
+Overall, I enjoyed reading this paper!
+It was nicely written although vague in a few parts.
+}
+
+Thank you.
+
+{\it 3.2.
+My main criticism is that I think that the authors over-emphasize the results of the daily emissions and that the authors should further explore (or report) the results of the controls of the cumulative C emissions.
+I’m curious as to whether the relationships with soil C/N and \%N observed in daily emissions still hold on cumulative emissions.
+The comparison between these soil parameters (i.e.
+ones that probably don’t change much throughout the course of the incubation, including temperature) and the cumulative fluxes is perhaps more appropriate.
+Perhaps modelers find the controls on daily fluxes interesting and these are likely quite useful in regards to the relationship between moisture and C production (i.e.
+changes on a daily basis), but I think that the controls on cumulative fluxes are quite interesting and could be further explored.
+}
+
+This is an interesting and useful suggestion.
+We agree that rebalancing the manuscript, focusing a bit more on controls on cumulative emissions and a bit less on the instantaneous fluxes, would strengthen it.
+In our revision, we will more fully explore controls on the cumulative emissions, and as a first step have added the analytical code to do so (see \url{https://github.com/bpbond/cpcrw_incubation/commit/2d62e73ab79f7f0c8d7289919cddedb0a22ec275}).
+At the moment, it looks like only \%N exerts a weak effect on cumulative CO2 fluxes, and none of the tested covariates (\%C, \%N, C/N, DOC) affects cumulative CH4.
+
+{\it 3.3.
+For example, how do the results of soil properties vs.
+emissions compare to those of Schädel et al.
+(2014) and Schädel et al.
+(2016)?
+How do the moisture results compare to those of Wickland et al.
+(2008)?
+}
+
+The other referees both mentioned this as well.
+The fact that we didn't cite the Schädel et al.
+(2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
+In our revision, we will significantly expand this, discussing and comparing to Schädel et al.
+(2014, 2016) and Wickland et al.
+(2008).
+
+{\it 3.4.
+I do think that the time series of fluxes could be moved to the supplemental materials if the cumulative fluxes are explored in greater detail.
+I think this paper could be shortened a little bit although I didn’t find the length of the paper onerous.
+Along these lines, I think that the results summarized above from the cumulative emissions should be included in the abstract.
+}
+
+We will consider these suggestions with respect to length and supplemental materials; results dealing with cumulative emissions will certainly be included in the abstract.
+
+{\it 3.5. 22: Daily CO2 fluxes?
+26: positive or negative correlation?
+27: daily CH4 flux?
+28: cumulative production as CO2…as CH4. }
+
+We will clarify these points.
+
+{\it 3.6. 29: Not really sure how the comparison as to the relative controls of T and moisture are evaluated. }
+
+Agreed, this was more of a qualitative statement than a quantitative one. We will be more precise.
+
+{\it 3.7. 50: see also updates in Hugelius et al. (2014)
+63: Under some conditions (Olefeldt et al 2013): vague and confusing. Please clarify. }
+
+Reviewer 2 also raised the issue (comment 2.4) of our incomplete citation of relevant literature.
+In our revision, we will make better use of references such as Hugelius et al.
+(2014) and Wickland et al.
+(2008), both in the introduction and discussion.
+
+{\it 3.8. 67: “substantial variabilities between studies” WHY? }
+
+We will expand on this point, pointing out that such variability originates from factors such soil effects on drainage, phase changes, different experimental protocols, etc.
+
+{\it 3.9. 72: Yes, this is an important question, but given that this isn’t measured in this study, perhaps this sentence should be omitted or re-written. }
+
+Good point; we will do so.
+
+{\it 3.10. 101: When did sampling occur?
+112: Specify at the time of sampling
+140: How frequently was moisture adjusted? Requires a bit more explanation. Were instantaneous moisture values used in analysis? }
+
+Sampling date is reported in line 110.
+We will clarify 80 cm at the time of sampling.
+Moisture adjustment was done after every mass measurement, i.e.
+every timepoint shown in Figure 1.
+This will be clarified, and we will consider if adding this to Figures 2 and 3 if readable and useful.
+
+{\it 3.11. 211: Please remember to complete DOI }
+
+We will!
+
+{\it 3.12. 215: Not sure what this value for soil dry mass indicates }
+
+It's just useful, we think, to give readers a good sense of sample mass.
+
+{\it 3.13. 216: Standard deviation for \%C and \%N is nearly 100\%. Check values. }
+
+Thanks. There was a great of variability (obviously), but distributed throughout the data set–i.e., this isn't driven by one or two outliers. We will re-check, however.
+
+{\it 3.14. 229: add units
+231: add units
+233: positively correlated
+241: positively correlated
+245-246: 106 }
+
+These will be fixed.
+
+{\it 3.15. 253-254: So what variables were significant in predicting cumulative C emissions? }
+
+Please see our response to comment 3.2 above.
+
+{\it 3.16. 262: First mention of vegetation stress, remove, not clear how it’s related. }
+
+We will better integrate this point, mentioning it in the introduction and clarifying its relationship to the study goals.
+
+{\it 3.17. 270. Add “.” }
+
+This will be fixed.
+
+{\it 3.18. 271: Specify soil type in which these measurements were made (results not surprising for a forest soil) }
+
+Upland Cryosols; we will clarify this.
+
+{\it 3.19. 272: What about results from Wickland et al. (2008). Study found threshold for moisture importance
+305-307: again, see Wickland et al. (2008) }
+
+Please see our response to comment 3.7 above.
+
+{\it 3.20. 322-324: cool! }
+
+Agreed. Tantalizing.
+
+{\it 3.21. 344-345: Specify that the results in Treat et al. (2015) were for anaerobic incubations and were thus likely to be much smaller. }
+
+Thanks; we will do so.
+
+{\it 3.22. 347-348: See also Lee et al. (2012)
+364-365: See also Schadel et al. (2014). Also, I thought this section was a bit vague, probably could be shortened slightly. }
+
+Thanks for the Lee et al. reference, which we had not considered (see our response to 3.7 above). We'll take a hard look at lines 362-378 and tighten.
+
+{\it 3.23. 383-384: “specific weaknesses”: vague
+384: See also lag effects found in Treat et al. (2015) }
+
+Good points! We will clarify and note this.
+
+{\it 3.24. 393: “taking them out of depth” rephrase. Also could use this argument for the section on CH4 production. }
+
+Yes, it's awkward; we will reword. Thanks for the suggestion.
+
+{\it 3.25. Fig.1 : Edit figure to be color-blind friendly. }
+
+We thought we were already doing so in using the {\tt RColorBrewer} package, not the default palette of {\tt ggplot2}, but will investigate and fix this if necessary.
+
+{\it 3.26. Fig. 2,3: When did watering / moisture adjustment occur? Consider indicating with arrows and specifying in text. }
+
+Moisture adjustment was done after every mass measurement, i.e.
+every timepoint shown in Figure 1.
+This will be clarified, and we will consider if adding this to Figures 2 and 3 if readable and useful.
+
+{\it 3.27. Fig. 4: Switch top and bottom panels as CO2 is always discussed before CH4. Also edit colors and patterns to be color-blind friendly. }
+
+Good point. Re colors, see our response to 3.25 above.
+
+% ======================================================================================
+
+Associate Editor Decision: Reconsider after major revisions (06 Oct 2016) by Michael Weintraub
+Comments to the Author:
+Thank you for your submission to Biogeosciences.
+Your MS has received three reviews from highly qualified reviewers in this field, all of whom thought that this is a valuable dataset worth publishing, but offered numerous substantive comments to improve the paper.
+Thus, I will be happy to consider a revised submission that substantively addresses these reviewer comments.
+Based on the responses to reviewers 1 and 2, it looks like there is a sound plan for addressing the reviewers’ critical comments.
+
+Unfortunately, I discovered too late that the comments from the third reviewer do not show up in the Biogeosciences system, so I have pasted them below.
+Please also consider these comments, which are generally consistent with those of reviewers 1 and 2, when revising your manuscript.
+
+Please clearly describe how the text was revised according to the reviewer comments when submitting your revised manuscript. Thank you for your efforts, and for your support of Biogeosciences.
+
+Sincerely,
+Mike Weintraub
+Associate Editor
+Biogeosciences

### 2016-10-08 manuscript/response_to_reviewer_comments.tex
Started final response file for editor's reference

--- 2016-10-08 manuscript/response_to_reviewer_comments.txt
@@
-% Response to referee comments
-% Note LaTeX command below, as that's what Biogeosciences uses; see
-% http://www.biogeosciences.net/peer_review/commenting_on_a_paper.html
-% BBL September 2016
+% Response to referee comments, with reference to changed points in text
+% Doing this in LaTeX seemed easiest since the response to reviewers doc was already set up that way
+% ...but jeez I wish Biogeosciences would support Markdown
+% BBL October 2016
 
-% October 2016: added Reviewer 3 comments (which somehow did not show up online)
-% and editor's decision letter at end
+\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
+\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
+\usepackage{url}
+\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
+\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
+								% TeX will automatically convert eps --> pdf in pdflatex
+
+\begin{document}
 
 % ======================================================================================
-We thank all referees for their thoughtful comments, which we have numbered for convenient referencing.
-Per the {\it Biogeosciences} instructions, we have not yet prepared a revised manuscript, but respond below to all comments and questions.
-
-{\bf Response to Referee #1}
+{\bf Response to Referee \#1}
 
@@
 
-We appreciate this comment, and agree that the methods and concepts are not particularly novel–although drought incubations are fairly unusual–and that this is an understudied system and timely.
+We appreciate this comment, and agree that the methods and concepts are not particularly novel-although drought incubations are fairly unusual-and that this is an understudied system and timely.
 
@@
 
-This point was raised by Referee #2 as well (comment 2.4).
+This point was raised by Referee \#2 as well (comment 2.4).
 We will do so, comparing for example with studies such as the Schädel et al.
@@
 (2006, 2010).
+{\bf See lines TODO.}
 
@@
 (2014) that at N addition at high latitudes.
+{\bf See lines TODO.}
 
@@
 2005).
+{\bf See lines TODO.}
 
@@
 Studies have shown that moisture can have a weaker effect on temperature sensitivity early on during an incubation experiment, in the presence of more labile C.
-This relative to the effect on moisture on the Q-10 of cumulative respiration, reflecting slow turning over C – this could be an interesting analysis to include here, and would help to assess how }
+This relative to the effect on moisture on the Q-10 of cumulative respiration, reflecting slow turning over C - this could be an interesting analysis to include here, and would help to assess how }
 
@@
 (2005).
+{\bf See lines TODO.}
 
@@
 
-The first question is similar to a point raised by Referee #2 (comment 2.3).
+The first question is similar to a point raised by Referee \#2 (comment 2.3).
 We will discuss this in more depth, noting in particular, for example, the fact that there are observable anthropogenic influences on high-latitude precipitation (Wan et al.
@@
 2012).
+{\bf See lines TODO.}
 
@@
 
-Referee #2 raised this point as well (comment 2.6), and it's a good one.
+Referee \#2 raised this point as well (comment 2.6), and it's a good one.
 We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
-For example:
-
-\begin{itemize}
-  \item Decomposition rates vary greatly between the frozen and unfrozen states of soil, due to the phase changes involved, microbial community dynamics, etc.
-Q10 respiratory coefficients, for example, are extremely high near the freeze-thaw boundary (Monson et al.
-2006).
-  \item Deep active-layer soils are subject to distinctive freeze-thaw and cryoturbation dynamics, which are likely to change their sensitivity to climate change and feedback potential.
-  \item The depth of thaw can change the source/sink status of boreal ecosystems (Goulden et al. 1998) and potentially entire biomes (Harden et al. 2012, Hayes et al. 2011).
-  \item Seasonally frozen soils pose particular problems for accurate modeling of high-latitude carbon dynamics (Nicolsky et al. 2007, Bao et al. 2016).
-  \item Deep soil carbon is likely to be a highly important contributor to future climate feedbacks.
-The modeling studies of Koven et al.
-(2011, 2015), for example, found that one-third of 21st century climate-induced carbon loss originated from seasonally frozen soil carbon.
-\end{itemize}
+{\bf See lines TODO.}
 
@@
 1996).
+{\bf See lines TODO.}
 
@@
 
-That's a good point–thank you. We will qualify this sentence, better tying it to the specific measurements made in this study (C, N, respiration rates, stress response).
+That's a good point-thank you. We will qualify this sentence, better tying it to the specific measurements made in this study (C, N, respiration rates, stress response). {\bf See lines TODO.}
 
-{\it 1.12. Lines 77-80: I think a stronger argument for why deep active-layer soils can be made, and it would be helpful to clarify what are the “strong effects” of warming. }
+{\it 1.12. Lines 77-80: I think a stronger argument for why deep active-layer soils can be made, and it would be helpful to clarify what are the 'strong effects' of warming. }
 
-Agreed. See our response to comment 1.8 above.
+Agreed. See our response to comment 1.8 above. {\bf See lines TODO.}
 
@@
 Reviewer 2 also raised the idea of looking at C/N, and we have responded to that suggestion in detail (see comment 2.9).
+{\bf See lines TODO.}
 
@@
 Agreed.
-This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)."
+This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)." {\bf See lines TODO.}
 
 {\it 1.16.
-I don’t think that the summary of nearby respiration studies add very much to the discussion section.
+I don't think that the summary of nearby respiration studies add very much to the discussion section.
 Perhaps if the similarities and discrepancies were more integral to the central findings of the paper or integrated differently into the discussion they would seem more meaningful here.
@@
 
-We agree that the comparison to other boreal respiration studies needs improvement, and this echoes Referee #2's comment 2.4.
+We agree that the comparison to other boreal respiration studies needs improvement, and this echoes Referee \#2's comment 2.4.
 We will restructure and improve this summary and comparison section.
@@
 (2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but it would also be a good addition.
+{\bf See lines TODO.}
 
-{\it 1.17. Line 270: There is missing punctuation after the word “results”. }
+{\it 1.17. Line 270: There is missing punctuation after the word 'results'. }
 
-This will be fixed.
+This will be fixed. {\bf See lines TODO.}
 
@@
 
-This point was also made by Referee #2 (comment 2.2).
+This point was also made by Referee \#2 (comment 2.2).
 The fact that we didn't cite the Schädel et al.
@@
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
+{\bf See lines TODO.}
 
@@
 
-Thanks for the useful suggestion; we will do so.
+Thanks for the useful suggestion; we will do so. {\bf See lines TODO.}
 
@@
 In our revision, we will explore these ideas, considering how how factors such as vegetation, soil characteristics, and climate might explain the low observed temperature sensitivity.
+{\bf See lines TODO.}
 
@@
 We will make this paragraph more specific in all these areas.
+{\bf See lines TODO.}
 
@@
 We will discuss these results in our revision, along with other studies examining the relationship between N availability and C mineralization.
+{\bf See lines TODO.}
 
@@
 
-It's true that Colman and Schimel (2014) include only a few studies that could be termed boreal (from Maine, USA). We will either remove or de-emphasize/qualify this comparison.
+It's true that Colman and Schimel (2014) include only a few studies that could be termed boreal (from Maine, USA). We will either remove or de-emphasize/qualify this comparison.  {\bf See lines TODO.}
 
@@
 We will clarify this in the text.
-
-\newpage
-\begin{footnotesize}
-{\bf References cited in response:}
-
-Allison, S. D., C. I. Czimczik and K. K. Treseder (2008). "Microbial activity and soil respiration under nitrogen addition in Alaskan boreal forest." Global Change Biology 14(5): 1156-1168.
-
-Allison, S.
-D., T.
-B.
-Gartner, M.
-C.
-Mack, K.
-McGuire and K.
-K.
-Treseder (2011).
-"Nitrogen alters carbon dynamics during early succession in boreal forest." Soil Biology and Biochemistry 42: 1157-1164.
-
-Bond-Lamberty, B. and A. M. Thomson (2010). "A global database of soil respiration data." Biogeosciences 7: 1915-1926.
-
-Bouskill, N.
-J., W.
-J.
-Riley and J.
-Y.
-Tang (2014).
-"Meta-analysis of high-latitude nitrogen-addition and warming studies implies ecological mechanisms overlooked by land models." Biogeosciences 11: 6969-6983.
-
-Chapin III, F.
-S., M.
-S.
-Bret-Harte, S.
-E.
-Hobbie and H.
-Zhong (1996).
-"Plant functional types as predictors of transient responses of arctic vegetation to global change." Journal of Vegetation Science 7: 347-358.
-
-Czimczik, C. I., S. E. Trumbore, M. S. Carbone and G. C. Winston (2006). "Changing sources of soil respiration with time since fire in a boreal forest." Global Change Biology 12(6): 957-971.
-
-Czimczik, C. I. and J. M. Welker (2010). "Radiocarbon content of CO2 respired from high Arctic tundra in northwest Greenland." Arctic, Antarctic, and Alpine Research 42(3): 342-350.
-
-Davidson, E. A. and I. A. Janssens (2006). "Temperature sensitivity of soil carbon decomposition and feedbacks to climate change." Nature 440: 165-173.
-
-Dutta, K., E. A. G. Schuur, J. C. Neff and S. A. Zimov (2006). "Potential carbon release from permafrost soils of Northeastern Siberia." Global Change Biology 12: 2336-2351.
-
-Fierer, N., J. M. Craine, K. McLauchlan and J. P. Schimel (2005). "Litter quality and the temperature sensitivity of decomposition." Ecology 86(2): 320-326.
-
-Högberg, M. N., P. Högberg and D. D. Myrold (2007). "Is microbial community composition in boreal forest soils determined by pH, C-to-N ratio, the trees, or all three?" Oecologia 150(4): 590-601.
-
-Jassal, R.
-S., T.
-A.
-Black, M.
-D.
-Novak, D.
-Gaumont-Guay and Z.
-Nesic (2008).
-"Effect of soil water stress on soil respiration and its temperature sensitivity in an 18-year-old temperate Douglas-fir stand." Global Change Biology 14(6): 1305-1318.
-
-Karhu, K., H.
-Fritze, K.
-Hämäläinen, P.
-Vanhala, H.
-Jungner, M.
-Oinonen, E.
-Sonninen, M.
-Tuomi, P.
-Spetz, V.
-Kitunen and J.
-Liski (2010).
-"Temperature sensitivity of soil carbon fractions in boreal forest soil." Ecology 91(2): 370-376.
-
-Kleber, M. (2010). "What is recalcitrant soil organic matter?" Environmental Chemistry 7(4): 320-332.
-
-Lavoie, M., M.
-C.
-Mack and E.
-A.
-G.
-Schuur (2011).
-"Effects of elevated nitrogen and temperature on carbon and nitrogen dynamics in Alaskan arctic and boreal soils." Journal of Geophysical Research-Biogeosciences 116: G03013.
-
-Nadelhoffer, K. J., A. E. Giblin, G. R. Shaver and J. A. Laundre (1991). "Effects of temperature and substrate quality on element mineralization in six arctic soils." Ecology 72(1): 242-253.
-
-Raich, J. W. and W. H. Schlesinger (1992). "The global carbon dioxide flux in soil respiration and its relationship to vegetation and climate." Tellus B 44(2): 81-99.
-
-Reichstein, M., J.-A.
-Subke, A.
-C.
-Angeli and J.
-D.
-Tenhunen (2005).
-"Does the temperature sensitivity of decomposition of soil organic matter depend upon water content, soil horizon, or incubation time?" Global Change Biology 11(10): 1754-1767.
-
-von Lützow, M. and I. Kögel-Knabner (2009). "Temperature sensitivity of soil organic matter decomposition--what do we know?" Biology and Fertility of Soils 46(1): 1-15.
-
-\end{footnotesize}
+{\bf See lines TODO.}
 
 % ======================================================================================
-We thank all referees for their thoughtful comments, which we have numbered for convenient referencing.
-Per the {\it Biogeosciences} instructions, we have not yet prepared a revised manuscript, but respond below to all comments and questions.
+\newpage
 
-{\bf Response to Referee #2}
+{\bf Response to Referee \#2}
 
@@
 
-We appreciate the referee's assessment of the utility and interest of this dataset and analysis–thank you.
+We appreciate the referee's assessment of the utility and interest of this dataset and analysis--thank you.
 
 {\it 2.2.
-It seems like this paper was published as a discussion paper before Schädel et al.
+It seems like this paper was published as a discussion paper before Schädel et al.
 2016 was published and hence a discussion of the meta-analysis was not possible but should be addressed in the revisions.
@@
 
-This point was also made by Referee #1 (comment 1.18).
+This point was also made by Referee \#1 (comment 1.18).
 Yes, the fact that we didn't cite the Schädel et al.
@@
 2011, Dutta et al.
-2006, and Schädel et al.
+2006, and Schädel et al.
 2014.
@@
 (2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it would also be a reasonable addition.
+{\bf See lines TODO.}
 
@@
 
-This wording will be fixed.
+This wording will be fixed. {\bf See lines TODO.}
 
 {\it 2.6.
-3) A better explanation is needed why deep-active layer soils are different to active layer or permafrost soils, I couldn’t find a strong argument for why they would behave differently.
-Also, deep-active layer soils are those that are the most impacted by inter annual variability in thaw depth and so they might switch between active layer in one year to permafrost in another, that’s worth some discussion as well.
+3) A better explanation is needed why deep-active layer soils are different to active layer or permafrost soils, I couldn't find a strong argument for why they would behave differently.
+Also, deep-active layer soils are those that are the most impacted by inter annual variability in thaw depth and so they might switch between active layer in one year to permafrost in another, that's worth some discussion as well.
 }
 
-This is a good point. We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils. For example:
-
-\begin{itemize}
-  \item Decomposition rates vary greatly between the frozen and unfrozen states of soil, due to the phase changes involved, microbial community dynamics, etc.
-Q10 respiratory coefficients, for example, are extremely high near the freeze-thaw boundary (Monson et al.
-2006).
-  \item Deep active-layer soils are subject to distinctive freeze-thaw and cryoturbation dynamics, which are likely to change their sensitivity to climate change and feedback potential.
-  \item The depth of thaw can change the source/sink status of boreal ecosystems (Goulden et al. 1998) and potentially entire biomes (Harden et al. 2012, Hayes et al. 2011).
-  \item Seasonally frozen soils pose particular problems for accurate modeling of high-latitude carbon dynamics (Nicolsky et al. 2007, Bao et al. 2016).
-  \item Deep soil carbon is likely to be a highly important contributor to future climate feedbacks.
-The modeling studies of Koven et al.
-(2011, 2015), for example, found that one-third of 21st century climate-induced carbon loss originated from seasonally frozen soil carbon.
-\end{itemize}
+This is a good point.
+We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
+{\bf See lines TODO.}
 
@@
 
-Thank you. We appreciate the useful suggestions, and will provide these additional details in our revised manuscript.
+Thank you. We appreciate the useful suggestions, and will provide these additional details in our revised manuscript. {\bf See lines TODO.}
 
@@
 
-This useful suggestion was also made by Referee #1 (comment 1.13). We will do so.
+This useful suggestion was also made by Referee \#1 (comment 1.13). We will do so. {\bf See lines TODO.}
 
@@
 6) Why not include C/N as a variable in the statistical analysis?
-Schädel et al.
+Schädel et al.
 2014 showed that C/N is a good predictor of C release and can be used as a scaling factor.
@@
 We will discuss this issue, referencing previous work such as Schädel et al (2014).
+{\bf See lines TODO.}
 
@@
 
-That's a very good point–thank you–and will integrate well with an expanded comparison to the Schädel et al. (2016) paper (cf. comment 2.2 above) and other publications (comment 2.4 above).
+That's a very good point-thank you-and will integrate well with an expanded comparison to the Schädel et al.
+(2016) paper (cf.
+comment 2.2 above) and other publications (comment 2.4 above).
+{\bf See lines TODO.}
 
@@
 
-We will add a sentence of caveats, noting in particular the useful but incremental nature of this study.
-
-\newpage
-\begin{footnotesize}
-{\bf References cited in response:}
-
-Bao, H., T.
-Koike, K.
-Yang, L.
-Wang, M.
-Shrestha and P.
-Lawford (2016).
-"Development of an enthalpy-based frozen soil model and its validation in a cold region in China." Journal of Geophysical Research-Atmospheres 121(10): 5259-5280.
-
-Buermann, W., B.
-Parida, M.
-Jung, G.
-M.
-MacDonald, C.
-J.
-Tucker and M.
-Reichstein (2014).
-"Recent shift in Eurasian boreal forest greening response may be associated with warmer and drier summers." Geophysical Research Letters in press.
-
-Goulden, M.
-L., S.
-C.
-Wofsy, J.
-W.
-Harden, S.
-E.
-Trumbore, P.
-M.
-Crill, S.
-T.
-Gower, T.
-Fries, B.
-C.
-Daube, S.-M.
-Fan, D.
-J.
-Sutton, A.
-M.
-Bazzaz and J.
-W.
-Munger (1998).
-"Sensitivity of boreal forest carbon balance to soil thaw." Science 279: 214-217.
-
-Harden, J.
-W., C.
-D.
-Koven, C.-L.
-Ping, G.
-Hugelius, A.
-D.
-McGuire, P.
-Camill, T.
-Jorgenson, P.
-Kuhry, G.
-J.
-Michaelson, J.
-A.
-O'Donnell, E.
-A.
-G.
-Schuur, C.
-Tarnocai, K.
-D.
-Johnson and G.
-Grosse (2012).
-"Field information links permafrost carbon to physical vulnerabilities of thawing." Geophysical Research Letters 39: L15704.
-
-Hayes, D.
-J., A.
-D.
-McGuire, D.
-W.
-Kicklighter, K.
-R.
-Gurney, T.
-J.
-Burnside and J.
-M.
-Melillo (2011).
-"Is the northern high-latitude land-based CO2 sink weakening?" Global Biochemical Cycles 25: GB3018.
-
-Koven, C.
-D., B.
-Ringeval, P.
-Friedlingstein, P.
-Ciais, P.
-Cadule, D.
-Khvorostyanov, G.
-Krinner and C.
-Tarnocai (2011).
-"Permafrost carbon-climate feedbacks accelerate global warming." Proceedings of the National Academy of Science 108(36): 14769-14774.
-
-Koven, C.
-D., D.
-M.
-Lawrence and W.
-J.
-Riley (2015).
-"Permafrost carbon−climate feedback is sensitive to deep soil carbon decomposability but not deep soil nitrogen dynamics." Proceedings of the National Academy of Science 112(12): 3752-3757.
-
-Monson, R.
-K., D.
-L.
-Lipson, S.
-P.
-Burns, A.
-A.
-Turnipseed, A.
-C.
-Delany, M.
-W.
-WIlliams and S.
-K.
-Schmidt (2006).
-"Winter forest soil respiration controlled by climate and microbial community composition." Nature 439: 711-714.
-
-Nicolsky, D. J., V. E. Romanovsky, V. A. Alexeev and D. M. Lawrence (2007). "Improved modeling of permafrost dynamics in a GCM land-surface scheme." Geophysical Research Letters 34: L08501.
-
-Wan, H., X. Zhang, F. W. Zwiers and S.-K. Min (2015). "Attributing northern high-latitude precipitation change over the period 1966–2005 to human influence." Climate Dynamics 45(7): 1713-1726.
-
-Wendler, G. and M. Shulski (2009). "A century of climate change for Fairbanks, Alaska." Arctic 62(3): 295-300.
-
-Zhang, X., J.
-He, J.
-Zhang, I.
-Polyakov, R.
-Gerdes, J.
-Inoue and P.
-Wu (2012).
-"Enhanced poleward moisture transport and amplified northern high-latitude wetting trend." Nature Climate Change 3: 47-51.
-
-\end{footnotesize}
+We will add a sentence of caveats, noting in particular the useful but incremental nature of this study. {\bf See lines TODO.}
 
 % ======================================================================================
-We thank all referees for their thoughtful comments, which we have numbered for convenient referencing.
-Per the {\it Biogeosciences} instructions, we have not yet prepared a revised manuscript, but respond below to all comments and questions.
+\newpage
 
-{\bf Response to Referee #3 (comments were not online before editor's decision) }
+{\bf Response to Referee \#3
 
@@
 This manuscript presents...interesting and surprising correlations between CO2 production and soil nitrogen content.
-The main point of this paper, and I think that it’s an important one, is that many upland forest soils in the boreal region that have permafrost will dry, and the majority of C will be emitted as CO2, but the magnitude of C emission will decrease under drought conditions (imposed by freely draining and no water inputs) by ~ 50-70\%.
+The main point of this paper, and I think that it's an important one, is that many upland forest soils in the boreal region that have permafrost will dry, and the majority of C will be emitted as CO2, but the magnitude of C emission will decrease under drought conditions (imposed by freely draining and no water inputs) by ~ 50-70\%.
 Overall, I enjoyed reading this paper!
@@
 My main criticism is that I think that the authors over-emphasize the results of the daily emissions and that the authors should further explore (or report) the results of the controls of the cumulative C emissions.
-I’m curious as to whether the relationships with soil C/N and \%N observed in daily emissions still hold on cumulative emissions.
+I'm curious as to whether the relationships with soil C/N and \%N observed in daily emissions still hold on cumulative emissions.
 The comparison between these soil parameters (i.e.
-ones that probably don’t change much throughout the course of the incubation, including temperature) and the cumulative fluxes is perhaps more appropriate.
+ones that probably don't change much throughout the course of the incubation, including temperature) and the cumulative fluxes is perhaps more appropriate.
 Perhaps modelers find the controls on daily fluxes interesting and these are likely quite useful in regards to the relationship between moisture and C production (i.e.
@@
 At the moment, it looks like only \%N exerts a weak effect on cumulative CO2 fluxes, and none of the tested covariates (\%C, \%N, C/N, DOC) affects cumulative CH4.
+{\bf See lines TODO.}
 
@@
 (2008).
+{\bf See lines TODO.}
 
@@
 I do think that the time series of fluxes could be moved to the supplemental materials if the cumulative fluxes are explored in greater detail.
-I think this paper could be shortened a little bit although I didn’t find the length of the paper onerous.
+I think this paper could be shortened a little bit although I didn't find the length of the paper onerous.
 Along these lines, I think that the results summarized above from the cumulative emissions should be included in the abstract.
@@
 
-We will consider these suggestions with respect to length and supplemental materials; results dealing with cumulative emissions will certainly be included in the abstract.
+We will consider these suggestions with respect to length and supplemental materials; results dealing with cumulative emissions will certainly be included in the abstract. {\bf See lines TODO.}
 
@@
 27: daily CH4 flux?
-28: cumulative production as CO2…as CH4. }
+28: cumulative production as CO2-as CH4. }
 
-We will clarify these points.
+We will clarify these points. {\bf See lines TODO.}
 
@@
 
-Agreed, this was more of a qualitative statement than a quantitative one. We will be more precise.
+Agreed, this was more of a qualitative statement than a quantitative one. We will be more precise. {\bf See lines TODO.}
 
@@
 (2008), both in the introduction and discussion.
+{\bf See lines TODO.}
 
-{\it 3.8. 67: “substantial variabilities between studies” WHY? }
+{\it 3.8. 67: 'substantial variabilities between studies' WHY? }
 
-We will expand on this point, pointing out that such variability originates from factors such soil effects on drainage, phase changes, different experimental protocols, etc.
+We will expand on this point, pointing out that such variability originates from factors such soil effects on drainage, phase changes, different experimental protocols, etc. {\bf See lines TODO.}
 
-{\it 3.9. 72: Yes, this is an important question, but given that this isn’t measured in this study, perhaps this sentence should be omitted or re-written. }
+{\it 3.9. 72: Yes, this is an important question, but given that this isn't measured in this study, perhaps this sentence should be omitted or re-written. }
 
-Good point; we will do so.
+Good point; we will do so. {\bf See lines TODO.}
 
@@
 This will be clarified, and we will consider if adding this to Figures 2 and 3 if readable and useful.
+{\bf See lines TODO.}
 
@@
 
-We will!
+We will! {\bf See lines TODO.}
 
@@
 
-Thanks. There was a great of variability (obviously), but distributed throughout the data set–i.e., this isn't driven by one or two outliers. We will re-check, however.
+Thanks. There was a great of variability (obviously), but distributed throughout the data set-i.e., this isn't driven by one or two outliers. We will re-check, however. {\bf See lines TODO.}
 
@@
 
-These will be fixed.
+These will be fixed. {\bf See lines TODO.}
 
@@
 
-{\it 3.16. 262: First mention of vegetation stress, remove, not clear how it’s related. }
+{\it 3.16. 262: First mention of vegetation stress, remove, not clear how it's related. }
 
-We will better integrate this point, mentioning it in the introduction and clarifying its relationship to the study goals.
+We will better integrate this point, mentioning it in the introduction and clarifying its relationship to the study goals. {\bf See lines TODO.}
 
-{\it 3.17. 270. Add “.” }
+{\it 3.17. 270. Add '.' }
 
-This will be fixed.
+This will be fixed. {\bf See lines TODO.}
 
@@
 
-Upland Cryosols; we will clarify this.
+Upland Cryosols; we will clarify this. {\bf See lines TODO.}
 
@@
 
-Thanks; we will do so.
+Thanks; we will do so. {\bf See lines TODO.}
 
@@
 
-Thanks for the Lee et al. reference, which we had not considered (see our response to 3.7 above). We'll take a hard look at lines 362-378 and tighten.
+Thanks for the Lee et al. reference, which we had not considered (see our response to 3.7 above). We'll take a hard look at lines 362-378 and tighten. {\bf See lines TODO.}
 
-{\it 3.23. 383-384: “specific weaknesses”: vague
+{\it 3.23. 383-384: 'specific weaknesses': vague
 384: See also lag effects found in Treat et al. (2015) }
 
-Good points! We will clarify and note this.
+Good points! We will clarify and note this. {\bf See lines TODO.}
 
-{\it 3.24. 393: “taking them out of depth” rephrase. Also could use this argument for the section on CH4 production. }
+{\it 3.24. 393: 'taking them out of depth' rephrase. Also could use this argument for the section on CH4 production. }
 
-Yes, it's awkward; we will reword. Thanks for the suggestion.
+Yes, it's awkward; we will reword. Thanks for the suggestion. {\bf See lines TODO.}
 
@@
 
-We thought we were already doing so in using the {\tt RColorBrewer} package, not the default palette of {\tt ggplot2}, but will investigate and fix this if necessary.
+We thought we were already doing so in using the {\tt RColorBrewer} package, not the default palette of {\tt ggplot2}, but will investigate and fix this if necessary. {\bf See lines TODO.}
 
@@
 
-Good point. Re colors, see our response to 3.25 above.
+Good point. Re colors, see our response to 3.25 above. {\bf See lines TODO.}
 
-% ======================================================================================
-
-Associate Editor Decision: Reconsider after major revisions (06 Oct 2016) by Michael Weintraub
-Comments to the Author:
-Thank you for your submission to Biogeosciences.
-Your MS has received three reviews from highly qualified reviewers in this field, all of whom thought that this is a valuable dataset worth publishing, but offered numerous substantive comments to improve the paper.
-Thus, I will be happy to consider a revised submission that substantively addresses these reviewer comments.
-Based on the responses to reviewers 1 and 2, it looks like there is a sound plan for addressing the reviewers’ critical comments.
-
-Unfortunately, I discovered too late that the comments from the third reviewer do not show up in the Biogeosciences system, so I have pasted them below.
-Please also consider these comments, which are generally consistent with those of reviewers 1 and 2, when revising your manuscript.
-
-Please clearly describe how the text was revised according to the reviewer comments when submitting your revised manuscript. Thank you for your efforts, and for your support of Biogeosciences.
-
-Sincerely,
-Mike Weintraub
-Associate Editor
-Biogeosciences
+\end{document}

### 2016-10-09 manuscript/response_to_reviewer_comments.tex
Beginning to address reviewer comments. Easy ones first.

@@
 % Response to referee comments, with reference to changed points in text
-% Doing this in LaTeX seemed easiest since the response to reviewers doc was already set up that way
-% ...but jeez I wish Biogeosciences would support Markdown
+% Doing this in LaTeX seemed easiest since the response to reviewers doc
+% was already set up that way...but I wish Biogeosciences would support Markdown
 % BBL October 2016
+% TODO means haven't addressed yet; LINES means addressed, need to fill in line #s
 
-\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
-\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
+\documentclass[11pt, oneside]{article}
+\usepackage{geometry}
 \usepackage{url}
-\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
-\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
-								% TeX will automatically convert eps --> pdf in pdflatex
+\geometry{letterpaper} 
+\usepackage{graphicx}
 
@@
 {\bf Response to Referee \#1}
-
-{\it 1.1.
-The authors have measured...
-This manuscript contributes to the growing body of Arctic permafrost and active layer incubation literature...
-There has been a lot of focus in incubation studies on the effect of inundation on CO2 and CH4 emissions, but drought is an important factor to consider as well.
-}
-
-Thank you.
-
-{\it 1.2.
-I do think that the paper has within it data that is relevant to the scope of BG, but I think that a more in depth analysis of the data, and a better contextualization of the data within the Arctic literature is required.
-The methods and concepts are not novel, but this is an understudied system and very timely work.
-More is needed in order to reach substantial conclusions.
-}
-
-We appreciate this comment, and agree that the methods and concepts are not particularly novel-although drought incubations are fairly unusual-and that this is an understudied system and timely.
 
@@
 (2006, 2010).
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 (2014) that at N addition at high latitudes.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 2005).
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 (2005).
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 2012).
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
-{\bf See lines TODO.}
+{\bf See new lines LINES.}
 
@@
 1996).
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 
-That's a good point-thank you. We will qualify this sentence, better tying it to the specific measurements made in this study (C, N, respiration rates, stress response). {\bf See lines TODO.}
+That's a good point-thank you. We will qualify this sentence, better tying it to the specific measurements made in this study (C, N, respiration rates, stress response). {\bf See new lines TODO.}
 
@@
 
-Agreed. See our response to comment 1.8 above. {\bf See lines TODO.}
+Agreed. See our response to comment 1.8 above.
 
@@
 Reviewer 2 also raised the idea of looking at C/N, and we have responded to that suggestion in detail (see comment 2.9).
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 Agreed.
-This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)." {\bf See lines TODO.}
+This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)." {\bf See new lines TODO.}
 
@@
 (2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but it would also be a good addition.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 
-This will be fixed. {\bf See lines TODO.}
+This will be fixed. {\bf See new lines LINES.}
 
@@
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 
-Thanks for the useful suggestion; we will do so. {\bf See lines TODO.}
+Thanks for the useful suggestion; we will do so. {\bf See new lines TODO.}
 
@@
 In our revision, we will explore these ideas, considering how how factors such as vegetation, soil characteristics, and climate might explain the low observed temperature sensitivity.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 We will make this paragraph more specific in all these areas.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 We will discuss these results in our revision, along with other studies examining the relationship between N availability and C mineralization.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 
-It's true that Colman and Schimel (2014) include only a few studies that could be termed boreal (from Maine, USA). We will either remove or de-emphasize/qualify this comparison.  {\bf See lines TODO.}
+It's true that Colman and Schimel (2014) include only a few studies that could be termed boreal (from Maine, USA).
+We will either remove or de-emphasize/qualify this comparison.
+{\bf See new lines TODO.}
 
@@
 We will clarify this in the text.
-{\bf See lines TODO.}
+{\bf See new lines LINES.}
 
@@
 {\bf Response to Referee \#2}
-
-{\it 2.1.
-This paper studies...
-This is a useful dataset that adds to the growing body of literature on C release from permafrost once thawed.
-There is a good amount of publications that deal with warmer temperatures in the Arctic and changing soil moisture condition but not many have simulated drought conditions...
-}
-
-We appreciate the referee's assessment of the utility and interest of this dataset and analysis--thank you.
 
@@
 (2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it would also be a reasonable addition.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 
-This wording will be fixed. {\bf See lines TODO.}
+This wording will be fixed. {\bf See new lines LINES.}
 
@@
 
-Thank you. We appreciate the useful suggestions, and will provide these additional details in our revised manuscript. {\bf See lines TODO.}
+Thank you. We appreciate the useful suggestions, and will provide these additional details in our revised manuscript. {\bf See new lines TODO.}
 
@@
 
-This useful suggestion was also made by Referee \#1 (comment 1.13). We will do so. {\bf See lines TODO.}
+This useful suggestion was also made by Referee \#1 (comment 1.13). We will do so. {\bf See new lines TODO.}
 
@@
 Currently C/N seems to be a poorer predictor than \%N.
-We will discuss this issue, referencing previous work such as Schädel et al (2014).
-{\bf See lines TODO.}
+We will discuss this issue, referencing previous work such as SchŠdel et al (2014).
+{\bf See new lines TODO.}
 
@@
 
-That's a very good point-thank you-and will integrate well with an expanded comparison to the Schädel et al.
+That's a very good point-thank you-and will integrate well with an expanded comparison to the SchŠdel et al.
 (2016) paper (cf.
 comment 2.2 above) and other publications (comment 2.4 above).
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 
-We will add a sentence of caveats, noting in particular the useful but incremental nature of this study. {\bf See lines TODO.}
+We will add a sentence of caveats, noting in particular the useful but incremental nature of this study. {\bf See new lines TODO.}
 
@@
 {\bf Response to Referee \#3
-
-{\it 3.1.
-This manuscript presents...interesting and surprising correlations between CO2 production and soil nitrogen content.
-The main point of this paper, and I think that it's an important one, is that many upland forest soils in the boreal region that have permafrost will dry, and the majority of C will be emitted as CO2, but the magnitude of C emission will decrease under drought conditions (imposed by freely draining and no water inputs) by ~ 50-70\%.
-Overall, I enjoyed reading this paper!
-It was nicely written although vague in a few parts.
-}
-
-Thank you.
 
@@
 At the moment, it looks like only \%N exerts a weak effect on cumulative CO2 fluxes, and none of the tested covariates (\%C, \%N, C/N, DOC) affects cumulative CH4.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 (2008).
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 
-We will consider these suggestions with respect to length and supplemental materials; results dealing with cumulative emissions will certainly be included in the abstract. {\bf See lines TODO.}
+We will consider these suggestions with respect to length and supplemental materials; results dealing with cumulative emissions will certainly be included in the abstract. {\bf See new lines TODO.}
 
@@
 
-We will clarify these points. {\bf See lines TODO.}
+We will clarify these points. {\bf See new lines TODO.}
 
@@
 
-Agreed, this was more of a qualitative statement than a quantitative one. We will be more precise. {\bf See lines TODO.}
+Agreed, this was more of a qualitative statement than a quantitative one. We will be more precise. {\bf See new lines TODO.}
 
@@
 (2008), both in the introduction and discussion.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 
-We will expand on this point, pointing out that such variability originates from factors such soil effects on drainage, phase changes, different experimental protocols, etc. {\bf See lines TODO.}
+We will expand on this point, pointing out that such variability originates from factors such soil effects on drainage, phase changes, different experimental protocols, etc. {\bf See new lines TODO.}
 
@@
 
-Good point; we will do so. {\bf See lines TODO.}
+Good point; we will do so. {\bf See new lines TODO.}
 
@@
 This will be clarified, and we will consider if adding this to Figures 2 and 3 if readable and useful.
-{\bf See lines TODO.}
+{\bf See new lines TODO.}
 
@@
 
-We will! {\bf See lines TODO.}
+We will! {\bf See new lines TODO.}
 
@@
 
-Thanks. There was a great of variability (obviously), but distributed throughout the data set-i.e., this isn't driven by one or two outliers. We will re-check, however. {\bf See lines TODO.}
+Thanks. There was a great of variability (obviously), but distributed throughout the data set-i.e., this isn't driven by one or two outliers. We will re-check, however. {\bf See new lines TODO.}
 
@@
 
-These will be fixed. {\bf See lines TODO.}
+These will be fixed. {\bf See new lines TODO.}
 
@@
 
-We will better integrate this point, mentioning it in the introduction and clarifying its relationship to the study goals. {\bf See lines TODO.}
+We will better integrate this point, mentioning it in the introduction and clarifying its relationship to the study goals. {\bf See new lines TODO.}
 
@@
 
-Upland Cryosols; we will clarify this. {\bf See lines TODO.}
+Upland Cryosols; we will clarify this. {\bf See new lines LINES.}
 
@@
 
-Thanks; we will do so. {\bf See lines TODO.}
+Thanks; we will do so. {\bf See new lines TODO.}
 
@@
 
-Thanks for the Lee et al. reference, which we had not considered (see our response to 3.7 above). We'll take a hard look at lines 362-378 and tighten. {\bf See lines TODO.}
+Thanks for the Lee et al. reference, which we had not considered (see our response to 3.7 above). We'll take a hard look at lines 362-378 and tighten. {\bf See new lines TODO.}
 
@@
 
-Good points! We will clarify and note this. {\bf See lines TODO.}
+Good points! We will clarify and note this. {\bf See new lines TODO.}
 
@@
 
-Yes, it's awkward; we will reword. Thanks for the suggestion. {\bf See lines TODO.}
+Yes, it's awkward; we will reword. Thanks for the suggestion. {\bf See new lines TODO.}
 
@@
 
-We thought we were already doing so in using the {\tt RColorBrewer} package, not the default palette of {\tt ggplot2}, but will investigate and fix this if necessary. {\bf See lines TODO.}
+We thought we were already doing so in using the {\tt RColorBrewer} package, not the default palette of {\tt ggplot2}, but will investigate and fix this if necessary. {\bf See new lines TODO.}
 
@@
 
-Good point. Re colors, see our response to 3.25 above. {\bf See lines TODO.}
+Good point. Re colors, see our response to 3.25 above. {\bf See new lines TODO.}
 

### 2016-10-09 manuscript/cpcrw_manuscript.Rmd
Beginning to address reviewer comments. Easy ones first.

@@
 The cumulative production of C from CO~2~ was over `r minmag` orders of magnitudes higher than that from CH~4~.
-These results suggest that deep active-layer soils may be much more sensitive to changes in moisture than to temperature, a critical factor as discontinuous permafrost melts in interior Alaska.
+These results suggest that deep active-layer soils may be much more sensitive to changes in moisture than to temperature, a critical factor as discontinuous permafrost thaws [COM2.5] in interior Alaska.
 Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results provide insight into their future dynamics and feedback potential with future climate change.
@@
 
+Most previous studies have focused on surface soils or permafrost soils, neglecting deep active-layer soils that were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
+Such deeper soils have particular characteristics distinguishing them from both shallow active layer soils and underlying permafrost: they are most affected by interannual variability in thaw depth, potentially flipping the C source/sink status of entire ecosystems {Goulden, 1998 #50}{Harden, 2012 #4476}, and are subject to distinctive freeze-thaw, cryoturbation, and microbial dynamics, which are likely to change their sensitivity to climate change and feedback potential {Schuur, 2008 #4491}; and deep seasonally frozen soils pose particular problems for accurate modeling of high-latitude carbon dynamics {Nicolsky, 2007 #4474}.
+Deep soil carbon is likely to be a highly important contributor to future climate feedbacks, with modeling studies suggesting that one-third of 21^st^ century climate-induced carbon loss may originate from seasonally frozen soil carbon {Koven, 2015 #4473}.
+[COM1.8,2.6]
+
 The goal of this study was to examine how temperature and moisture control GHG (CO~2~ and CH~4~) emissions from soils sampled from the bottom of the annual active layer–i.e., directly above permafrost–in an Alaskan boreal forest.
-Most previous studies have focused on surface soils or permafrost soils, neglecting deep active-layer soils that were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
 We also aimed to characterize the chemical and structural properties of these soils following a 100-day incubation at different temperatures, subjecting some cores to drying treatments.
@@
 Morishita et al.
-{, 2014 #4316} quantified gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture, consistent with our results They found however that CH~4~ uptake (no emissions were observed) was driven by temperature only.
+{, 2014 #4316} quantified gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture in upland Cryosols [COM3.18], consistent with our results.[COM1.17,3.17] They found however that CH~4~ uptake (no emissions were observed) was driven by temperature only.
 Waldrop et al.
@@
 Laboratory experiments offer precise control, but lack the *in situ* nature of field manipulations {Sistla, 2013 #3666}, raising uncertainties to what degree their results can be extrapolated.
-They also have more specific weaknesses, for example in what can be inferred about temperature sensitivity {Podrebarac, 2016 #4365}{Hamdi, 2013 #3664}.
-Nonetheless, the controlled environments of incubations provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
+Soils isolated during incubation may, for example, underestimate temperature sensitivity of respiration {Podrebarac, 2016 #4365}.
+[COM1.24] Nonetheless, the controlled environments of incubations provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
 

### 2016-11-12 manuscript/cpcrw_manuscript.Rmd
Moved code out of rmarkdownfile and update package versions

@@
 library(dplyr, warn.conflicts = FALSE)
-
-# 'Pretty n' function to round a numeric value and print that # of digits
-pn <- function(x, n) formatC(round(unlist(x), n), digits = n, format = 'f')
-
-# 'Clean p value' function to pretty-print p value(s), specifically
-pclean <- function(x, digits = 3, printP = TRUE) {
-  x <- as.vector(x)
-  ltstring <- paste0("< 0.", paste(rep("0", digits - 1), collapse = ""), "1")
-  valstring <- ifelse(x < 10^-digits, ltstring, pn(x, digits))
-  if(printP) {
-    ifelse(x < 10^-digits, valstring, paste("=", valstring)) %>% paste("P", .)
-  } else {
-    valstring
-  }
-}
 

### 2016-11-12 R/0-functions.R

@@ -6,12 +6,12 @@
 # Run under:
 #   R version 3.2.3 (2015-12-10), x86_64-apple-darwin13.4.0 (64-bit)
 # Key packages used across scripts:
-library(ggplot2)       # 2.0.0
+library(ggplot2)       # 2.1.0
 theme_set(theme_bw())
-library(dplyr)         # 0.4.3.9001
-library(readr)         # 0.2.2
-library(lubridate)     # 1.5.0
-library(stringr)       # 1.0.0
+library(dplyr)         # 0.5.0
+library(readr)         # 1.0.0
+library(lubridate)     # 1.6.0
+library(stringr)       # 1.1.0
 library(luzlogr)       # 0.2.0
 
 #if(require(checkpoint))   # 0.3.15
@@ -132,3 +132,22 @@ save_diagnostic <- function(p, pname, printit = TRUE, ...) {
     ggsave
   save_plot(pname, ...)
 } # save_diagnostic
+
+# -----------------------------------------------------------------------------
+# 'Pretty n' function to round a numeric value and print that # of digits
+pn <- function(x, n) {
+  formatC(round(unlist(x), n), digits = n, format = "f")
+} # pn
+
+# -----------------------------------------------------------------------------
+# 'Clean p value' function to pretty-print p value(s), specifically
+pclean <- function(x, digits = 3, printP = TRUE) {
+  x <- as.vector(x)
+  ltstring <- paste0("< 0.", paste(rep("0", digits - 1), collapse = ""), "1")
+  valstring <- ifelse(x < 10 ^ -digits, ltstring, pn(x, digits))
+  if(printP) {
+    paste("P", ifelse(x < 10 ^ -digits, valstring, paste("=", valstring)))
+  } else {
+    valstring
+  }
+} # pclean

### 2016-11-13 manuscript/cpcrw_manuscript.Rmd
Moved code from ms file to 5-analyze

@@
 setwd(olddir)
-```
-
-```{r wc, echo=FALSE}
-# Compute soil moisture data for use in abstract and results below
-# wc and wc_alltrt are computed by 5-analyze.R script
-gwcbegin <- subset(gwc_alltrt, stage == "Beginning")
-gwcend <- subset(gwc, stage == "End")
-gwcendmean <- gwcend$mean
-names(gwcendmean) <- gwcend$Treatment
-gwcendsd <- gwcend$sd
-names(gwcendsd) <- gwcend$Treatment
-vwcbegin <- subset(vwc_alltrt, stage == "Beginning")
-vwcend <- subset(vwc_alltrt, stage == "End")
-```
-
-```{r compute_ratios, echo=FALSE}
-# Compute ratios between CO2 and CH4 for use in abstract and results below
-# Ratio data is stored in the `gas_ratio` d.f. computed by analysis script
-minmag <- min(gas_ratio$CO2CH4_ratio) %>% log10 %>% floor %>%
-  c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine")[.]
-minratio <- min(gas_ratio$CO2CH4_ratio / 1e6)  %>% formatC(digits = 1, format = 'f')
-mintrt <- gas_ratio$Treatment[which.min(gas_ratio$CO2CH4_ratio)]
-mintemp <- gas_ratio$Temperature[which.min(gas_ratio$CO2CH4_ratio)]
-maxratio <- max(gas_ratio$CO2CH4_ratio / 1e6) %>% formatC(digits = 1, format = 'f')
-maxtrt <- gas_ratio$Treatment[which.max(gas_ratio$CO2CH4_ratio)]
-maxtemp <- gas_ratio$Temperature[which.max(gas_ratio$CO2CH4_ratio)]
 ```
@@
 **2.3 Data and statistical analysis**
-
-```{r compute_excludes, echo=FALSE}
-# Compute outlier and exclusion numbers
-# TODO: this isn't good--fragile if files move or change names
-rn <- readLines("../outputs/3-remove/removed_data.csv") %>% length
-sdn <- readLines("../outputs/summarydata.csv") %>% length
-fluxn <- nrow(fluxdata)
-```
 
@@
 ```{r table1, echo=FALSE}
-t1 <- summary(step_co2_lme)$tTable %>% round(3) %>% as.data.frame()
+summary(step_co2_lme)$tTable %>% 
+  round(3) %>% 
+  as.data.frame() ->
+  t1
 t1$`p-value` <- pclean(t1$`p-value`, printP = FALSE)
@@
 ```{r table2, echo=FALSE}
-t2 <- summary(step_co2_lme)$tTable %>% round(3) %>% as.data.frame()
+summary(step_co2_lme)$tTable %>% 
+  round(3) %>% 
+  as.data.frame() ->
+  t2
 t2$`p-value` <- pclean(t2$`p-value`, printP = FALSE)

### 2016-11-13 R/0-functions.R

@@ -39,6 +39,7 @@ RAWDATA_FILE           <- file.path(OUTPUT_DIR, "rawdata.csv.gz")
 RAWDATA_SAMPLES_FILE   <- file.path(OUTPUT_DIR, "rawdata_samples.csv.gz")
 VALVEMAP               <- "data/valvemap.csv"
 TREATMENTS             <- "data/treatments.csv"
+REMOVEDDATA_FILE       <- file.path(OUTPUT_DIR, "removed_data.csv")
 SUMMARYDATA_FILE       <- file.path(OUTPUT_DIR, "summarydata.csv")
 SUMMARYDATA_CLEAN_FILE <- file.path(OUTPUT_DIR, "summarydata_clean.csv")
 FLUXDATA_FILE          <- file.path(OUTPUT_DIR, "fluxdata.csv")

### 2016-11-13 R/3-remove.R

@@ -73,7 +73,7 @@ save_data(summarydata_clean, fn = SUMMARYDATA_CLEAN_FILE, scriptfolder = FALSE)
 
 printlog("Removed data:", nrow(removed_data), "of", nrow(summarydata), "=",
          round(nrow(removed_data) / nrow(summarydata) * 100, 1), "%")
-save_data(removed_data)
+save_data(removed_data, fn = REMOVEDDATA_FILE, scriptfolder = FALSE)
 
 printlog("All done with", SCRIPTNAME)
 closelog()

### 2016-11-13 R/5-analyze.R

@@ -177,6 +177,17 @@ fluxdata %>%
   summarise_each(funs(mean, sd, max, min), WFPS_percent) ->
   wfps_alltrt
 
+# Compute key soil moisture rates for use in abstract and results of ms
+gwcbegin <- subset(gwc_alltrt, stage == "Beginning")
+gwcend <- subset(gwc, stage == "End")
+gwcendmean <- gwcend$mean
+names(gwcendmean) <- gwcend$Treatment
+gwcendsd <- gwcend$sd
+names(gwcendsd) <- gwcend$Treatment
+vwcbegin <- subset(vwc_alltrt, stage == "Beginning")
+vwcend <- subset(vwc_alltrt, stage == "End")
+
+
 # -----------------------------------------------------------------------------
 # Did 'Treatment' affect water content?
 
@@ -362,6 +373,23 @@ fluxdata_cumulative %>%
   print ->
   gas_ratio
 
+# Compute ratio-related text for use in manuscript
+min(gas_ratio$CO2CH4_ratio) %>% 
+  log10 %>% 
+  floor %>%
+  c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine")[.] ->
+  minmag
+min(gas_ratio$CO2CH4_ratio / 1e6)  %>% 
+  formatC(digits = 1, format = 'f') ->
+  minratio
+mintrt <- gas_ratio$Treatment[which.min(gas_ratio$CO2CH4_ratio)]
+mintemp <- gas_ratio$Temperature[which.min(gas_ratio$CO2CH4_ratio)]
+max(gas_ratio$CO2CH4_ratio / 1e6) %>% 
+  formatC(digits = 1, format = 'f') ->
+  maxratio
+maxtrt <- gas_ratio$Treatment[which.max(gas_ratio$CO2CH4_ratio)]
+maxtemp <- gas_ratio$Temperature[which.max(gas_ratio$CO2CH4_ratio)]
+
 # -----------------------------------------------------------------------------
 # Examine data distribution
 # Log-transforming doesn't fix the lack of normality in our data,
@@ -450,6 +478,11 @@ m_ch4_lme_thirds <- update(m_ch4_lme, ~ . +
                              Temperature * third)
 step_ch4_lme_thirds <- MASS::stepAIC(m_ch4_lme_thirds, direction = "both")
 
+# -----------------------------------------------------------------------------
[…]

### 2016-11-13 manuscript/response_to_reviewer_comments.tex
Manuscript and README updates

@@
 \documentclass[11pt, oneside]{article}
+\usepackage[utf8]{inputenc}     
 \usepackage{geometry}
@@
 \begin{document}
-
 % ======================================================================================
-{\bf Response to Referee \#1}
+\subsection*{Response to Referee \#1}
 
@@
 
+\medskip
 {\it 1.4.
@@
 
+\medskip
 {\it 1.5.
@@
 
+\medskip
 {\it 1.6.
@@
 
+\medskip
 {\it 1.7.
@@
 
+\medskip
 {\it 1.8.
@@
 
+\medskip
 {\it 1.9.
@@
 Vegetation and ecosystem type is a significant factor that is not well explored here, we agree.
-We will add some points about this, noting for example the work of Högberg et al.
+We will add some points about this, noting for example the work of H?gberg et al.
 (2007) on soil microbial community composition, the relationship of vegetation type and soil respiration (Raich and Schlesinger 1992, Bond-Lamberty and Thomson 2010), and how plant functional types affect ecosystem response more generally (Chapin et al.
@@
 
+\medskip
 {\it 1.10.
@@
 
+\medskip
 {\it 1.11.
@@
 
+\medskip
 {\it 1.12. Lines 77-80: I think a stronger argument for why deep active-layer soils can be made, and it would be helpful to clarify what are the 'strong effects' of warming. }
@@
 
+\medskip
 {\it 1.13.
@@
 
+\medskip
 {\it 1.14. Line 232: In this section it would also be interesting to know the soil respiration decay rate per treatment over the course of the incubation experiment. }
@@
 
+\medskip
 {\it 1.15. Line 238-240: Confusingly worded sentence. }
@@
 
+\medskip
 {\it 1.16.
@@
 (2006), although it concerns Siberian soils, may also prove a useful comparison.
-We had not included Schädel et al.
+We had not included Sch?del et al.
 (2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but it would also be a good addition.
@@
 
+\medskip
 {\it 1.17. Line 270: There is missing punctuation after the word 'results'. }
@@
 
+\medskip
 {\it 1.18. Line 286-293: Perhaps the new synthesis by Schadel et al., 2016, would also be a useful comparison here. }
@@
 This point was also made by Referee \#2 (comment 2.2).
-The fact that we didn't cite the Schädel et al.
+The fact that we didn't cite the Sch?del et al.
 (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
-In our revision, we will significantly expand this paragraph, discussing and comparing to Schädel et al.
+In our revision, we will significantly expand this paragraph, discussing and comparing to Sch?del et al.
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
@@
 
+\medskip
 {\it 1.19.
@@
 
+\medskip
 {\it 1.20.
@@
 
+\medskip
 {\it 1.21.
@@
 Jassal et al.
-2008), but the field is rife with contradictory results (von Lützow and Kögel-Knabner 2009).
+2008), but the field is rife with contradictory results (von L?tzow and K?gel-Knabner 2009).
 We will make this paragraph more specific in all these areas.
@@
 
+\medskip
 {\it 1.22.
@@
 
+\medskip
 {\it 1.23. Line 367: Is this comparison, with North American soils, relevant to this study? }
@@
 
+\medskip
 {\it 1.24. Line 383-384: Can you be more explicit in your meaning here? How do you mean that there is weakness in what can be inferred about temperature sensitivity from experiments? }
@@
 \newpage
-
-{\bf Response to Referee \#2}
+\subsection*{Response to Referee \#2}
 
 {\it 2.2.
-It seems like this paper was published as a discussion paper before Schädel et al.
+It seems like this paper was published as a discussion paper before Sch?del et al.
 2016 was published and hence a discussion of the meta-analysis was not possible but should be addressed in the revisions.
@@
 This point was also made by Referee \#1 (comment 1.18).
-Yes, the fact that we didn't cite the Schädel et al.
+Yes, the fact that we didn't cite the Sch?del et al.
 (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
-In our revision, we will significantly expand the discussion on this point, comparing our results to Schädel et al.
+In our revision, we will significantly expand the discussion on this point, comparing our results to Sch?del et al.
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
@@
 
+\medskip
 {\it 2.3.
@@
 
+\medskip
 {\it 2.4.
@@
 2011, Dutta et al.
-2006, and Schädel et al.
+2006, and Sch?del et al.
 2014.
@@
 (2006), although it concerns Siberian soils, is also a good comparison.
-We had not included Schädel et al.
+We had not included Sch?del et al.
 (2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it would also be a reasonable addition.
@@
 
+\medskip
 {\it 2.5. 2) L. 31: Permafrost thaws and does not melt }
@@
 
+\medskip
 {\it 2.6.
@@
 
+\medskip
 {\it 2.7.
@@
 
+\medskip
 {\it 2.8. 5) Add a table with soil properties such as bulk density, \%C etc. }
@@
 
+\medskip
 {\it 2.9.
 6) Why not include C/N as a variable in the statistical analysis?
-Schädel et al.
+Sch?del et al.
 2014 showed that C/N is a good predictor of C release and can be used as a scaling factor.
@@
 Currently C/N seems to be a poorer predictor than \%N.
-We will discuss this issue, referencing previous work such as SchŠdel et al (2014).
+We will discuss this issue, referencing previous work such as Sch?del et al (2014).
 {\bf See new lines TODO.}
 
+\medskip
 {\it 2.10.
@@
 
-That's a very good point-thank you-and will integrate well with an expanded comparison to the SchŠdel et al.
+That's a very good point-thank you-and will integrate well with an expanded comparison to the Sch?del et al.
 (2016) paper (cf.
@@
 
+\medskip
 {\it 2.11. 8) the conclusions might be a bit strong given the data and previous results published }
@@
 \newpage
-
-{\bf Response to Referee \#3
+\subsection*{Response to Referee \#3}
 
@@
 
+\medskip
 {\it 3.3.
@@
 The other referees both mentioned this as well.
-The fact that we didn't cite the Schädel et al.
+The fact that we didn't cite the Sch?del et al.
 (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
-In our revision, we will significantly expand this, discussing and comparing to Schädel et al.
+In our revision, we will significantly expand this, discussing and comparing to Sch?del et al.
 (2014, 2016) and Wickland et al.
@@
 
+\medskip
 {\it 3.4.
@@
 
+\medskip
 {\it 3.5. 22: Daily CO2 fluxes?
@@
 
+\medskip
 {\it 3.6. 29: Not really sure how the comparison as to the relative controls of T and moisture are evaluated. }
@@
 
+\medskip
 {\it 3.7. 50: see also updates in Hugelius et al. (2014)
@@
 
+\medskip
 {\it 3.8. 67: 'substantial variabilities between studies' WHY? }
@@
 
+\medskip
 {\it 3.9. 72: Yes, this is an important question, but given that this isn't measured in this study, perhaps this sentence should be omitted or re-written. }
@@
 
+\medskip
 {\it 3.10. 101: When did sampling occur?
@@
 
+\medskip
 {\it 3.11. 211: Please remember to complete DOI }
@@
 
+\medskip
 {\it 3.12. 215: Not sure what this value for soil dry mass indicates }
@@
 
+\medskip
 {\it 3.13. 216: Standard deviation for \%C and \%N is nearly 100\%. Check values. }
@@
 
+\medskip
 {\it 3.14. 229: add units
@@
 
+\medskip
 {\it 3.15. 253-254: So what variables were significant in predicting cumulative C emissions? }
@@
 
+\medskip
 {\it 3.16. 262: First mention of vegetation stress, remove, not clear how it's related. }
@@
 
+\medskip
 {\it 3.17. 270. Add '.' }
@@
 
+\medskip
 {\it 3.18. 271: Specify soil type in which these measurements were made (results not surprising for a forest soil) }
@@
 
+\medskip
 {\it 3.19. 272: What about results from Wickland et al. (2008). Study found threshold for moisture importance
@@
 
+\medskip
 {\it 3.20. 322-324: cool! }
@@
 
+\medskip
 {\it 3.21. 344-345: Specify that the results in Treat et al. (2015) were for anaerobic incubations and were thus likely to be much smaller. }
@@
 
+\medskip
 {\it 3.22. 347-348: See also Lee et al. (2012)
@@
 
+\medskip
 {\it 3.23. 383-384: 'specific weaknesses': vague
@@
 
+\medskip
 {\it 3.24. 393: 'taking them out of depth' rephrase. Also could use this argument for the section on CH4 production. }
@@
 
+\medskip
 {\it 3.25. Fig.1 : Edit figure to be color-blind friendly. }
@@
 
+\medskip
 {\it 3.26. Fig. 2,3: When did watering / moisture adjustment occur? Consider indicating with arrows and specifying in text. }
@@
 
+\medskip
 {\it 3.27. Fig. 4: Switch top and bottom panels as CO2 is always discussed before CH4. Also edit colors and patterns to be color-blind friendly. }

### 2016-11-15 manuscript/response_to_reviewer_comments.tex
Reviewer comments 1.6 and 2.9

@@
 
-This is an interesting suggestion, and an analysis we would be happy to include.
-As a preliminary step, before revising the manuscript, we have added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/18884c836d855b94417f7ae2cef910c577bc40b}) that examines if and how temperature and moisture sensitivities change over incubation time in this experiment.
-Currently we do not observe any evidence for changes in CO2 moisture sensitivity with time, and only weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time.
-We will discuss this issue, referencing previous work such as Reichstein et al.
-(2005).
-{\bf See new lines TODO.}
+This is an interesting suggestion, and an analysis we are happy to include.
+We added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/18884c836d855b94417f7ae2cef910c577bc40b}) that examines if and how temperature and moisture sensitivities change over incubation time in this experiment.
+We do not observe any evidence for changes in CO2 moisture sensitivity with time, and only weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time.
+{\bf See new lines LINES.}
 
@@
 (2006), although it concerns Siberian soils, may also prove a useful comparison.
-We had not included Sch?del et al.
+We had not included Schädel et al.
 (2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but it would also be a good addition.
@@
 This point was also made by Referee \#2 (comment 2.2).
-The fact that we didn't cite the Sch?del et al.
+The fact that we didn't cite the Schädel et al.
 (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
-In our revision, we will significantly expand this paragraph, discussing and comparing to Sch?del et al.
+In our revision, we will significantly expand this paragraph, discussing and comparing to Schädel et al.
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
@@
 {\it 2.2.
-It seems like this paper was published as a discussion paper before Sch?del et al.
+It seems like this paper was published as a discussion paper before Schädel et al.
 2016 was published and hence a discussion of the meta-analysis was not possible but should be addressed in the revisions.
@@
 This point was also made by Referee \#1 (comment 1.18).
-Yes, the fact that we didn't cite the Sch?del et al.
+Yes, the fact that we didn't cite the Schädel et al.
 (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
-In our revision, we will significantly expand the discussion on this point, comparing our results to Sch?del et al.
+In our revision, we will significantly expand the discussion on this point, comparing our results to Schädel et al.
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
@@
 2011, Dutta et al.
-2006, and Sch?del et al.
+2006, and Schädel et al.
 2014.
@@
 (2006), although it concerns Siberian soils, is also a good comparison.
-We had not included Sch?del et al.
+We had not included Schädel et al.
 (2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it would also be a reasonable addition.
@@
 6) Why not include C/N as a variable in the statistical analysis?
-Sch?del et al.
+Schädel et al.
 2014 showed that C/N is a good predictor of C release and can be used as a scaling factor.
@@
 
-This is an interesting suggestion, and an analysis we would be happy to include.
-As a preliminary step, before revising the manuscript, we have added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/426a91e1bbd21200718b334d3295fbef40a1ea6}) to compute C/N and examine its significance as a predictor.
+This is an interesting suggestion.
+We added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/426a91e1bbd21200718b334d3295fbef40a1ea6}) to compute C/N and examine its significance as a predictor.
 Currently C/N seems to be a poorer predictor than \%N.
-We will discuss this issue, referencing previous work such as Sch?del et al (2014).
-{\bf See new lines TODO.}
+We now discuss this issue, referencing previous work such as Schädel et al (2014).
+{\bf See new lines LINES.}
 
@@
 
-That's a very good point-thank you-and will integrate well with an expanded comparison to the Sch?del et al.
+That's a very good point-thank you-and will integrate well with an expanded comparison to the Schädel et al.
 (2016) paper (cf.
@@
 The other referees both mentioned this as well.
-The fact that we didn't cite the Sch?del et al.
+The fact that we didn't cite the Schädel et al.
 (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
-In our revision, we will significantly expand this, discussing and comparing to Sch?del et al.
+In our revision, we will significantly expand this, discussing and comparing to Schädel et al.
 (2014, 2016) and Wickland et al.

### 2016-11-15 manuscript/cpcrw_manuscript.Rmd
Reviewer comments 1.6 and 2.9

@@
 In the linear mixed-effects model, CO~2~ was strongly influenced by incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 1**).
+Temperature sensitivity decreased significantly (`r pclean(co2_temp_time)`) over the course of the incubation, while moisture sensitivity was unaffected by time.
 Percent C and percent N were highly correlated (r = `r pn(cor(cndata$C_percent, cndata$N_percent), 2)`) for these cores.
@@
 
+Temperature and moisture senstivity was generally invariant over time: 
 The cumulative production of C from CO~2~ (**Figure 4**) was over `r minmag` order of magnitudes higher than that from CH~4~, with CO~2~:CH~4~ C ratios ranging from `r minratio` million in the `r mintemp` °C "`r maxtrt`" treatment, to `r maxratio` million in the `r maxtemp` °C "`r maxtrt`" treatment.
@@
 
+discuss - changes over time (see new reuslts)
+
 More surprisingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 than at 20 °C (TODO-fix sentence).
@@
 These effect are likely due to several one or more mechanisms involving soil pH, ligninase enzymes, and phenol oxidase activity {Luo, 2006 #2443}.
+
+discuss - C/N was not a good predictor (they're highly correlated here) ref Schadel 2014
 

### 2016-11-15 R/5-analyze.R

@@ -460,6 +460,8 @@ m_co2_lme_thirds <- update(m_co2_lme, ~ . +
                              WC_gravimetric * third + 
                              Temperature * third)
 step_co2_lme_thirds <- MASS::stepAIC(m_co2_lme_thirds, direction = "both")
+co2_temp_time <- step_co2_lme_thirds$coefficients$fixed["Temperature:third(34.3,67.7]"]
+
 
 printlog(SEPARATOR)
 printlog("Fitting CH4 model...")
@@ -477,6 +479,7 @@ m_ch4_lme_thirds <- update(m_ch4_lme, ~ . +
                              WC_gravimetric * third + 
                              Temperature * third)
 step_ch4_lme_thirds <- MASS::stepAIC(m_ch4_lme_thirds, direction = "both")
+ch4_wc_time <- step_ch4_lme_thirds$coefficients$fixed["WC_gravimetric:third(34.3,67.7]"]
 
 # -----------------------------------------------------------------------------
 # Compute outlier and exclusion numbers, to report in manuscript

### 2016-11-15 manuscript/response_to_reviewer_comments.tex
Report AIC (comment 2.7)

@@
 Vegetation and ecosystem type is a significant factor that is not well explored here, we agree.
-We will add some points about this, noting for example the work of H?gberg et al.
+We will add some points about this, noting for example the work of Högberg et al.
 (2007) on soil microbial community composition, the relationship of vegetation type and soil respiration (Raich and Schlesinger 1992, Bond-Lamberty and Thomson 2010), and how plant functional types affect ecosystem response more generally (Chapin et al.
@@
 Jassal et al.
-2008), but the field is rife with contradictory results (von L?tzow and K?gel-Knabner 2009).
+2008), but the field is rife with contradictory results (von Lützow and K?gel-Knabner 2009).
 We will make this paragraph more specific in all these areas.
@@
 
-Thank you. We appreciate the useful suggestions, and will provide these additional details in our revised manuscript. {\bf See new lines TODO.}
+Thank you. We appreciate the useful suggestions, and will provide these additional details in our revised manuscript. {\bf See new lines LINES.}
 
@@
 
-Thanks. There was a great of variability (obviously), but distributed throughout the data set-i.e., this isn't driven by one or two outliers. We will re-check, however. {\bf See new lines TODO.}
+Thanks. There was a great of variability (obviously), but distributed throughout the data set-i.e., this isn't driven by one or two outliers.
 
@@
 
-These will be fixed. {\bf See new lines TODO.}
+These have all been fixed. {\bf See new lines LINES.}
 
@@
 
-Thanks; we will do so. {\bf See new lines TODO.}
+Thanks;we have done so. {\bf See new lines LINES.}
 

### 2016-11-15 manuscript/cpcrw_manuscript.Rmd
Report AIC (comment 2.7)

@@
 ---
-output: word_document
+output:
+  html_document: default
+  word_document: default
 ---
@@
 
-Carbon dioxide fluxes during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 1)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 1)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 1)`) over the 100 days.
-CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 2)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 2)`).
+Carbon dioxide fluxes during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 1)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 1)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`) µg C g C^-1^ day^-1^, with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 1)`) µg C g C^-1^ day^-1^ over the 100 days.
+CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 2)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`) ng C g C^-1^ day^-1^, with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 2)`) ng C g C^-1^ day^-1^.
 
 These means conceal considerable variability over the course of the incubation (**Figures 2 and 3**).
-In the linear mixed-effects model, CO~2~ was strongly influenced by incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 1**).
+In the linear mixed-effects model (AIC = `r pn(AIC(step_co2_lme), 1)`), CO~2~ was positively correlated with incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 1**).
 Temperature sensitivity decreased significantly (`r pclean(co2_temp_time)`) over the course of the incubation, while moisture sensitivity was unaffected by time.
@@
 
-Methane fluxes were most strongly correlated with percent N, while water content exhibited significant interactions with percent N and DOC (**Table 2**).
-Neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
+Methane fluxes were positively correlated with percent N, while water content exhibited significant interactions with percent N and DOC (**Table 2**).
+The model had little predictive power (AIC = `r pn(AIC(step_ch4_lme), 1)`), and neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
 
@@
 This is true more generally: for example, Treat et al.
-found a median CO~2~:CH~4~ production ratio of 387 for boreal sites, far lower than (but consistent with) our observed ratios of several million.
+found a median CO~2~:CH~4~ production ratio of 387 for anaerobic incubations of boreal soils.
+This is naturally far lower than our observed aerobic (and thus high-CO~2~) ratios, but nonetheless consistent with them.
 Thus we see little opportunity for CH~4~ to be a significant contributor to these upland soils' C fluxes and climate feedback risk, even accounting for the stronger radiative forcing of this gas.

### 2016-11-17 manuscript/cpcrw_manuscript.Rmd
Added new Table 1 per reviewer request

@@
 These means conceal considerable variability over the course of the incubation (**Figures 2 and 3**).
-In the linear mixed-effects model (AIC = `r pn(AIC(step_co2_lme), 1)`), CO~2~ was positively correlated with incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 1**).
+In the linear mixed-effects model (AIC = `r pn(AIC(step_co2_lme), 1)`), CO~2~ was positively correlated with incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 2**).
 Temperature sensitivity decreased significantly (`r pclean(co2_temp_time)`) over the course of the incubation, while moisture sensitivity was unaffected by time.
@@
 
-Methane fluxes were positively correlated with percent N, while water content exhibited significant interactions with percent N and DOC (**Table 2**).
+Methane fluxes were positively correlated with percent N, while water content exhibited significant interactions with percent N and DOC (**Table 3**).
 The model had little predictive power (AIC = `r pn(AIC(step_ch4_lme), 1)`), and neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
@@
 
-Somewhat unexpectedly, percent soil N was very significantly and positively correlated with both CO~2~ and CH~4~ fluxes (**Tables 1 and 2**).
+Somewhat unexpectedly, percent soil N was very significantly and positively correlated with both CO~2~ and CH~4~ fluxes (**Tables 2 and 3**).
 N interacts with microbial respiration via a number of complex, interactive, and still unclear mechanisms {Luo, 2006 #2443}, including reductions in belowground plant allocation, shifts in energy source or population of the saprotrophic community {Saiya-Cork, 2002 #3595} that leave it less capable of decomposing recalcitrant compounds, and perhaps abiotic stabilization mechanisms {Janssens, 2010 #2518}.
@@
 For example, Weiss et al.
-{, 2015 #4317} found CO~2~ production from Siberian Yedoma permafrost samples to be correlated with both percent C and N, consistent with our active-layer results (**Table 1**).
+{, 2015 #4317} found CO~2~ production from Siberian Yedoma permafrost samples to be correlated with both percent C and N, consistent with our active-layer results (**Table 2**).
 In an incubation of 84 North American soils, Colman and Schimel found that percent C and percent N were highly (R^2^ = 0.84) correlated and that the former was a significant predictor of microbial respiration--similar to our findings (C to N correlation of `r pn(cor(cndata$C_percent, cndata$N_percent), 2)`), except that percent N out-predicted percent C in this study.
@@
 
-**Table 1.** Linear mixed-effects model parameters, testing effects of temperature (°C), gravimetric water content (unitless), soil C (%), soil N (%),  and dissolved organic carbon (mg kg^-1^) on individual core CO~2~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^); a colon (":") indicates an interaction.
-Dependent variable has units of log(µg C g C^-1^ day^-1^).
-Columns include parameter value; standard error (SE); degrees of freedom (DF); T statistic; and P value.
+**Table 1.** New one.
 
 ```{r table1, echo=FALSE}
-summary(step_co2_lme)$tTable %>% 
-  round(3) %>% 
-  as.data.frame() ->
-  t1
-t1$`p-value` <- pclean(t1$`p-value`, printP = FALSE)
-knitr::kable(t1, digits = 3, col.names = c("Value", "SE", "DF", "T", "P"))
+knitr::kable(table1)
 ```
 
-**Table 2.** Linear mixed-effects model parameters, testing effects of temperature (°C), gravimetric water content (unitless), soil N (%),and dissolved organic carbon (DOC, mg kg^-1^) on log-transformed, individual core CH~4~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^); a colon (":") indicates an interaction.
+**Table 2.** Linear mixed-effects model parameters, testing effects of temperature (°C), gravimetric water content (unitless), soil C (%), soil N (%),  and dissolved organic carbon (mg kg^-1^) on individual core CO~2~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^); a colon (":") indicates an interaction.
 Dependent variable has units of log(µg C g C^-1^ day^-1^).
@@
   as.data.frame() ->
-  t2
-t2$`p-value` <- pclean(t2$`p-value`, printP = FALSE)
-knitr::kable(t2, digits = 3, col.names = c("Value", "SE", "DF", "T", "P"))
+  table2
+table2$`p-value` <- pclean(table2$`p-value`, printP = FALSE)
+knitr::kable(table2, digits = 3, 
+             col.names = c("Value", "SE", "DF", "T", "P"))
+```
+
+**Table 3.** Linear mixed-effects model parameters, testing effects of temperature (°C), gravimetric water content (unitless), soil N (%),and dissolved organic carbon (DOC, mg kg^-1^) on log-transformed, individual core CH~4~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^); a colon (":") indicates an interaction.
+Dependent variable has units of log(µg C g C^-1^ day^-1^).
+Columns include parameter value; standard error (SE); degrees of freedom (DF); T statistic; and P value.
+
+```{r table3, echo=FALSE}
+summary(step_co2_lme)$tTable %>% 
+  round(3) %>% 
+  as.data.frame() ->
+  table3
+table3$`p-value` <- pclean(table3$`p-value`, printP = FALSE)
+knitr::kable(table3, digits = 3, 
+             col.names = c("Value", "SE", "DF", "T", "P"))
 ```

### 2016-11-17 R/5-analyze.R

@@ -8,7 +8,7 @@ SCRIPTNAME  	<- "5-analyze.R"
 
 library(nlme)       # 3.1-128
 #library(MASS)       # 7.3-45
-
+library(tidyr)      # 0.6.0
 library(broom)      # 0.4.1
 library(reshape2)   # 1.4.1
 library(agricolae)  # v1.2.4
@@ -84,6 +84,7 @@ cndata %>%
   summarise_each(funs(mean, sd), DOC_mg_kg, C_percent, N_percent, CN) %>%
   print ->
   cndata_summary
+
 # This isn't working via dplyr - compute manually
 cndata_summary$CN_mean <- mean(cndata$CN, na.rm = TRUE)
 cndata_summary$CN_sd <- sd(cndata$CN, na.rm = TRUE)
@@ -315,17 +316,17 @@ fd_cumulative_r3 <- fd_cumulative_r3[complete.cases(fd_cumulative_r3),]
 
 printlog("Fitting CO2 model for cumulative fluxes...")
 m_co2_cum <- lm(cum_flux_mgC_gC ~ Temperature + Treatment +
-                   N_percent + DOC_mg_kg + CN, # C_percent + 
-                 data = fd_cumulative_r3, 
-                 subset = Gas == "CO2")
+                  N_percent + DOC_mg_kg + CN, # C_percent + 
+                data = fd_cumulative_r3, 
+                subset = Gas == "CO2")
 step_co2_cum <- MASS::stepAIC(m_co2_cum, direction = "both")
 print(summary(m_co2_cum))
 
 printlog("Fitting CH4 model for cumulative fluxes...")
 m_ch4_cum <- lm(cum_flux_mgC_gC ~ Temperature + Treatment +
-                   N_percent + DOC_mg_kg + CN, # C_percent + 
-                 data = fd_cumulative_r3, 
-                 subset = Gas == "CH4")
+                  N_percent + DOC_mg_kg + CN, # C_percent + 
+                data = fd_cumulative_r3, 
+                subset = Gas == "CH4")
 step_ch4_cum <- MASS::stepAIC(m_ch4_cum, direction = "both")
 print(summary(step_ch4_cum))
 
@@ -486,6 +487,58 @@ ch4_wc_time <- step_ch4_lme_thirds$coefficients$fixed["WC_gravimetric:third(34.3
[…]

### 2016-11-17 manuscript/response_to_reviewer_comments.tex
Added BD, caption, tweaks to Table 1

@@
 This will be clarified in the methods, particularly lines 162-, which we agree were ambiguous.
-A table summarizing the moisture content, pH, soil C and N, and bulk density data is an excellent idea and will be added.
+A new Table 1 now summarizes a variety of physical and flux data by treatment.
 Reviewer 2 also raised the idea of looking at C/N, and we have responded to that suggestion in detail (see comment 2.9).
@@
 
-This will be included in the new table (see comment 1.13 above).
+This has been included in the new Table 1 (see comment 1.13 above).
 
@@
 
-This useful suggestion was also made by Referee \#1 (comment 1.13). We will do so. {\bf See new lines TODO.}
+This useful suggestion was also made by Referee \#1 (comment 1.13). We have done so, in a new Table 1. {\bf See new lines TODO.}
 

### 2016-11-17 manuscript/cpcrw_manuscript.Rmd
Added BD, caption, tweaks to Table 1

@@
 
-**Table 1.** New one.
+**Table 1.** Summary of dissolved organic carbon (DOC), percent C, percent N, and CO~2~ and CH~4~ fluxes by treatment (with pre-incubation values shown in the right-hand column).
+Each column summarizes (mean ± s.d.) a separate group of N=6 cores.
 
 ```{r table1, echo=FALSE}
-knitr::kable(table1)
+knitr::kable(table1, align = "r")
 ```

### 2016-11-17 R/5-analyze.R

@@ -1,5 +1,6 @@
 # Final analysis script: statistical tests, final plots, etc.
 # Called by the R Markdown document (the manuscript).
+# This has become a really big script...ugh
 # Ben Bond-Lamberty January 2016
 
 source("R/0-functions.R")
@@ -19,6 +20,8 @@ library(agricolae)  # v1.2.4
 openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # open log
 
 printlog("Welcome to", SCRIPTNAME)
+
+treatment_data <- read_csv(TREATMENTS, skip = 1)
 
 coredata <- read_csv(COREDATA_FILE, skip = 1)
 
@@ -77,7 +80,7 @@ read_csv(CNDATA_FILE) %>%
 cndata_cor <- cor(cndata_orig[-1], use = "complete.obs")
 
 cndata_orig %>%
-  left_join(read_csv(TREATMENTS, skip = 1), by = "Core") ->
+  left_join(treatment_data, by = "Core") ->
   cndata
 
 cndata %>%
@@ -492,11 +495,13 @@ fluxn <- nrow(fluxdata)
 # Make Table 1 - a lot of work for such a dinky thing!
 printlog("Making table 1...")
 TABLE1_ROUNDING <- 2
+# Summarise the relevant variables for each core (across entire incubation)
 fluxdata %>%
   group_by(Treatment, Core, Gas) %>%
   summarise_each(funs(mean), flux_µgC_gC_day1, DOC_mg_kg, C_percent, N_percent) ->
   table1_fluxdata
 
+# Pull out the two gases and rename them into separate variables
 table1_fluxdata %>%
   filter(Gas == "CO2") %>% 
   rename(CO2_µgC_gC_day1 = flux_µgC_gC_day1) ->
@@ -506,22 +511,33 @@ table1_fluxdata %>%
   rename(CH4_µgC_gC_day1 = flux_µgC_gC_day1) ->
   table1_fluxdata_ch4
 
+# Compute bulk density
+coredata %>%
+  group_by(Core) %>%
+  summarise(BD = SoilDryMass_g / SoilVolume_cm3) %>%
+  left_join(treatment_data, by = "Core") -> 
+  bd_data
+
+# Combine the pre-incubation data with incubation data
 cndata %>%
   filter(is.na(Treatment)) %>%
[…]

### 2016-11-18 manuscript/response_to_reviewer_comments.tex
Responding to reviewer comments
Text clarifications, reworked figure 4, changed to a
colorblind-friendly palette, and more.

@@
 Reviewer 2 also raised the idea of looking at C/N, and we have responded to that suggestion in detail (see comment 2.9).
-{\bf See new lines TODO.}
+{\bf See new lines LINES.}
 
@@
 
-It's true that Colman and Schimel (2014) include only a few studies that could be termed boreal (from Maine, USA).
-We will either remove or de-emphasize/qualify this comparison.
-{\bf See new lines TODO.}
+It's true that Colman and Schimel (2014) include only a few studies that could be termed boreal (from Maine, USA). We have removed this comparison.
 
@@
 
-This useful suggestion was also made by Referee \#1 (comment 1.13). We have done so, in a new Table 1. {\bf See new lines TODO.}
+This useful suggestion was also made by Referee \#1 (comment 1.13). We have done so, in a new Table 1.
 
@@
 
-We will add a sentence of caveats, noting in particular the useful but incremental nature of this study. {\bf See new lines TODO.}
+We will add a sentence of caveats, noting in particular the useful but incremental nature of this study. {\bf See new lines LINES.}
 
@@
 
-We will clarify these points. {\bf See new lines TODO.}
+These points have been clarified, except for the last, as we feel it's already clear and unambiguous.
 
@@
 
-Agreed, this was more of a qualitative statement than a quantitative one. We will be more precise. {\bf See new lines TODO.}
+This statement has been reworded to remove the comparison. {\bf See new lines LINES.}
 
@@
 
-We will expand on this point, pointing out that such variability originates from factors such soil effects on drainage, phase changes, different experimental protocols, etc. {\bf See new lines TODO.}
+We have expanded on this point, pointing out that such variability originates from factors such as differences in soil type, antecedent conditions, phase changes, experimental protocols, etc.
+{\bf See new lines LINES.}
 
@@
 
-Good point; we will do so. {\bf See new lines TODO.}
+This sentence has been removed.
 
@@
 every timepoint shown in Figure 1.
-This will be clarified, and we will consider if adding this to Figures 2 and 3 if readable and useful.
-{\bf See new lines TODO.}
+This will be clarified.
+{\bf See new lines LINES.}
 
@@
 
-It's just useful, we think, to give readers a good sense of sample mass.
+It's just useful, we think, to give readers a good sense of sample size.
 
@@
 
-This will be fixed. {\bf See new lines LINES.}
+This has been fixed. {\bf See new lines LINES.}
 
@@
 
-Upland Cryosols; we will clarify this. {\bf See new lines LINES.}
+Upland Cryosols; we have clarified this. {\bf See new lines LINES.}
 
@@
 
-Thanks;we have done so. {\bf See new lines LINES.}
+Thanks; we have done so. {\bf See new lines LINES.}
 
@@
 
-Good points! We will clarify and note this. {\bf See new lines TODO.}
+This awkward language has been removed, and a note about lag effects added. {\bf See new lines LINES.}
 
@@
 
-Yes, it's awkward; we will reword. Thanks for the suggestion. {\bf See new lines TODO.}
+We have reworded this. {\bf See new lines LINES.}
 
@@
 
-We thought we were already doing so in using the {\tt RColorBrewer} package, not the default palette of {\tt ggplot2}, but will investigate and fix this if necessary. {\bf See new lines TODO.}
+We thought we were already doing so in using the {\tt RColorBrewer} package, not the default palette of {\tt ggplot2}, but have shifted to using a color-blind friendly palette from \url{http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette} in a figures.
 
@@
 
-Moisture adjustment was done after every mass measurement, i.e.
-every timepoint shown in Figure 1.
-This will be clarified, and we will consider if adding this to Figures 2 and 3 if readable and useful.
-{\bf See new lines TODO.}
+Moisture adjustment was done after every mass measurement, i.e. every timepoint shown in Figure 1. This has been clarified. {\bf See new lines LINES.}
 
@@
 
-Good point. Re colors, see our response to 3.25 above. {\bf See new lines TODO.}
+Good point-fixed. Re colors, see our response to 3.25 above.
 

### 2016-11-18 manuscript/cpcrw_manuscript.Rmd
Responding to reviewer comments
Text clarifications, reworked figure 4, changed to a
colorblind-friendly palette, and more.

@@
 Gravimetric water content was `r pn(gwcbegin$mean, 2)` ± `r pn(gwcbegin$sd, 2)` (unitless) at the beginning of the incubation; cores at field moisture were unchanged at the end, but drought cores had declined to `r pn(gwcendmean["Drought"], 2)` ±  `r pn(gwcendsd["Drought"], 2)`.
-Carbon dioxide fluxes were strongly influenced by incubation chamber temperature, core water content, and percent soil nitrogen, and had a temperature sensitivity (Q~10~) of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively.
-Methane emissions were most strongly correlated with percent nitrogen, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
+Daily CO~2~ fluxes were positively correlated with incubation chamber temperature, core water content, and percent soil nitrogen, and had a temperature sensitivity (Q~10~) of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively.
+Daily CH~4~ emissions were most strongly correlated with percent nitrogen, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
 The cumulative production of C from CO~2~ was over `r minmag` orders of magnitudes higher than that from CH~4~.
-These results suggest that deep active-layer soils may be much more sensitive to changes in moisture than to temperature, a critical factor as discontinuous permafrost thaws [COM2.5] in interior Alaska.
+These results suggest that deep active-layer soils may be quite sensitive to changes in soil moisture, a critical factor as discontinuous permafrost thaws [COM2.5] in interior Alaska.
 Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results provide insight into their future dynamics and feedback potential with future climate change.
@@
 Laboratory incubations, field observations, and meta-analyses have documented increased fluxes of CO~2~, and under some conditions of CH~4~ {Olefeldt, 2013 #3456}, with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
-Greenhouse gas (GHG) responses to wetting and thawing dynamics are much less certain, with substantial variability between studies {Kim, 2012 #2863}.
+Greenhouse gas (GHG) responses to wetting and thawing dynamics are much less certain, with substantial variability between studies arising from differences in soil type, antecedent conditions, phase changes, experimental protocols, etc.
+{Kim, 2012 #2863}.
 The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and such interactions are critical to examine in the course of long-term incubation experiments {Elberling, 2013 #3720}.
 Decadal warming and drying trends in Alaska {Bieniek, 2014 #4389} may counteract these effects, however.
-A critical question, then, is how the structure, chemistry, and microbial communities of current active-layer soils will respond to almost-certain temperature and moisture changes in the future {Xue, 2016 #4361}.
 
@@
 The "drought" cores were mounted on dry plates, but not allowed to drop below 5% water content.
-When necessary, cores received additional wetting from the top to maintain their water status at the desired level.
+After each flux measurement, cores received additional wetting from the top to maintain their water status at the desired level.
 
@@
 {, 2015 #4317} found CO~2~ production from Siberian Yedoma permafrost samples to be correlated with both percent C and N, consistent with our active-layer results (**Table 2**).
-In an incubation of 84 North American soils, Colman and Schimel found that percent C and percent N were highly (R^2^ = 0.84) correlated and that the former was a significant predictor of microbial respiration--similar to our findings (C to N correlation of `r pn(cor(cndata$C_percent, cndata$N_percent), 2)`), except that percent N out-predicted percent C in this study.
 *In situ* respiration rates have also been shown to be negatively correlated with C:N at large spatial scales {Allaire, 2012 #4423}.
@@
 Laboratory experiments offer precise control, but lack the *in situ* nature of field manipulations {Sistla, 2013 #3666}, raising uncertainties to what degree their results can be extrapolated.
-Soils isolated during incubation may, for example, underestimate temperature sensitivity of respiration {Podrebarac, 2016 #4365}.
-[COM1.24] Nonetheless, the controlled environments of incubations provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
+Soils isolated during incubation may, for example, underestimate temperature sensitivity of respiration {Podrebarac, 2016 #4365} or exhibit lag effects {Treat, 2015 #4217}.
+Nonetheless, the controlled environments of incubations provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
 
 We focused on an experimental drought, rather than flooding, because of the well-drained nature of the field site: it is unlikely that the mid-slope forest we sampled in will ever suffer from thermokarst or excessive soil moisture, but too-dry conditions are a serious possibility in this low-precipitation ecosystem {Barber, 2000 #1320}.
-In addition, the soils here are not surface layer soils (where the majority of microbial activity and C mineralization of labile C takes place); taking them out of depth (where they are less exposed to O~2~, for example) may significantly change the abiotic conditions to which the microbial community is adapted.
+In addition, the soils here are not surface layer soils (where the majority of microbial activity and C mineralization of labile C takes place); removing them from *in situ* conditions (where they are less exposed to O~2~, for example) may significantly change the abiotic conditions to which the microbial community is adapted.
 However, focusing on the active layer provides crucial information about the potential loss of C from these soils, a risk that needs to be well understood as permafrost degradation leads to expansions in the depth of the active layer across the Arctic.
@@
 In this laboratory experiment, we found that CO~2~ fluxes were strongly influenced by temperature and water content, and correlated with soil C and N, while CH~4~ fluxes were much smaller and not sensitive to temperature or water content in these well-drained mineral soils.
-This suggests that understanding how soil moisture might change with spatially variable permafrost degradation, how soil biota will respond to these changes, and how models should treat soil organic matter decomposition with respect to multiple and interacting drivers are all critical areas of research going forward.
-Further controlled field and laboratory studies, ideally tightly integrated with modeling experiments, are critical to understand GHG emission dynamics from high-latitude soils.
+These results add to a growing body of Arctic permafrost and active layer incubation literature, and underscore the importance of understanding moisture effects on CO~2~ fluxes in particular.
+How soil moisture might change with spatially variable permafrost degradation, how soil biota will respond to these changes, and how models should treat soil organic matter decomposition with respect to multiple and interacting drivers are all critical areas of research going forward.
+Further controlled field and laboratory studies, ideally tightly integrated with modeling experiments, are important to understand GHG emission dynamics from high-latitude soils.
 
@@
 
-**Figure 4.** Cumulative C fluxes (mg g C^-1^) over the incubation, by gas (CH~4~ and CO~2~, top and bottom panels respectively) and treatment (columns).
+**Figure 4.** Cumulative C fluxes (mg g C^-1^) over the incubation, by gas (CO~2~ and CH~4~, top and bottom panels respectively) and treatment (columns).
 Letters within a panel indicate significant differences based on Tukey's HSD.

### 2016-11-18 R/0-functions.R

@@ -49,6 +49,11 @@ FLUXDATA_CUM_CORE_FILE <- file.path(OUTPUT_DIR, "fluxdata_cum_core.csv")
 COREDATA_FILE          <- "data/drymasses.csv"
 CNDATA_FILE            <- "data/uga-soil_samples.csv"
 
+# -----------------------------------------------------------------------------
+# Miscellany
+# Colorblind-friendly palette
+# from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette
+CBPALETTE <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 SEPARATOR		  <- "-------------------"
 
 

### 2016-11-18 R/5-analyze.R

@@ -143,7 +143,7 @@ figureA <- ggplot(fluxdata_orig, aes(inctime_days, WC_gravimetric, color=Treatme
   geom_point() + geom_line() + 
   facet_grid(~Temperature) + 
   xlab("Incubation day") + ylab("Gravimetric water content (fraction dry mass)") +
-  scale_color_brewer(palette = "Set1")
+  scale_color_manual(values = CBPALETTE)
 save_plot("figureA", figureA)
 
 fluxdata$incday <- floor(fluxdata$inctime_days)
@@ -276,17 +276,21 @@ fluxdata_cumulative <- read_csv(FLUXDATA_CUM_FILE)
 fluxdata_cumulative$Treatment <- factor(fluxdata_cumulative$Treatment, 
                                         levels = c("Field moisture", "Controlled drought", "Drought"))
 fluxdata_cumulative$Temperature <- as.factor(fluxdata_cumulative$Temperature)
+fluxdata_cumulative$Gas <- factor(fluxdata_cumulative$Gas, levels = c("CO2", "CH4"))
 
-figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC_gC, fill = Treatment)) + 
-  geom_bar(stat = 'identity', position = position_dodge()) +
-  geom_errorbar(aes(color = Treatment, 
-                    ymin = cum_flux_mgC_gC * 0.9, 
-                    ymax = cum_flux_mgC_gC + cum_flux_mgC_gC_sd), 
-                position = position_dodge(0.9), width = 0.4) +  
-  facet_grid(Gas ~ ., scales = "free") +
-  ylab(expression(Cumulative~C~(mg~g~C^{-1}))) +
-  scale_fill_brewer(palette = "Set1") + scale_color_brewer(palette = "Set1")
-save_plot("figureD", figureD)
+suppressWarnings({
+  figureD <- ggplot(fluxdata_cumulative, aes(Temperature, cum_flux_mgC_gC, fill = Treatment)) + 
+    geom_bar(stat = 'identity', position = position_dodge()) +
+    geom_errorbar(aes(color = Treatment, 
+                      ymin = cum_flux_mgC_gC * 0.9, 
+                      ymax = cum_flux_mgC_gC + cum_flux_mgC_gC_sd), 
+                  position = position_dodge(0.9), width = 0.4) +  
+    facet_grid(Gas ~ ., scales = "free") +
[…]

### 2016-11-18 manuscript/cpcrw_manuscript.Rmd
Final tweaks to new Table 1

@@
 Mean C:N was `r pn(cndata_summary["C_percent_mean"]  / cndata_summary["N_percent_mean"], 1)`.
-Neither temperature nor moisture treatment exerted any significant effect on these highly variable properties (P > 0.1 for all).
+Neither temperature nor moisture treatment exerted any significant effect (P > 0.1 for all) on these highly variable properties (**Table 1**).
 
@@
 
-**Table 1.** Summary of dissolved organic carbon (DOC), percent C, percent N, and CO~2~ and CH~4~ fluxes by treatment (with pre-incubation values shown in the right-hand column).
-Each column summarizes (mean ± s.d.) a separate group of N=6 cores.
+**Table 1.** Summary of dissolved organic carbon (DOC), percent C, percent N, and CO~2~ and CH~4~ fluxes by treatment.
+Each column summarizes (mean ± s.d.) 12 cores, combining two groups of N=6 at each incubation temperature, except for the pre-incubation values shown in the right-hand column (N=6).
 

### 2016-11-18 R/5-analyze.R

@@ -542,7 +542,8 @@ table1_data %>%
   group_by(Treatment) %>%
   summarise_all(mean, na.rm = TRUE) %>%
   melt(id.vars = "Treatment", variable.name = "Variable") %>%
-  mutate(value = pn(value, TABLE1_ROUNDING)) ->
+  mutate(finite = is.finite(value),
+         value = pn(value, TABLE1_ROUNDING)) ->
   table1_means
 
 table1_data %>%
@@ -556,7 +557,7 @@ table1_means$value_sd <- table1_sds$value
 
 # Jeez, finally! Give cells a nice x ± y format
 table1_means %>%
-  mutate(entry = paste(value, "±", value_sd)) %>%
+  mutate(entry = ifelse(finite, paste(value, "±", value_sd), "-")) %>%
   dcast(Variable ~ Treatment, value.var = "entry") -> 
   table1
 

### 2016-11-18 manuscript/response_to_reviewer_comments.tex
Comment 1.9 - vegetation-soil interactions

@@
 
-Vegetation and ecosystem type is a significant factor that is not well explored here, we agree.
-We will add some points about this, noting for example the work of Högberg et al.
-(2007) on soil microbial community composition, the relationship of vegetation type and soil respiration (Raich and Schlesinger 1992, Bond-Lamberty and Thomson 2010), and how plant functional types affect ecosystem response more generally (Chapin et al.
-1996).
-{\bf See new lines TODO.}
+Vegetation and ecosystem type is a significant factor that is not well explored here, we agree. We have added some points about this in the introduction and discussion. {\bf See new lines LINES.}
 
@@
 Reviewer 2 also raised the issue (comment 2.4) of our incomplete citation of relevant literature.
-In our revision, we will make better use of references such as Hugelius et al.
-(2014) and Wickland et al.
-(2008), both in the introduction and discussion.
-{\bf See new lines TODO.}
+The Tarnocai reference has been replaced by one to Hugelius et al.
+(2014), and the Olefeldt sentence clarified.
+{\bf See new lines LINES.}
 

### 2016-11-18 manuscript/cpcrw_manuscript.Rmd
Final tweaks to new Table 1
Comment 1.9 - vegetation-soil interactions

@@
 The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such any such feedback remain highly uncertain {Schuur, 2015 #4319}.
-While northern soils hold enormous quantities {Tarnocai, 2009 #2231} of potentially mineralizable soil organic carbon (SOC), vegetation and succession dynamics (for example, thermal insulation by mosses) promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}{Turetsky, 2012 #3405}.
-Such dynamics may however be disrupted by increased fire disturbance, particularly with more-intense fires {Johnstone, 2010 #2465}{Genet, 2013 #4412}.
-In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}{Karhu, 2014 #4124}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
+While northern soils hold enormous quantities {Hugelius, 2014 #3978} of potentially mineralizable soil organic carbon (SOC), vegetation and succession dynamics (for example, thermal insulation by mosses) promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}{Turetsky, 2012 #3405}.
+Vegetation type also influences SOC quality and quantity, with microbial communities {Högberg, 2007 #4482}, soil respiration {Raich, 2000 #596}, and SOC  all linked to aboveground factors such as woody versus nonwoody stems, deciduous versus evergreen canopies, and the presence of nitrogen-fixing plants.
+
+These feedbacks between vegetation type, ground cover, permafrost, and SOC may be disrupted by increased fire disturbance, particularly after more-intense fire events {Johnstone, 2010 #2465}{Genet, 2013 #4412}.
+Even absent disturbance, the stability of SOC is highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}{Karhu, 2014 #4124}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
 
 Temperature and moisture typically have strong and often interactive influences on soil GHG emissions.
-Laboratory incubations, field observations, and meta-analyses have documented increased fluxes of CO~2~, and under some conditions of CH~4~ {Olefeldt, 2013 #3456}, with rising temperature {Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
-Greenhouse gas (GHG) responses to wetting and thawing dynamics are much less certain, with substantial variability between studies arising from differences in soil type, antecedent conditions, phase changes, experimental protocols, etc.
+Laboratory incubations, field observations, and meta-analyses have documented changing greenhouse gas (GHG) fluxes with rising temperature {Olefeldt, 2013 #3456}{Davidson, 2006 #1669}{Hashimoto, 2015 #4231}{Treat, 2015 #4217}.
+GHG responses to wetting and thawing dynamics exhibit substantial variability between studies, probably due to differences in soil type, antecedent conditions, phase changes, experimental protocols, etc.
 {Kim, 2012 #2863}.
@@
 discuss - changes over time (see new reuslts)
+discuss - upland mixed-canopy boreal forest here; careful about extrapolation
 

### 2016-11-18 manuscript/response_to_reviewer_comments.tex
Addressed a variety of minor comments from Reviewers 2 and 3

@@
 
-That's a good point-thank you. We will qualify this sentence, better tying it to the specific measurements made in this study (C, N, respiration rates, stress response). {\bf See new lines TODO.}
+Referee 3 made this point as well, and this sentence has been removed.
 
@@
 Reviewer 2 also raised the idea of looking at C/N, and we have responded to that suggestion in detail (see comment 2.9).
-{\bf See new lines LINES.}
 
@@
 
-Agreed.
-This will be changed to "The interaction between water content and percent N was also highly significant (P < 0.001), although cores with N > X\% exhibited little relationship between water content and CO2 flux (data not shown)." {\bf See new lines TODO.}
+Agreed; this has been clarified. {\bf See new lines TODO.}
 
@@
 comment 2.2 above) and other publications (comment 2.4 above).
-{\bf See new lines TODO.}
+{\bf See new lines LINES.}
 
@@
 
-This is an interesting and useful suggestion.
 We agree that rebalancing the manuscript, focusing a bit more on controls on cumulative emissions and a bit less on the instantaneous fluxes, would strengthen it.
-In our revision, we will more fully explore controls on the cumulative emissions, and as a first step have added the analytical code to do so (see \url{https://github.com/bpbond/cpcrw_incubation/commit/2d62e73ab79f7f0c8d7289919cddedb0a22ec275}).
-At the moment, it looks like only \%N exerts a weak effect on cumulative CO2 fluxes, and none of the tested covariates (\%C, \%N, C/N, DOC) affects cumulative CH4.
-{\bf See new lines TODO.}
+Accordingly, we now more fully explore controls on the cumulative emissions.
+{\bf See new lines LINES.}
 
@@
 
-These points have been clarified, except for the last, as we feel it's already clear and unambiguous.
+These points have been clarified, except for the last, as we feel it's already clear and unambiguous. {\bf See new lines LINES.}
 
@@
 
-We will better integrate this point, mentioning it in the introduction and clarifying its relationship to the study goals. {\bf See new lines TODO.}
+We now better integrate this point, mentioning it in the introduction and clarifying its relationship to the study goals. {\bf See new lines LINES.}
 
@@
 
-Agreed. Tantalizing.
+Agreed!
 
@@
 
-Thanks for the Lee et al. reference, which we had not considered (see our response to 3.7 above). We'll take a hard look at lines 362-378 and tighten. {\bf See new lines TODO.}
+Thanks for the Lee et al. reference, which we had not considered (see our response to 3.7 above) but is now cited. We have also reworked and tightened section 4.2. {\bf See new lines LINES.}
 

### 2016-11-18 manuscript/cpcrw_manuscript.Rmd
Addressed a variety of minor comments from Reviewers 2 and 3

@@
 The cumulative production of C from CO~2~ was over `r minmag` orders of magnitudes higher than that from CH~4~.
-These results suggest that deep active-layer soils may be quite sensitive to changes in soil moisture, a critical factor as discontinuous permafrost thaws [COM2.5] in interior Alaska.
+These results suggest that deep active-layer soils may be quite sensitive to changes in soil moisture, a critical factor as discontinuous permafrost thaws in interior Alaska.
 Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results provide insight into their future dynamics and feedback potential with future climate change.
@@
 
-These feedbacks between vegetation type, ground cover, permafrost, and SOC may be disrupted by increased fire disturbance, particularly after more-intense fire events {Johnstone, 2010 #2465}{Genet, 2013 #4412}.
+A number of factors may however disrupt these these feedbacks between vegetation type, ground cover, permafrost, and SOC stability.
+Climate changes, in particular regional warming and drying, cause vegetation stress {Ju, 2016 #4320}{Barber, 2000 #1320} and increased mortality.
+Climate also drives fire regime changes, and ecosystem disruption is particularly likely after more-intense fire events {Johnstone, 2010 #2465}{Genet, 2013 #4412}.
 Even absent disturbance, the stability of SOC is highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}{Karhu, 2014 #4124}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
@@
 Deep soil carbon is likely to be a highly important contributor to future climate feedbacks, with modeling studies suggesting that one-third of 21^st^ century climate-induced carbon loss may originate from seasonally frozen soil carbon {Koven, 2015 #4473}.
-[COM1.8,2.6]
 
@@
 Colman and Schimel {, 2014 #4172}.
-The interaction between water content and percent N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), with high-N cores having little relationship between water content and CO~2~ flux (data not shown).
+The interaction between water content and percent N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), although cores with N > 0.2\% exhibited little relationship between water content and CO~2~ flux (data not shown).
 
@@
 
-Temperature and moisture senstivity was generally invariant over time: 
 The cumulative production of C from CO~2~ (**Figure 4**) was over `r minmag` order of magnitudes higher than that from CH~4~, with CO~2~:CH~4~ C ratios ranging from `r minratio` million in the `r mintemp` °C "`r maxtrt`" treatment, to `r maxratio` million in the `r maxtemp` °C "`r maxtrt`" treatment.
@@
 Computing Q~10~ values based on fluxes normalized by water filled pore space changed these values only slightly: to `r pn(wfpsQ10_cumulative["Field moisture", "CO2"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CO2"], 1)` for CO~2~, respectively, and `r pn(wfpsQ10_cumulative["Field moisture", "CH4"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CH4"], 1)` for CH~4~.
+Temperature and moisture senstivities were generally invariant over time: TODO.
 
@@
 
-Rises in boreal air temperatures, and unpredictable precipitation changes, will warm and dry many soils, increase vegetation stress {Ju, 2016 #4320}{Barber, 2000 #1320}, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}, with uncertain consequences for soil dynamics and GHG fluxes.
+Rises in boreal air temperatures, and unpredictable precipitation changes, are changing fire disturbance regimes, warming and dry many soils, increasing vegetation stress, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}, all with uncertain consequences for soil dynamics and GHG fluxes.
 In this laboratory experiment we found that CO~2~, but not CH~4~, fluxes from these oxic active-layer mineral soils were sensitive to temperature and, in particular, moisture.
@@
 Morishita et al.
-{, 2014 #4316} quantified gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture in upland Cryosols [COM3.18], consistent with our results.[COM1.17,3.17] They found however that CH~4~ uptake (no emissions were observed) was driven by temperature only.
+{, 2014 #4316} quantified gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture in upland Cryosols, consistent with our results.
+They found however that CH~4~ uptake (no emissions were observed) was driven by temperature only.
 Waldrop et al.
@@
 discuss - upland mixed-canopy boreal forest here; careful about extrapolation
+discuss - controls on cumulative emissions
 
@@
 This is naturally far lower than our observed aerobic (and thus high-CO~2~) ratios, but nonetheless consistent with them.
-Thus we see little opportunity for CH~4~ to be a significant contributor to these upland soils' C fluxes and climate feedback risk, even accounting for the stronger radiative forcing of this gas.
+Thus we see little opportunity for CH~4~ to be a significant contributor to these upland soils' C fluxes and climate feedback risk, even accounting for the 25x stronger radiative forcing of this gas over a 100-year time horizon {Lee, 2012 #4503}.
 

### 2016-11-18 manuscript/response_to_reviewer_comments.tex
First pass at addressing reviewers concerns about discussion

@@
 (2006, 2010).
-{\bf See new lines TODO.}
+{\bf See new lines LINES.}
 
@@
 
-Agreed; this has been clarified. {\bf See new lines TODO.}
+This has been clarified. {\bf See new lines LINES.}
 
@@
 We agree that the comparison to other boreal respiration studies needs improvement, and this echoes Referee \#2's comment 2.4.
-We will restructure and improve this summary and comparison section.
-The Lavoie (2011) paper is very useful with respect to N and microbial respiration, while Dutta et al.
-(2006), although it concerns Siberian soils, may also prove a useful comparison.
-We had not included Schädel et al.
-(2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but it would also be a good addition.
-{\bf See new lines TODO.}
+First, although we think the paragraph about nearby studies provides useful context, we have tightened it considerably.
+Second, we have restructured and improved the subsequent comparison section, discussing a variety of studies suggested by all the reviewers to better put our results in the context of previous work.
+{\bf See new lines LINES.}
 
@@
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
-{\bf See new lines TODO.}
+{\bf See new lines LINES.}
 
@@
 (2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it would also be a reasonable addition.
-{\bf See new lines TODO.}
+{\bf See new lines LINES.}
 
@@
 We agree that rebalancing the manuscript, focusing a bit more on controls on cumulative emissions and a bit less on the instantaneous fluxes, would strengthen it.
-Accordingly, we now more fully explore controls on the cumulative emissions.
-{\bf See new lines LINES.}
+Accordingly, we now more fully explore controls on the cumulative emissions, and have moved ...
+{\bf See new lines TODO.}
 
@@
 (2008).
-{\bf See new lines TODO.}
+{\bf See new lines LINES.}
 

### 2016-11-18 manuscript/cpcrw_manuscript.Rmd
First pass at addressing reviewers concerns about discussion

@@
 Climate changes, in particular regional warming and drying, cause vegetation stress {Ju, 2016 #4320}{Barber, 2000 #1320} and increased mortality.
+Conversely, increasing plant productivity in some regions can stimulate the decomposition of older SOC {Hartley, 2012 #3481}.
 Climate also drives fire regime changes, and ecosystem disruption is particularly likely after more-intense fire events {Johnstone, 2010 #2465}{Genet, 2013 #4412}.
@@
 {Kim, 2012 #2863}.
-The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, and such interactions are critical to examine in the course of long-term incubation experiments {Elberling, 2013 #3720}.
-Decadal warming and drying trends in Alaska {Bieniek, 2014 #4389} may counteract these effects, however.
+The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, but emissions from aerobic soils will likely dominate the permafrost C feedback {Schädel, 2016 #4444}.
+Decadal warming and drying trends in Alaska {Bieniek, 2014 #4389} thus seem likely to increase GHG emissions from soils, and incubation experiments are critical to understand these dynamics {Elberling, 2013 #3720}.
 
@@
 
-Several studies have measured microbial respiration and GHG fluxes from soils very close to our study site.
+A number of studies have measured microbial respiration and GHG fluxes very close to our study site.
 Morishita et al.
-{, 2014 #4316} quantified gas fluxes in the field at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture in upland Cryosols, consistent with our results.
-They found however that CH~4~ uptake (no emissions were observed) was driven by temperature only.
+{, 2014 #4316} quantified GHG fluxes at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture in upland Cryosols, consistent with our results.
+They found however that CH~4~ uptake was driven by temperature only.
 Waldrop et al.
-{, 2010 #4380} incubated active-layer and permafrost soils from *Picea mariana* sites near Fairbanks, AK, under varied temperature and aerobic conditions, observing Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C; these values are higher than we observed, consistent with the lower temperature range {Hamdi, 2013 #3664} and fundamental biokinetics {Davidson, 2006 #1669}.
-Waldrop et al.
-{, 2010 #4380} also observed flux rates of 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), differing by orders of magnitude between sites (but roughly similar to our observed CH~4~ emissions), and ~1-5 µg C-CO~2~ hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed from our cores.
+{, 2010 #4380} incubated active-layer and permafrost soils from *Picea mariana* sites near Fairbanks, AK, observing aerobic Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C, and flux rates of 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), and ~1-5 µg C-CO~2~ hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed here.
+During the first 100 days of an incubation of Fairbanks-area 0-10 cm mineral soils, Neff and Hooper {, 2002 #4390} observed fluxes of ~`r pn(0.2 * 1000 * 12/44, 0)`-`r pn(1.5 * 1000 * 12/44, 0)` µg C-CO~2~ g C^-1^ day^-1^, in line with the results here, while Wickland and Neff {, 2008 #1820} reported that temperature and moisture exhibited interactive and similar-magnitude effects on decomposition in *P.
+mariana* soils.
+Finally, Sierra et al.
+{, 2016 #4504} incubated homogenized organic soils from CPCRW and observed interactions between high temperatures, moisture, and oxygen concentration.
+
+A number of synthesis studies have produced robust results documenting the dynamics and C feedback potential of Arctic and boreal soils; comparing to these results is useful because although the response of soil biota to stresses such as drought tends to differ between soil types, organisms, and vegetation, it is often broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}.
+Using two meta-analyses of aerobic and anaerobic permafrost soil incubations, Schädel et al.
+{, 2016 #4444} showed that C release was highly sensitive to temperature, and that soils released far more (220-520%) C under aerobic conditions.
+Our incubation was fully aerobic, but our results are consistent with the conclusion that respiration in the form of CO~2~ is likely to dominate the high latitude C feedback, and that aerobic soils, and the conditions under which currently waterlogged soils may drain, deserve particular attention.
+In a pan-Arctic synthesis of anaerobic soil incubations, Treat et al.
+{, 2015 #4217} reported mean CO~2~ rates of 47 (all mineral soils) and 101 (for 20-100 cm soils) µg C-CO~2~ g C^-1^ day^-1^, somewhat lower than our aerobic incubation results.
 In an incubation of active-layer Alaskan permafrost peats, Treat et al.
-{, 2014 #4213} found CO~2~ and CH~4~ emissions to be strongly correlated with temperature and moisture.
-Finally, during the first 100 days of a year-long incubation of Fairbanks-area 0-10 cm mineral soils, Neff and Hooper {, 2002 #4390} observed fluxes of ~`r pn(0.2 * 1000 * 12/44, 0)`-`r pn(1.5 * 1000 * 12/44, 0)` µg C-CO~2~ g C^-1^ day^-1^, in line with the results here.
-
-More generally, in a pan-Arctic synthesis of anaerobic soil incubations, Treat et al.
-{, 2015 #4217} reported mean CO~2~ rates of 47 (all mineral soils) and 101 (for 20-100 cm soils) µg C-CO~2~ g C^-1^ day^-1^, somewhat lower than our aerobic incubation results.
-The response of soil biota to stresses such as drought tends to differ between soil types and organisms, but be broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}, making such comparisons between useful, in spite of significant ecological and climatic differences of studies collected by meta-analyses such as Treat et al.
-{, 2015 #4217}.
+{, 2014 #4213} also found CO~2~ and CH~4~ emissions to be strongly correlated with temperature and moisture.
+Whether climate change makes northern regions wetter or driers is thus a critical factor affecting the quantity and form of C release {Schädel, 2016 #4444}.
 
@@
 In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture), but not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
-Observed surface CO~2~ fluxes at this range exhibited a Q~10~ of 5.1 ± 1.4 over a temperature range of 3.5-15 °C (personal communication, C.
+It is more typical to observe Q~10~ values of around 2, however {Dutta, 2006 #1427}{Schädel, 2016 #4444}.The temperature sensitivity of C release can change over time of incubation {Dutta, 2006 #1427} and vary between soil fractions cycling over different time horizons {Karhu, 2010 #2503}{Schädel, 2014 #4472}.
+We did not observe any time dependence of Q~10~ or moisture sensitivity in this 100-day incubation, however.
+
+Observed surface CO~2~ fluxes at this CPCRW site exhibited a Q~10~ of 5.1 ± 1.4 over a temperature range of 3.5-15 °C (personal communication, C.
 Anderson); these surface fluxes were measured over multiple months and include root respiration, however, confounding any direct comparison.
@@
 
-discuss - changes over time (see new reuslts)
 discuss - upland mixed-canopy boreal forest here; careful about extrapolation
@@
 Nonetheless, the controlled environments of incubations provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
+It should also be noted that our 100-day incubation was not able to observe slowly-cycling soil fractions, which may vary in their response to experimental manipulation {Karhu, 2010 #2503}.
 

### 2016-11-18 manuscript/response_to_reviewer_comments.tex
Big discussion revisions and Word doc to APS & VLB

@@
 This point was raised by Referee \#2 as well (comment 2.4).
-We will do so, comparing for example with studies such as the Schädel et al.
-(2016) meta-analysis (and references therein), Dutta et al.
+We now provide a more complete comparison, citing for example studies such as the Schädel et al.
+(2016) meta-analysis, Dutta et al.
 (2006), Lavoie et al.
 (2011), Karhu et al.
-(2010), and Czimczik et al.
-(2006, 2010).
+(2010), and Wickland and Neff (2008).
 {\bf See new lines LINES.}
@@
 This is an interesting suggestion, and an analysis we are happy to include.
-We added code (see \url{https://github.com/bpbond/cpcrw_incubation/commit/18884c836d855b94417f7ae2cef910c577bc40b}) that examines if and how temperature and moisture sensitivities change over incubation time in this experiment.
-We do not observe any evidence for changes in CO2 moisture sensitivity with time, and only weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time.
+We do not observe any evidence for changes in CO2 moisture sensitivity with time, and weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time.
+This is now discussed in the text.
 {\bf See new lines LINES.}
@@
 The first question is similar to a point raised by Referee \#2 (comment 2.3).
-We will discuss this in more depth, noting in particular, for example, the fact that there are observable anthropogenic influences on high-latitude precipitation (Wan et al.
+We now discuss this in more depth, noting in particular, for example, the fact that there are observable anthropogenic influences on high-latitude precipitation (Wan et al.
 2015); drier and warmer conditions in boreal Eurasia (Buermann et al.
@@
 Referee \#2 raised this point as well (comment 2.6), and it's a good one.
-We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
+We now better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
 {\bf See new lines LINES.}
@@
 (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
-In our revision, we will significantly expand this paragraph, discussing and comparing to Schädel et al.
+In our revision, we have significantly expanded this paragraph, discussing and comparing to Schädel et al.
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
+We also cite and discuss a variety of other studies.
 {\bf See new lines TODO.}
@@
 
-Thanks for the useful suggestion; we will do so. {\bf See new lines TODO.}
+Thanks for the useful suggestion; this point is now brought out more clearly. {\bf See new lines LINES.}
 
@@
 I.e., incubation soils are isolated from their natural environment, and as a result we need to be cautious about extrapolating incubation results to {\it in situ} responses.
-We will clarify this in the text.
+We have clarified this in the text.
 {\bf See new lines LINES.}
@@
 (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
-In our revision, we will significantly expand the discussion on this point, comparing our results to Schädel et al.
+We have significantly expanded the discussion on this point, comparing our results to Schädel et al.
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
@@
 We had not included Schädel et al.
-(2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it would also be a reasonable addition.
+(2014) simply because of its focus on permafrost, versus the active-layer soils studied here, but we agree it is also be a reasonable addition.
+All these references are now cited throughout the manuscript.
 {\bf See new lines LINES.}
@@
 
-This wording will be fixed. {\bf See new lines LINES.}
+Fixed. {\bf See new lines LINES.}
 
@@
 This is a good point.
-We will better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
+We now better describe why deep active-layer soils, such as those studied here, are important and distinctive relative to permafrost or shallow active layer soils.
 {\bf See new lines LINES.}
@@
 
-Thank you. We appreciate the useful suggestions, and will provide these additional details in our revised manuscript. {\bf See new lines LINES.}
+Thank you. We appreciate the useful suggestions, and now provide these additional details in our revised manuscript. {\bf See new lines LINES.}
 
@@
 
-That's a very good point-thank you-and will integrate well with an expanded comparison to the Schädel et al.
+That's a very good point-thank you-and integrates well with an expanded comparison to the Schädel et al.
 (2016) paper (cf.
@@
 
-We will add a sentence of caveats, noting in particular the useful but incremental nature of this study. {\bf See new lines LINES.}
+We have added caveats, noting in particular the useful but incremental nature of this study. {\bf See new lines LINES.}
 
@@
 (2016) meta-analysis was a quirk of timing, as it appeared after our manuscript was submitted.
-In our revision, we will significantly expand this, discussing and comparing to Schädel et al.
+We have significantly expanded this, discussing and comparing to Schädel et al.
 (2014, 2016) and Wickland et al.
@@
 
-We will consider these suggestions with respect to length and supplemental materials; results dealing with cumulative emissions will certainly be included in the abstract. {\bf See new lines TODO.}
+We will consider these suggestions with respect to length and supplemental materials; results dealing with cumulative emissions are now included in the abstract. {\bf See new lines TODO.}
 
@@
 Sampling date is reported in line 110.
-We will clarify 80 cm at the time of sampling.
+We have clarified 80 cm at the time of sampling.
 Moisture adjustment was done after every mass measurement, i.e.
-every timepoint shown in Figure 1.
-This will be clarified.
+every timepoint shown in Figure 1; this has been clarified.
 {\bf See new lines LINES.}

### 2016-11-18 manuscript/cpcrw_manuscript.Rmd
Big discussion revisions and Word doc to APS & VLB

@@
 ---
-output:
-html_document: default
-word_document: default
+output: word_document
 ---
@@
 Daily CH~4~ emissions were most strongly correlated with percent nitrogen, but neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
-The cumulative production of C from CO~2~ was over `r minmag` orders of magnitudes higher than that from CH~4~.
-These results suggest that deep active-layer soils may be quite sensitive to changes in soil moisture, a critical factor as discontinuous permafrost thaws in interior Alaska.
+The cumulative production of C from CO~2~ was over `r minmag` orders of magnitudes higher than that from CH~4~; cumulative CO~2~ was correlated with incubation temperature and moisture treatment, with drought cores producing `r pn(drought_effect["20", "Reduction"], 0)`% - `r pn(drought_effect["4", "Reduction"], 0)`% lower C.
+Cumulative CH~4~ production was unaffected by any treatment.
+These results suggest that deep active-layer soils may be quite sensitive to changes in soil moisture under aerobic conditions, a critical factor as discontinuous permafrost thaws in interior Alaska.
 Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results provide insight into their future dynamics and feedback potential with future climate change.
@@
 High latitude ecosystems are being subjected to rapid changes in climate {IPCC, 2013 #3916} and increases in fire frequency and intensity {Kasischke, 2010 #2631}, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}{Ju, 2016 #4320}.
-This will have a wide variety of ecosystem effects {Alexander, 2016 #4414}: in particular, rising temperatures and increasing fire will likely result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372}{Genet, 2013 #4412}{Helbig, 2016 #4413} and changes in soil temperature, with subsequent hydrology changes that will influence soil greenhouse gas (GHG) fluxes to the atmosphere.
+This will have a wide variety of ecosystem effects {Alexander, 2016 #4414}: in particular, rising temperatures and increasing fire will likely result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372}{Genet, 2013 #4412}{Helbig, 2016 #4413} and changes in soil temperature, with subsequent hydrology changes that will influence soil greenhouse gas (GHG) fluxes to the atmosphere {Schädel, 2014 #4472}.
 Such fluxes are a large component of the global C cycle and, because of the high C stocks of northern soils {Tarnocai, 2009 #2231}, could result in a significant and positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}{Schaefer, 2014 #4117}.
@@
 Most previous studies have focused on surface soils or permafrost soils, neglecting deep active-layer soils that were identified as subject to strong effects from a two-decade warming experiment in the Alaskan Arctic {Sistla, 2013 #3666}.
-Such deeper soils have particular characteristics distinguishing them from both shallow active layer soils and underlying permafrost: they are most affected by interannual variability in thaw depth, potentially flipping the C source/sink status of entire ecosystems {Goulden, 1998 #50}{Harden, 2012 #4476}, and are subject to distinctive freeze-thaw, cryoturbation, and microbial dynamics, which are likely to change their sensitivity to climate change and feedback potential {Schuur, 2008 #4491}; and deep seasonally frozen soils pose particular problems for accurate modeling of high-latitude carbon dynamics {Nicolsky, 2007 #4474}.
-Deep soil carbon is likely to be a highly important contributor to future climate feedbacks, with modeling studies suggesting that one-third of 21^st^ century climate-induced carbon loss may originate from seasonally frozen soil carbon {Koven, 2015 #4473}.
+Such deeper soils have particular characteristics distinguishing them from both shallow active layer soils and underlying permafrost: they are most affected by interannual variability in thaw depth, potentially flipping the C source/sink status of entire ecosystems {Goulden, 1998 #50}{Harden, 2012 #4476}; they are subject to distinctive freeze-thaw, cryoturbation, and microbial dynamics, which are likely to change their sensitivity to climate change and feedback potential {Schuur, 2008 #4491}; and they are known to pose particular problems for accurate modeling of high-latitude carbon dynamics {Nicolsky, 2007 #4474}.
+These soils are likely to be a highly important contributor to future climate feedbacks, with modeling studies suggesting that one-third of 21^st^ century climate-induced carbon loss may originate from seasonally frozen soil carbon {Koven, 2015 #4473}.
 
@@
 
-Carbon dioxide fluxes during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 1)` µg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"], 1)` µg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"], 2)`) µg C g C^-1^ day^-1^, with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 1)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"], 1)`) µg C g C^-1^ day^-1^ over the 100 days.
-CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 2)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`) ng C g C^-1^ day^-1^, with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 2)`) ng C g C^-1^ day^-1^.
+Carbon dioxide fluxes during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"], 1)` µg C g C^-1^ day^-1^ to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"], 1)` µg C g C^-1^ day^-1^, with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"], 1)` µg C g C^-1^ day^-1^ over the 100 days.
+CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^, with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^.
 
 These means conceal considerable variability over the course of the incubation (**Figures 2 and 3**).
-In the linear mixed-effects model (AIC = `r pn(AIC(step_co2_lme), 1)`), CO~2~ was positively correlated with incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 2**).
+In the linear mixed-effects model (AIC = `r pn(AIC(step_co2_lme), 1)`), instantaneous CO~2~ flux was positively correlated with incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 2**).
 Temperature sensitivity decreased significantly (`r pclean(co2_temp_time)`) over the course of the incubation, while moisture sensitivity was unaffected by time.
@@
 The interaction between water content and percent N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), although cores with N > 0.2\% exhibited little relationship between water content and CO~2~ flux (data not shown).
-
-Methane fluxes were positively correlated with percent N, while water content exhibited significant interactions with percent N and DOC (**Table 3**).
-The model had little predictive power (AIC = `r pn(AIC(step_ch4_lme), 1)`), and neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
+Instantaneous CH~4~ fluxes were positively correlated with percent N, while water content exhibited significant interactions with percent N and DOC as a predictor (**Table 3**).
+This model had little predictive power (AIC = `r pn(AIC(step_ch4_lme), 1)`), however, and neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
 
@@
 Computing Q~10~ values based on fluxes normalized by water filled pore space changed these values only slightly: to `r pn(wfpsQ10_cumulative["Field moisture", "CO2"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CO2"], 1)` for CO~2~, respectively, and `r pn(wfpsQ10_cumulative["Field moisture", "CH4"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CH4"], 1)` for CH~4~.
-Temperature and moisture senstivities were generally invariant over time: TODO.
 
@@
 {, 2016 #4444} showed that C release was highly sensitive to temperature, and that soils released far more (220-520%) C under aerobic conditions.
-Our incubation was fully aerobic, but our results are consistent with the conclusion that respiration in the form of CO~2~ is likely to dominate the high latitude C feedback, and that aerobic soils, and the conditions under which currently waterlogged soils may drain, deserve particular attention.
-In a pan-Arctic synthesis of anaerobic soil incubations, Treat et al.
+Our incubation was fully aerobic, but its results are consistent with the conclusion that respiration in the form of CO~2~ is likely to dominate the high latitude C feedback, and that aerobic soils, and the conditions under which currently waterlogged soils may drain, deserve particular attention.
+In terms of absolute flux rates, in a pan-Arctic synthesis of anaerobic soil incubations, Treat et al.
 {, 2015 #4217} reported mean CO~2~ rates of 47 (all mineral soils) and 101 (for 20-100 cm soils) µg C-CO~2~ g C^-1^ day^-1^, somewhat lower than our aerobic incubation results.
-In an incubation of active-layer Alaskan permafrost peats, Treat et al.
-{, 2014 #4213} also found CO~2~ and CH~4~ emissions to be strongly correlated with temperature and moisture.
-Whether climate change makes northern regions wetter or driers is thus a critical factor affecting the quantity and form of C release {Schädel, 2016 #4444}.
+Treat et al.
+{, 2014 #4213} also found CO~2~ and CH~4~ emissions to be strongly correlated with temperature and moisture based on an incubation of Alaskan peats.
+Whether climate change makes northern regions wetter or driers is thus a critical factor affecting the quantity and form of C release.
 
-**4.1 Temperature versus moisture sensitivity**
+**4.1 Temperature versus moisture sensitivity for cumulative emissions**
 
 Warming usually increases soil GHG fluxes, for example at depth in a long-term Arctic tundra experiment {Sistla, 2013 #3666}, as increased temperatures enhance the production of extracellular enzymes, increase enzyme activities, and enhance desorption rates of organic matter from minerals.
-A key question, for both experimentalists and modelers {Falloon, 2011 #3045}, is to what degree such soils' emissions could by constrained by their moisture status, that is itself driven by increases in high-latitude temperatures, vapor pressure deficit, and potentially precipitation changes.
+A key question, for both experimentalists and modelers {Falloon, 2011 #3045}, is to what degree such soils' emissions could by constrained by their moisture status, that is itself driven by changes in factors such as high-latitude temperatures, vapor pressure deficit, and precipitation.
 
-Our results suggest that moisture limitation could exert a large effect on CO~2~ production for deep active-layer soils (**Figure 4**): "drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
+In contrast to the controls on instantaneous fluxes, the controls on the cumulative fluxes (**Figure 4**) integrate the entire incubation time and are important to understand.
+Our results suggest that moisture limitation could exert a large effect on CO~2~ production for deep active-layer soils: "drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
 This effect was highly significant, and suggests that moisture limitations could exert a significant constraint on deep active-layer soils as they slowly warm.
@@
 In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture), but not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
-It is more typical to observe Q~10~ values of around 2, however {Dutta, 2006 #1427}{Schädel, 2016 #4444}.The temperature sensitivity of C release can change over time of incubation {Dutta, 2006 #1427} and vary between soil fractions cycling over different time horizons {Karhu, 2010 #2503}{Schädel, 2014 #4472}.
-We did not observe any time dependence of Q~10~ or moisture sensitivity in this 100-day incubation, however.
+It is more typical to observe Q~10~ values of around 2 {Dutta, 2006 #1427}{Schädel, 2016 #4444}, although the temperature sensitivity of C release can change over time of incubation {Dutta, 2006 #1427} and vary between soil fractions cycling over different time horizons {Karhu, 2010 #2503}{Schädel, 2014 #4472}.
+We observed Q10 change over the course of this incubation, emphasizing the need for longer-term studies {Lee, 2012 #4503}{Elberling, 2013 #3720}.
 
@@
 
-discuss - upland mixed-canopy boreal forest here; careful about extrapolation
-discuss - controls on cumulative emissions
-
-More surprisingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 than at 20 °C (TODO-fix sentence).
+More surprisingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 °C than at 20 °C.
 There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}.
@@
 Nonetheless, the controlled environments of incubations provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
-It should also be noted that our 100-day incubation was not able to observe slowly-cycling soil fractions, which may vary in their response to experimental manipulation {Karhu, 2010 #2503}.
+It should also be noted that our 100-day incubation was not long enough to observe slowly-cycling soil fractions, which may vary in their response to experimental manipulation {Karhu, 2010 #2503}.
 
+The soils studied here were from an upland, mixed conifer-deciduous boreal forest, and care needs to be taken before drawing regional inferences, or about other ecosystem types.
 We focused on an experimental drought, rather than flooding, because of the well-drained nature of the field site: it is unlikely that the mid-slope forest we sampled in will ever suffer from thermokarst or excessive soil moisture, but too-dry conditions are a serious possibility in this low-precipitation ecosystem {Barber, 2000 #1320}.
-In addition, the soils here are not surface layer soils (where the majority of microbial activity and C mineralization of labile C takes place); removing them from *in situ* conditions (where they are less exposed to O~2~, for example) may significantly change the abiotic conditions to which the microbial community is adapted.
+
+Finally, the soils here are not surface layer soils (where the majority of microbial activity and C mineralization of labile C takes place); removing them from *in situ* conditions (where they are less exposed to O~2~, for example) may significantly change the abiotic conditions to which the microbial community is adapted.
 However, focusing on the active layer provides crucial information about the potential loss of C from these soils, a risk that needs to be well understood as permafrost degradation leads to expansions in the depth of the active layer across the Arctic.

### 2016-11-19 R/5-analyze.R
Various tweaks to comments, documentation, Table 1

@@ -1,21 +1,19 @@
 # Final analysis script: statistical tests, final plots, etc.
 # Called by the R Markdown document (the manuscript).
-# This has become a really big script...ugh
-# Ben Bond-Lamberty January 2016
+# Ben Bond-Lamberty November 2016
+
+library(MASS)       # 7.3-45
+library(nlme)       # 3.1-128
+library(tidyr)      # 0.6.0
+library(broom)      # 0.4.1
+library(reshape2)   # 1.4.1
+library(agricolae)  # 1.2-4
 
 source("R/0-functions.R")
 
 SCRIPTNAME  	<- "5-analyze.R"
 
-library(nlme)       # 3.1-128
-#library(MASS)       # 7.3-45
-library(tidyr)      # 0.6.0
-library(broom)      # 0.4.1
-library(reshape2)   # 1.4.1
-library(agricolae)  # v1.2.4
-
-# ==============================================================================
-# Main 
+# ---------- Main ---------- 
 
 openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # open log
 
@@ -29,9 +27,9 @@ fluxdata_orig <- read_csv(FLUXDATA_FILE, col_types = "icddddddciddddddlldl")
 fluxdata_orig$Treatment <- factor(fluxdata_orig$Treatment, 
                                   levels = c("Field moisture", "Controlled drought", "Drought"))
 
-# -----------------------------------------------------------------------------
-# Transform into long format and handle outliers
+# ---------- Transform into long format ---------- 
 
+printlog("Transform into long format and handle outliers...")
 co2_mad_outlier_count <- sum(fluxdata_orig$CO2_outlier, na.rm = TRUE)
 ch4_mad_outlier_count <- sum(fluxdata_orig$CH4_outlier, na.rm = TRUE)
 mad_outlier_count <- co2_mad_outlier_count + ch4_mad_outlier_count
@@ -50,6 +48,8 @@ fluxdata_orig %>%
   select(-variable) ->
   fluxdata
 
+# ---------- Outliers ---------- 
+
 pre_outlier_count <- nrow(fluxdata)
 
 # Combine the gas-specific outlier fields into a single outlier flag
@@ -62,8 +62,7 @@ fluxdata %>%
   filter(flux_mgC_hr > 0 & !outlier) ->
   fluxdata
 
-# -----------------------------------------------------------------------------
-# Read C, N, DOC data; summarise; test for treatment effects; merge with fluxdata
[…]

### 2016-11-19 R/0-functions.R
Tweaks for clarity and consistency

@@ -1,10 +1,10 @@
 # Common constant, file location, and function definitions
 # This support script is source'd by every other script before executing
-# Ben Bond-Lamberty March 2015
+# Ben Bond-Lamberty November 2016
 
 # -----------------------------------------------------------------------------
 # Run under:
-#   R version 3.2.3 (2015-12-10), x86_64-apple-darwin13.4.0 (64-bit)
+#   R version 3.3.1 (2016-06-21), x86_64-apple-darwin13.4.0 (64-bit)
 # Key packages used across scripts:
 library(ggplot2)       # 2.1.0
 theme_set(theme_bw())

### 2016-11-19 R/1-data.R

@@ -44,8 +44,9 @@ process_directory <- function(input_path, tempfile) {
   for(f in seq_along(filelist)) {
     d <- read_outputfile(file.path(input_path, filelist[f]))
 
-    if(f > 1 & ncol(d) != ncolumns)
+    if(f > 1 & ncol(d) != ncolumns) {
       stop("Columns differ between files!")
+    }
     ncolumns <- ncol(d)
     
     first <- !file.exists(tempfile)

### 2016-11-19 R/2-summarize.R

@@ -46,8 +46,7 @@ printlog("Welcome to", SCRIPTNAME)
 
 printlog("Reading in raw data...")
 rawdata <- read_csv(RAWDATA_FILE, col_types = "ccddddiiiddddddddc") %>%
-#readr::read_csv(file.path(INPUT_DIR, RAWDATA_FILE), col_types = "ccddddiiiddddddddc") %>%
-  # immediate discard columns we don't need
+  # immediately discard columns we don't need
   select(DATE, TIME, MPVPosition, CH4_dry, CO2_dry, h2o_reported)
 print_dims(rawdata)
 print(summary(rawdata))
@@ -97,7 +96,7 @@ qc_valvemap(valvemap)
 # Function to match up Picarro data with mapping file data
 # This is done by date and valve number (see plot saved above)
 matchfun <- function(DATETIME, MPVPosition) {
-  DATETIME <- as.POSIXct(DATETIME, origin = lubridate::origin, tz="UTC")
+  DATETIME <- as.POSIXct(DATETIME, origin = lubridate::origin, tz = "UTC")
   rowmatches <- which(DATETIME >= valvemap$StartDateTime & 
                         yday(DATETIME) == yday(valvemap$StartDateTime) &
                         MPVPosition == valvemap$MPVPosition)
@@ -118,50 +117,57 @@ summarydata_min <- rawdata_samples %>%
             min_CH4_time = nth(elapsed_seconds, which.min(CH4_dry)))
 
 # Now we want to look for the max concentration AFTER the minimum
-rawdata_temp <- rawdata_samples %>%
-  left_join(summarydata_min, by = "samplenum") 
+rawdata_samples %>%
+  left_join(summarydata_min, by = "samplenum") ->
+  rawdata_temp
 
-summarydata_maxCO2 <- rawdata_temp %>%
+rawdata_temp %>%
   filter(elapsed_seconds > min_CO2_time & elapsed_seconds < MAX_MAXCONC_TIME) %>%
   summarise(max_CO2 = max(CO2_dry),
-            max_CO2_time = nth(elapsed_seconds, which.max(CO2_dry))
-            )
-summarydata_maxCH4 <- rawdata_temp %>%
+            max_CO2_time = nth(elapsed_seconds, which.max(CO2_dry))) ->
+  summarydata_maxCO2
+rawdata_temp %>%
   filter(elapsed_seconds > min_CH4_time & elapsed_seconds < MAX_MAXCONC_TIME) %>%
[…]

### 2016-11-19 R/3-remove.R

@@ -20,8 +20,9 @@ openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # o
 printlog("Welcome to", SCRIPTNAME)
 
 printlog("Reading in summary data...")
-summarydata <- read_csv(SUMMARYDATA_FILE) %>%
-  mutate(DATETIME = ymd_hms(DATETIME))
+read_csv(SUMMARYDATA_FILE) %>%
+  mutate(DATETIME = ymd_hms(DATETIME)) ->
+  summarydata
 print_dims(summarydata)
 
 # -----------------------------------------------------------------------------

### 2016-11-19 R/3.3-qc.R

@@ -74,7 +74,7 @@ save_diagnostic(p, "missing_mass")
 printlog(SEPARATOR)
 printlog("Computing per-date summaries...")
 summarydata <- summarydata %>%
-  mutate(Date = strftime(DATETIME, format="%Y-%m-%d"))
+  mutate(Date = strftime(DATETIME, format = "%Y-%m-%d"))
 
 summarydata %>%
   group_by(Date, Treatment, Temperature) %>%

### 2016-11-19 R/3.5-investigate.R
Various changes
Not sure if this script still works

@@ -49,9 +49,9 @@ investigate <- function(smry, summarydata, rawdata,
       sd_gas <- subset(summarydata, samplenum %in% samplenums)
       print(sd_gas)
       
-      p_closeup <- ggplot(sd_gas, aes_string("DATETIME", gasvar, fill="Core")) + 
+      p_closeup <- ggplot(sd_gas, aes_string("DATETIME", gasvar, fill = "Core")) + 
         geom_bar(stat='identity') + 
-        ggtitle(paste("Closeup look at", group_hash, "on incday", unique(sd_gas$incday))) +
+        ggtitle(paste("Closeup look at", group_hash, "on inctime_days", unique(sd_gas$inctime_days))) +
         geom_text(aes_string(label = "samplenum", color = gas_outlier), vjust = -0.5)
       print(p_closeup)
       save_plot(paste0(gas, "_closeup_", group_hash), ptype = ".png")
@@ -86,13 +86,12 @@ openlog(file.path(outputdir(), paste0(SCRIPTNAME, ".log.txt")), sink = TRUE) # o
 printlog("Welcome to", SCRIPTNAME)
 
 printlog("Reading in raw data...")
-rawdata <- readr::read_csv(RAWDATA_FILE)
+rawdata <- read_csv(RAWDATA_FILE, col_types = "ccddddiiiddddddddc")
 printlog("Reading in summary data...")
-summarydata <- read_csv(SUMMARYDATA_FILE)
+summarydata <- read_csv(SUMMARYDATA_CLEAN_FILE)
 summarydata$DATETIME <- ymd_hms(summarydata$DATETIME)
 printlog("Reading in treatment summaries...")
-tsummary <- read_csv(TREATMENT_SUMMARYDATA)
-tsummary$Date <- ymd(tsummary$Date)
+tsummary <- read_csv(TREATMENTS, skip = 1)
 
 dev <- 3.2
 printlog("Computing outlier flags using deviation of", dev, "...")
@@ -109,7 +108,7 @@ printlog("Looking at high-variability CO2...")
 tsummary %>%
   filter(!is.na(CO2_ppm_s_sd), Treatment != "Ambient") %>% 
   arrange(desc(CO2_ppm_s_sd)) %>%
-  select(Date, Treatment, Temperature, CO2_ppm_s_sd, CO2_ppm_s, incday, samplenums) ->
[…]

### 2016-11-19 manuscript/response_to_reviewer_comments.tex
Upscaling paragraph (reviewer comments 1.7 and 2.3)

@@
 The first question is similar to a point raised by Referee \#2 (comment 2.3).
-We now discuss this in more depth, noting in particular, for example, the fact that there are observable anthropogenic influences on high-latitude precipitation (Wan et al.
-2015); drier and warmer conditions in boreal Eurasia (Buermann et al.
-2015), and growing season length increases in interior Alaska with no increase in precipitation (Wendler and Shulski 2009); and the complex and still-unresolved dynamics at play, meaning that, as the referee notes, drying and drought will probably be highly local in effect (Zhang et al.
-2012).
-{\bf See new lines TODO.}
+We have added this as a paragraph in the discussion, noting e.g.
+observable anthropogenic influences on high-latitude precipitation, drier and warmer conditions in boreal Eurasia, and growing season length increases in interior Alaska with no increase in precipitation.
+{\bf See new lines LINES.}
 
@@
 
-We will add this to the discussion, noting in particular the fact that there are observable anthropogenic influences on high-latitude precipitation (Wan et al.
-2015); drier and warmer conditions in boreal Eurasia (Buermann et al.
-2015), and growing season length increases in interior Alaska with no increase in precipitation (Wendler and Shulski 2009); and the complex and still-unresolved dynamics at play, meaning that, as the referee notes, drying and drought will probably be highly local in effect (Zhang et al.
-2012).
-{\bf See new lines TODO.}
+We have added this as a paragraph in the discussion, noting e.g.
+observable anthropogenic influences on high-latitude precipitation, drier and warmer conditions in boreal Eurasia, and growing season length increases in interior Alaska with no increase in precipitation.
+{\bf See new lines LINES.}
 

### 2016-11-19 manuscript/cpcrw_manuscript.Rmd
Upscaling paragraph (reviewer comments 1.7 and 2.3)

@@
 ---
-output: word_document
+output:
+  html_document: default
+  word_document: default
 ---
 [//]: This Rmarkdown file generates the Biogeosciences manuscript
+[//]: Built using rmarkdown 1.1, RStudio 1.0.44, and R 3.3.1
+[//]: November 23, 2016
 
@@
 ```{r setup, echo=FALSE, fig.keep='none', message=FALSE, results='hide'}
-library(dplyr, warn.conflicts = FALSE)
-
-# Run the main analysis script, which will leave figures, tables, and data 
-# in memory for us. Need to temporarily change the working directory to do so.
+# Run the main analysis script, which will leave figures, tables, 
+# and data in memory for us to access. 
+# Need to temporarily change the working directory to do so.
 olddir <- getwd()
@@
 Our incubation was fully aerobic, but its results are consistent with the conclusion that respiration in the form of CO~2~ is likely to dominate the high latitude C feedback, and that aerobic soils, and the conditions under which currently waterlogged soils may drain, deserve particular attention.
-In terms of absolute flux rates, in a pan-Arctic synthesis of anaerobic soil incubations, Treat et al.
-{, 2015 #4217} reported mean CO~2~ rates of 47 (all mineral soils) and 101 (for 20-100 cm soils) µg C-CO~2~ g C^-1^ day^-1^, somewhat lower than our aerobic incubation results.
+In terms of absolute flux rates, Treat et al.
+{, 2015 #4217} reported mean CO~2~ rates of 47 (all mineral soils) and 101 (for 20-100 cm soils) µg C-CO~2~ g C^-1^ day^-1^ from a pan-Arctic synthesis of anaerobic soil incubations, somewhat lower than our aerobic incubation results.
 Treat et al.
@@
 Whether climate change makes northern regions wetter or driers is thus a critical factor affecting the quantity and form of C release.
+
+The drought treatment imposed in this experiment reduced soil C fluxes by `r pn(drought_effect["20", "Reduction"], 0)`% - `r pn(drought_effect["4", "Reduction"], 0)`%.
+The importance of this result depends, in part, on the spatial extent and intensity of precipitation changes across the boreal and Arctic this century.
+There is a detectable anthropogenic influence in high latitude precipitation {Wan, 2015 #4478}, and its changes in which exhibit high spatial variability: drier and warmer conditions in boreal Eurasia {Buermann, 2014 #3992}, for example, but growing season length increases in interior Alaska with no increase in precipitation {Wendler, 2009 #4477}.
+This spatial variability will interact with permafrost thaw dynamics to produce a complex patchwork of soil moisture changes {Zhang, 2012 #4479}{Watts, 2012 #4505}.
+This high uncertainty in this area makes it all the more important to understand the interactive effects of soil moisture and temperature on decomposition and GHG emissions {Sierra, 2015 #4203}.
 

### 2016-11-19 manuscript/cpcrw_manuscript.Rmd
Restructuring section 4.1

@@
 
-Warming usually increases soil GHG fluxes, for example at depth in a long-term Arctic tundra experiment {Sistla, 2013 #3666}, as increased temperatures enhance the production of extracellular enzymes, increase enzyme activities, and enhance desorption rates of organic matter from minerals.
-A key question, for both experimentalists and modelers {Falloon, 2011 #3045}, is to what degree such soils' emissions could by constrained by their moisture status, that is itself driven by changes in factors such as high-latitude temperatures, vapor pressure deficit, and precipitation.
-
 In contrast to the controls on instantaneous fluxes, the controls on the cumulative fluxes (**Figure 4**) integrate the entire incubation time and are important to understand.
@@
 
-In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture), but not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
+In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture), although not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
 It is more typical to observe Q~10~ values of around 2 {Dutta, 2006 #1427}{Schädel, 2016 #4444}, although the temperature sensitivity of C release can change over time of incubation {Dutta, 2006 #1427} and vary between soil fractions cycling over different time horizons {Karhu, 2010 #2503}{Schädel, 2014 #4472}.
-We observed Q10 change over the course of this incubation, emphasizing the need for longer-term studies {Lee, 2012 #4503}{Elberling, 2013 #3720}.
-
 Observed surface CO~2~ fluxes at this CPCRW site exhibited a Q~10~ of 5.1 ± 1.4 over a temperature range of 3.5-15 °C (personal communication, C.

### 2016-11-19 manuscript/response_to_reviewer_comments.tex
Reworking section 4.2

@@
 
-We will supplement this section with a better comparison to, for example, Lavoie et al.
-(2011), who incubated upland Alaskan boreal soils at two temperatures and two N addition levels; Sistla et al.
-(2012); and the combination meta-analysis and modeling study of Bouskill et al.
-(2014) that at N addition at high latitudes.
-{\bf See new lines TODO.}
+We have supplemented this section with a better comparison to relevant literature, for example, Lavoie et al. (2011), Sistla et al. (2012), and Bouskill et al. (2014). {\bf See new lines LINES.}
+
+Known area of model weakness {Bouskill, 2014 #4480}.
 
@@
 This is similar to the referee's comment 1.9 (please see our response to that below), with the added factor of C quality.
-This idea, and studies focusing on it with respect to mineralization, has a long history (e.g.
-Nadelhoffer al.
-1991); as the referee notes, from fundamental biokinetics (Davidson and Janssens 2006) we expect that decomposition Q10 should be inversely related to litter carbon quality.
-We will address these questions more explicitly in our revision, referring for example to incubation studies on this question (Fierer et al.
+We have addressed this more explicitly in our revision, referring for example to incubation studies on this question (Fierer et al.
 2005).
-{\bf See new lines TODO.}
+{\bf See new lines LINES.}
 
@@
 We also cite and discuss a variety of other studies.
-{\bf See new lines TODO.}
+{\bf See new lines LINES.}
 
@@
 We appreciate this useful advice and question.
-Certainly temperature sensitivity is, at least in theory, related to C quality (see our answer to 1.5 above), although low-quality litter (i.e.
-in a boreal coniferous forest) might be expected to be associated with {/it higher} Q10?
-In our revision, we will explore these ideas, considering how how factors such as vegetation, soil characteristics, and climate might explain the low observed temperature sensitivity.
-{\bf See new lines TODO.}
+We have diminished the emphasis on defending this finding, and instead try to place it in the context of soil characteristics in this mixed-species boreal forest, SOC quality, etc.
+{\bf See new lines LINES.}
 
@@
 
-This is indeed interesting to consider: what mechanisms might produce a Q10 {\it increase} under drought conditions?
+This is interesting to consider: what mechanisms might produce a Q10 {\it increase} under drought conditions?
 This is opposite to what is usually observed (e.g.
 Jassal et al.
-2008), but the field is rife with contradictory results (von Lützow and K?gel-Knabner 2009).
-We will make this paragraph more specific in all these areas.
-{\bf See new lines TODO.}
+2008), but the field is rife with contradictory results (von Lützow and Kögel-Knabner 2009).
+We have made this paragraph more specific in this area.
+{\bf See new lines LINES.}
 
@@
 (2014), which focused specifically on high latitudes, would be a useful addition.
-We don't know of any studies examining the N limitation of deep soils at this site, but Allison et al.
-(2008, 2011) used fertilization and incubation experiments to examine how Alaskan boreal biota (plants and microbes) responded to N addition.
+We don't know of any studies examining the N limitation of deep soils at this site.
 We will discuss these results in our revision, along with other studies examining the relationship between N availability and C mineralization.
-{\bf See new lines TODO.}
+{\bf See new lines LINES.}
 

### 2016-11-19 manuscript/cpcrw_manuscript.Rmd
Reworking section 4.2

@@
 There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}.
+In addition, conditions such as drought can change the amount and quality of DOC available to microbes (TODO), but we observed no DOC changes between treatments here.
 Unlike surface soils, deep active layer soils, which store large quantities of soil C {Mueller, 2015 #4435}, are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils may be older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}.
-Thus, it is not surprising that these more stable C compounds would be metabolized by processes that have been reported to be less temperature-sensitive.
+Thus, it may not be surprising that these more stable C compounds would be metabolized by processes that have been reported to be less temperature-sensitive.
 
@@
 Somewhat unexpectedly, percent soil N was very significantly and positively correlated with both CO~2~ and CH~4~ fluxes (**Tables 2 and 3**).
-N interacts with microbial respiration via a number of complex, interactive, and still unclear mechanisms {Luo, 2006 #2443}, including reductions in belowground plant allocation, shifts in energy source or population of the saprotrophic community {Saiya-Cork, 2002 #3595} that leave it less capable of decomposing recalcitrant compounds, and perhaps abiotic stabilization mechanisms {Janssens, 2010 #2518}.
+Nitrogen interacts with microbial respiration via a number of complex, interactive, and still unclear mechanisms {Luo, 2006 #2443}, including reductions in belowground plant allocation, shifts in energy source or population of the saprotrophic community {Saiya-Cork, 2002 #3595} that leave it less capable of decomposing recalcitrant compounds, and perhaps abiotic stabilization mechanisms {Janssens, 2010 #2518}.
 Meta-analyses have generally shown negative to neutral effects of N deposition on microbial biomass {Treseder, 2008 #4421} and respiration {Ramirez, 2012 #4422}, and total soil respiration across ecosystems and biomes {Janssens, 2010 #2518}{Zhou, 2014 #3845}.
-These effect are likely due to several one or more mechanisms involving soil pH, ligninase enzymes, and phenol oxidase activity {Luo, 2006 #2443}.
-
-discuss - C/N was not a good predictor (they're highly correlated here) ref Schadel 2014
-
-These conclusions have generally come from examining the effect of anthropogenic N deposition or experimental N amendments.
-Individual studies, including this one, have however observed positive correlations between ambient soil N and microbial respiration.
+These effect are likely due to several one or more mechanisms involving soil pH, ligninase enzymes, and phenol oxidase activity {Luo, 2006 #2443}, and incubation results examining N effects can be highly variable {Lavoie, 2011 #3060}{Sistla, 2012 #4481}.
+Some studies have however observed positive correlations between ambient soil N and microbial respiration.
 For example, Weiss et al.
 {, 2015 #4317} found CO~2~ production from Siberian Yedoma permafrost samples to be correlated with both percent C and N, consistent with our active-layer results (**Table 2**).
+
+The C:N ratio was not a good predictor of GHG fluxes in this study, although this ratio is often found to be a good predictor in meta-analyses {Sistla, 2012 #4481}{Schädel, 2014 #4472}.
 *In situ* respiration rates have also been shown to be negatively correlated with C:N at large spatial scales {Allaire, 2012 #4423}.
-Percent C and N both varied widely in our soil cores, even though they were collected within tens of meters of each other, suggesting that active-layer SOC response to temperature and moisture may also be highly spatially variable, even in a visually uniform boreal forest.
-Spatially explicit analyses of soil properties, temperatures {Bond-Lamberty, 2005 #1072}, and respiration {Allaire, 2012 #4423} are likely necessary to accurately constrain and predict soil fluxes in this ecosystem.
+Percent C and N both varied widely in our soil cores, and were highly correlated with each other (**Table 1**), even though the cores were collected within tens of meters of each other.
+This suggests that active-layer SOC response to temperature and moisture may also be highly spatially variable, even in a mixed-species boreal forest that we expected, *a priori*, to provide spatial variation in litter and SOC quality {Fierer, 2005 #2598}.
+Spatially explicit analyses of soil biochemistry, temperatures {Bond-Lamberty, 2005 #1072}, and respiration {Allaire, 2012 #4423} are likely necessary to accurately constrain and predict soil fluxes in this ecosystem.
 

### 2016-11-20 manuscript/response_to_reviewer_comments.tex
Posted data, added DOI
Tweaks for concision and clarity

@@
 
-Known area of model weakness {Bouskill, 2014 #4480}.
-
 \medskip
@@
 
-This is an interesting suggestion, and an analysis we are happy to include.
+This is an interesting suggestion.
 We do not observe any evidence for changes in CO2 moisture sensitivity with time, and weak changes in CO2 temperature sensitivity; CH4 emissions show a weak decline in moisture sensitivity with time.
@@
 
-We will! {\bf See new lines TODO.}
+Done. {\bf See new lines LINES.}
 
@@
 
-We thought we were already doing so in using the {\tt RColorBrewer} package, not the default palette of {\tt ggplot2}, but have shifted to using a color-blind friendly palette from \url{http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette} in a figures.
+We thought we were already doing so in using the {\tt RColorBrewer} package, not the default palette of {\tt ggplot2}, but have shifted to using a color-blind friendly palette from \url{http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette} in all figures.
 

### 2016-11-20 manuscript/cpcrw_manuscript.Rmd
Posted data, added DOI
Tweaks for concision and clarity

@@
 Rapid climatic changes, rising air temperatures, and increased fires are expected to drive permafrost degradation and alter soil carbon (C) cycling in many high-latitude ecosystems.
-How these soils will respond to changes in their temperature, moisture, and overlying vegetation is highly uncertain, but critical to understand given the large soil C stocks in these regions.
+How these soils will respond to changes in their temperature, moisture, and overlying vegetation is uncertain, but critical to understand given the large soil C stocks in these regions.
 We used a laboratory experiment to examine how temperature and moisture control CO~2~ and CH~4~ emissions from mineral soils sampled from the bottom of the annual active layer, i.e.
@@
 Cumulative CH~4~ production was unaffected by any treatment.
-These results suggest that deep active-layer soils may be quite sensitive to changes in soil moisture under aerobic conditions, a critical factor as discontinuous permafrost thaws in interior Alaska.
+These results suggest that deep active-layer soils may be sensitive to changes in soil moisture under aerobic conditions, a critical factor as discontinuous permafrost thaws in interior Alaska.
 Deep but unfrozen high-latitude soils have been shown to be strongly affected by long-term experimental warming, and these results provide insight into their future dynamics and feedback potential with future climate change.
@@
 High latitude ecosystems are being subjected to rapid changes in climate {IPCC, 2013 #3916} and increases in fire frequency and intensity {Kasischke, 2010 #2631}, notably in northwestern North America and Alaska {Hinzman, 2005 #4315}{Ju, 2016 #4320}.
-This will have a wide variety of ecosystem effects {Alexander, 2016 #4414}: in particular, rising temperatures and increasing fire will likely result in permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372}{Genet, 2013 #4412}{Helbig, 2016 #4413} and changes in soil temperature, with subsequent hydrology changes that will influence soil greenhouse gas (GHG) fluxes to the atmosphere {Schädel, 2014 #4472}.
-Such fluxes are a large component of the global C cycle and, because of the high C stocks of northern soils {Tarnocai, 2009 #2231}, could result in a significant and positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}{Schaefer, 2014 #4117}.
+This will have a wide variety of ecosystem effects {Alexander, 2016 #4414}: in particular, rising temperatures and increasing fire will likely result in changes in soil temperature and permafrost degradation {Pastick, 2015 #4311}{Zhang, 2015 #4372}{Genet, 2013 #4412}{Helbig, 2016 #4413}, with subsequent hydrology changes that will influence soil greenhouse gas (GHG) fluxes to the atmosphere {Schädel, 2014 #4472}.
+Such fluxes are a large component of the global C cycle and could result in a significant and positive climate feedback {Treat, 2015 #4217}{Koven, 2011 #3092}{Schaefer, 2014 #4117}.
 
 The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such any such feedback remain highly uncertain {Schuur, 2015 #4319}.
-While northern soils hold enormous quantities {Hugelius, 2014 #3978} of potentially mineralizable soil organic carbon (SOC), vegetation and succession dynamics (for example, thermal insulation by mosses) promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}{Turetsky, 2012 #3405}.
+While northern soils hold enormous quantities {Hugelius, 2014 #4162} of potentially mineralizable soil organic carbon (SOC), vegetation and succession dynamics (for example, thermal insulation by mosses) promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}{Turetsky, 2012 #3405}.
 Vegetation type also influences SOC quality and quantity, with microbial communities {Högberg, 2007 #4482}, soil respiration {Raich, 2000 #596}, and SOC  all linked to aboveground factors such as woody versus nonwoody stems, deciduous versus evergreen canopies, and the presence of nitrogen-fixing plants.
-
 A number of factors may however disrupt these these feedbacks between vegetation type, ground cover, permafrost, and SOC stability.
@@
 Conversely, increasing plant productivity in some regions can stimulate the decomposition of older SOC {Hartley, 2012 #3481}.
-Climate also drives fire regime changes, and ecosystem disruption is particularly likely after more-intense fire events {Johnstone, 2010 #2465}{Genet, 2013 #4412}.
-Even absent disturbance, the stability of SOC is highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}{Karhu, 2014 #4124}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.
+Climate also drives fire regime changes, and ecosystem disruption is particularly likely after intense fires {Johnstone, 2010 #2465}{Genet, 2013 #4412}.
+Even absent disturbance, the stability of SOC is highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}{Karhu, 2014 #4124}, and its protection (whether by organomineral sorption, chemical lability, or physical location) from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}{Schmidt, 2011 #3135}.
 
@@
 The anaerobic conditions common following permafrost thaw are expected to lower CO~2~ emissions but increase those of CH~4~ {Treat, 2015 #4217;Treat, 2014 #4213}, but emissions from aerobic soils will likely dominate the permafrost C feedback {Schädel, 2016 #4444}.
-Decadal warming and drying trends in Alaska {Bieniek, 2014 #4389} thus seem likely to increase GHG emissions from soils, and incubation experiments are critical to understand these dynamics {Elberling, 2013 #3720}.
+Decadal warming and drying trends in Alaska {Bieniek, 2014 #4389} thus seem likely to increase GHG emissions from soils, and laboratory incubation experiments are critical to understand these dynamics {Elberling, 2013 #3720}.
 
@@
 We also aimed to characterize the chemical and structural properties of these soils following a 100-day incubation at different temperatures, subjecting some cores to drying treatments.
-We hypothesized that (i) CO~2~ would be the dominant pathway for C loss in these largely aerobic soils; (ii) soils maintained at field moisture and high temperature would lose more C-CO~2~ than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity; and (iii) core CH~4~ fluxes would be sensitive only to temperature, as no anaerobic conditions were imposed on the cores.
+We hypothesized that (i) CO~2~ would be the dominant pathway for C loss in these largely aerobic soils; (ii) soils maintained at field moisture and high (20˚C) temperature would lose more C-CO~2~ than cores incubated at 4˚C, due to increased aerobic and anaerobic microbial activity; and (iii) core CH~4~ fluxes would be small and sensitive only to temperature, as no anaerobic conditions were imposed on the cores.
 
@@
 The forest was at least 90 years old {cf.
-\\Morishita, 2014 #4316} according to tree rings taken at the stem base of several of the largest white spruce.
+\\Morishita, 2014 #4316} according to tree cores taken at the base of several of the largest white spruce.
 The soil is characterized as a poorly-drained silt loam, and on average had ~20 cm of organic material over the mineral soil.
@@
 On 18 August 2015 cores were placed into one of two growth chambers (Conviron Control Systems BDW80, Winnipeg, Canada) maintained at 4 and 20 °C temperatures and 70% relative humidity and allowed to equilibrate for two weeks.
-Starting on 31 August 2015 we measured the cores' mass and GHG (CH~4~ and CO~2~) emissions four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation.
+Starting on 31 August 2015 we measured the cores' mass and GHG emissions four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation.
 Throughout the incubation, cores had a 200 µm mesh screen fit to the base and were mounted on porous ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action.
 The "drought" cores were mounted on dry plates, but not allowed to drop below 5% water content.
-After each flux measurement, cores received additional wetting from the top to maintain their water status at the desired level.
+After each flux measurement, cores received additional wetting from the top to maintain their desired water status.
 
@@
 Soil volume was calculated as the total core volume minus the volume of particles >2 mm diameter, with the latter determined by water displacement.
-Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic C.
+Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic carbon (DOC).
 Samples were combusted in an oxygen atmosphere at 1350 °C, and measured for gaseous C and N using a Elementar Vario Max CNS.
@@
 Each core’s respiration flux (*F*) was then calculated as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where *V* is the core-specific system volume, *M* the core dry mass as determined at the end of the incubation, *P~a~* atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), *R* the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and *T* the chamber air temperature (K) at time of measurement.
-The final respiration rate was expressed on a dry soil mass basis (µg C g soil^-1^ day^-1^).
+The final respiration rate was expressed on a soil C basis (µg or ng C g C^-1^ day^-1^).
 
@@
 This experiment was run as an 'open experiment' {Bond-Lamberty, 2016 #4388} with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
-The summarized flux data backing the main results have been archived at [*DOI to be filled in*].
+The summarized flux data backing the main results have been archived under the Digital Object Identifier m9.figshare.4240436.v1.
 
@@
 CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^, with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^.
+These means conceal considerable variability over the course of the incubation (**Figures 2 and 3**).
 
-These means conceal considerable variability over the course of the incubation (**Figures 2 and 3**).
 In the linear mixed-effects model (AIC = `r pn(AIC(step_co2_lme), 1)`), instantaneous CO~2~ flux was positively correlated with incubation chamber temperature, core gravimetric water content, and percent soil N (all P < 0.05, and the latter two P < 0.001; **Table 2**).
@@
 The cumulative flux numbers above result in CO~2~ temperature sensitivity (Q~10~) values of `r pn(Q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(Q10_cumulative["Drought", "CO2"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CH~4~ were `r pn(Q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(Q10_cumulative["Drought", "CH4"], 1)`.
-Computing Q~10~ values based on fluxes normalized by water filled pore space changed these values only slightly: to `r pn(wfpsQ10_cumulative["Field moisture", "CO2"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CO2"], 1)` for CO~2~, respectively, and `r pn(wfpsQ10_cumulative["Field moisture", "CH4"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CH4"], 1)` for CH~4~.
+Computing Q~10~ values based on fluxes normalized by water filled pore space changed these values only slightly: to `r pn(wfpsQ10_cumulative["Field moisture", "CO2"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CO2"], 1)` for CO~2~, for the field moisture and drought treatments respectively, and `r pn(wfpsQ10_cumulative["Field moisture", "CH4"], 1)` and `r pn(wfpsQ10_cumulative["Drought", "CH4"], 1)` for CH~4~.
 
@@
 
-Rises in boreal air temperatures, and unpredictable precipitation changes, are changing fire disturbance regimes, warming and dry many soils, increasing vegetation stress, degrade permafrost and deepen the active layer {Schuur, 2015 #4319}, all with uncertain consequences for soil dynamics and GHG fluxes.
+Rises in boreal air temperatures, and unpredictable precipitation changes, will change fire disturbance regimes, warm and dry many soils, increase vegetation stress, degrade permafrost, and deepen the active layer {Schuur, 2015 #4319}, all with uncertain consequences for soil dynamics and GHG fluxes.
 In this laboratory experiment we found that CO~2~, but not CH~4~, fluxes from these oxic active-layer mineral soils were sensitive to temperature and, in particular, moisture.
@@
 {, 2014 #4316} quantified GHG fluxes at CPCRW and nearby forests, and found CO~2~ production to be correlated with both temperature and moisture in upland Cryosols, consistent with our results.
-They found however that CH~4~ uptake was driven by temperature only.
 Waldrop et al.
 {, 2010 #4380} incubated active-layer and permafrost soils from *Picea mariana* sites near Fairbanks, AK, observing aerobic Q~10~ values of 9.0 (active layer) and 2.3 (permafrost) from -5 to 5 °C, and flux rates of 0.001-0.10 µmol CH~4~ day^-1^ g^-1^ (~`r pn(0.001 * 16 / 0.012 / 1000, 3)`-`r pn(0.1 * 16 / 0.012 / 1000, 3)` ng C g C^-1^ day^-1^), and ~1-5 µg C-CO~2~ hr^-1^ g^-1^ (~`r pn(1 * 24 / 0.012, 0)`-`r pn(5 * 24 / 0.012, 0)` µg C g C^-1^ day^-1^), considerably higher than the CO~2~ rates observed here.
-During the first 100 days of an incubation of Fairbanks-area 0-10 cm mineral soils, Neff and Hooper {, 2002 #4390} observed fluxes of ~`r pn(0.2 * 1000 * 12/44, 0)`-`r pn(1.5 * 1000 * 12/44, 0)` µg C-CO~2~ g C^-1^ day^-1^, in line with the results here, while Wickland and Neff {, 2008 #1820} reported that temperature and moisture exhibited interactive and similar-magnitude effects on decomposition in *P.
+During the first 100 days of an incubation of Fairbanks-area 0-10 cm mineral soils, Neff and Hooper {, 2002 #4390} observed fluxes of ~`r pn(0.2 * 1000 * 12/44, 0)`-`r pn(1.5 * 1000 * 12/44, 0)` µg C-CO~2~ g C^-1^ day^-1^, in line with the results here, while Wickland and Neff {, 2008 #1820} reported that temperature and moisture exhibited interactive effects, of similar magnitude, on decomposition in *P.
 mariana* soils.
-Finally, Sierra et al.
-{, 2016 #4504} incubated homogenized organic soils from CPCRW and observed interactions between high temperatures, moisture, and oxygen concentration.
 
-A number of synthesis studies have produced robust results documenting the dynamics and C feedback potential of Arctic and boreal soils; comparing to these results is useful because although the response of soil biota to stresses such as drought tends to differ between soil types, organisms, and vegetation, it is often broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}.
+A number of synthesis studies have documented dynamics and C feedback potential of Arctic and boreal soils more generally; comparing to these results is useful because although the response of soil biota to stresses such as drought tends to differ between soil types, organisms, and vegetation, it is often broadly similar across biomes and climatic conditions {Manzoni, 2012 #3545}.
 Using two meta-analyses of aerobic and anaerobic permafrost soil incubations, Schädel et al.
@@
 The importance of this result depends, in part, on the spatial extent and intensity of precipitation changes across the boreal and Arctic this century.
-There is a detectable anthropogenic influence in high latitude precipitation {Wan, 2015 #4478}, and its changes in which exhibit high spatial variability: drier and warmer conditions in boreal Eurasia {Buermann, 2014 #3992}, for example, but growing season length increases in interior Alaska with no increase in precipitation {Wendler, 2009 #4477}.
+There is a detectable anthropogenic influence in high latitude precipitation changes {Wan, 2015 #4478}, but these changes are inconsistent: drier and warmer conditions in boreal Eurasia {Buermann, 2014 #3992}, for example, but growing season length increases in interior Alaska with no increase in precipitation {Wendler, 2009 #4477}.
 This spatial variability will interact with permafrost thaw dynamics to produce a complex patchwork of soil moisture changes {Zhang, 2012 #4479}{Watts, 2012 #4505}.
-This high uncertainty in this area makes it all the more important to understand the interactive effects of soil moisture and temperature on decomposition and GHG emissions {Sierra, 2015 #4203}.
-
-**4.1 Temperature versus moisture sensitivity for cumulative emissions**
-
-In contrast to the controls on instantaneous fluxes, the controls on the cumulative fluxes (**Figure 4**) integrate the entire incubation time and are important to understand.
-Our results suggest that moisture limitation could exert a large effect on CO~2~ production for deep active-layer soils: "drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
-This effect was highly significant, and suggests that moisture limitations could exert a significant constraint on deep active-layer soils as they slowly warm.
-Such moisture constraints are thought to be already exerting effects on vegetation and soil fluxes at large scales {Ju, 2016 #4320}{Bond-Lamberty, 2012 #3480}, but our understanding of the interactive effects involved is poor.
-
-In contrast, the temperature sensitivities observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture), although not unprecedented in comparison to a wide range of other laboratory soil incubations {Hamdi, 2013 #3664}.
-It is more typical to observe Q~10~ values of around 2 {Dutta, 2006 #1427}{Schädel, 2016 #4444}, although the temperature sensitivity of C release can change over time of incubation {Dutta, 2006 #1427} and vary between soil fractions cycling over different time horizons {Karhu, 2010 #2503}{Schädel, 2014 #4472}.
-Observed surface CO~2~ fluxes at this CPCRW site exhibited a Q~10~ of 5.1 ± 1.4 over a temperature range of 3.5-15 °C (personal communication, C.
-Anderson); these surface fluxes were measured over multiple months and include root respiration, however, confounding any direct comparison.
-It is also important to note that while increased temperature does not always drive C mineralization rates in forest mineral soils {Giardina, 2000 #2159}, it is linked with increases in soil moisture content can lead to changes in microbial community structure and GHG fluxes {Xue, 2016 #4361}.
-
-More surprisingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 °C than at 20 °C.
-There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}.
-In addition, conditions such as drought can change the amount and quality of DOC available to microbes (TODO), but we observed no DOC changes between treatments here.
-Unlike surface soils, deep active layer soils, which store large quantities of soil C {Mueller, 2015 #4435}, are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils may be older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}.
-Thus, it may not be surprising that these more stable C compounds would be metabolized by processes that have been reported to be less temperature-sensitive.
+The high uncertainty in this area makes it all the more important to understand the interactive effects of soil moisture and temperature on decomposition and GHG emissions {Sierra, 2015 #4203}.
 
@@
 
+**4.1 Temperature versus moisture sensitivity for cumulative emissions**
+
+The cumulative GHG fluxes (**Figure 4**) integrate the entire 100-day incubation, eliminating the day-to-day variablity of instantaneous fluxes and are thus more generalizable.
+Our results suggest that moisture limitation could exert a large effect on CO~2~ production for deep active-layer soils: "drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
+This effect was highly significant, and suggests that moisture limitations could exert a significant constraint on deep active-layer soils as they slowly warm.
+Such moisture constraints are thought to be already exerting effects on vegetation and soil fluxes at large scales {Ju, 2016 #4320}{Bond-Lamberty, 2012 #3480}, but our understanding of the interactive effects involved is poor.
+
+The Q~10~ values observed in this experiment were low (all less than 2.0, even when controlling for changes in soil moisture).
+Temperature sensitivities of ~2 are more typical {Dutta, 2006 #1427}{Schädel, 2016 #4444}, although the temperature sensitivity of C release can change over time of incubation {Dutta, 2006 #1427} and vary between soil fractions cycling over different time horizons {Karhu, 2010 #2503}{Schädel, 2014 #4472}.
+Observed surface CO~2~ fluxes at this CPCRW site exhibited a Q~10~ of 5.1 ± 1.4 over a temperature range of 3.5-15 °C (personal communication, C.
+Anderson); these surface fluxes were measured over multiple months and include root respiration, however, preventing any direct comparison.
+While increased temperature does not always drive C mineralization rates in forest mineral soils {Giardina, 2000 #2159}, it is linked with increases in soil moisture content can lead to changes in microbial community structure and GHG fluxes {Xue, 2016 #4361}.
+
+Interestingly, Q~10~ values were lower in the drought treatment cores, a mathematical consequence of the fact that drought restricted CO~2~ respiration more at 4 °C than at 20 °C.
+There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}.
+Conditions such as drought can change the amount and quality of DOC available to microbes (TODO), but we observed no DOC changes between treatments here.
+Deep active layer soils store large quantities of soil C {Mueller, 2015 #4435} but are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils is older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}.
+Thus, it may not be surprising that these more stable C compounds would be metabolized by processes that have been reported to be less temperature-sensitive.
+
 **4.2 Soil nitrogen**
@@
 
-The C:N ratio was not a good predictor of GHG fluxes in this study, although this ratio is often found to be a good predictor in meta-analyses {Sistla, 2012 #4481}{Schädel, 2014 #4472}.
+The C:N ratio was not a significant predictor of GHG fluxes in this study, although this ratio has been found to be important in meta-analyses {Sistla, 2012 #4481}{Schädel, 2014 #4472}.
 *In situ* respiration rates have also been shown to be negatively correlated with C:N at large spatial scales {Allaire, 2012 #4423}.
@@
 Soils isolated during incubation may, for example, underestimate temperature sensitivity of respiration {Podrebarac, 2016 #4365} or exhibit lag effects {Treat, 2015 #4217}.
+It should also be noted that our 100-day incubation was not long enough to observe slowly-cycling soil fractions, which may vary in their response to experimental manipulation {Karhu, 2010 #2503}.
 Nonetheless, the controlled environments of incubations provide an important way to elucidate the key mechanisms controlling GHG from high-latitude soils {Schuur, 2015 #4319}.
-It should also be noted that our 100-day incubation was not long enough to observe slowly-cycling soil fractions, which may vary in their response to experimental manipulation {Karhu, 2010 #2503}.
 
 The soils studied here were from an upland, mixed conifer-deciduous boreal forest, and care needs to be taken before drawing regional inferences, or about other ecosystem types.
-We focused on an experimental drought, rather than flooding, because of the well-drained nature of the field site: it is unlikely that the mid-slope forest we sampled in will ever suffer from thermokarst or excessive soil moisture, but too-dry conditions are a serious possibility in this low-precipitation ecosystem {Barber, 2000 #1320}.
+We focused on an experimental drought, rather than flooding, because of the well-drained nature of the field site: it is unlikely that the mid-slope forest we sampled in will ever suffer from thermokarst or excessive soil moisture, but too-dry conditions are a serious possibility in this relatively low-precipitation ecosystem {Barber, 2000 #1320}.
 

### 2016-11-23 R/5-analyze.R
Figure 3 units fix

@@ -217,12 +217,13 @@ fluxdata %>%
             samplenums = paste(samplenum, collapse = " ")) ->
   fluxdata_figsBC
 
-figsBC <- function(fd) {
+figsBC <- function(fd, multiplier = 1) {
   fd <- filter(fd, !is.na(incday), !is.na(flux_sd), !is.na(flux))
-  ggplot(fd, aes(incday, flux)) + 
+  ggplot(fd, aes(incday, flux * multiplier)) + 
     geom_point() + 
     facet_grid(Temperature ~ Treatment) + 
-    geom_errorbar(aes(ymin = flux - flux_sd, ymax = flux + flux_sd)) +
+    geom_errorbar(aes(ymin = flux * multiplier - flux_sd * multiplier,
+                      ymax = flux * multiplier + flux_sd * multiplier)) +
     xlab("Incubation day") + ylab(expression(µg~C~g~C^{-1}~day^{-1}))
 }
 
@@ -232,8 +233,9 @@ fluxdata_figsBC %>%
   figureB
 fluxdata_figsBC %>%
   filter(Gas == "CH4") %>%
-  figsBC ->
+  figsBC(multiplier = 1000) ->
   figureC
+figureC <- figureC + ylab(expression(ng~C~g~C^{-1}~day^{-1}))
 
 save_plot("figureB", figureB)
 save_plot("figureC", figureC)

### 2016-11-23 manuscript/cpcrw_manuscript.Rmd
VLB and APS comments; fixed Table 1 and Figure 3

@@
 ---
-[//]: This Rmarkdown file generates the Biogeosciences manuscript
-[//]: Built using rmarkdown 1.1, RStudio 1.0.44, and R 3.3.1
-[//]: November 23, 2016
+<!---
+This Rmarkdown file generates the Biogeosciences manuscript
+Built using rmarkdown 1.1, RStudio 1.0.44, and R 3.3.1
+November 23, 2016
+-->
 
@@
 The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such any such feedback remain highly uncertain {Schuur, 2015 #4319}.
-While northern soils hold enormous quantities {Hugelius, 2014 #4162} of potentially mineralizable soil organic carbon (SOC), vegetation and succession dynamics (for example, thermal insulation by mosses) promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}{Turetsky, 2012 #3405}.
+While northern soils hold enormous quantities of potentially mineralizable soil organic carbon (SOC) {Hugelius, 2014 #4162}, vegetation and succession dynamics (for example, thermal insulation by mosses) promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}{Turetsky, 2012 #3405}.
 Vegetation type also influences SOC quality and quantity, with microbial communities {Högberg, 2007 #4482}, soil respiration {Raich, 2000 #596}, and SOC  all linked to aboveground factors such as woody versus nonwoody stems, deciduous versus evergreen canopies, and the presence of nitrogen-fixing plants.
@@
 Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic carbon (DOC).
-Samples were combusted in an oxygen atmosphere at 1350 °C, and measured for gaseous C and N using a Elementar Vario Max CNS.
-DOC was measured using a Shimadzu 5000 TOC Analyzer.
+Samples were combusted in an oxygen atmosphere at 1350 °C, and measured for gaseous C and N using an Elementar Vario Max CNS (Langenselbold, Germany).
+DOC was measured using a Shimadzu 5000 TOC Analyzer (Columbia, Maryland, USA).
 
@@
 This experiment was run as an 'open experiment' {Bond-Lamberty, 2016 #4388} with all analysis code, data (from raw instrument data to final summaries), diagnostics, etc., available at https://github.com/bpbond/cpcrw_incubation.
-The summarized flux data backing the main results have been archived under the Digital Object Identifier m9.figshare.4240436.v1.
+The summarized flux data backing the main results have been archived under the Digital Object Identifier 10.6084/m9.figshare.4240436.v1.
 
@@
 CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^ to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^, with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 2)` ng C g C^-1^ day^-1^.
-These means conceal considerable variability over the course of the incubation (**Figures 2 and 3**).
+These means conceal considerable variability over the course of the incubation (**Table 1, Figures 2 and 3**).
 
@@
 The interaction between water content and percent N was also highly significant (`r pclean(summary(step_co2_lme)$tTable["WC_gravimetric:N_percent", "p-value"])`), although cores with N > 0.2\% exhibited little relationship between water content and CO~2~ flux (data not shown).
-Instantaneous CH~4~ fluxes were positively correlated with percent N, while water content exhibited significant interactions with percent N and DOC as a predictor (**Table 3**).
+Instantaneous CH~4~ fluxes were positively correlated with percent N, while water content exhibited significant interactions with percent N and DOC as a predictor (**Table A1**).
 This model had little predictive power (AIC = `r pn(AIC(step_ch4_lme), 1)`), however, and neither temperature nor water content was a significant first-order predictor of CH~4~ fluxes.
@@
 
-The cumulative GHG fluxes (**Figure 4**) integrate the entire 100-day incubation, eliminating the day-to-day variablity of instantaneous fluxes and are thus more generalizable.
+The cumulative GHG fluxes (**Figure 4**) integrate the entire 100-day incubation, eliminating the day-to-day variability of instantaneous fluxes and are thus more generalizable.
 Our results suggest that moisture limitation could exert a large effect on CO~2~ production for deep active-layer soils: "drought" cores' cumulative production was `r pn(drought_effect["4", "Reduction"], 0)`% (4 °C) and `r pn(drought_effect["20", "Reduction"], 0)`% (20 °C) lower than the cores kept at field moisture.
@@
 There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}.
-Conditions such as drought can change the amount and quality of DOC available to microbes (TODO), but we observed no DOC changes between treatments here.
+Conditions such as drought can change the amount and quality of DOC available to microbes {e.g., \Lundquist, 1999 #4506}, but we observed no DOC changes between treatments here.
 Deep active layer soils store large quantities of soil C {Mueller, 2015 #4435} but are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils is older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}.
@@
 *In situ* respiration rates have also been shown to be negatively correlated with C:N at large spatial scales {Allaire, 2012 #4423}.
-Percent C and N both varied widely in our soil cores, and were highly correlated with each other (**Table 1**), even though the cores were collected within tens of meters of each other.
+Percent C and N both varied widely in our soil cores (**Table 1**), and were highly correlated with each other, even though the cores were collected within tens of meters of each other.
 This suggests that active-layer SOC response to temperature and moisture may also be highly spatially variable, even in a mixed-species boreal forest that we expected, *a priori*, to provide spatial variation in litter and SOC quality {Fierer, 2005 #2598}.
@@
 **Table 1.** Summary of dissolved organic carbon (DOC), percent C, percent N, and CO~2~ and CH~4~ fluxes by treatment.
-Each column summarizes (mean ± s.d.) 12 cores, combining two groups of N=6 at each incubation temperature, except for the pre-incubation values shown in the right-hand column (N=6).
+The "Field moisture" and "Drought" columns summarize (mean ± s.d.) 12 cores, combining two groups of N=6 at each incubation temperature, while the "Controlled drought" and "Pre-incubation" columns are N=6.
+
+<!---
+TODO: Table 1 comes in from script with units like "mg/kg", not "mg kg-1"
+This is the one change-by-hand thing in this Rmarkdown file  :(
+-->
 
@@
 knitr::kable(table2, digits = 3, 
-             col.names = c("Value", "SE", "DF", "T", "P"))
-```
-
-**Table 3.** Linear mixed-effects model parameters, testing effects of temperature (°C), gravimetric water content (unitless), soil N (%),and dissolved organic carbon (DOC, mg kg^-1^) on log-transformed, individual core CH~4~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^); a colon (":") indicates an interaction.
-Dependent variable has units of log(µg C g C^-1^ day^-1^).
-Columns include parameter value; standard error (SE); degrees of freedom (DF); T statistic; and P value.
-
-```{r table3, echo=FALSE}
-summary(step_co2_lme)$tTable %>% 
-  round(3) %>% 
-  as.data.frame() ->
-  table3
-table3$`p-value` <- pclean(table3$`p-value`, printP = FALSE)
-knitr::kable(table3, digits = 3, 
              col.names = c("Value", "SE", "DF", "T", "P"))
@@
 
-**Figure 4.** Cumulative C fluxes (mg g C^-1^) over the incubation, by gas (CO~2~ and CH~4~, top and bottom panels respectively) and treatment (columns).
+**Figure 4.** Cumulative mass-normalized C fluxes (mg g C^-1^) over the incubation, by gas (CO~2~ and CH~4~, top and bottom panels respectively), treatment (columns), and temperatures (x-axis, °C).
 Letters within a panel indicate significant differences based on Tukey's HSD.
@@
 ```
+
+**Table A1.** Linear mixed-effects model parameters, testing effects of temperature (°C), gravimetric water content (unitless), soil N (%),and dissolved organic carbon (DOC, mg kg^-1^) on log-transformed, individual core CH~4~ fluxes (+`r FLUX_ADDITION` µg C g C^-1^ day^-1^); a colon (":") indicates an interaction.
+Dependent variable has units of log(µg C g C^-1^ day^-1^).
+Columns include parameter value; standard error (SE); degrees of freedom (DF); T statistic; and P value.
+
+```{r table3, echo=FALSE}
+summary(step_co2_lme)$tTable %>% 
+  round(3) %>% 
+  as.data.frame() ->
+  table3
+table3$`p-value` <- pclean(table3$`p-value`, printP = FALSE)
+knitr::kable(table3, digits = 3, 
+             col.names = c("Value", "SE", "DF", "T", "P"))
+```

### 2016-11-23 R/5-analyze.R

@@ -529,7 +529,7 @@ cndata %>%
          `N (%)` = N_percent,
          `BD (g/cm3)` = BD,
          `CO2 (µg C/g C/day)` = CO2_µgC_gC_day1, 
-         `CH4 (µg C/g C/day)` = CH4_µgC_gC_day1) ->
+         `CH4 (ng C/g C/day)` = CH4_µgC_gC_day1) ->
   table1_data
 
 # Compute the mean and s.d. for each treatment
@@ -559,6 +559,10 @@ table1_means %>%
          entry = ifelse(finite, paste(value, "±", value_sd), "-")) %>%
   dcast(Variable ~ Treatment, value.var = "entry") -> 
   table1
+
+# I'm not sure where this number came from (bulk density is not in uga_soil_samples.csv)
+# but Peyton added it to Table 1. Assume it's good
+table1[which(table1$Variable == "BD (g/cm3)"), "Pre-incubation"] <- "1.13 ± 0.29"
 
 printlog("All done with", SCRIPTNAME)
 closelog()

### 2016-11-23 manuscript/response_to_reviewer_comments.tex
Final changes, esp. line numbers in response doc

@@
 (2010), and Wickland and Neff (2008).
-{\bf See new lines LINES.}
+{\bf See new lines 308-337, 368-372, 401-408, 427-432.}
 
@@
 
-We have supplemented this section with a better comparison to relevant literature, for example, Lavoie et al. (2011), Sistla et al. (2012), and Bouskill et al. (2014). {\bf See new lines LINES.}
+We have supplemented this section with a better comparison to relevant literature, for example, Lavoie et al. (2011), Sistla et al. (2012), and Bouskill et al. (2014). {\bf See new lines 391-408.}
 
@@
 2005).
-{\bf See new lines LINES.}
+{\bf See new lines 414-420.}
 
@@
 This is now discussed in the text.
-{\bf See new lines LINES.}
+{\bf See new lines 260-261, 366-371.}
 
@@
 observable anthropogenic influences on high-latitude precipitation, drier and warmer conditions in boreal Eurasia, and growing season length increases in interior Alaska with no increase in precipitation.
-{\bf See new lines LINES.}
+{\bf See new lines 326-337.}
 
@@
 We now better describe why deep active-layer soils, such as those studied here, and important and distinctive relative to permafrost or shallow active layer soils.
-{\bf See new lines LINES.}
+{\bf See new lines 87-98.}
 
@@
 
-Vegetation and ecosystem type is a significant factor that is not well explored here, we agree. We have added some points about this in the introduction and discussion. {\bf See new lines LINES.}
+Vegetation and ecosystem type is a significant factor that is not well explored here, we agree.
+We have added some points about this in the introduction and discussion.
+{\bf See new lines 51-63, 433-439.}
 
@@
 
-This has been clarified. {\bf See new lines LINES.}
+This has been clarified. {\bf See new lines 264-266.}
 
@@
 Second, we have restructured and improved the subsequent comparison section, discussing a variety of studies suggested by all the reviewers to better put our results in the context of previous work.
-{\bf See new lines LINES.}
+{\bf See new lines 295-307 for nearby studies, and expanded context in lines 308-337, 368-372, 401-408, 427-432.}
 
@@
 
-This will be fixed. {\bf See new lines LINES.}
+This will be fixed. {\bf See new line 298.}
 
@@
 We also cite and discuss a variety of other studies.
-{\bf See new lines LINES.}
+{\bf See new lines 46, 81, 313, 368-371, 411.}
 
@@
 
-Thanks for the useful suggestion; this point is now brought out more clearly. {\bf See new lines LINES.}
+Thanks for the useful suggestion; this point is now brought out more clearly. {\bf See new lines 333-337, 356-365, 454-458.}
 
@@
 We have diminished the emphasis on defending this finding, and instead try to place it in the context of soil characteristics in this mixed-species boreal forest, SOC quality, etc.
-{\bf See new lines LINES.}
+{\bf See new lines 366-390.}
 
@@
 We have made this paragraph more specific in this area.
-{\bf See new lines LINES.}
+{\bf See new lines 378-390.}
 
@@
 We don't know of any studies examining the N limitation of deep soils at this site.
-We will discuss these results in our revision, along with other studies examining the relationship between N availability and C mineralization.
-{\bf See new lines LINES.}
+We now discuss these results in our revision, along with other studies examining the relationship between N availability and C mineralization.
+{\bf See new lines 392-408.}
 
@@
 We have clarified this in the text.
-{\bf See new lines LINES.}
+{\bf See new lines 425-426.}
 
@@
 (2016) in depth, particularly their findings of higher aerobic than anaerobic respiration; respiration dominance of CO2 versus CH4; and Q10 values.
-{\bf See new lines LINES.}
+{\bf See new lines 46, 81, 313, 368-371, 411.}
 
@@
 All these references are now cited throughout the manuscript.
-{\bf See new lines LINES.}
+{\bf See new lines 308-337, 368-372, 401-408, 427-432.}
 
@@
 
-Fixed. {\bf See new lines LINES.}
+Fixed. {\bf See new line 33.}
 
@@
 We now better describe why deep active-layer soils, such as those studied here, are important and distinctive relative to permafrost or shallow active layer soils.
-{\bf See new lines LINES.}
+{\bf See new lines 87-98.}
 
@@
 
-Thank you. We appreciate the useful suggestions, and now provide these additional details in our revised manuscript. {\bf See new lines LINES.}
+Thank you. We appreciate the useful suggestions, and now provide these additional details in our revised manuscript. {\bf See new lines 257-270.}
 
@@
 We now discuss this issue, referencing previous work such as Schädel et al (2014).
-{\bf See new lines LINES.}
+{\bf See new lines 409-420.}
 
@@
 comment 2.2 above) and other publications (comment 2.4 above).
-{\bf See new lines LINES.}
+{\bf See new lines 350-354.}
 
@@
 
-We have added caveats, noting in particular the useful but incremental nature of this study. {\bf See new lines LINES.}
+We have added caveats, noting in particular the useful but incremental nature of this study. {\bf See new lines 449-460.}
 
@@
 We agree that rebalancing the manuscript, focusing a bit more on controls on cumulative emissions and a bit less on the instantaneous fluxes, would strengthen it.
-Accordingly, we now more fully explore controls on the cumulative emissions, and have moved ...
-{\bf See new lines TODO.}
+Accordingly, we now more fully explore controls on the cumulative emissions, and have moved the table summarizing the instantaneous CH4 flux model, to an appendix.
+{\bf See new lines 826-832.}
 
@@
 (2008).
-{\bf See new lines LINES.}
+{\bf See new lines 46, 81, 313, 368-371, 411, and line 306 for Wickland.}
 
@@
 
-We will consider these suggestions with respect to length and supplemental materials; results dealing with cumulative emissions are now included in the abstract. {\bf See new lines TODO.}
+We have moved one table to supplementary material (see response to comment 3.2 above), and now summarize cumulative emissions results in the abstract. {\bf See new lines 27-31.}
 
@@
 
-These points have been clarified, except for the last, as we feel it's already clear and unambiguous. {\bf See new lines LINES.}
+These points have been clarified, except for the last, as we feel it's already clear and unambiguous. {\bf See new lines 20-36.}
 
@@
 
-This statement has been reworded to remove the comparison. {\bf See new lines LINES.}
+This statement has been reworded to remove the comparison. {\bf See new lines 31-33.}
 
@@
 (2014), and the Olefeldt sentence clarified.
-{\bf See new lines LINES.}
+{\bf See new lines 52 and 74.}
 
@@
 We have expanded on this point, pointing out that such variability originates from factors such as differences in soil type, antecedent conditions, phase changes, experimental protocols, etc.
-{\bf See new lines LINES.}
+{\bf See new lines 75-77.}
 
@@
 every timepoint shown in Figure 1; this has been clarified.
-{\bf See new lines LINES.}
+{\bf See new lines 164-165.}
 
@@
 
-Done. {\bf See new lines LINES.}
+Done. {\bf See new lines 234-235.}
 
@@
 
-These have all been fixed. {\bf See new lines LINES.}
+These have all been fixed. {\bf See new lines 252-270.}
 
@@
 
-We now better integrate this point, mentioning it in the introduction and clarifying its relationship to the study goals. {\bf See new lines LINES.}
+We now better integrate this point, mentioning it in the introduction and clarifying its relationship to the study goals. {\bf See new lines 58-62, 291, 310-315.}
 
@@
 
-This has been fixed. {\bf See new lines LINES.}
+This has been fixed. {\bf See new line 298.}
 
@@
 
-Upland Cryosols; we have clarified this. {\bf See new lines LINES.}
+Upland Cryosols; we have clarified this. {\bf See new line 298.}
 
@@
 
-Thanks; we have done so. {\bf See new lines LINES.}
+Thanks; we have done so. {\bf See new line 321.}
 
@@
 
-Thanks for the Lee et al. reference, which we had not considered (see our response to 3.7 above) but is now cited. We have also reworked and tightened section 4.2. {\bf See new lines LINES.}
+Thanks for the Lee et al. reference, which we had not considered (see our response to 3.7 above) but is now cited. We have also reworked and tightened section 4.2. {\bf See new lines 354, 391-420.}
 
@@
 
-This awkward language has been removed, and a note about lag effects added. {\bf See new lines LINES.}
+This awkward language has been removed, and a note about lag effects added. {\bf See new lines 425-429.}
 
@@
 
-We have reworded this. {\bf See new lines LINES.}
+We have reworded this. {\bf See new lines 440-447.}
 
@@
 
-Moisture adjustment was done after every mass measurement, i.e. every timepoint shown in Figure 1. This has been clarified. {\bf See new lines LINES.}
+Moisture adjustment was done after every mass measurement, i.e. every timepoint shown in Figure 1. This has been clarified. {\bf See new lines 164-165.}
 

### 2016-11-23 manuscript/cpcrw_manuscript.Rmd
Final changes, esp. line numbers in response doc

@@
 There is evidence that climate warming changes the microbial decay dynamics of soil organic C compounds generally considered to be stable {Frey, 2013 #4062}{Bond-Lamberty, 2016 #4371}.
-Conditions such as drought can change the amount and quality of DOC available to microbes {e.g., \Lundquist, 1999 #4506}, but we observed no DOC changes between treatments here.
+Conditions such as drought can change the amount and quality of DOC available to microbes {\Lundquist, 1999 #4506}, but we observed no DOC changes between treatments here.
 Deep active layer soils store large quantities of soil C {Mueller, 2015 #4435} but are not subject to abundant inputs of fresh C from vegetation, so the starting quality of the native soil C in active layer soils is older, more microbially processed, and dominated by more stable "heavy" organic C {Karlsson, 2011 #4436}.
@@
 
-**Table 1.** Summary of dissolved organic carbon (DOC), percent C, percent N, and CO~2~ and CH~4~ fluxes by treatment.
+**Table 1.** Summary of dissolved organic carbon (DOC), percent C, percent N, bulk density (BD), and CO~2~ and CH~4~ fluxes by treatment.
 The "Field moisture" and "Drought" columns summarize (mean ± s.d.) 12 cores, combining two groups of N=6 at each incubation temperature, while the "Controlled drought" and "Pre-incubation" columns are N=6.
